### 2019-07-23 chapters/3-background.tex
Initial commit

\chapter{Background}\label{chap:background}
Explain the general context of your work.
Explain mathematical background required and introduce notation.
\input{figures/background/algBackpropagation}
\input{figures/background/figTikz}

\section{JavaScript}
\subsection{TypeCoercion}
\subsection{NPM Packages}

\section{TypeScript}
\subsection{Declaration Files}
\subsection{Definitely Typedd}

\section{Jalangi}

### 2019-07-23 chapters/4-approach.tex
Initial commit

\chapter{Approach}\label{chap:approach}
The approach starts with the problem definition and continues with what you have done. Try to give an intuition first and describe everything with words and then be more formal like `Let $g$ be ...'.

Strive to make the following items crystal clear to your readers.
\begin{itemize}
\item What is the problem you are treating?
\item What are the research questions you want to answer in this work?
\item What are your own contributions and what is work of others? If
  you are working in a team, it is important to demarcate against the
  contributions of others in your team.
\end{itemize}

Describe each major technical problem and how it was solved.

\section{Problem Definition}
\todo{JavaScript popularity has increased dramatically over the last years. TypeScript has evolved naturally as the professional way of developing code in JavaScript, providing the robustness that ervery compiled language has. Generating the declaration files manually is very time consuming and error prone.}

\todo{NodeJS, frameworks like React, Angular, Vue. The migation to Single Page Apps and Serverless, etc.}

\todo{Info from Stackoverflow and Github}

\todo{Mention DefinitelyTyped repository. Explicar claramente que el objetivo de esta tesis es generar }

\todo{El objetivo es solucionar el problema de definitely typed!!! Por eso se usa como objetivo que el input sea un node module}

\section{TypeScript Declaration Files Generation Method}
\todo{Explain overview process for generating declaration file.  Explain that the focus of this thesis was to focus on the runtime information and use the symbolic executor as a 'refinement' of the test cases. Use 'public' document of the Master project}

This Thesis introduces a \textit{TypeScript Declaration Files Generation Method}, which generates a valid TypeScript Declaration File for a specific NPM package, as explained in Fig. \ref{fig:tsd_generation_method_block_diagram}.

\input{figures/approach/typescriptDeclarationFilesGenerationMethodBlockDiagram}

Following the goal of easing and automating the generation of TypeScript Declaration Files for existings JS Libraries, the method receives the name of the JS Library published to the NPM Registry as an input argument and returns the corresponding TypeScript declaration file.

The method consists of a continuous refinement process. It generates a TypeScript Declaration File based on data flow and type information gathered at run-time from a code base that executes the JS Library. A Symbolic Execution Engine expands the code base in every iteration by generating new test cases that explore different execution paths.

The file generated after each iteration is a valid and fully functional declaration file, suitable for usage in development environments. It contains no errors, matches the structure of the corresponding JS Library and could be uploaded to the DefinitelyTyped Repository. It is therefore intended to be integrated within the build process of the JS Library.

\todo{Decir que para la tesis se decidio no inlcuir el feedback loop. Solo se hizo la parte de runtime, etc.}

\subsection{Overview}
As explained before, the DefinitelyTyped Repository is the official repository for finding declaration files for npm packages. Therefore, the \textit{TypeScript Declaration Files Generation Method} is intended to be used on existing, published npm packages. The generated output TypeScript declaration file is a valid file which can be used for development and uploaded to the DefinitelyTyped Repository.

The method uses an instrumentation framework to gather data flow information and type information at runtime. A second independent block uses the gathered information to generate a TypeScript Declaration File. The signatures obtained from the generated declaration file are used to generate pre and post conditions on a Symbolic Execution Engine in order to build different test cases that explore new execution paths. The code from the test cases is then added to the existing code base. The declaration file is generated again using an expanded code base.

A feedback loop refines iteratively the outcome by combining run-time information gathering and symbolic analysis: It performs Run-Time Analysis to generate a declaration file using run-time information and a Symbolic Execution Engine to refine and expand the code base being used to gather run-time information.

Running the TypeScript Declaration File Generation Tool with the same run-time information as input will always produce the same result. If new execution paths are covered by the test cases, new run-time information will be gathered and the declaration file will change accordingly. In order to obtain a sound analysis, this feedback loop would be iterated until there are no differences among the generated declaration files. However, by trading soundness for scalability, the overall procedure can be stopped either using a loop bound, a time bound or an heuristics mechanism. Exploring and investigating the proper stopping criteria is beyond the scope of this work.

Finally, the generated declaration files are fully functional and valid files. They can be used in a TypeScript project without further modifications

\subsection{Implementation}
\todo{Decir que es un framework. Decir que es tipo una pipeline donde todo se conecta con lo otro. Explicar que por mas que reciba un string como input}

\section{Initial code base}

\section{Run-time Information Gathering} \label{run-time-information-gathering}
\subsection{Instrumentation}
\subsubsection{Interactions}
\todo{inputValue}
\todo{getField}
\todo{putField}
\todo{methodCall}
\todo{usedAsArgument}
\todo{convertedTo}

\subsubsection{Wrapper Objects}
\todo{Proxy Objects}
\todo{Primitives}
\todo{Null}
\todo{Undefined}
\subsection{Output Format}
\todo{Explicar que es un JSON con todas las interactions. Explicar estructura de cómo se forma el JSON}
\subsection{Blacklisted Modules}
\todo{Incluir grafico con los blacklisted modules}

\subsection{Implementation}
\todo{NodeJS \& Docker. Mostrar cómo se invoca, etc.}

\section{TypeScript Declaration File Generation}
\todo{Tres tipos de modulos en TypeScript}
\todo{Traceids y que las signatures de las funciones no se unen, sino que se acumulan segun traceId. Obviamente las iguales no se escriben de vuelta.}
\todo{El nombre de las interfaces se crea con el nombre de la variable}
\todo{El tipo se extrae del valor en runtime.}
\subsection{Implementation}
\todo{TypeScript \& Docker. Mostrar cómo se invoca, etc.}

\section{Evaluation}
\subsection{Parsing}
\subsubsection{TypeScript Compiler API}
\subsection{Comparison}
\todo{Number of functions}
\todo{Number of parameters}
\todo{Interfaces}
\todo{methodCall}
\todo{usedAsArgument}
\todo{Syntax \& Semantic Errors}

### 2019-07-23 chapters/5-experiments.tex
Initial commit

\chapter{Experiments}\label{chap:experiments}

Experiments often include benchmarks.
You need to describe you benchmarking setup: parameters of the machine used (architecture, memory, etc), parameters of the software used (OS version, libraries used and their version), the benchmark programs.
The programs need not be part of the thesis, but they should be described and available separately.

If you measure run times and evaluate your findings statistically, then there is great potential for errors.
Consult SIGPLAN's checklist for empirical evaluations \url{http://www.sigplan.org/Resources/EmpiricalEvaluation/} for advice.

\input{figures/experiments/figExample}
\input{figures/experiments/tabAccuracy}

\section{Definitely Typed Modules}

\section{Code Examples Extraction}
\subsection{Repositories URL}
\subsection{Readme Files}
\subsection{Code Examples}

\section{Code Execution}

\section{Run-time Information Gathering}

\section{Declaration Files Generation}

\section{Evaluation}

\section{JS Operators Usage}

### 2019-07-23 chapters/6-conclusions.tex
Initial commit

\chapter{Conclusion}\label{chap:conclusion}

Brief wrapup of the achievements of this thesis.

Directions for future work (if applicable).

\section{Future Work}

\subsection{Type Inference}
\todo{Deberia ser hecho en otro modulo que no sea el runtime information. El runtime information deberia exponer unicamente los datos que le hacen falta al otro para inferir los tipos.}
\subsubsection{Variable Names}
\todo{Se puede extraer info del nombre de las variables y del nombre de los metodos, etc. La developer's intention.}

\todo{Tambien se puede identificar patterns de programacion donde siempre los tipos son los mismos.}

\subsubsection{Operators}
\todo{Explicar el tema de la type inference con los operators}

\subsection{Uploading Results to Definitely Typed Repository}

### 2019-07-23 figures/approach/typescriptDeclarationFilesGenerationMethodBlockDiagram.tex
Initial commit

\begin{figure}[t]
\begin{centering}
    {\includegraphics[scale=0.9]{figures/approach/typescript_declaration_files_generation_method_block_diagram.pdf}}
    \caption[TypeScript Declaration Files Generation Method]{\textbf{TypeScript Declaration Files Generation Method} - Initial code base is retrieved from the npm package's repository. A valid TypeScript Declaration File is generated using run-time information. A Symbolic Execution Engine creates test cases based on the generated Declaration File and via a feedback loop enriches the code base until the stopping criteria is reached. The final TypeScript Declaration File gets returned.}
    \label{fig:tsd_generation_method_block_diagram}
\end{centering}
\end{figure}

### 2019-07-25 figures/approach/evaluation/evaluationDiagram.tex
Add evaluation diagram

\begin{figure}[t]
\begin{centering}
    {\includegraphics[scale=0.9]{figures/approach/evaluation/evaluation-diagram.pdf}}
    \caption[Evaluation against DefinitelyTyped Repository]{\textbf{Evaluation of generated declaration files against DefinitelyTyped Repository} - A parser transforms the generated declaration file and the equivalent file in the DefinitelyTyped repository into a JSON file using the TypeScript Compiler API \citep{typescript-compiler-api}. Comparison is then performed on the JSON files, i.e. not on the declaration files.}
    \label{fig:evaluation-diagram}
\end{centering}
\end{figure}

### 2019-07-25 figures/approach/typescript-declaration-files-generation-method/typescriptDeclarationFilesGenerationMethodBlockDiagram.tex
Add evaluation diagram

--- 2019-07-23 figures/approach/typescriptDeclarationFilesGenerationMethodBlockDiagram.tex
@@
 \begin{centering}
-    {\includegraphics[scale=0.9]{figures/approach/typescript_declaration_files_generation_method_block_diagram.pdf}}
+    {\includegraphics[scale=0.9]{figures/approach/typescript-declaration-files-generation-method/typescript_declaration_files_generation_method_block_diagram.pdf}}
     \caption[TypeScript Declaration Files Generation Method]{\textbf{TypeScript Declaration Files Generation Method} - Initial code base is retrieved from the npm package's repository.

### 2019-07-25 chapters/4-approach.tex
Add evaluation diagram and modifies text

@@
 
-\input{figures/approach/typescriptDeclarationFilesGenerationMethodBlockDiagram}
+\input{figures/approach/typescript-declaration-files-generation-method/typescriptDeclarationFilesGenerationMethodBlockDiagram.tex}
 
@@
 
+The feedback loop is not included in the implementation.
+Exploring new execution paths by creating new test cases will refine the declaration files, i.e.
+new types will be assigned to different variables, new interfaces will be detected, existing interfaces will be defined more accurately by detecting new properties, etc.
+From an incremental point of view, actually generating the declaration files from run-time information has to be developed first.
+A declaration file must exist first so that it can be refined refined afterwards.
+
 \todo{Decir que para la tesis se decidio no inlcuir el feedback loop. Solo se hizo la parte de runtime, etc.}
 
-\subsection{Overview}
+\subsection{Implementation}
 As explained before, the DefinitelyTyped Repository is the official repository for finding declaration files for npm packages.
@@
 
-The method uses an instrumentation framework to gather data flow information and type information at runtime.
+The method uses Jalangi as instrumentation framework to gather data flow information and type information at runtime.
 A second independent block uses the gathered information to generate a TypeScript Declaration File.
@@
 
-\subsection{Implementation}
 \todo{Decir que es un framework. Decir que es tipo una pipeline donde todo se conecta con lo otro. Explicar que por mas que reciba un string como input}
@@
 \section{Evaluation}
+\input{figures/approach/evaluation/evaluationDiagram.tex}
 \subsection{Parsing}

### 2019-07-25 chapters/4-approach.tex
Add blacklisted modules figure

@@
 
-This Thesis introduces a \textit{TypeScript Declaration Files Generation Method}, which generates a valid TypeScript Declaration File for a specific NPM package, as explained in Fig.
-\ref{fig:tsd_generation_method_block_diagram}.
+This Thesis introduces a \textit{TypeScript Declaration Files Generation Method}, which generates a valid TypeScript Declaration File for a specific NPM package, as explained in \figref{fig:tsd_generation_method_block_diagram}.
 
@@
 \subsection{Blacklisted Modules}
+\input{figures/approach/blacklisted-modules/blacklistedModules.tex}
 \todo{Incluir grafico con los blacklisted modules}

### 2019-07-25 figures/approach/blacklisted-modules/blacklistedModules.tex
Add blacklisted modules figure

\newsavebox{\mintedbox}

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.45\textwidth}
			\jscode{code/blacklisted-modules/foo.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[foo/index.js]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.45\textwidth}
			\jscode{code/blacklisted-modules/A.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[A/index.js]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.48\textwidth}
			\jscode{code/blacklisted-modules/D.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[D/index.js]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.45\textwidth}
			\jsoncode{code/blacklisted-modules/blacklist.json}
		\end{minipage}
	\end{lrbox}
	\subfloat[blacklist.json]{\usebox{\mintedbox}}

	\subfloat{\includegraphics[width=1\textwidth]{figures/approach/blacklisted-modules/blacklisted-modules.pdf}}

	\caption[Non-instrumented node modules]{\textbf{Non-instrumented node modules} - Dependencies tree for module \textit{foo}. Modules \textit{A} and \textit{B} are included in the blacklist. Module \textit{D} is instrumented anyway, even though its father is not. Node native modules, like \textit{fs} or \textit{http}, are excluded anyway by Jalangi.}
	\label{fig:blacklisted-modules}
\end{figure}

### 2019-07-25 code/blacklisted-modules/A.js

@@ -1,0 +1,4 @@
+const C = require("C");
+const D = require("D");
+
+// Code ...

### 2019-07-25 code/blacklisted-modules/D.js

@@ -1,0 +1,4 @@
+const fs = require("fs");
+const http = require("http");
+
+// Code ...

### 2019-07-25 code/blacklisted-modules/foo.js

@@ -1,0 +1,8 @@
+const A = require("A");
+const B = require("B");
+
+function foo() {
+	// Do something ...
+}
+
+module.exports = foo;

### 2019-07-25 figures/approach/typescript-declaration-files-generation-method/typescriptDeclarationFilesGenerationMethodBlockDiagram.tex
Add blacklisted modules figure
Add textwidth to size of figures

--- 2019-07-23 figures/approach/typescriptDeclarationFilesGenerationMethodBlockDiagram.tex
@@
-\begin{figure}[t]
+\begin{figure}[h]
 \begin{centering}
-    {\includegraphics[scale=0.9]{figures/approach/typescript_declaration_files_generation_method_block_diagram.pdf}}
+    {\includegraphics[width=0.9\textwidth]{figures/approach/typescript-declaration-files-generation-method/typescript_declaration_files_generation_method_block_diagram.pdf}}
     \caption[TypeScript Declaration Files Generation Method]{\textbf{TypeScript Declaration Files Generation Method} - Initial code base is retrieved from the npm package's repository.

### 2019-07-27 chapters/5-experiments.tex
Add figures with results of JS operators usage

@@
 \section{JS Operators Usage}
+\todo{Agregar que es una matriz simetrica.
+Que era para hacer un analisis mas global.
+Para algunos operators se puede inferir que el orden de los tipos sí afecta.
+Por ejemplo para instance of donde obviamente importa el orden.
+Pero para == por ejemplo generalmente se pone a == null y no null == a}.
+\input{figures/experiments/operators/relational-operators/relationalOperators.tex}
+\input{figures/experiments/operators/additive-operators/additiveOperators.tex}
+\input{figures/experiments/operators/multiplicative-operators/multiplicativeOperators.tex}
+\input{figures/experiments/operators/equality-operators/equalityOperators.tex}
+\input{figures/experiments/operators/conditional-operators/conditionalOperators.tex}

### 2019-07-27 figures/experiments/operators/additive-operators/additiveOperators.tex
Add figures with results of JS operators usage

\begin{figure}[h]
	\centering

	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			title=Operator ${+}$,
			width=0.45\textwidth,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/additive-operators/operator_+.dat};
		\end{axis}
	\end{tikzpicture}
	%
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${-}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/additive-operators/operator_-.dat};
		\end{axis}
	\end{tikzpicture}

	\centering
	\hspace{0.1\textwidth}
	\begin{tikzpicture}
		\begin{axis}[
			hide axis,
			scale only axis,
			height=0pt,
			width=0pt,
			colormap/hot,
			colorbar horizontal,
			point meta min=0,
			point meta max=100,
			colorbar style={
				title=\%,
				width=0.7\textwidth,
				xtick={0, 10, 20, 30, 40, ..., 100}
			}]
			\addplot [draw=none] coordinates {(0,0)};
		\end{axis}
	\end{tikzpicture}
	\caption[Additive operators]{\textbf{Type distribution for additive operators} - Both operators are mainly used with the \textit{number-number} combination. Operator $+$ is also used for pure string concatenation. However, due to JS Type Coercion, $+$ is also used for \textit{string} and \textit{number} concatenation}
\end{figure}

### 2019-07-27 figures/experiments/operators/conditional-operators/conditionalOperators.tex
Add figures with results of JS operators usage

\begin{figure}[h]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
			title=Conditional operators,
			width=1\textwidth,
			ybar=0pt,
			ymax=100,
			enlargelimits=0.15,
			legend style={at={(0.5,-0.2)}, anchor=north,legend columns=-1},
			symbolic x coords={number,string,undefined,object,null,boolean,function,array},
			xtick=data,
			nodes near coords, 
			nodes near coords align={vertical},
			x tick label style={rotate=90,anchor=east},
		]
		\addplot coordinates {
			(number, 7.91)
			(string, 1.11)
			(undefined, 5.50)
			(object, 1.09)
			(null, 1.06)
			(boolean, 82.63)
			(function, 0.43)
			(array, 0.28)
		};
		\end{axis}
	\end{tikzpicture}
	\caption[Conditional operators]{\textbf{Type distribution for conditional operators} - It includes all operators that trigger a condition check before branching: ${if-then-else}$, ${switch-case}$, ${while}$, ${for}$, ${||}$, ${\&\&}$, ${ ? : }$.
	}
\end{figure}

### 2019-07-27 figures/experiments/operators/equality-operators/equalityOperators.tex
Add figures with results of JS operators usage

\begin{figure}[h]
	\centering

	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			title=Operator ${==}$,
			width=0.45\textwidth,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/equality-operators/operator_==.dat};
		\end{axis}
	\end{tikzpicture}
	%
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${!=}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/equality-operators/operator_!=.dat};
		\end{axis}
	\end{tikzpicture}

	\centering
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${===}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/equality-operators/operator_===.dat};
		\end{axis}
	\end{tikzpicture}
	%
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${!==}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/equality-operators/operator_!==.dat};
		\end{axis}
	\end{tikzpicture}

	\centering
	\hspace{0.1\textwidth}
	\begin{tikzpicture}
		\begin{axis}[
			hide axis,
			scale only axis,
			height=0pt,
			width=0pt,
			colormap/hot,
			colorbar horizontal,
			point meta min=0,
			point meta max=100,
			colorbar style={
				title=\%,
				width=0.7\textwidth,
				xtick={0, 10, 20, 30, 40, ..., 100}
			}]
			\addplot [draw=none] coordinates {(0,0)};
		\end{axis}
	\end{tikzpicture}
	\caption[Relational operators]{
		\textbf{Type distribution for equality operators} - Operator ${==}$ is mainly used for number comparison. String comparison is performed by using ${===}$ and ${!==}$ operators. Comparison against \textit{null} is performed by using both ${!=}$ and ${!==}$ operators
	}
\end{figure}

### 2019-07-27 figures/experiments/operators/multiplicative-operators/multiplicativeOperators.tex
Add figures with results of JS operators usage

\begin{figure}[h]
	\centering

	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			title=Operator ${*}$,
			width=0.45\textwidth,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/multiplicative-operators/operator_star.dat};
		\end{axis}
	\end{tikzpicture}
	%
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${/}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/multiplicative-operators/operator_division.dat};
		\end{axis}
	\end{tikzpicture}

	\centering
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${\%}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/multiplicative-operators/operator_percentage.dat};
		\end{axis}
	\end{tikzpicture}

	\centering
	\hspace{0.1\textwidth}
	\begin{tikzpicture}
		\begin{axis}[
			hide axis,
			scale only axis,
			height=0pt,
			width=0pt,
			colormap/hot,
			colorbar horizontal,
			point meta min=0,
			point meta max=100,
			colorbar style={
				title=\%,
				width=0.7\textwidth,
				xtick={0, 10, 20, 30, 40, ..., 100}
			}]
			\addplot [draw=none] coordinates {(0,0)};
		\end{axis}
	\end{tikzpicture}
	\caption[Relational operators]{\textbf{Type distribution for multiplicative operators} - All multiplicative operators are mainly used with the \textit{number-number} combination}
\end{figure}

### 2019-07-27 figures/experiments/operators/relational-operators/relationalOperators.tex
Add figures with results of JS operators usage

\begin{figure}[h]
	\centering

	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			title=Operator ${<}$,
			width=0.45\textwidth,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/relational-operators/operator_<.dat};
		\end{axis}
	\end{tikzpicture}
	%
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${>}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/relational-operators/operator_>.dat};
		\end{axis}
	\end{tikzpicture}

	\centering
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${<=}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/relational-operators/operator_<=.dat};
		\end{axis}
	\end{tikzpicture}
	%
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${>=}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/relational-operators/operator_>=.dat};
		\end{axis}
	\end{tikzpicture}

	\centering
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${instanceof}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/relational-operators/operator_instanceof.dat};
		\end{axis}
	\end{tikzpicture}
	%
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${in}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/relational-operators/operator_in.dat};
		\end{axis}
	\end{tikzpicture}

	\centering
	\hspace{0.1\textwidth}
	\begin{tikzpicture}
		\begin{axis}[
			hide axis,
			scale only axis,
			height=0pt,
			width=0pt,
			colormap/hot,
			colorbar horizontal,
			point meta min=0,
			point meta max=100,
			colorbar style={
				title=\%,
				width=0.7\textwidth,
				xtick={0, 10, 20, 30, 40, ..., 100}
			}]
			\addplot [draw=none] coordinates {(0,0)};
		\end{axis}
	\end{tikzpicture}
	\caption[Relational operators]{\textbf{Type distribution for relational operators} - Operators ${<}$, ${>}$, ${<=}$, ${>=}$ are mainly used with the \textit{number-number} combination}
\end{figure}

### 2019-07-28 figures/background/stackoverflowSurvey-ProgrammingLanguagesEvolution.tex
Add figures with Stackoverflow Developer Survey Information

\begin{figure}[h]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			width=1\textwidth,
			ylabel=Percentage of developers,
			title={JS, TS, Python and PHP Popularity Evolution},
			enlargelimits=0.05,
			legend style={at={(0.5,-0.15)},
				anchor=north,legend columns=-1},
			xticklabels={
				2011,
				2012,
				2013,
				2014,
				2015,
				2016,
				2017,
				2018,
				2019
			},
			xtick={0,...,8},
			grid=major,
		]

		% JavaScript
		\addplot coordinates {
			(0,50.45)
			(1,50.08)
			(2,56.60)
			(3,58.90)
			(4,54.40)
			(5,55.4)
			(6,66.7)
			(7,71.5)
			(8,69.7)
		};

		% TypeScript
		\addplot coordinates {
			(0,0)
			(1,0)
			(2,0)
			(3,0)
			(4,0)
			(5,0)
			(6,11.3)
			(7,18.3)
			(8,23.5)
		};

		% Python
		\addplot coordinates {
			(0,20.46)
			(1,19.39)
			(2,21.9)
			(3,23.4)
			(4,23.8)
			(5,24.9)
			(6,27.6)
			(7,37.9)
			(8,39.4)
		};

		% PHP
		\addplot coordinates {
			(0,28.02)
			(1,30.84)
			(2,34.8)
			(3,28.9)
			(4,29.7)
			(5,25.9)
			(6,27.2)
			(7,31.4)
			(8,25.8)
		};
		\legend{JavaScript,TypeScript,Python,PHP}
		\end{axis}
	\end{tikzpicture}
	\caption[Conditional operators]{\textbf{JS, TS, Python and PHP Popularity Evolution 2011 - 2019} - According to Stackoverflow's Developer Survey, for the seventh year in a row, JavaScript is the most commonly used programming language. While JavaScript's popularity remained constant for the last 3 years, TypeScript's popularity is increasing every year. There is no data for TypeScript before 2017.}
\end{figure}

### 2019-07-29 figures/experiments/overallFunnel.tex
Add funnel with number of modules for each stage of the experiment

\definecolor{myellow}{RGB}{228,212,0}
\definecolor{mgreen}{RGB}{5,104,57}

\newcommand\funnel[3]{%
\pgfmathsetmacro\mwid{(0.3+\val*0.06)}
\pgfmathsetmacro\mradius{(\val*0.01 + 1)}
\pgfmathsetmacro\mheight{(\val*0.003 + 0.4)}
\pgfmathsetmacro\marc{\mwid-.4}
    \begin{scope}[%
        shift={(0,#1)}, 
        line width=.05pt, 
        %x=5mm, 
        %scale=1.\xi,
        yshift=\xi*4
        ]
    \draw[black,bottom color=#2, top color=#2] (-\mwid,0) -- (-\mwid+.4,-\mheight) arc (190:350:\marc cm and \mradius mm) -- (\mwid,0);
    \draw[black,fill=#3] (0,0) ellipse (\mwid cm and \mradius mm);
    \path (-\mwid,0) -- (-\mwid+.4,-\mheight) coordinate[midway] (a\xi);
    \end{scope}
}

\begin{figure}[h]
	\hspace*{-0.28\textwidth}
	\centering
	\begin{tikzpicture}
		\foreach \val
				[%
				count=\xi starting from 0, 
				evaluate=\xi as \shadecolor using int(25*\xi),
				evaluate=\xi as \coord using int(\xi-12)
				]
			in {
				4.05,
				9.37,
				24.02,
				37.49,
				71.19,
				82.50,
				100.00
			}{
				\funnel{\coord}{mgreen!\shadecolor !myellow}{mgreen!\shadecolor !myellow}
			}

		\node[left=0.02\textwidth of a0] {Generated Declaration Files};
		\node[right=0.07\textwidth of a0] {\textbf{244}};

		\node[left=0.02\textwidth of a1] {Run-time Information};
		\node[right=0.12\textwidth of a1] {\textbf{565}};

		\node[left=0.02\textwidth of a2] {Working Examples};
		\node[right=0.26\textwidth of a2] {\textbf{1448}};

		\node[left=0.02\textwidth of a3] {Code Examples};
		\node[right=0.38\textwidth of a3] {\textbf{2260}};

		\node[left=0.02\textwidth of a4] {README file};
		\node[right=0.70\textwidth of a4] {\textbf{4292}};

		\node[left=0.02\textwidth of a5] {Github Repository};
		\node[right=0.80\textwidth of a5] {\textbf{4974}};

		\node[align=right,left=0.02\textwidth of a6, text width=0.25\textwidth] {Definitely Typed Modules};
		\node[right=0.97\textwidth of a6] {\textbf{6029}};

	\end{tikzpicture}
	\caption[Number of analyzed modules for each stage of the experiment]{\textbf{Number of analyzed modules for each stage of the experiment} - A TypeScript Declaration File was generated for only 244 modules, out of 6029 modules in the DefinitelyTyped Repository. It was possible to gather valid run-time information for only 25\% of the modules for which a Code Example was extracted.}
\end{figure}

### 2019-08-05 chapters/3-background.tex
Add todo messages
Add figures with Stackoverflow Developer Survey Information
Add funnel with number of modules for each stage of the experiment
Add code example for run-time information gathering

@@
 Explain mathematical background required and introduce notation.
-\input{figures/background/algBackpropagation}
-\input{figures/background/figTikz}
 
 \section{JavaScript}
+\todo{Agregar img con data de la survey de stackoverflow indicando la popularidad de JS. Agarrar data de encuestas viejas y hacer una comparación de agunas techs, tipo JS, PHP, Python, Java, etc.}
+
+\todo{Talvez agregar}
+\input{figures/background/stackoverflowSurvey-ProgrammingLanguages2019.tex}
+
+\input{figures/background/stackoverflowSurvey-ProgrammingLanguagesEvolution.tex}
+
+\todo{Agregar tema serverless. Que NodeJS tiene el 75\% de los deploys}
+
 \subsection{TypeCoercion}
+\todo{Hablar de WTFS}
+\todo{Aca incluir texto de what does it mean to be a JS Type}
 \subsection{NPM Packages}
+\todo{Agregar info de NPM by numbers! Cuantos downloads, cuantos packages, etc.}
+\todo{https://snyk.io/blog/npm-passes-the-1-millionth-package-milestone-what-can-we-learn/}
 
 \section{TypeScript}
-\subsection{Declaration Files}
+\subsection{Declaration Files} \label{sec:declaration-files-background}
 \subsection{Definitely Typedd}
 
-\section{Jalangi}
+\section{Jalangi} \label{sec:jalangi}

### 2019-08-05 chapters/4-approach.tex
Add todo messages
Add code example for run-time information gathering

--- 2019-07-23 chapters/4-approach.tex
@@
 \chapter{Approach}\label{chap:approach}
-The approach starts with the problem definition and continues with what you have done. Try to give an intuition first and describe everything with words and then be more formal like `Let $g$ be ...'.
-
-Strive to make the following items crystal clear to your readers.
-\begin{itemize}
-\item What is the problem you are treating?
-\item What are the research questions you want to answer in this work?
-\item What are your own contributions and what is work of others? If
-  you are working in a team, it is important to demarcate against the
-  contributions of others in your team.
-\end{itemize}
-
-Describe each major technical problem and how it was solved.
 
 \section{Problem Definition}
-\todo{JavaScript popularity has increased dramatically over the last years.
-TypeScript has evolved naturally as the professional way of developing code in JavaScript, providing the robustness that ervery compiled language has.
-Generating the declaration files manually is very time consuming and error prone.}
+In spite of the fact that JavaScript was originally intended for simple web-pages scripting, it's popularity has increased over the last years, being the most commonly used programming language among professional developers in 2018.
+Development of robust and high-quality frameworks such as jQuery, AngularJS, React or Vue.js contributed to JavaScript being the leading programming language for client-side development.
 
-\todo{NodeJS, frameworks like React, Angular, Vue. The migation to Single Page Apps and Serverless, etc.}
+Moreover, several back-end applications are being written in JavaScript using Node.js, thus unifying the programming languages of client and back-end code.
+Companies with architectures based on concepts such as serverless and microservices are moving traditional back-end code written in Java, Python or PHP to the cloud and replacing it by small compute services, like AWS Lambda or Google Cloud Functions, or directly by cloud managed services.
+It is very common that development teams decide to write the compute services in Node.js, thus reducing the technological stack.
 
-\todo{Info from Stackoverflow and Github}
+However, JavaScript was not intented for developing large-scale applications.
+Lack of good interface descriptions using types and unintuitive type coercion have been slowing development of JavaScript applications.
 
-\todo{Mention DefinitelyTyped repository. Explicar claramente que el objetivo de esta tesis es generar }
+The TypeScript programming language is a superset of JavaScript with optional type annotations that compiles to plain JavaScript.
+It has become a widely used alternative for JavaScript.
+It enables IDEs to perform code intelligence tasks such as code completion, code navigation or refactoring and it rules out possible run-time errors through static type checking.
 
-\todo{El objetivo es solucionar el problema de definitely typed!!! Por eso se usa como objetivo que el input sea un node module}
+Aware that types are important for developing maintainable software systems, several industrial developers started using TypeScript in JavaScript projects.
+However, despite TypeScript's popularity, many new and existing JavaScript Libraries are still being written in pure JavaScript.
+
+The possibility of using exissting JS libraries in TypeScript projects is critical for a smooth transition and for its adoption in industrial development.
+TypeScript enables it through \textit{declaration files}, a typed description of the library's API.
+A public repository, DefinitelyTyped \citep{definitely-typed-repository}, contains declaration files for more than 6000 JS Libraries.
+71\% of the top 1000 depended-upon packages have their corresponding declaration file in DefinitelyTyped, as explained in \chapref{chap:background}.
+
+Unfortunately, such files are written and maintained manually, which is tedious, error prone and highly time consuming.
+The evolution of the declaration files does not follow in some cases the continous development of the corresponding libraries.
+As a result, some declaration files end up being out of date, thus hindering their usability.
+Mismatches between the declaration files and the actual JavaScript Library code have a negative impact on developers.
+Type checking messages and IDE code intelligence features are not accurate, thus increasing developing costs, generating actually more difficulties than before migrating to TypeScript.
+
+Declaration files are critical for TypeScript projects.
+TypeScript's popularity increase and the continuous growth of packages uploaded to the NPM registry, surpassing \textit{1 million} packages by June 2019, indicate that mismatches between declaration files and JavaScript implementations will likely grow.
+This Thesis tackles the declaration file generation problem and proposes a solution for assiting programmers in creating TypeScript Declaration Files from JavaScript code and keeping them up to date after.
+
+\todo{Agregar una imagen de un modulo simple con codigo JavaScript y que del otro lado salga el declaration file.}
 
 \section{TypeScript Declaration Files Generation Method}
-\todo{Explain overview process for generating declaration file.
-Explain that the focus of this thesis was to focus on the runtime information and use the symbolic executor as a 'refinement' of the test cases.
-Use 'public' document of the Master project}
 
-This Thesis introduces a \textit{TypeScript Declaration Files Generation Method}, which generates a valid TypeScript Declaration File for a specific NPM package, as explained in Fig.
-\ref{fig:tsd_generation_method_block_diagram}.
+We introduce a \textit{TypeScript Declaration Files Generation Method}, which generates a valid TypeScript Declaration File for a specific JavaScript Library uploaded to the NPM Registry, as explained in \figref{fig:tsd_generation_method_block_diagram}.
 
-\input{figures/approach/typescriptDeclarationFilesGenerationMethodBlockDiagram}
-
-Following the goal of easing and automating the generation of TypeScript Declaration Files for existings JS Libraries, the method receives the name of the JS Library published to the NPM Registry as an input argument and returns the corresponding TypeScript declaration file.
+\input{figures/approach/typescript-declaration-files-generation-method/typescriptDeclarationFilesGenerationMethodBlockDiagram.tex}
 
 The method consists of a continuous refinement process.
-It generates a TypeScript Declaration File based on data flow and type information gathered at run-time from a code base that executes the JS Library.
-A Symbolic Execution Engine expands the code base in every iteration by generating new test cases that explore different execution paths.
+It generates a TypeScript Declaration File based on data flow and type information gathered at run-time from a code base that executes the JavaScript Library.
+The code base is expanded by a Symbolic Execution Engine that generates test cases using the generated type definitions.
+As a result, new execution paths are explored, gathering new run-time information and thus refining the declaration file in every iteration.
 
-The file generated after each iteration is a valid and fully functional declaration file, suitable for usage in development environments.
-It contains no errors, matches the structure of the corresponding JS Library and could be uploaded to the DefinitelyTyped Repository.
-It is therefore intended to be integrated within the build process of the JS Library.
+Following the goal of easing and automating the generation of TypeScript Declaration Files for existings JavaScript Libraries, the method receives the name of the JavaScript Library published to the NPM Registry as an input argument and returns the corresponding TypeScript declaration file.
 
-\todo{Decir que para la tesis se decidio no inlcuir el feedback loop. Solo se hizo la parte de runtime, etc.}
+The declaration file returned by the method is valid and fully functional, making it suitable for being used within the development process.
+It contains no errors and matches the structure of the JavaScript Library under analysis, so that the JavaScript code generated after compilation runs.
+The conducted experiments included tests that consisted on replacing a specific type definition from DefinitelyTyped \citep{definitely-typed-repository} with the one generated in the experiments: TypeScript compilation was sucessful, the generated JavaScript code ran without errors and code intelligence features performed by IDEs like code completion worked as expected.
 
-\subsection{Overview}
-As explained before, the DefinitelyTyped Repository is the official repository for finding declaration files for npm packages.
-Therefore, the \textit{TypeScript Declaration Files Generation Method} is intended to be used on existing, published npm packages.
+The feedback loop was not included in the implementation, though.
+As explained, exploring new execution paths by creating new test cases will refine the declaration files: new types will be assigned to different variables, new interfaces will be detected and existing interfaces will be defined more accurately by detecting new properties.
+From an incremental point of view, generating the declaration files from run-time information has to be developed in a first place.
+A declaration file must exist first so that it can be refined afterwards.
+
+For that reason, the decision was to keep the focus on building an end-to-end architecture that supports the generation of valid declaration files for a specific JavaScript Library.
+The proposed architecture supports a future addition of the refinement loop, without modifying the existing blocks.
+The Symbolic Execution Engine will expand the existing code base by generating new test cases.
+The actual generation of declaration files remains unmodified.
+
+\subsection{Implementation}
+As explained before, the DefinitelyTyped Repository is the official repository for finding declaration files for existing JavaScript Libraries.
+The \textit{TypeScript Declaration Files Generation Method} is intended to be used on existing, published npm packages.
 The generated output TypeScript declaration file is a valid file which can be used for development and uploaded to the DefinitelyTyped Repository.
 
-The method uses an instrumentation framework to gather data flow information and type information at runtime.
-A second independent block uses the gathered information to generate a TypeScript Declaration File.
+Code examples that execute the JavaScript Library are needed to extract the run-time information via code instrumentation.
+It is achieved by retrieving the examples provided in the README files of the repositories of the different libraries.
+This is generally the place where developers explicitely show how to use their code.
+It showed to be an appropriate and pragmatic way of extracting the developer's intention and providing an useful initial code base with meaningful examples, thus avoiding a possible cold start problem.
+
+The examples and the code base of the library are instrumented with Jalangi \citep{DBLP:conf/sigsoft/SenKBG13}\citep{DBLP:conf/sigsoft/SenKBG13a} to gather data flow information and type information at runtime.
+Jalangi is a dynamic analysis configurable framework that provides several analysis modules that were extended as needed to retrieve the required run-time information, which is then saved to an output JSON file.
+This block is written in JavaScript and it runs in Node.js within a Docker container.
+
+A second independent block uses the run-time information to generate a TypeScript Declaration File.
+It infers the overall structure of the JS Library, the interfaces and the types from the JSON file.
+This block is a Node.js application written in TypeScript that also runs within a Docker container.
+
 The signatures obtained from the generated declaration file are used to generate pre and post conditions on a Symbolic Execution Engine in order to build different test cases that explore new execution paths.
-The code from the test cases is then added to the existing code base.
-The declaration file is generated again using an expanded code base.
-
-A feedback loop refines iteratively the outcome by combining run-time information gathering and symbolic analysis: It performs Run-Time Analysis to generate a declaration file using run-time information and a Symbolic Execution Engine to refine and expand the code base being used to gather run-time information.
+The code from the test cases is then added to the existing code base and the declaration file is generated again using an expanded code base.
 
@@
 If new execution paths are covered by the test cases, new run-time information will be gathered and the declaration file will change accordingly.
-In order to obtain a sound analysis, this feedback loop would be iterated until there are no differences among the generated declaration files.
+In order to obtain a sound analysis, the feedback loop would be iterated until the generated declaration files in consecutive iterations do not differ.
 However, by trading soundness for scalability, the overall procedure can be stopped either using a loop bound, a time bound or an heuristics mechanism.
-Exploring and investigating the proper stopping criteria is beyond the scope of this work.
 
-Finally, the generated declaration files are fully functional and valid files. They can be used in a TypeScript project without further modifications
-
-\subsection{Implementation}
-\todo{Decir que es un framework. Decir que es tipo una pipeline donde todo se conecta con lo otro. Explicar que por mas que reciba un string como input}
+\todo{Agregar una imagen mostrando que es un unico script que hace todo. Mostrar imagenes de docker mostrando en qué está escrito cada cosa.}
 
 \section{Initial code base}
+To extract run-time information of a JavaScript Library, it is necessary, by definition, to actually execute the code.
+It is not enough to instrument the source files of the library.
+The analysis modules provided by Jalangi to gather information are only triggered if the instrumented code gets executed.
+
+We decided to extract the code examples from the Readme files of the repository associated to the NPM Package.
+Readme files are usually used by developers for briefly describing what the code does, what problem it solves, how to install the application, how to build the code, etc.
+It is very common that developers provide code examples in the readme files to show how the code works and how to use it.
+This is specially true for NPM Packages, which are in general created to solve a specific problem of JavaScript development.
+
+Obtaining code examples for a specific NPM Package is achieved in three steps:
+\begin{enumerate}
+  \item Obtaining the repository from the package.
+  \item Retrieving the README file from the repository.
+  \item Extract the code examples from the README file.
+\end{enumerate}
+
+\subsection{Repository}
+The \mintinline{bash}{npm view} command shows data about a package.
+It is possible to retrieve a specific field of the package registry by specifying fhe field name after the package descriptor.
+For example to retrieve information about the \textit{lodash} package, the following command can be used: \mintinline{bash}{npm view lodash}.
+
+For this particular case, only the url of the repository associated with the package is needed.
+This can be achieved by adding the correponding field name: \mintinline{bash}{npm view <PACKAGE> repository.url}.
+
+\begin{bashinline}
+$ npm view npm repository.url
+git+https://github.com/npm/cli.git
+
+$ npm view lodash repository.url
+git+https://github.com/lodash/lodash.git
+
+$ npm view jquery repository.url$
+git+https://github.com/jquery/jquery.git
+\end{bashinline}
+
+\subsection{Readme files}
+The url for downloading a raw file from a Github Repository is the following:
+
+\begin{bashinline}
+https://raw.githubusercontent.com/THE-REPOSITORY/master/PATH-TO-FILE
+\end{bashinline}
+
+After the repository url is extracted from the npm package, the url for downloading the readme file needs to be constructed performing some string manipulation.
+Finally, the readme file is obtained with a simple \mintinline{bash}{curl} or \mintinline{bash}{wget} command.
+
+\begin{bashinline}
+$ npm view abs repository.url
+git+ssh://git@github.com/IonicaBizau/abs.git
+
+$ curl -L -o readme.md --fail \
+    https://raw.githubusercontent.com/IonicaBizau/abs/master/README.md
+\end{bashinline}
+
+\subsection{Code examples}
+Readme files are written using Markdown\footnote{https://www.markdownguide.org}, a very common lightweight and simple markup language.
+GitHub or BitBucket will convert the readme files into HTML automatically and render them when accessing the repository through a web browser.
+
+It is very common to write code examples within code blocks indicating the programming language, so that it gets highlighted with the specific syntax.
+The Markdown identifiers for JavaScript are \mintinline{bash}{js} or \mintinline{bash}{javascript}.
+The code examples are finally retrieved by filtering the content within the corresponding code blocks.
+
+\begin{textinline}
+$ cat example-readme.md
+# Description
+This is an example that shows how to use console.log().
+
+```js
+var a = "Hello World!";
+
+function f(s) {
+	console.log(s);
+}
+
+f(a);
+```
+
+$ cat example-readme.md | \
+  sed 's/```javascript/```js/g' | \
+  sed -n '/^```js/,/^```/ p' > example.js
+
+$ node example.js
+Hello World!
+\end{textinline}
 
 \section{Run-time Information Gathering} \label{run-time-information-gathering}
-\subsection{Instrumentation}
-\subsubsection{Interactions}
+As shown in a simple example in \figref{fig:run-time-information-gathering-simple-example}, run-time analysis will gather information such as:
+
+\begin{enumerate}
+  \item Function \mintinline{bash}{f} got invoked with parameters \mintinline{bash}{a} and \mintinline{bash}{b} with types \mintinline{bash}{string} and \mintinline{bash}{number}.
+  \item Property or method \mintinline{bash}{foo} of parameter \mintinline{bash}{a} of function \mintinline{bash}{f} was accessed within the function.
+  \item Parameter \mintinline{bash}{a} of function \mintinline{bash}{f} was used as operand for operator \mintinline{bash}{==}.
+\end{enumerate}
+
+The dynamic analysis framework used for gathering this kind of information is Jalangi.
+As explained in \secref{sec:jalangi}, the configurable analysis modules enable programming custom callbacks that get triggered with virtually any JavaScript event.
+The events that are observed are:
+\begin{enumerate}
+  \item Binary operations, like \mintinline{bash}{==}, \mintinline{bash}{+} or \mintinline{bash}{===}.
+  \item Variable declaration.
+  \item Function, method, or constructor invocation.
+  \item Access to an object's property.
+  \item Unary operations, like \mintinline{bash}{!} or \mintinline{bash}{typeof}.
+\end{enumerate}
+
+The implementation stores these observations as entities called \textit{interactions}.
+As explained later in \secref{run-time-analysis}, they are used for translating, modifying and aggregating Jalangi's raw event information in order to get an application specific data representation.
+
+The main goal of the run-time analysis is to gather information useful for determining a function's signature.
+Therefore, observations need to get associated with a function's argument.
+The implementation wraps variables around Proxy Objects to store tracking data, hereby enabling to associate the observation with a specific argument.
+The effect is reverted using Jalangi's callbacks to modify the behaviour of the affected operators, such as \mintinline{bash}{===} or \mintinline{bash}{typeof}.
+
+Jalangi fails to instrument some JavaScript Libraries.
+No effort was made in trying to fix failing instrumentation.
+Failing JS Libraries were ommited from the analysis by providing a list with the names of the modules that should not be instrumented.
+
+Each function is given an unique key, which is used as the primary key of a map where the run-time information is stored.
+Each function has an array of arguments and each interaction gets associated with an argument.
+The map is finally returned as a JSON file that can be used for later processing.
+
+The tool is written in JavaScript and runs in Node.js within a Docker container.
+
+\input{figures/approach/run-time-information-gathering/simple-example.tex}
+
+\subsection{Analysis} \label{run-time-analysis}
 \todo{inputValue}
+\input{figures/approach/run-time-information-gathering/analysis/input-value.tex}
+\input{figures/approach/run-time-information-gathering/analysis/get-field.tex}
+\input{figures/approach/run-time-information-gathering/analysis/method-call.tex}
+\input{figures/approach/run-time-information-gathering/analysis/used-as-argument.tex}
+
 \todo{getField}
@@
 \todo{usedAsArgument}
-\todo{convertedTo}
+\todo{operatorInteraction}
 
-\subsubsection{Wrapper Objects}
+\todo{followingInteractions}
+\todo{traces}
+
+\subsection{Wrapper Objects}
 \todo{Proxy Objects}
@@
 \todo{Undefined}
+
 \subsection{Output Format}
 \todo{Explicar que es un JSON con todas las interactions. Explicar estructura de cómo se forma el JSON}
+
 \subsection{Blacklisted Modules}
+\input{figures/approach/blacklisted-modules/blacklistedModules.tex}
 \todo{Incluir grafico con los blacklisted modules}

### 2019-08-05 chapters/6-conclusions.tex
Add todo message for merging declaration files
Add code example for run-time information gathering

@@
 
+\draft{Una de las conclusiones mas importantes es que la type inference al final no es tan importante.
+Lo dificil es generar el declaration file como corresponde, detectar las interfaces, etc.
+El tipo de las variables puede, en principio, inferirse de los ejemplos de los developers.
+Es una aproximación bastante acertada y se atacan los otros problemas.}
+
+\draft{Decir que todos los bloques pueden mejorarse independientemente: Extraer mejor los readme files, mejorar la instrumentation usando Babel para quedarse siempre en ES5...
+mejorar la generation de declaration files, etc.}
+
 \section{Future Work}
@@
 \todo{Deberia ser hecho en otro modulo que no sea el runtime information. El runtime information deberia exponer unicamente los datos que le hacen falta al otro para inferir los tipos.}
+\subsubsection{Return values}
+\todo{El return value se podria incluir en el analysis de Jalangi, detectando las interactions que tiene etc etc.
+Esto no es tan facil porque no se puede simplemente inspeccionar el return value, sino que hay que saber para qué se usa.
+Los ejemplos podrían ser útiles}.
+
 \subsubsection{Variable Names}
@@
 
+\todo{MERGE Declaration Files - Ver si se agrega}
+
 \subsubsection{Operators}

### 2019-08-05 figures/approach/run-time-information-gathering/analysis/get-field.tex
Add code example for run-time information gathering

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.4\textwidth}
			\jscode{code/run-time-information-gathering/simple-examples/get-field.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[Instrumented JS code]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.58\textwidth}
			\jsoncode{code/run-time-information-gathering/simple-examples/output-get-field.json}
		\end{minipage}
	\end{lrbox}
	\subfloat[Run-time information output]{\usebox{\mintedbox}}
	\caption[\textit{getField} interaction example]{\textbf{\textit{getField} interaction example} - The accessed properties of argument \textit{a} with their type are recorded. Several additional fields and entries of the final output JSON file were ommited for simplicity.}
	\label{fig:run-time-information-gathering-get-field}
\end{figure}

### 2019-08-05 figures/approach/run-time-information-gathering/analysis/input-value.tex
Add code example for run-time information gathering

--- 2019-08-05 figures/approach/run-time-information-gathering/analysis/get-field.tex
@@
 		\begin{minipage}{0.4\textwidth}
-			\jscode{code/run-time-information-gathering/simple-examples/get-field.js}
+			\jscode{code/run-time-information-gathering/simple-examples/input-value.js}
 		\end{minipage}
@@
 		\begin{minipage}{0.58\textwidth}
-			\jsoncode{code/run-time-information-gathering/simple-examples/output-get-field.json}
+			\jsoncode{code/run-time-information-gathering/simple-examples/output-input-value.json}
 		\end{minipage}
@@
 	\subfloat[Run-time information output]{\usebox{\mintedbox}}
-	\caption[\textit{getField} interaction example]{\textbf{\textit{getField} interaction example} - The accessed properties of argument \textit{a} with their type are recorded.
+	\caption[\textit{inputValue} interaction example]{\textbf{\textit{inputValue} interaction example} - The type of the arguments of function \textit{foo()} is recorded.
 Several additional fields and entries of the final output JSON file were ommited for simplicity.}
-	\label{fig:run-time-information-gathering-get-field}
+	\label{fig:run-time-information-gathering-input-value}
 \end{figure}

### 2019-08-05 figures/approach/run-time-information-gathering/analysis/method-call.tex
Add code example for run-time information gathering

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.4\textwidth}
			\jscode{code/run-time-information-gathering/simple-examples/method-call.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[Instrumented JS code]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.58\textwidth}
			\jsoncode{code/run-time-information-gathering/simple-examples/output-method-call.json}
		\end{minipage}
	\end{lrbox}
	\subfloat[Run-time information output]{\usebox{\mintedbox}}
	\caption[\textit{methodCall} interaction example]{\textbf{\textit{methodCall} interaction example} - The invoked method of argument \textit{a} is recorded. Method information can be accessed through the reference \mintinline{bash}{functionId}, for which, in this example, the \mintinline{bash}{inputValue} interactions are stored. Main entry for \mintinline{bash}{functionId_3} has been ommited. Several additional fields and entries of the final output JSON file were also ommited for simplicity.}
	\label{fig:run-time-information-gathering-method-call}
\end{figure}

### 2019-08-05 figures/approach/run-time-information-gathering/analysis/used-as-argument.tex
Add code example for run-time information gathering

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.3\textwidth}
			\jscode{code/run-time-information-gathering/simple-examples/used-as-argument.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[Instrumented JS code]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.68\textwidth}
			\jsoncode{code/run-time-information-gathering/simple-examples/output-used-as-argument.json}
		\end{minipage}
	\end{lrbox}
	\subfloat[Run-time information output]{\usebox{\mintedbox}}
	\caption[\textit{usedAsArgument} interaction example]{\textbf{\textit{usedAsArgument} interaction example} - Reporting when an argument is used as an argument of another function provides more accurate information. Several additional fields and entries of the final output JSON file were ommited for simplicity.}
	\label{fig:run-time-information-gathering-used-as-argument}
\end{figure}

### 2019-08-05 figures/approach/run-time-information-gathering/simple-example.tex
Add code example for run-time information gathering

--- 2019-08-05 figures/approach/run-time-information-gathering/analysis/input-value.tex
@@
 		\begin{minipage}{0.4\textwidth}
-			\jscode{code/run-time-information-gathering/simple-examples/input-value.js}
+			\jscode{code/run-time-information-gathering/simple-examples/simple-example.js}
 		\end{minipage}
@@
 		\begin{minipage}{0.58\textwidth}
-			\jsoncode{code/run-time-information-gathering/simple-examples/output-input-value.json}
+			\jsoncode{code/run-time-information-gathering/simple-examples/output-simple-example.json}
 		\end{minipage}
@@
 	\subfloat[Run-time information output]{\usebox{\mintedbox}}
-	\caption[\textit{inputValue} interaction example]{\textbf{\textit{inputValue} interaction example} - The type of the arguments of function \textit{foo()} is recorded.
+	\caption[Simple run-time information gathering example]{\textbf{Simple run-time information gathering example} - Example code showing \textit{inputValue} and \textit{getField} interactions.
 Several additional fields and entries of the final output JSON file were ommited for simplicity.}
-	\label{fig:run-time-information-gathering-input-value}
+	\label{fig:run-time-information-gathering-simple-example}
 \end{figure}

### 2019-08-05 code/run-time-information-gathering/simple-examples/get-field.js

@@ -1,0 +1,10 @@
+function foo(a) {
+	return a.bar + a.baz;
+}
+
+foo(
+	{
+		bar: "hello",
+		baz: "world"
+	}
+); 

### 2019-08-05 code/run-time-information-gathering/simple-examples/input-value.js

@@ -1,0 +1,5 @@
+function foo(a, b) {
+	return a + b;
+}
+
+foo("hello", "world");

### 2019-08-05 code/run-time-information-gathering/simple-examples/method-call.js

@@ -1,0 +1,14 @@
+function foo(a) {
+	return a.doSomething(
+		1,
+		2
+	);
+}
+
+var myObj = {
+	doSomething: function (x, y) {
+		return x + y;
+	}
+};
+
+foo(myObj);

### 2019-08-05 code/run-time-information-gathering/simple-examples/simple-example.js

@@ -1,0 +1,10 @@
+function foo(a, b) {
+	return a.bar + b;
+}
+
+foo(
+	{
+		bar: 1
+	},
+	2
+); // -> 3

### 2019-08-05 code/run-time-information-gathering/simple-examples/used-as-argument.js

@@ -1,0 +1,18 @@
+function foo(a) {
+	if (a.x) {
+		// ...
+	}
+
+	bar(a);
+}
+
+function bar(a) {
+	if (a.y) {
+		// ...
+	}
+}
+
+foo({
+	x: 200,
+	y: 300
+});

### 2019-08-06 figures/approach/run-time-information-gathering/analysis/following-interactions.tex
Add following interactions and traces examples

--- 2019-08-05 figures/approach/run-time-information-gathering/analysis/input-value.tex
@@
 		\begin{minipage}{0.4\textwidth}
-			\jscode{code/run-time-information-gathering/simple-examples/input-value.js}
+			\jscode{code/run-time-information-gathering/simple-examples/following-interactions.js}
 		\end{minipage}
@@
 		\begin{minipage}{0.58\textwidth}
-			\jsoncode{code/run-time-information-gathering/simple-examples/output-input-value.json}
+			\jsoncode{code/run-time-information-gathering/simple-examples/output-following-interactions.json}
 		\end{minipage}
@@
 	\subfloat[Run-time information output]{\usebox{\mintedbox}}
-	\caption[\textit{inputValue} interaction example]{\textbf{\textit{inputValue} interaction example} - The type of the arguments of function \textit{foo()} is recorded.
+	\caption[\textit{followingInteractions} interaction example]{\textbf{\textit{followingInteractions} interaction example} - Observations on return values of \mintinline{bash}{getField} and \mintinline{bash}{methodCall} interactions get associated to the leading interaction.
 Several additional fields and entries of the final output JSON file were ommited for simplicity.}
-	\label{fig:run-time-information-gathering-input-value}
+	\label{fig:run-time-information-gathering-following-interactions}
 \end{figure}

### 2019-08-06 figures/approach/run-time-information-gathering/analysis/traces.tex
Add following interactions and traces examples

--- 2019-08-05 figures/approach/run-time-information-gathering/analysis/input-value.tex
@@
 	\begin{lrbox}{\mintedbox}
-		\begin{minipage}{0.4\textwidth}
-			\jscode{code/run-time-information-gathering/simple-examples/input-value.js}
+		\begin{minipage}{0.46\textwidth}
+			\jscode{code/run-time-information-gathering/simple-examples/traces.js}
 		\end{minipage}
@@
 	\begin{lrbox}{\mintedbox}
-		\begin{minipage}{0.58\textwidth}
-			\jsoncode{code/run-time-information-gathering/simple-examples/output-input-value.json}
+		\begin{minipage}{0.52\textwidth}
+			\jsoncode{code/run-time-information-gathering/simple-examples/output-traces.json}
 		\end{minipage}
@@
 	\subfloat[Run-time information output]{\usebox{\mintedbox}}
-	\caption[\textit{inputValue} interaction example]{\textbf{\textit{inputValue} interaction example} - The type of the arguments of function \textit{foo()} is recorded.
+	\caption[\textit{traces} example]{\textbf{\textit{traces} example} - Interactions on different invocations of a function are stored with different \mintinline{bash}{traces}.
 Several additional fields and entries of the final output JSON file were ommited for simplicity.}
-	\label{fig:run-time-information-gathering-input-value}
+	\label{fig:run-time-information-gathering-traces}
 \end{figure}

### 2019-08-06 code/run-time-information-gathering/simple-examples/following-interactions.js

@@ -1,0 +1,15 @@
+function foo(a) {
+	var b = a.x;
+
+	if (b.y) {
+		// ...
+	}
+}
+
+var myObj = {
+	x: {
+		y: "bar"
+	}
+};
+
+foo(myObj);

### 2019-08-06 code/run-time-information-gathering/simple-examples/traces.js

@@ -1,0 +1,19 @@
+function foo(a) {
+	var b;
+
+	if (Math.random() < 0.5) {
+		b = a.someField;
+	} else {
+		b = a.otherField;
+	}
+
+	return b;
+}
+
+var myObj = {
+	someField: 1,
+	otherField: 2
+};
+
+foo(myObj);
+foo(myObj);

### 2019-08-06 chapters/4-approach.tex
Add following interactions and traces examples
Add text about function container and argument container

@@
 The implementation stores these observations as entities called \textit{interactions}.
-As explained later in \secref{run-time-analysis}, they are used for translating, modifying and aggregating Jalangi's raw event information in order to get an application specific data representation.
+As explained later in \secref{sec:run-time-analysis}, they are used for translating, modifying and aggregating Jalangi's raw event information in order to get an application specific data representation.
 
@@
 
-\subsection{Analysis} \label{run-time-analysis}
-\todo{inputValue}
+\subsection{Analysis} \label{sec:run-time-analysis}
+The provided analysis module implements the following Jalangi callbacks, which were explained in detail in \secref{sec:jalangi}:
+\begin{itemize}
+  \item binaryPre()
+  \item declare()
+  \item getFieldPre()
+  \item putFieldPre()
+  \item functionEnter()
+  \item functionExit()
+  \item invokeFun()
+  \item invokeFunPre()
+  \item unaryPre()
+  \item write()
+\end{itemize}
+
+They enable observation of all JS operations considered relevant.
+Every invoked function is given an unique identifier and information for each argument of every function gets stored.
+The entities \mintinline{bash}{FunctionContainer}, \mintinline{bash}{ArgumentContainer} and \mintinline{bash}{Interaction} are introduced to support the representation of the gathered run-time information.
+
+Run-time information is stored in a map that contains objects of type \mintinline{bash}{FunctionContainer}.
+This entity contains static information about the function as well as implementation specific data needed for generating the right type of declaration file, as explained in \secref{sec:declaration-files-background}.
+Each field of the entity is explained in \coderef{code:function-container}.
+
+\begin{code}
+  \jscode{code/run-time-information-gathering/analysis/function-container.js}
+  \caption[FunctionContainer object]{\textbf{\mintinline{bash}{FunctionContainer} object} - \mintinline{bash}{FunctionContainer} is the main entity in the map where the run-time information gets gathered.
+It contains implementation specific information for the declaration files generation.
+Entities of type \mintinline{bash}{ArgumentContainer} get appended to the \mintinline{bash}{args} property.}
+  \label{code:function-container}
+\end{code}
+
+Moreover, as explained in \coderef{code:argument-container}, the \mintinline{bash}{ArgumentContainer} object works as a container for all interactions recorded for that argument.
+Additionally, it contains simple information about the argument itself, like the index or the name given in the function declaration.
+
+\begin{code}
+  \jscode{code/run-time-information-gathering/analysis/argument-container.js}
+  \caption[ArgumentContainer object]{\textbf{\mintinline{bash}{ArgumentContainer} object} - The \mintinline{bash}{ArgumentContainer} object works mainly as a container for all relevant observations regarding that argument.}
+  \label{code:argument-container}
+\end{code}
+
+Finally, observations on the arguments of a function are stored as objects of type \mintinline{bash}{Interaction}.
+They have to be always associated to an \mintinline{bash}{ArgumentContainer} object or to another \mintinline{bash}{Interaction}.
+They are explained in detail in \secref{sec:run-time-interactions}.
+
+\subsection{Interactions} \label{sec:run-time-interactions}
+
+\draft{enables the developer to create custom analysis modules by implementing the callbacks to virtually every JavaScript. Jalangi provides a way of writing a custom analysis by implementing}
+
+\subsubsection{\textit{inputValue} interaction}
 \input{figures/approach/run-time-information-gathering/analysis/input-value.tex}
+
+\subsubsection{\textit{getField} interaction}
 \input{figures/approach/run-time-information-gathering/analysis/get-field.tex}
+
+\subsubsection{\textit{methodCall} interaction}
 \input{figures/approach/run-time-information-gathering/analysis/method-call.tex}
+
+\subsubsection{\textit{usedAsArgument} interaction}
 \input{figures/approach/run-time-information-gathering/analysis/used-as-argument.tex}
 
-\todo{getField}
-\todo{putField}
-\todo{methodCall}
-\todo{usedAsArgument}
-\todo{operatorInteraction}
+\subsubsection{\textit{operator} interaction}
 
-\todo{followingInteractions}
-\todo{traces}
+\subsubsection{\textit{followingInteractions}}
+\input{figures/approach/run-time-information-gathering/analysis/following-interactions.tex}
+
+\subsubsection{Traces}
+\input{figures/approach/run-time-information-gathering/analysis/traces.tex}
 

### 2019-08-06 code/run-time-information-gathering/analysis/argument-container.js

@@ -1,0 +1,24 @@
+{
+	/*
+		Index of the argument.
+		function foo(a, b) {
+			...
+		}
+	*/
+	"argumentIndex": 0,
+
+	/*
+		Name of the argument as written in the function declaration.
+		function foo(a, b) {
+			...
+		}
+	*/
+	"argumentName": "a",
+
+	/*
+		Array of interactions associated to the argument.
+	*/
+	"interactions": [
+		
+	]	
+}

### 2019-08-06 code/run-time-information-gathering/analysis/function-container.js

@@ -1,0 +1,55 @@
+{
+	/* 
+		Unique id given to the function when first invoked.
+	*/
+	"functionId": "functionId_XXX",
+
+	/*
+		Name of the function or method as indicated in the code:
+			- function foo(...) {}
+			- {
+				foo: function(...) {} 
+			}
+	*/
+	"functionName": "foo",
+
+	/*
+		Boolean indicating whether the function is being used as a constructor: 'new foo()'.
+	*/
+	"isConstructor": false, 
+
+	/*
+		Map containing entities of type 'ArgumentContainer'
+	*/
+	"args": {},
+
+	/*
+		Array of objects containing the type and the traceId for that specific return value.
+	*/
+	"returnTypeOfs": [
+		{
+			"typeOf": "number",
+			"traceId": "trace__1"
+		}
+	],
+
+	/*
+		Name of the module to which the function belongs to:
+			const myModule = require('myModule');
+			myModule.foo()
+	*/
+	"requiredModule": "myModule",
+
+	/*
+		Boolean indicating whether the function is the actual function being exported by the module:
+
+		true:
+			const foo = require('myModule');
+			foo()
+		
+		false:
+			const myModule = require('myModule');
+			myModule.foo()
+	*/
+	"isExported": true
+}

### 2019-08-07 chapters/4-approach.tex
Add explanation for interactions

@@
 \subsection{Interactions} \label{sec:run-time-interactions}
+Jalangi enables the developer to create custom analysis modules by implementing the callbacks that are relevant for the application.
+However, the callbacks belong to an abstraction layer that is not aware of the actual implementation.
 
-\draft{enables the developer to create custom analysis modules by implementing the callbacks to virtually every JavaScript. Jalangi provides a way of writing a custom analysis by implementing}
+Observations of the arguments are stored as \mintinline{bash}{Interactions}.
+They represent the relevant observation data that needs to be extracted for the purpose of best describing a function in order to generate a TypeScript declaration file.
+
+The object \mintinline{bash}{Interaction} is associated to an \mintinline{bash}{ArgumentContainer}.
+If the \mintinline{bash}{ArgumentContainer} does not exist when the Jalangi callback is triggerd, the \mintinline{bash}{Interaction} is not saved.
+The main goal of gathering run-time information is to describe a function and its arguments.
+Thus, callbacks triggered on variables that are not arguments of functions are ignored, as shown in \coderef{code:ignored-observations}.
+The mechanism to map an observation with an \textit{ArgumentContainer} is described in detail in \secref{sec:run-time-wrapper-objects}
+
+\begin{code}
+  \jscode{code/run-time-information-gathering/analysis/ignored-observations.js}
+  \caption[Ignored observations]{\textbf{Ignored observations} - Variable \mintinline{bash}{b} gets declared within the scope of function \mintinline{bash}{foo}.
+Therefore, no observation on it gets stored.}
+  \label{code:ignored-observations}
+\end{code}
 
 \subsubsection{\textit{inputValue} interaction}
+It is the most simple interaction.
+As shown in \figref{fig:run-time-information-gathering-input-value}, it gets triggered when a function is invoked and it stores the input value of a specific argument.
+If a function gets invoked several times, different interactions get stored, each one of them with a different \mintinline{bash}{traceId}.
+
 \input{figures/approach/run-time-information-gathering/analysis/input-value.tex}
@@
 \subsubsection{\textit{getField} interaction}
+This interaction gets triggered whenever a property of an object gets accessed, either if the get field operation is \mintinline{bash}{obj.field} or \mintinline{bash}{obj['field']}.
+As shown in \figref{fig:run-time-information-gathering-get-field}, the name of the accessed field and the type of the actual property get stored.
+
+It should be noted that the field \mintinline{bash}{returnTypeOf} represents the type of the accessed property at the time when the interaction gets triggered.
+The accessed property may somehow get modified afterwards, which will not modify the existing interaction.
+If the property gets accessed again, a new interaction with a new \mintinline{bash}{returnTypeOf} value will be added, as shown in \coderef{code:get-field-changing-at-runtime}.
+
+\begin{code}
+  \jscode{code/run-time-information-gathering/analysis/get-field-changing-at-runtime.js}
+  \caption[Modified accessed property]{\textbf{Modified accessed property} - Variable \mintinline{bash}{b} gets declared within the scope of function \mintinline{bash}{foo}.
+Therefore, no observation on it gets stored.}
+  \label{code:get-field-changing-at-runtime}
+\end{code}
+
 \input{figures/approach/run-time-information-gathering/analysis/get-field.tex}
@@
 \subsubsection{\textit{methodCall} interaction}
+This interaction gets triggered with the invocation of an object's method.
+As explained in \figref{fig:run-time-information-gathering-method-call}, information about the method itself is stored as another \mintinline{bash}{FunctionContainer} in the overall map.
+The reference to the \mintinline{bash}{FunctionContainer} is achieved through the properties \mintinline{bash}{functionId} and \mintinline{bash}{traceIdInTargetFunction}.
+
 \input{figures/approach/run-time-information-gathering/analysis/method-call.tex}
+
+\todo{ADD PUT FIELD INTERACTION}
 
@@
 \subsubsection{\textit{operator} interaction}
+\input{figures/approach/run-time-information-gathering/analysis/operator.tex}
 
@@
 
-\subsection{Wrapper Objects}
+\subsection{Wrapper Objects} \label{sec:run-time-wrapper-objects}
 \todo{Proxy Objects}

### 2019-08-07 figures/approach/run-time-information-gathering/analysis/operator.tex
Add explanation for interactions

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.34\textwidth}
			\jscode{code/run-time-information-gathering/simple-examples/operator.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[Instrumented JS code]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.62\textwidth}
			\jsoncode{code/run-time-information-gathering/simple-examples/output-operator.json}
		\end{minipage}
	\end{lrbox}
	\subfloat[Run-time information output]{\usebox{\mintedbox}}
	\caption[\textit{operator} interaction example]{\textbf{\textit{operator} interaction example} - The interaction records the operator and the type of both operands. The field \mintinline{bash}{operandForInteraction} indicates which operand triggered the interaction. The example shows how that value is \mintinline{bash}{left} for one argument and \mintinline{bash}{right} for the other one. Several additional fields and entries of the final output JSON file were ommited for simplicity.}
	\label{fig:run-time-information-gathering-following-interactions}
\end{figure}

### 2019-08-07 code/run-time-information-gathering/analysis/get-field-changing-at-runtime.js

@@ -1,0 +1,21 @@
+function foo(a) {
+	/**
+	 * Interaction saved with returnTypeOf: string
+	 */
+	if (a.bar) {
+		// ...
+	}
+
+	a.bar = 123;
+
+	/**
+	 * Interaction saved with returnTypeOf: number
+	 */
+	if (a.bar) {
+		// ...
+	}
+}
+
+foo({
+	bar: "hello"
+});

### 2019-08-07 code/run-time-information-gathering/analysis/ignored-observations.js

@@ -1,0 +1,12 @@
+function foo(a) {
+	// code ...
+
+	var b = {};
+	if (b.something) {
+		// ...
+	}
+
+	// code ...
+}
+
+foo(1);

### 2019-08-07 code/run-time-information-gathering/simple-examples/operator.js

@@ -1,0 +1,5 @@
+function foo(a, b) {
+	return a + b;
+}
+
+foo(1, 2);

### 2019-08-07 chapters/4-approach.tex
Add usedAsArgument and operator interactions

@@
   \item getFieldPre()
-  \item putFieldPre()
+  \item setFieldPre()
   \item functionEnter()
@@
 
-\todo{ADD PUT FIELD INTERACTION}
+\subsubsection{\textit{setField} interaction}
+The trigger for this interaction is writing an object's property, either by accessing the field as \mintinline{bash}{obj.field = 1} or \mintinline{bash}{obj['field'] = 1}.
+It can be seen in \figref{fig:run-time-information-gathering-set-field} that the interaction contains the name of the property to be written and the type of the value to be stored.
+
+\input{figures/approach/run-time-information-gathering/analysis/set-field.tex}
 
 \subsubsection{\textit{usedAsArgument} interaction}
+The run-time information is stored in a map with one \mintinline{bash}{FunctionContainer} entry per function.
+As seen in \figref{fig:run-time-information-gathering-used-as-argument}, an argument of function \mintinline{bash}{foo()} is used as parameter to invoke a second function \mintinline{bash}{bar()}.
+Interactions triggered within function \mintinline{bash}{bar()} are going to be stored within the \mintinline{bash}{FunctionContainer} corresponding to function \mintinline{bash}{bar()}, i.e.
+not the one associated to function \mintinline{bash}{foo()}.
+
+However, for the example provided in \figref{fig:run-time-information-gathering-used-as-argument} it is relevant to also associate interactions triggered within function \mintinline{bash}{bar()} to the function \mintinline{bash}{foo()}, since they can  be useful for describing the argument of \mintinline{bash}{foo()} more accurately.
+
+Therefore, \mintinline{bash}{usedAsArgument} interaction is introduced.
+It gets triggered when an argument is used in the invocation of a consecutive function.
+It serves the purpose of connecting related interactions.
+As shown in \figref{fig:run-time-information-gathering-used-as-argument}, it mainly contains information for identifying the target \mintinline{bash}{FunctionContainer}.
+
 \input{figures/approach/run-time-information-gathering/analysis/used-as-argument.tex}
@@
 \subsubsection{\textit{operator} interaction}
+Identifying if an argument is used with a specific operator is very important for infering the type of that variable.
+Relational operators like \mintinline{bash}{<, >, <=, >=} are mainly used with type \mintinline{bash}{number}.
+On the other side, operator \mintinline{bash}{+} is mainly used with \mintinline{bash}{number} or \mintinline{bash}{string}.
+Empirical evidence of how operators are used is presented in \secref{sec:experiments-js-operators-usage}.
+
+The \mintinline{bash}{operator} interaction stores the operator being used and the type of the left hand and right hand operands.
+It also indicates which operand is the one that triggered the interaction.
+The supported operators are the ones that trigger Jalangi's \mintinline{bash}{binaryPre()} callback.
+Conditionals are merged into one operator name \mintinline{bash}{conditional}, since Jalangi's \mintinline{bash}{conditional()} callback does not offer a way of identifying the operator for this cases.
+Both callbacks are explained in section \secref{sec:jalangi}.
+
+\figref{fig:run-time-information-gathering-operator} provides an example for operator \mintinline{bash}{+}.
+
 \input{figures/approach/run-time-information-gathering/analysis/operator.tex}

### 2019-08-07 figures/approach/run-time-information-gathering/analysis/set-field.tex
Add usedAsArgument and operator interactions

--- 2019-08-05 figures/approach/run-time-information-gathering/analysis/method-call.tex
@@
 		\begin{minipage}{0.4\textwidth}
-			\jscode{code/run-time-information-gathering/simple-examples/method-call.js}
+			\jscode{code/run-time-information-gathering/simple-examples/set-field.js}
 		\end{minipage}
@@
 		\begin{minipage}{0.58\textwidth}
-			\jsoncode{code/run-time-information-gathering/simple-examples/output-method-call.json}
+			\jsoncode{code/run-time-information-gathering/simple-examples/output-set-field.json}
 		\end{minipage}
@@
 	\subfloat[Run-time information output]{\usebox{\mintedbox}}
-	\caption[\textit{methodCall} interaction example]{\textbf{\textit{methodCall} interaction example} - The invoked method of argument \textit{a} is recorded.
-Method information can be accessed through the reference \mintinline{bash}{functionId}, for which, in this example, the \mintinline{bash}{inputValue} interactions are stored.
-Main entry for \mintinline{bash}{functionId_3} has been ommited.
+	\caption[\textit{setField} interaction example]{\textbf{\textit{setField} interaction example} - When a property of an argument gets written, the name of the written property and the type are recorded.
 Several additional fields and entries of the final output JSON file were also ommited for simplicity.}
-	\label{fig:run-time-information-gathering-method-call}
+	\label{fig:run-time-information-gathering-set-field}
 \end{figure}

### 2019-08-07 code/run-time-information-gathering/simple-examples/set-field.js

@@ -1,0 +1,13 @@
+function foo(a) {
+	// ...
+
+	a.otherField = 2;
+
+	// ...
+
+	return a;
+}
+
+foo({
+	someField: 1
+});

### 2019-08-09 chapters/4-approach.tex
Add followingInteractions and traces

@@
 It also indicates which operand is the one that triggered the interaction.
+\figref{fig:run-time-information-gathering-operator} provides an example for operator \mintinline{bash}{+}.
+
 The supported operators are the ones that trigger Jalangi's \mintinline{bash}{binaryPre()} callback.
@@
 Both callbacks are explained in section \secref{sec:jalangi}.
-
-\figref{fig:run-time-information-gathering-operator} provides an example for operator \mintinline{bash}{+}.
+Right hand operand is left as \mintinline{bash}{undefined} for that case.
 
@@
 \subsubsection{\textit{followingInteractions}}
+If a property of an argument is accessed the Jalangi callback \mintinline{bash}{getField()} gets triggered and a \mintinline{bash}{getField} interaction gets associated to the corresponding \mintinline{bash}{ArgumentContainer}.
+In order to describe the interface of the argument more accurately, it is also relevant to recursively record if properties of the return values are accessed, as explained in \figref{fig:run-time-information-gathering-following-interactions}.
+
+This was implemented by not only mapping the observation with an \mintinline{bash}{ArgumentContainer} but also with a \mintinline{bash}{getField} or \mintinline{bash}{methodCall} interaction.
+Both \mintinline{bash}{ArgumentContainer} and \mintinline{bash}{Interaction} implement the \mintinline[breaklines]{javascript}{addInteraction} method, as shown in \coderef{code:add-interaction-example}
+
+\begin{code}
+  \jscode{code/run-time-information-gathering/analysis/add-interaction-example.js}
+  \caption[addInteraction method]{\textbf{addInteraction method} - Example showing how interactions can be associated to \mintinline{bash}{ArgumentContainer} or another \mintinline{bash}{Interaction}.}
+  \label{code:add-interaction-example}
+\end{code}
+
 \input{figures/approach/run-time-information-gathering/analysis/following-interactions.tex}
@@
 \subsubsection{Traces}
+Different executions of a function are identified through different \mintinline{bash}{traceIds}.
+Each interaction and return value of a \mintinline{bash}{FunctionContainer} stores the corresponding \mintinline{bash}{traceId}.
+The \mintinline{bash}{traceId} field enables identifying correlations between input types and return values, as explained in \secref{sec:typescript-overloading}.
+An example is provided in \figref{fig:run-time-information-gathering-traces}.
+
 \input{figures/approach/run-time-information-gathering/analysis/traces.tex}
@@
 \todo{El tipo se extrae del valor en runtime.}
+\subsection{Overloading} \label{sec:typescript-overloading}
 \subsection{Implementation}

### 2019-08-09 code/run-time-information-gathering/analysis/add-interaction-example.js

@@ -1,0 +1,7 @@
+let a = new ArgumentContainer();
+
+let i1 = new MethodCallInteraction();
+a.addInteraction(i1);
+
+let i2 = new GetFieldInteraction();
+i1.addInteraction(i2);

### 2019-08-09 chapters/4-approach.tex
Finish Wrapper Objects

@@
   \jscode{code/run-time-information-gathering/analysis/add-interaction-example.js}
-  \caption[addInteraction method]{\textbf{addInteraction method} - Example showing how interactions can be associated to \mintinline{bash}{ArgumentContainer} or another \mintinline{bash}{Interaction}.}
+  \caption[addInteraction method]{\textbf{addInteraction method} - Example showing how interactions can be associated with \mintinline{bash}{ArgumentContainer} or with another \mintinline{bash}{Interaction}.}
   \label{code:add-interaction-example}
@@
 \subsection{Wrapper Objects} \label{sec:run-time-wrapper-objects}
-\todo{Proxy Objects}
-\todo{Primitives}
-\todo{Null}
-\todo{Undefined}
+An \mintinline{bash}{Interaction} needs to be associated to an \mintinline{bash}{ArgumentContainer} or to another \mintinline{bash}{Interaction}.
+This means that when Jalangi triggers a callback, it must be determined if that callback is relevant or not, i.e.
+if that callback refers to an argument of an invoked function.
+However, this can not be determined by only using the information provided in the different callbacks.
+
+However, Jalangi exposes the actual value affected by the callback.
+For example, for the \mintinline{bash}{getField()} callback triggered by operation \mintinline{javascript}{myObj.prop}, Jalangi does provide the object \mintinline{javascript}{myObj} in the callback.
+
+The solution consists in wrapping every input argument around a Wrapper Object when the function gets invoked.
+The Wrapper Object will store the reference to the corresponding \mintinline{bash}{ArgumentContainer} or \mintinline{bash}{Interaction}.
+Jalangi callbacks will then get triggered on the Wrapper Object instead.
+A mapping with the corresponding \mintinline{bash}{ArgumentContainer} or \mintinline{bash}{Interaction} is then straightforward, since the reference can be directly retrieved from the Wrapper Object.
+
+Jalangi enables to modify the behaviour of the original code.
+Using \mintinline{bash}{invokFunPre} and \mintinline{javascript}{declare} callbacks, a function invocation is detected and the arguments of the function are modified in run-time.
+
+Every function invocation generates new Wrapper Objects, which means that a Wrapper Object represents the input argument for a specific invocation.
+\mintinline{bash}{TraceIds} are also stored within these objects.
+
+The solution extends a similar concept already implemented by Jalangi: \mintinline{bash}{ShadowObjects}.
+It is a unique object associated to another object which can store meta-information.
+However, a \mintinline{bash}{ShadowObject} can not be created for primitives values, which is a major drawback.
+It is indeed relevant to capture, for example, if a \mintinline{bash}{number} is being used with a particular operator or if the property \mintinline{bash}{length} of a \mintinline{bash}{string} is being accessed.
+
+The implemented solution wraps objects around JavaScript Proxies.
+Primitive values like \mintinline{bash}{number} or \mintinline{bash}{string} are converted to their native wapper objects, i.e.
+\mintinline{bash}{Number} and \mintinline{bash}{String}.
+Finally, \mintinline{bash}{null} and \mintinline{bash}{undefined} are converted to empty objects with modified \mintinline{bash}{valueOf()} and \mintinline{bash}{toString()} methods.
+
+The behaviour of the original code is being explicitely modified.
+Therefore, original values are used instead for affected operators like \mintinline{bash}{===, !==, typeof}, for which wrapper objects will not work as expected.
+For \mintinline{bash}{null} and \mintinline{bash}{undefined} wrappers, the original values are used for conditional operators, since the result of the abstract operation \mintinline{bash}{ToBoolean} on an object is always \mintinline{text}{true}, as defined in the ECMAScript Language Specification \citep{ecma-script}.
+This approach could have been used for every operation by using the original value in the affected Jalangi callbacks.
+However, the implementation is complex and error prone.
+Instead, it was decided to modify the corresponding methods or specific operators so that they are compliant with JavaScript Type Coercion, which is explained in detail in {sec:background-js-type-coercion}.
 

### 2019-08-10 chapters/3-background.tex
Add todo about declaration files
Finish Wrapper Objects
Add examples for generation of declaration files

@@
 
-\subsection{TypeCoercion}
+\subsection{TypeCoercion} \label{sec:background-js-type-coercion}
 \todo{Hablar de WTFS}
 \todo{Aca incluir texto de what does it mean to be a JS Type}
-\subsection{NPM Packages}
+\subsection{NPM Packages} \label{sec:background-npm-packages}
 \todo{Agregar info de NPM by numbers! Cuantos downloads, cuantos packages, etc.}
@@
 
+\todo{Agregar info de cómo se resuelve el modulo y dónde se busca, etc -> Mirar link Requiring modules in Node.js: Everything you need to know}
+
 \section{TypeScript}
 \subsection{Declaration Files} \label{sec:declaration-files-background}
+\todo{Hablar de los tipos de declaration files que hay}
 \subsection{Definitely Typedd}

### 2019-08-10 chapters/4-approach.tex
Add examples for generation of declaration files

@@
 
-\subsection{Output Format}
-\todo{Explicar que es un JSON con todas las interactions. Explicar estructura de cómo se forma el JSON}
+\subsection{Output}
+Information is stored in a map containing \mintinline{bash}{FunctionContainer} entries.
+When the executions stops, Jalangi's callback \mintinline{bash}{endExecution()} is called and the information is saved as a JSON to a file.
 
 \subsection{Blacklisted Modules}
+When instrumenting a JS Library, Jalangi will try to instrument all required modules by resolving and finding the modules.
+Unfortunately, Jalangi's native behaviour is to stop the complete analysis if there is an error in the instrumentation of any of the required modules or in the script itself.
+
+The workaround introduced by the implementation modifies Jalangi's behaviour for failing instrumentation.
+For a given required module, it first tries to instrument it.
+If the instrumentation if unsuccesful, it simply skips the instrumentation and returns the original uninstrumented module.
+
+As expected, no Jalangi callbacks will be triggered for that module's code.
+However, this workaround enables to at least still get run-time information of other required modules by not breaking the complete analysis.
+As explained in \figref{fig:blacklisted-modules}, instead of stopping the analysis, failing modules are skipped but modules required by them do get instrumented anyway.
+Modifying Jalangi's instrumentation algorithm is out of the scope of this work.
+
+The application can be invoked by passing an optional argument with the path to a a JSON file containing the list of modules that should not be instrumented.
+
 \input{figures/approach/blacklisted-modules/blacklistedModules.tex}
-\todo{Incluir grafico con los blacklisted modules}
-
-\subsection{Implementation}
-\todo{NodeJS \& Docker. Mostrar cómo se invoca, etc.}
 
 \section{TypeScript Declaration File Generation}
+The actual generation of the declaration file is the next step in the pipeline after gathering the run-time information, as shown in \figref{fig:tsd_generation_method_block_diagram}.
+It uses the JSON output file containing the run-time information and generates a TypeScript declaration file.
+
+\input{figures/approach/typescript-declaration-files-generation-method/examples/generation-declaration-file.tex}
+
+\input{figures/approach/typescript-declaration-files-generation-method/examples/generation-declaration-file-with-interfaces.tex}
+
+TypeScript provides templates 
 \todo{Tres tipos de modulos en TypeScript}
@@
 \subsection{Overloading} \label{sec:typescript-overloading}
+\subsection{Type inference}
 \subsection{Implementation}

### 2019-08-10 figures/approach/typescript-declaration-files-generation-method/examples/code-intelligence-with-interfaces.tex
Add examples for generation of declaration files

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.48\textwidth}
			\jscode{code/typescript-declaration-file-generation/example-with-interfaces/index.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[index.js - Entry point to be executed with \mintinline{bash}{node index.js}]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.5\textwidth}
			\jscode{code/typescript-declaration-file-generation/example-with-interfaces/printer.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[printer.js - Module for which the TS Declaration File is going to be generated]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/typescript-declaration-file-generation/example-with-interfaces/printer.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[printer/index.d.ts]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.7\textwidth}
			\tscode{code/typescript-declaration-file-generation/example-with-interfaces/index.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[index.ts]{\usebox{\mintedbox}}

	\subfloat{\includegraphics[width=1\textwidth]{figures/approach/typescript-declaration-files-generation-method/examples/example-with-interfaces/code-intelligence-vscode.png}}

	\caption[Declaration file generation example]{\textbf{Declaration file generation example} - Declaration file for the \mintinline{bash}{calculator} module is generated using the \mintinline{bash}{module-class.d.ts} template. Declaration file can be used within a TypeScript project: code intelligence features work as expected and the generated JavaScript file after compilation can be correctly executed.}
	\label{fig:run-time-information-gathering-get-field}
\end{figure}

### 2019-08-10 figures/approach/typescript-declaration-files-generation-method/examples/generation-declaration-file-with-interfaces.tex
Add examples for generation of declaration files

--- 2019-08-10 figures/approach/typescript-declaration-files-generation-method/examples/code-intelligence-with-interfaces.tex
@@
 	\begin{lrbox}{\mintedbox}
-		\begin{minipage}{0.48\textwidth}
+		\begin{minipage}{0.45\textwidth}
 			\jscode{code/typescript-declaration-file-generation/example-with-interfaces/index.js}
@@
 	\begin{lrbox}{\mintedbox}
-		\begin{minipage}{0.5\textwidth}
+		\begin{minipage}{0.54\textwidth}
 			\jscode{code/typescript-declaration-file-generation/example-with-interfaces/printer.js}
@@
 	\begin{lrbox}{\mintedbox}
-		\begin{minipage}{0.8\textwidth}
+		\begin{minipage}{0.51\textwidth}
 			\tscode{code/typescript-declaration-file-generation/example-with-interfaces/printer.d.ts}
@@
 	\subfloat[printer/index.d.ts]{\usebox{\mintedbox}}
-
+	\hfill
 	\begin{lrbox}{\mintedbox}
-		\begin{minipage}{0.7\textwidth}
+		\begin{minipage}{0.48\textwidth}
 			\tscode{code/typescript-declaration-file-generation/example-with-interfaces/index.ts}
@@
 
-	\subfloat{\includegraphics[width=1\textwidth]{figures/approach/typescript-declaration-files-generation-method/examples/example-with-interfaces/code-intelligence-vscode.png}}
-
-	\caption[Declaration file generation example]{\textbf{Declaration file generation example} - Declaration file for the \mintinline{bash}{calculator} module is generated using the \mintinline{bash}{module-class.d.ts} template.
-Declaration file can be used within a TypeScript project: code intelligence features work as expected and the generated JavaScript file after compilation can be correctly executed.}
+	\caption[Interfaces inference]{\textbf{Interfaces inference} - Declaration file for the \mintinline{bash}{printer} module is generated using the \mintinline{bash}{module-function-d-ts} template.
+Interface of the input argument is generated correctly.
+The name of the interface is always generated as \mintinline{bash}{'I__ + argumentName'}, in this case \mintinline{bash}{I__input}.}
 	\label{fig:run-time-information-gathering-get-field}

### 2019-08-10 figures/approach/typescript-declaration-files-generation-method/examples/generation-declaration-file.tex
Add examples for generation of declaration files

--- 2019-08-10 figures/approach/typescript-declaration-files-generation-method/examples/code-intelligence-with-interfaces.tex
@@
 		\begin{minipage}{0.48\textwidth}
-			\jscode{code/typescript-declaration-file-generation/example-with-interfaces/index.js}
+			\jscode{code/typescript-declaration-file-generation/index.js}
 		\end{minipage}
@@
 		\begin{minipage}{0.5\textwidth}
-			\jscode{code/typescript-declaration-file-generation/example-with-interfaces/printer.js}
+			\jscode{code/typescript-declaration-file-generation/calculator.js}
 		\end{minipage}
 	\end{lrbox}
-	\subfloat[printer.js - Module for which the TS Declaration File is going to be generated]{\usebox{\mintedbox}}
-
-	\begin{lrbox}{\mintedbox}
-		\begin{minipage}{0.8\textwidth}
-			\tscode{code/typescript-declaration-file-generation/example-with-interfaces/printer.d.ts}
-		\end{minipage}
-	\end{lrbox}
-	\subfloat[printer/index.d.ts]{\usebox{\mintedbox}}
+	\subfloat[calculator.js - Module for which the TS Declaration File is going to be generated]{\usebox{\mintedbox}}
 
@@
 		\begin{minipage}{0.7\textwidth}
-			\tscode{code/typescript-declaration-file-generation/example-with-interfaces/index.ts}
+			\tscode{code/typescript-declaration-file-generation/calculator.d.ts}
+		\end{minipage}
+	\end{lrbox}
+	\subfloat[calculator/index.d.ts]{\usebox{\mintedbox}}
+
+	\begin{lrbox}{\mintedbox}
+		\begin{minipage}{0.7\textwidth}
+			\tscode{code/typescript-declaration-file-generation/index.ts}
 		\end{minipage}
@@
 	\subfloat[index.ts]{\usebox{\mintedbox}}
-
-	\subfloat{\includegraphics[width=1\textwidth]{figures/approach/typescript-declaration-files-generation-method/examples/example-with-interfaces/code-intelligence-vscode.png}}
-
 	\caption[Declaration file generation example]{\textbf{Declaration file generation example} - Declaration file for the \mintinline{bash}{calculator} module is generated using the \mintinline{bash}{module-class.d.ts} template.

### 2019-08-10 code/typescript-declaration-file-generation/calculator.js

@@ -1,0 +1,9 @@
+module.exports = function () {
+	this.sum = function (a, b) {
+		return a + b;
+	};
+
+	this.multiply = function (a, b) {
+		return a * b;
+	};
+};

### 2019-08-10 code/typescript-declaration-file-generation/example-with-interfaces/index.js

@@ -1,0 +1,12 @@
+var printer = require("./printer");
+
+var o = {
+	firstName: "John",
+	lastName: "Doe",
+	streetName: "Some Street",
+	houseNumber: "1234"
+};
+
+printer(o);
+
+// Prints: 'John Doe - Some Street 1234'

### 2019-08-10 code/typescript-declaration-file-generation/example-with-interfaces/index.ts

@@ -1,0 +1,8 @@
+import printer from './printer';
+
+printer({
+	firstName: 'John',
+	lastName: 'Doe',
+	streetName: 'Some Street',
+	houseNumber: '1234'
+});

### 2019-08-10 code/typescript-declaration-file-generation/example-with-interfaces/printer.d.ts

@@ -1,0 +1,12 @@
+export = Printer;
+
+declare function Printer(input: Printer.I__input): void;
+declare namespace Printer {
+	export interface I__input {
+		'firstName': string;
+		'lastName': string;
+		'streetName': string;
+		'houseNumber': string;
+	}
+
+}

### 2019-08-10 code/typescript-declaration-file-generation/example-with-interfaces/printer.js

@@ -1,0 +1,8 @@
+module.exports = function (input) {
+	var s = input.firstName + " " +
+		input.lastName + " - " +
+		input.streetName + " " +
+		input.houseNumber;
+
+	console.log(s);
+};

### 2019-08-10 code/typescript-declaration-file-generation/index.js

@@ -1,0 +1,6 @@
+var Calculator = require("./calculator");
+
+var c = new Calculator();
+
+c.sum(1, 1);
+c.multiply(10, 20);

### 2019-08-10 code/typescript-declaration-file-generation/index.ts

@@ -1,0 +1,6 @@
+import Calculator from './calculator';
+
+let calculator = new Calculator();
+
+calculator.sum(1, 2);
+calculator.multiply(10, 10);

### 2019-08-10 figures/approach/typescript-declaration-files-generation-method/examples/generation-declaration-file-with-nested-interfaces.tex
Add example for nested interfaces

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.7\textwidth}
			\jscode{code/typescript-declaration-file-generation/example-with-nested-interfaces/printer.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[printer.js]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.9\textwidth}
			\tscode{code/typescript-declaration-file-generation/example-with-nested-interfaces/printer.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[printer/index.d.ts]{\usebox{\mintedbox}}

	\caption[Nested interfaces]{\textbf{Nested interfaces} - Sub-interfaces for fields \mintinline{bash}{name} and \mintinline{bash}{address} are declared as separated interfaces and then associated to the main \mintinline{bash}{I__employee} interface.}
	\label{fig:run-time-information-gathering-get-field}
\end{figure}

### 2019-08-10 code/typescript-declaration-file-generation/example-with-nested-interfaces/printer.d.ts

@@ -1,0 +1,20 @@
+export = Printer;
+
+declare function Printer(employee: Printer.I__employee): void;
+declare namespace Printer {
+	export interface I__name {
+		'firstName': string;
+		'lastName': string;
+	}
+
+	export interface I__address {
+		'streetName': string;
+		'houseNumber': string;
+	}
+
+	export interface I__employee {
+		'name': I__name;
+		'address': I__address;
+	}
+
+}

### 2019-08-10 code/typescript-declaration-file-generation/example-with-nested-interfaces/printer.js

@@ -1,0 +1,8 @@
+module.exports = function (employee) {
+	var s = employee.name.firstName + " " +
+		employee.name.lastName + " - " +
+		employee.address.streetName + " " +
+		employee.address.houseNumber;
+
+	console.log(s);
+};

### 2019-08-11 chapters/4-approach.tex
Add example for nested interfaces
Finish TypeScript Declaration File Generation section

@@
 
-We introduce a \textit{TypeScript Declaration Files Generation Method}, which generates a valid TypeScript Declaration File for a specific JavaScript Library uploaded to the NPM Registry, as explained in \figref{fig:tsd_generation_method_block_diagram}.
+We introduce a \textit{TypeScript Declaration Files Generation Method}, \mintinline{text}{dts-generate}, which generates a valid TypeScript Declaration File for a specific JavaScript Library uploaded to the NPM Registry, as explained in \figref{fig:tsd_generation_method_block_diagram}.
 
@@
 
+The command line interface was inspired in the \mintinline{text}{dts-gen} \citep{dts-gen} package, a declaration files generator developed by Microsoft which is meant to be used as a starting point for writing declaration files.
+It can be seen in \coderef{code:dts-generate-example} that invoking the package is very simple and the only required argument is the name of the module published to the npm registry.
+
+\begin{code}
+  \begin{bashinline}
+$ ./dts-generate abs
+$ cat output/abs/index.d.ts 
+export = Abs;
+
+declare function Abs(input: string): string;
+
+    \end{bashinline}
+  \caption[Declaration file generation example]{\textbf{dts-generate usage} - Example of how to generate a declaration file for module \mintinline{text}{abs}.}
+  \label{code:dts-generate-example}
+\end{code}
+
 \todo{Agregar una imagen mostrando que es un unico script que hace todo. Mostrar imagenes de docker mostrando en qué está escrito cada cosa.}
 
-\section{Initial code base}
+\section{Initial code base} \label{sec:initial-code-base}
 To extract run-time information of a JavaScript Library, it is necessary, by definition, to actually execute the code.
@@
 \subsection{Repository}
-The \mintinline{bash}{npm view} command shows data about a package.
+The \mintinline{text}{npm view} command shows data about a package.
 It is possible to retrieve a specific field of the package registry by specifying fhe field name after the package descriptor.
-For example to retrieve information about the \textit{lodash} package, the following command can be used: \mintinline{bash}{npm view lodash}.
+For example to retrieve information about the \textit{lodash} package, the following command can be used: \mintinline{text}{npm view lodash}.
 
 For this particular case, only the url of the repository associated with the package is needed.
-This can be achieved by adding the correponding field name: \mintinline{bash}{npm view <PACKAGE> repository.url}.
+This can be achieved by adding the correponding field name: \mintinline{text}{npm view <PACKAGE> repository.url}.
 
@@
 After the repository url is extracted from the npm package, the url for downloading the readme file needs to be constructed performing some string manipulation.
-Finally, the readme file is obtained with a simple \mintinline{bash}{curl} or \mintinline{bash}{wget} command.
+Finally, the readme file is obtained with a simple \mintinline{text}{curl} or \mintinline{text}{wget} command.
 
@@
 It is very common to write code examples within code blocks indicating the programming language, so that it gets highlighted with the specific syntax.
-The Markdown identifiers for JavaScript are \mintinline{bash}{js} or \mintinline{bash}{javascript}.
+The Markdown identifiers for JavaScript are \mintinline{text}{js} or \mintinline{text}{javascript}.
 The code examples are finally retrieved by filtering the content within the corresponding code blocks.
@@
 \begin{enumerate}
-  \item Function \mintinline{bash}{f} got invoked with parameters \mintinline{bash}{a} and \mintinline{bash}{b} with types \mintinline{bash}{string} and \mintinline{bash}{number}.
-  \item Property or method \mintinline{bash}{foo} of parameter \mintinline{bash}{a} of function \mintinline{bash}{f} was accessed within the function.
-  \item Parameter \mintinline{bash}{a} of function \mintinline{bash}{f} was used as operand for operator \mintinline{bash}{==}.
+  \item Function \mintinline{text}{f} got invoked with parameters \mintinline{text}{a} and \mintinline{text}{b} with types \mintinline{text}{string} and \mintinline{text}{number}.
+  \item Property or method \mintinline{text}{foo} of parameter \mintinline{text}{a} of function \mintinline{text}{f} was accessed within the function.
+  \item Parameter \mintinline{text}{a} of function \mintinline{text}{f} was used as operand for operator \mintinline{text}{==}.
 \end{enumerate}
@@
 \begin{enumerate}
-  \item Binary operations, like \mintinline{bash}{==}, \mintinline{bash}{+} or \mintinline{bash}{===}.
+  \item Binary operations, like \mintinline{text}{==}, \mintinline{text}{+} or \mintinline{text}{===}.
   \item Variable declaration.
@@
   \item Access to an object's property.
-  \item Unary operations, like \mintinline{bash}{!} or \mintinline{bash}{typeof}.
+  \item Unary operations, like \mintinline{text}{!} or \mintinline{text}{typeof}.
 \end{enumerate}
@@
 The implementation wraps variables around Proxy Objects to store tracking data, hereby enabling to associate the observation with a specific argument.
-The effect is reverted using Jalangi's callbacks to modify the behaviour of the affected operators, such as \mintinline{bash}{===} or \mintinline{bash}{typeof}.
+The effect is reverted using Jalangi's callbacks to modify the behaviour of the affected operators, such as \mintinline{text}{===} or \mintinline{text}{typeof}.
 
@@
 Every invoked function is given an unique identifier and information for each argument of every function gets stored.
-The entities \mintinline{bash}{FunctionContainer}, \mintinline{bash}{ArgumentContainer} and \mintinline{bash}{Interaction} are introduced to support the representation of the gathered run-time information.
+The entities \mintinline{text}{FunctionContainer}, \mintinline{text}{ArgumentContainer} and \mintinline{text}{Interaction} are introduced to support the representation of the gathered run-time information.
 
-Run-time information is stored in a map that contains objects of type \mintinline{bash}{FunctionContainer}.
+Run-time information is stored in a map that contains objects of type \mintinline{text}{FunctionContainer}.
 This entity contains static information about the function as well as implementation specific data needed for generating the right type of declaration file, as explained in \secref{sec:declaration-files-background}.
@@
   \jscode{code/run-time-information-gathering/analysis/function-container.js}
-  \caption[FunctionContainer object]{\textbf{\mintinline{bash}{FunctionContainer} object} - \mintinline{bash}{FunctionContainer} is the main entity in the map where the run-time information gets gathered.
+  \caption[FunctionContainer object]{\textbf{\mintinline{text}{FunctionContainer} object} - \mintinline{text}{FunctionContainer} is the main entity in the map where the run-time information gets gathered.
 It contains implementation specific information for the declaration files generation.
-Entities of type \mintinline{bash}{ArgumentContainer} get appended to the \mintinline{bash}{args} property.}
+Entities of type \mintinline{text}{ArgumentContainer} get appended to the \mintinline{text}{args} property.}
   \label{code:function-container}
@@
 
-Moreover, as explained in \coderef{code:argument-container}, the \mintinline{bash}{ArgumentContainer} object works as a container for all interactions recorded for that argument.
+Moreover, as explained in \coderef{code:argument-container}, the \mintinline{text}{ArgumentContainer} object works as a container for all interactions recorded for that argument.
 Additionally, it contains simple information about the argument itself, like the index or the name given in the function declaration.
@@
   \jscode{code/run-time-information-gathering/analysis/argument-container.js}
-  \caption[ArgumentContainer object]{\textbf{\mintinline{bash}{ArgumentContainer} object} - The \mintinline{bash}{ArgumentContainer} object works mainly as a container for all relevant observations regarding that argument.}
+  \caption[ArgumentContainer object]{\textbf{\mintinline{text}{ArgumentContainer} object} - The \mintinline{text}{ArgumentContainer} object works mainly as a container for all relevant observations regarding that argument.}
   \label{code:argument-container}
@@
 
-Finally, observations on the arguments of a function are stored as objects of type \mintinline{bash}{Interaction}.
-They have to be always associated to an \mintinline{bash}{ArgumentContainer} object or to another \mintinline{bash}{Interaction}.
+Finally, observations on the arguments of a function are stored as objects of type \mintinline{text}{Interaction}.
+They have to be always associated to an \mintinline{text}{ArgumentContainer} object or to another \mintinline{text}{Interaction}.
 They are explained in detail in \secref{sec:run-time-interactions}.
@@
 
-Observations of the arguments are stored as \mintinline{bash}{Interactions}.
+Observations of the arguments are stored as \mintinline{text}{Interactions}.
 They represent the relevant observation data that needs to be extracted for the purpose of best describing a function in order to generate a TypeScript declaration file.
 
-The object \mintinline{bash}{Interaction} is associated to an \mintinline{bash}{ArgumentContainer}.
-If the \mintinline{bash}{ArgumentContainer} does not exist when the Jalangi callback is triggerd, the \mintinline{bash}{Interaction} is not saved.
+The object \mintinline{text}{Interaction} is associated to an \mintinline{text}{ArgumentContainer}.
+If the \mintinline{text}{ArgumentContainer} does not exist when the Jalangi callback is triggerd, the \mintinline{text}{Interaction} is not saved.
 The main goal of gathering run-time information is to describe a function and its arguments.
@@
   \jscode{code/run-time-information-gathering/analysis/ignored-observations.js}
-  \caption[Ignored observations]{\textbf{Ignored observations} - Variable \mintinline{bash}{b} gets declared within the scope of function \mintinline{bash}{foo}.
+  \caption[Ignored observations]{\textbf{Ignored observations} - Variable \mintinline{text}{b} gets declared within the scope of function \mintinline{text}{foo}.
 Therefore, no observation on it gets stored.}
@@
 As shown in \figref{fig:run-time-information-gathering-input-value}, it gets triggered when a function is invoked and it stores the input value of a specific argument.
-If a function gets invoked several times, different interactions get stored, each one of them with a different \mintinline{bash}{traceId}.
+If a function gets invoked several times, different interactions get stored, each one of them with a different \mintinline{text}{traceId}.
 
@@
 \subsubsection{\textit{getField} interaction}
-This interaction gets triggered whenever a property of an object gets accessed, either if the get field operation is \mintinline{bash}{obj.field} or \mintinline{bash}{obj['field']}.
+This interaction gets triggered whenever a property of an object gets accessed, either if the get field operation is \mintinline{text}{obj.field} or \mintinline{text}{obj['field']}.
 As shown in \figref{fig:run-time-information-gathering-get-field}, the name of the accessed field and the type of the actual property get stored.
 
-It should be noted that the field \mintinline{bash}{returnTypeOf} represents the type of the accessed property at the time when the interaction gets triggered.
+It should be noted that the field \mintinline{text}{returnTypeOf} represents the type of the accessed property at the time when the interaction gets triggered.
 The accessed property may somehow get modified afterwards, which will not modify the existing interaction.
-If the property gets accessed again, a new interaction with a new \mintinline{bash}{returnTypeOf} value will be added, as shown in \coderef{code:get-field-changing-at-runtime}.
+If the property gets accessed again, a new interaction with a new \mintinline{text}{returnTypeOf} value will be added, as shown in \coderef{code:get-field-changing-at-runtime}.
 
@@
   \jscode{code/run-time-information-gathering/analysis/get-field-changing-at-runtime.js}
-  \caption[Modified accessed property]{\textbf{Modified accessed property} - Variable \mintinline{bash}{b} gets declared within the scope of function \mintinline{bash}{foo}.
+  \caption[Modified accessed property]{\textbf{Modified accessed property} - Variable \mintinline{text}{b} gets declared within the scope of function \mintinline{text}{foo}.
 Therefore, no observation on it gets stored.}
@@
 This interaction gets triggered with the invocation of an object's method.
-As explained in \figref{fig:run-time-information-gathering-method-call}, information about the method itself is stored as another \mintinline{bash}{FunctionContainer} in the overall map.
-The reference to the \mintinline{bash}{FunctionContainer} is achieved through the properties \mintinline{bash}{functionId} and \mintinline{bash}{traceIdInTargetFunction}.
+As explained in \figref{fig:run-time-information-gathering-method-call}, information about the method itself is stored as another \mintinline{text}{FunctionContainer} in the overall map.
+The reference to the \mintinline{text}{FunctionContainer} is achieved through the properties \mintinline{text}{functionId} and \mintinline{text}{traceIdInTargetFunction}.
 
@@
 \subsubsection{\textit{setField} interaction}
-The trigger for this interaction is writing an object's property, either by accessing the field as \mintinline{bash}{obj.field = 1} or \mintinline{bash}{obj['field'] = 1}.
+The trigger for this interaction is writing an object's property, either by accessing the field as \mintinline{text}{obj.field = 1} or \mintinline{text}{obj['field'] = 1}.
 It can be seen in \figref{fig:run-time-information-gathering-set-field} that the interaction contains the name of the property to be written and the type of the value to be stored.
@@
 \subsubsection{\textit{usedAsArgument} interaction}
-The run-time information is stored in a map with one \mintinline{bash}{FunctionContainer} entry per function.
-As seen in \figref{fig:run-time-information-gathering-used-as-argument}, an argument of function \mintinline{bash}{foo()} is used as parameter to invoke a second function \mintinline{bash}{bar()}.
-Interactions triggered within function \mintinline{bash}{bar()} are going to be stored within the \mintinline{bash}{FunctionContainer} corresponding to function \mintinline{bash}{bar()}, i.e.
-not the one associated to function \mintinline{bash}{foo()}.
+The run-time information is stored in a map with one \mintinline{text}{FunctionContainer} entry per function.
+As seen in \figref{fig:run-time-information-gathering-used-as-argument}, an argument of function \mintinline{text}{foo()} is used as parameter to invoke a second function \mintinline{text}{bar()}.
+Interactions triggered within function \mintinline{text}{bar()} are going to be stored within the \mintinline{text}{FunctionContainer} corresponding to function \mintinline{text}{bar()}, i.e.
+not the one associated to function \mintinline{text}{foo()}.
 
-However, for the example provided in \figref{fig:run-time-information-gathering-used-as-argument} it is relevant to also associate interactions triggered within function \mintinline{bash}{bar()} to the function \mintinline{bash}{foo()}, since they can  be useful for describing the argument of \mintinline{bash}{foo()} more accurately.
+However, for the example provided in \figref{fig:run-time-information-gathering-used-as-argument} it is relevant to also associate interactions triggered within function \mintinline{text}{bar()} to the function \mintinline{text}{foo()}, since they can  be useful for describing the argument of \mintinline{text}{foo()} more accurately.
 
-Therefore, \mintinline{bash}{usedAsArgument} interaction is introduced.
+Therefore, \mintinline{text}{usedAsArgument} interaction is introduced.
 It gets triggered when an argument is used in the invocation of a consecutive function.
 It serves the purpose of connecting related interactions.
-As shown in \figref{fig:run-time-information-gathering-used-as-argument}, it mainly contains information for identifying the target \mintinline{bash}{FunctionContainer}.
+As shown in \figref{fig:run-time-information-gathering-used-as-argument}, it mainly contains information for identifying the target \mintinline{text}{FunctionContainer}.
 
@@
 Identifying if an argument is used with a specific operator is very important for infering the type of that variable.
-Relational operators like \mintinline{bash}{<, >, <=, >=} are mainly used with type \mintinline{bash}{number}.
-On the other side, operator \mintinline{bash}{+} is mainly used with \mintinline{bash}{number} or \mintinline{bash}{string}.
+Relational operators like \mintinline{text}{<, >, <=, >=} are mainly used with type \mintinline{text}{number}.
+On the other side, operator \mintinline{text}{+} is mainly used with \mintinline{text}{number} or \mintinline{text}{string}.
 Empirical evidence of how operators are used is presented in \secref{sec:experiments-js-operators-usage}.
 
-The \mintinline{bash}{operator} interaction stores the operator being used and the type of the left hand and right hand operands.
+The \mintinline{text}{operator} interaction stores the operator being used and the type of the left hand and right hand operands.
 It also indicates which operand is the one that triggered the interaction.
-\figref{fig:run-time-information-gathering-operator} provides an example for operator \mintinline{bash}{+}.
+\figref{fig:run-time-information-gathering-operator} provides an example for operator \mintinline{text}{+}.
 
-The supported operators are the ones that trigger Jalangi's \mintinline{bash}{binaryPre()} callback.
-Conditionals are merged into one operator name \mintinline{bash}{conditional}, since Jalangi's \mintinline{bash}{conditional()} callback does not offer a way of identifying the operator for this cases.
+The supported operators are the ones that trigger Jalangi's \mintinline{text}{binaryPre()} callback.
+Conditionals are merged into one operator name \mintinline{text}{conditional}, since Jalangi's \mintinline{text}{conditional()} callback does not offer a way of identifying the operator for this cases.
 Both callbacks are explained in section \secref{sec:jalangi}.
-Right hand operand is left as \mintinline{bash}{undefined} for that case.
+Right hand operand is left as \mintinline{text}{undefined} for that case.
 
@@
 \subsubsection{\textit{followingInteractions}}
-If a property of an argument is accessed the Jalangi callback \mintinline{bash}{getField()} gets triggered and a \mintinline{bash}{getField} interaction gets associated to the corresponding \mintinline{bash}{ArgumentContainer}.
+If a property of an argument is accessed the Jalangi callback \mintinline{text}{getField()} gets triggered and a \mintinline{text}{getField} interaction gets associated to the corresponding \mintinline{text}{ArgumentContainer}.
 In order to describe the interface of the argument more accurately, it is also relevant to recursively record if properties of the return values are accessed, as explained in \figref{fig:run-time-information-gathering-following-interactions}.
 
-This was implemented by not only mapping the observation with an \mintinline{bash}{ArgumentContainer} but also with a \mintinline{bash}{getField} or \mintinline{bash}{methodCall} interaction.
-Both \mintinline{bash}{ArgumentContainer} and \mintinline{bash}{Interaction} implement the \mintinline[breaklines]{javascript}{addInteraction} method, as shown in \coderef{code:add-interaction-example}
+This was implemented by not only mapping the observation with an \mintinline{text}{ArgumentContainer} but also with a \mintinline{text}{getField} or \mintinline{text}{methodCall} interaction.
+Both \mintinline{text}{ArgumentContainer} and \mintinline{text}{Interaction} implement the \mintinline[breaklines]{javascript}{addInteraction} method, as shown in \coderef{code:add-interaction-example}
 
@@
   \jscode{code/run-time-information-gathering/analysis/add-interaction-example.js}
-  \caption[addInteraction method]{\textbf{addInteraction method} - Example showing how interactions can be associated with \mintinline{bash}{ArgumentContainer} or with another \mintinline{bash}{Interaction}.}
+  \caption[addInteraction method]{\textbf{addInteraction method} - Example showing how interactions can be associated with \mintinline{text}{ArgumentContainer} or with another \mintinline{text}{Interaction}.}
   \label{code:add-interaction-example}
@@
 \subsubsection{Traces}
-Different executions of a function are identified through different \mintinline{bash}{traceIds}.
-Each interaction and return value of a \mintinline{bash}{FunctionContainer} stores the corresponding \mintinline{bash}{traceId}.
-The \mintinline{bash}{traceId} field enables identifying correlations between input types and return values, as explained in \secref{sec:typescript-overloading}.
+Different executions of a function are identified through different \mintinline{text}{traceIds}.
+Each interaction and return value of a \mintinline{text}{FunctionContainer} stores the corresponding \mintinline{text}{traceId}.
+The \mintinline{text}{traceId} field enables identifying correlations between input types and return values, as explained in \secref{sec:typescript-overloading}.
 An example is provided in \figref{fig:run-time-information-gathering-traces}.
@@
 \subsection{Wrapper Objects} \label{sec:run-time-wrapper-objects}
-An \mintinline{bash}{Interaction} needs to be associated to an \mintinline{bash}{ArgumentContainer} or to another \mintinline{bash}{Interaction}.
+An \mintinline{text}{Interaction} needs to be associated to an \mintinline{text}{ArgumentContainer} or to another \mintinline{text}{Interaction}.
 This means that when Jalangi triggers a callback, it must be determined if that callback is relevant or not, i.e.
@@
 However, Jalangi exposes the actual value affected by the callback.
-For example, for the \mintinline{bash}{getField()} callback triggered by operation \mintinline{javascript}{myObj.prop}, Jalangi does provide the object \mintinline{javascript}{myObj} in the callback.
+For example, for the \mintinline{text}{getField()} callback triggered by operation \mintinline{javascript}{myObj.prop}, Jalangi does provide the object \mintinline{javascript}{myObj} in the callback.
 
 The solution consists in wrapping every input argument around a Wrapper Object when the function gets invoked.
-The Wrapper Object will store the reference to the corresponding \mintinline{bash}{ArgumentContainer} or \mintinline{bash}{Interaction}.
+The Wrapper Object will store the reference to the corresponding \mintinline{text}{ArgumentContainer} or \mintinline{text}{Interaction}.
 Jalangi callbacks will then get triggered on the Wrapper Object instead.
-A mapping with the corresponding \mintinline{bash}{ArgumentContainer} or \mintinline{bash}{Interaction} is then straightforward, since the reference can be directly retrieved from the Wrapper Object.
+A mapping with the corresponding \mintinline{text}{ArgumentContainer} or \mintinline{text}{Interaction} is then straightforward, since the reference can be directly retrieved from the Wrapper Object.
 
 Jalangi enables to modify the behaviour of the original code.
-Using \mintinline{bash}{invokFunPre} and \mintinline{javascript}{declare} callbacks, a function invocation is detected and the arguments of the function are modified in run-time.
+Using \mintinline{text}{invokFunPre} and \mintinline{javascript}{declare} callbacks, a function invocation is detected and the arguments of the function are modified in run-time.
 
 Every function invocation generates new Wrapper Objects, which means that a Wrapper Object represents the input argument for a specific invocation.
-\mintinline{bash}{TraceIds} are also stored within these objects.
+\mintinline{text}{TraceIds} are also stored within these objects.
 
-The solution extends a similar concept already implemented by Jalangi: \mintinline{bash}{ShadowObjects}.
+The solution extends a similar concept already implemented by Jalangi: \mintinline{text}{ShadowObjects}.
 It is a unique object associated to another object which can store meta-information.
-However, a \mintinline{bash}{ShadowObject} can not be created for primitives values, which is a major drawback.
-It is indeed relevant to capture, for example, if a \mintinline{bash}{number} is being used with a particular operator or if the property \mintinline{bash}{length} of a \mintinline{bash}{string} is being accessed.
+However, a \mintinline{text}{ShadowObject} can not be created for primitives values, which is a major drawback.
+It is indeed relevant to capture, for example, if a \mintinline{text}{number} is being used with a particular operator or if the property \mintinline{text}{length} of a \mintinline{text}{string} is being accessed.
 
 The implemented solution wraps objects around JavaScript Proxies.
-Primitive values like \mintinline{bash}{number} or \mintinline{bash}{string} are converted to their native wapper objects, i.e.
-\mintinline{bash}{Number} and \mintinline{bash}{String}.
-Finally, \mintinline{bash}{null} and \mintinline{bash}{undefined} are converted to empty objects with modified \mintinline{bash}{valueOf()} and \mintinline{bash}{toString()} methods.
+Primitive values like \mintinline{text}{number} or \mintinline{text}{string} are converted to their native wapper objects, i.e.
+\mintinline{text}{Number} and \mintinline{text}{String}.
+Finally, \mintinline{text}{null} and \mintinline{text}{undefined} are converted to empty objects with modified \mintinline{text}{valueOf()} and \mintinline{text}{toString()} methods.
 
 The behaviour of the original code is being explicitely modified.
-Therefore, original values are used instead for affected operators like \mintinline{bash}{===, !==, typeof}, for which wrapper objects will not work as expected.
-For \mintinline{bash}{null} and \mintinline{bash}{undefined} wrappers, the original values are used for conditional operators, since the result of the abstract operation \mintinline{bash}{ToBoolean} on an object is always \mintinline{text}{true}, as defined in the ECMAScript Language Specification \citep{ecma-script}.
+Therefore, original values are used instead for affected operators like \mintinline{text}{===, !==, typeof}, for which wrapper objects will not work as expected.
+For \mintinline{text}{null} and \mintinline{text}{undefined} wrappers, the original values are used for conditional operators, since the result of the abstract operation \mintinline{text}{ToBoolean} on an object is always \mintinline{text}{true}, as defined in the ECMAScript Language Specification \citep{ecma-script}.
 This approach could have been used for every operation by using the original value in the affected Jalangi callbacks.
@@
 \subsection{Output}
-Information is stored in a map containing \mintinline{bash}{FunctionContainer} entries.
-When the executions stops, Jalangi's callback \mintinline{bash}{endExecution()} is called and the information is saved as a JSON to a file.
+Information is stored in a map containing \mintinline{text}{FunctionContainer} entries.
+When the executions stops, Jalangi's callback \mintinline{text}{endExecution()} is called and the information is saved as a JSON to a file.
 
@@
 The actual generation of the declaration file is the next step in the pipeline after gathering the run-time information, as shown in \figref{fig:tsd_generation_method_block_diagram}.
-It uses the JSON output file containing the run-time information and generates a TypeScript declaration file.
+It is a lightweight, simple and fast application, which does not interact with the actual JavaScript module at run-time.
+Instead, it uses the JSON output file containing the run-time information and generates a TypeScript declaration file which is use ready to be used within a TypeScript project.
+The tool itself is written in TypeScript and runs within a Docker container in NodeJS.
 
-\input{figures/approach/typescript-declaration-files-generation-method/examples/generation-declaration-file.tex}
+\subsection{Templates}
+TypeScript provides templates for writing declaration files, as explained in \secref{sec:declaration-files-background}.
+Each template corresponds to a different way of exporting a JS module.
+The code uses different fields from the \mintinline{text}{FunctionContainer} entity in the run-time information to detect how the module is being used:
+
+\begin{itemize}
+  \item \mintinline{text}{requiredModule}: The name of the module that exported the corresponding function.
+It is useful for processing only functions exported by the module for which the declaration files is being generated.
+  \item \mintinline{text}{isExported}: Boolean that indicates whether the function is exported when importing the module.
+  \item \mintinline{text}{isConstructor}: Boolean that indicates whether the function was used as a constructor through the \mintinline{text}{new} operator.
+  \item \mintinline{text}{constructedBy}: Indicates the \mintinline{text}{functionId} that constructed the object invoking the corresponding function.
+\end{itemize}
+
+The following templates were implemented:
+
+\begin{enumerate}
+  \item \mintinline{text}{module.d.ts}: It is used when the exported module is a JavaScript object containing several properties and methods.
+It is the most common used template: more than 70\% of the modules uploaded to DefinitelyTyped are based on this template.
+The code checks if the exported module is neither a function nor a constructor, i.e.
+if properties \mintinline{text}{isExported} and \mintinline{text}{isConstructor} are \mintinline{text}{false}.
+An example is provided in \figref{fig:ts-declaration-file-module-template}.
+
+  \item \mintinline{text}{module-class.d.ts}: As shown in \figref{fig:ts-declaration-file-module-class-template}, this template is used when the exported function is supposed to be used as a constructor with the \mintinline{text}{new} operator.
+The code will use this template when both the fields \mintinline{text}{isExported} and \mintinline{text}{isConstructor} are true.
+It should be noted that the module exports only the constructor function.
+Methods belonging to the class are detected using the \mintinline{text}{constructedBy} field from the run-time data.
+
+  \item \mintinline{text}{module-function.d.ts}: It is used when the module exports only a function.
+The field \mintinline{text}{isExported} field is used to determine whether this templates should be used or not.
+An example is provided in \figref{fig:ts-declaration-file-module-function-template}.
+If further interfaces or types are required, they are written within a namespace named like the exported function.
+\end{enumerate}
+
+\input{figures/approach/typescript-declaration-files-generation-method/examples/templates/module-template.tex}
+
+\input{figures/approach/typescript-declaration-files-generation-method/examples/templates/module-class-template.tex}
+
+\input{figures/approach/typescript-declaration-files-generation-method/examples/templates/module-function-template.tex}
+
+JavaScript construtors where explained in \secref{sec:background-javascript}.
+There are built-in functions that are not able to be used as a constructors.
+For example, if \mintinline{text}{Math} is intended to be used as a constructor with the \mintinline{text}{new} operator, the following error will arise: \mintinline{text}{TypeError: Math is not a constructor}.
+ES6 arrow functions are not allowed to used as a constructor either.
+However, for those functions that can be both used as constructors or called as regular functions, there is no difference between them that actually indicate whether a function is intended to be used with the \mintinline{text}{new} operator or not.
+The choice of using the \mintinline{text}{module-function.d.ts} or the \mintinline{text}{module-class.d.ts} template is made after executing the code and depends on how the function was used.
+
+\coderef{code:typescript-templates-function-as-constructor} exposes a scenario for which a module could be infered as \mintinline{text}{module-function.d.ts} or \mintinline{text}{module-class.d.ts} depending only on how it was executed.
+Thus, examples that execute the module under analysis by truly representing the developer's intention are crucial for an accurate generation of declaration files and inference of types and interfaces.
+As explained in \secref{sec:initial-code-base}, this was achieved by extracting the code examples written by the developers from the repositories.
+
+\begin{code}
+  \jscode{code/typescript-declaration-file-generation/templates/function-as-constructor.js}
+  \caption[Choice between module-function.d.ts or module-class.d.ts]{\textbf{Choice between module-function.d.ts or module-class.d.ts} - If the exported module is used without the \mintinline{text}{new} operator, the \mintinline{text}{module-function.d.ts} template will be used.
+If used with the \mintinline{text}{new} operator, the \mintinline{text}{module-class.d.ts} template will be applied.}
+  \label{code:typescript-templates-function-as-constructor}
+\end{code}
+
+\subsection{Interfaces}
+Interfaces are created by exploring \mintinline{text}{getField} and \mintinline{text}{methodCall} interactions from the run-time information.
+The code will gather the interactions for a specific argument and build the interface by incrementally adding new properties.
+Interactions within the \mintinline{text}{followingInteractions} field are recursively traversed, building a new interface in each recursion level.
+
+An example is provided in \figref{fig:ts-declaration-file-generation-interfaces}. The name given to the interface is based on the name of the input argument.
+
+It can be also seen in \figref{fig:ts-declaration-file-generation-nested-interfaces} that nested interfaces are declared as independent interfaces and then declared as types of the prior interface.
 
@@
 
-TypeScript provides templates 
-\todo{Tres tipos de modulos en TypeScript}
-\todo{Traceids y que las signatures de las funciones no se unen, sino que se acumulan segun traceId. Obviamente las iguales no se escriben de vuelta.}
-\todo{El nombre de las interfaces se crea con el nombre de la variable}
+\input{figures/approach/typescript-declaration-files-generation-method/examples/generation-declaration-file-with-nested-interfaces.tex}
+
+\subsection{Type inference}
 \todo{El tipo se extrae del valor en runtime.}
-\subsection{Overloading} \label{sec:typescript-overloading}
-\subsection{Type inference}
 \subsection{Implementation}
-\todo{TypeScript \& Docker. Mostrar cómo se invoca, etc.}
+The application runs within a Docker container.
+\coderef{code:declaration-file-generation-docker} explains how to build and run the application and how to extract the declaration file from the container.
+
+\begin{code}
+\begin{bashinline}
+$ ./build/build.sh
+$ docker run --name generate-declaration-file \
+  -v PATH_TO_RUN_TIME_INFORMATION_JSON:/tmp/output.json \
+  tsd-generator \
+  --module-name calculator \
+  -i /tmp/output.json
+
+$ docker cp generate-declaration-file:/usr/local/app/output/. /tmp/ts-declaration-file
+$ docker rm generate-declaration-file
+$ cat /tmp/ts-declaration-file/calculator/index.d.ts
+export function sum(a: number, b: number): number;
+export function multiply(a: number, b: number): number;
+\end{bashinline}
+  \caption[Declaration file generation]{\textbf{Declaration file generation} - The Docker image needs to be built after cloning the repository.
+Afterwards, the declaration file is generated by the run-time information and then retrieved from the container.}
+  \label{code:declaration-file-generation-docker}
+\end{code}
 

### 2019-08-11 chapters/5-experiments.tex
Add funnel with number of modules for each stage of the experiment
Add usedAsArgument and operator interactions
Finish TypeScript Declaration File Generation section

@@
 
-\input{figures/experiments/figExample}
-\input{figures/experiments/tabAccuracy}
+\input{figures/experiments/overallFunnel.tex}
 
 \section{Definitely Typed Modules}
+\todo{Poner cómo es la distribución de los tipos de declaration files.}
+
+\draft{
+module-class	503	8,27%
+module	4379	71,99%
+module-function	1201	19,74%
+}
+
+\draft{Chequear modulo 'acl' por ejempo que tiene errores y esta exportado como module pero deberia ser exportado como module-class. Habia otro modulo pero no lo recuerdo.}
 
@@
 
-\section{JS Operators Usage}
+\section{JS Operators Usage} \label{sec:experiments-js-operators-usage}
 \todo{Agregar que es una matriz simetrica.

### 2019-08-11 chapters/55-related.tex
Finish TypeScript Declaration File Generation section

\chapter{Related Work}\label{chap:relatedwork}
Give a brief overview of the work relevant for your thesis.
Closely related work should be compared in depth: what are the pros and cons of their work compared to yours?

\todo{dts-gen package: https://github.com/microsoft/dts-gen}

### 2019-08-11 figures/approach/typescript-declaration-files-generation-method/examples/templates/module-class-template.tex
Finish TypeScript Declaration File Generation section

--- 2019-08-10 figures/approach/typescript-declaration-files-generation-method/examples/generation-declaration-file.tex
@@
 		\begin{minipage}{0.48\textwidth}
-			\jscode{code/typescript-declaration-file-generation/index.js}
+			\jscode{code/typescript-declaration-file-generation/templates/module-class/index.js}
 		\end{minipage}
@@
 		\begin{minipage}{0.5\textwidth}
-			\jscode{code/typescript-declaration-file-generation/calculator.js}
+			\jscode{code/typescript-declaration-file-generation/templates/module-class/calculator.js}
 		\end{minipage}
@@
 		\begin{minipage}{0.7\textwidth}
-			\tscode{code/typescript-declaration-file-generation/calculator.d.ts}
+			\tscode{code/typescript-declaration-file-generation/templates/module-class/calculator.d.ts}
 		\end{minipage}
@@
 		\begin{minipage}{0.7\textwidth}
-			\tscode{code/typescript-declaration-file-generation/index.ts}
+			\tscode{code/typescript-declaration-file-generation/templates/module-class/index.ts}
 		\end{minipage}
 	\end{lrbox}
-	\subfloat[index.ts]{\usebox{\mintedbox}}
-	\caption[Declaration file generation example]{\textbf{Declaration file generation example} - Declaration file for the \mintinline{bash}{calculator} module is generated using the \mintinline{bash}{module-class.d.ts} template.
-Declaration file can be used within a TypeScript project: code intelligence features work as expected and the generated JavaScript file after compilation can be correctly executed.}
-	\label{fig:run-time-information-gathering-get-field}
+	\subfloat[index.ts - TypeScript file using the generated declaration file]{\usebox{\mintedbox}}
+	\caption[module-class.d.ts template example]{\textbf{module-class.d.ts template example} - Declaration file for the \mintinline{bash}{calculator} module is generated using the \mintinline{bash}{module-class.d.ts} template.
+Exported module can be used as a constructor within a TypeScript project.}
+	\label{fig:ts-declaration-file-module-class-template}
 \end{figure}

### 2019-08-11 figures/approach/typescript-declaration-files-generation-method/examples/templates/module-function-template.tex
Finish TypeScript Declaration File Generation section

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.9\textwidth}
			\jscode{code/typescript-declaration-file-generation/templates/module-function/index.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[index.js - Entry point to be executed with \mintinline{bash}{node index.js}]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.9\textwidth}
			\jscode{code/typescript-declaration-file-generation/templates/module-function/snake-to-camel.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[snake-to-camel.js - Module for which the TS Declaration File is going to be generated]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/typescript-declaration-file-generation/templates/module-function/index.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[snake-to-camel/index.d.ts]{\usebox{\mintedbox}}

	\caption[module-function.d.ts template example]{\textbf{module-function.d.ts template example} - Declaration file for the \mintinline{bash}{snake-to-camel} module is generated using the \mintinline{bash}{module-function.d.ts} template.}
	\label{fig:ts-declaration-file-module-function-template}
\end{figure}

### 2019-08-11 figures/approach/typescript-declaration-files-generation-method/examples/templates/module-template.tex
Finish TypeScript Declaration File Generation section

--- 2019-08-10 figures/approach/typescript-declaration-files-generation-method/examples/generation-declaration-file.tex
@@
 	\begin{lrbox}{\mintedbox}
-		\begin{minipage}{0.48\textwidth}
-			\jscode{code/typescript-declaration-file-generation/index.js}
+		\begin{minipage}{0.7\textwidth}
+			\jscode{code/typescript-declaration-file-generation/templates/module/index.js}
 		\end{minipage}
@@
 	\subfloat[index.js - Entry point to be executed with \mintinline{bash}{node index.js}]{\usebox{\mintedbox}}
-	\hfill
+
 	\begin{lrbox}{\mintedbox}
-		\begin{minipage}{0.5\textwidth}
-			\jscode{code/typescript-declaration-file-generation/calculator.js}
+		\begin{minipage}{0.7\textwidth}
+			\jscode{code/typescript-declaration-file-generation/templates/module/calculator.js}
 		\end{minipage}
@@
 	\begin{lrbox}{\mintedbox}
-		\begin{minipage}{0.7\textwidth}
-			\tscode{code/typescript-declaration-file-generation/calculator.d.ts}
+		\begin{minipage}{0.9\textwidth}
+			\tscode{code/typescript-declaration-file-generation/templates/module/index.d.ts}
 		\end{minipage}
@@
 
-	\begin{lrbox}{\mintedbox}
-		\begin{minipage}{0.7\textwidth}
-			\tscode{code/typescript-declaration-file-generation/index.ts}
-		\end{minipage}
-	\end{lrbox}
-	\subfloat[index.ts]{\usebox{\mintedbox}}
-	\caption[Declaration file generation example]{\textbf{Declaration file generation example} - Declaration file for the \mintinline{bash}{calculator} module is generated using the \mintinline{bash}{module-class.d.ts} template.
-Declaration file can be used within a TypeScript project: code intelligence features work as expected and the generated JavaScript file after compilation can be correctly executed.}
-	\label{fig:run-time-information-gathering-get-field}
+	\caption[module.d.ts template example]{\textbf{module.d.ts template example} - Declaration file for the \mintinline{bash}{calculator} module is generated using the \mintinline{bash}{module.d.ts} template.}
+	\label{fig:ts-declaration-file-module-template}
 \end{figure}

### 2019-08-11 code/typescript-declaration-file-generation/templates/function-as-constructor.js

@@ -1,0 +1,15 @@
+function ToUpperCase(a) {
+	this.run = function () {
+		return a.toUpperCase();
+	};
+
+	return this.run(a);
+}
+
+ToUpperCase("hello");
+// "HELLO"
+
+
+var t = new ToUpperCase("world");
+t.run();
+// "WORLD"

### 2019-08-11 code/typescript-declaration-file-generation/templates/module-function/index.d.ts

@@ -1,0 +1,3 @@
+export = SnakeToCamel;
+
+declare function SnakeToCamel(s: string): string;

### 2019-08-11 code/typescript-declaration-file-generation/templates/module-function/index.js

@@ -1,0 +1,5 @@
+var snakeToCamel = require("./snake-to-camel");
+
+var camelCaseString = snakeToCamel("my-awesome-snake-string");
+
+// 'myAwesomeSnakeString'

### 2019-08-11 code/typescript-declaration-file-generation/templates/module-function/snake-to-camel.js

@@ -1,0 +1,9 @@
+function snakeToCamel(s) {
+	return s.replace(/([-_][a-z])/g, function (group) {
+		return group.toUpperCase()
+			.replace('-', '')
+			.replace('_', '');
+	});
+}
+
+module.exports = snakeToCamel;

### 2019-08-11 code/typescript-declaration-file-generation/templates/module/calculator.js

@@ -1,0 +1,8 @@
+module.exports = {
+	sum: function (a, b) {
+		return a + b;
+	},
+	multiply: function (a, b) {
+		return a * b;
+	}
+};

### 2019-08-11 code/typescript-declaration-file-generation/templates/module/index.d.ts

@@ -1,0 +1,2 @@
+export function sum(a: number, b: number): number;
+export function multiply(a: number, b: number): number;

### 2019-08-11 code/typescript-declaration-file-generation/templates/module/index.js

@@ -1,0 +1,4 @@
+var c = require("./calculator");
+
+c.sum(1, 2);
+c.sum(10, 20);

### 2019-08-13 chapters/4-approach.tex
Add content to the TypeInference section

@@
 
-\section{Problem Definition}
+\section{Problem Definition} \label{sec:problem-definition}
 In spite of the fact that JavaScript was originally intended for simple web-pages scripting, it's popularity has increased over the last years, being the most commonly used programming language among professional developers in 2018.
-Development of robust and high-quality frameworks such as jQuery, AngularJS, React or Vue.js contributed to JavaScript being the leading programming language for client-side development.
+Development of robust and high-quality frameworks such as jQuery\footnote{https://jquery.com}, AngularJS\footnote{https://angularjs.org}, React\footnote{https://reactjs.org} or Vue.js\footnote{https://vuejs.org} contributed to JavaScript being the leading programming language for client-side development.
 
 Moreover, several back-end applications are being written in JavaScript using Node.js, thus unifying the programming languages of client and back-end code.
-Companies with architectures based on concepts such as serverless and microservices are moving traditional back-end code written in Java, Python or PHP to the cloud and replacing it by small compute services, like AWS Lambda or Google Cloud Functions, or directly by cloud managed services.
+Companies with architectures based on concepts such as serverless and micro-services are moving traditional back-end code written in Java, Python or PHP to the cloud and replacing it by small compute services, like AWS Lambda or Google Cloud Functions, or directly by cloud managed services.
 It is very common that development teams decide to write the compute services in Node.js, thus reducing the technological stack.
 
-However, JavaScript was not intented for developing large-scale applications.
+However, JavaScript was not intended for developing large-scale applications.
 Lack of good interface descriptions using types and unintuitive type coercion have been slowing development of JavaScript applications.
@@
 
-Aware that types are important for developing maintainable software systems, several industrial developers started using TypeScript in JavaScript projects.
+Types are important for developing maintainable software systems.
+Therefore, several industrial developers started using TypeScript in JavaScript projects.
 However, despite TypeScript's popularity, many new and existing JavaScript Libraries are still being written in pure JavaScript.
 
-The possibility of using exissting JS libraries in TypeScript projects is critical for a smooth transition and for its adoption in industrial development.
+The possibility of using existing JS libraries in TypeScript projects is critical for a smooth transition and for its adoption in industrial development.
 TypeScript enables it through \textit{declaration files}, a typed description of the library's API.
-A public repository, DefinitelyTyped \citep{definitely-typed-repository}, contains declaration files for more than 6000 JS Libraries.
+A public repository, DefinitelyTyped, contains declaration files for more than 6000 JS Libraries \citep{definitely-typed-repository}.
 71\% of the top 1000 depended-upon packages have their corresponding declaration file in DefinitelyTyped, as explained in \chapref{chap:background}.
@@
 Unfortunately, such files are written and maintained manually, which is tedious, error prone and highly time consuming.
-The evolution of the declaration files does not follow in some cases the continous development of the corresponding libraries.
+The evolution of the declaration files does not follow in some cases the continuous development of the corresponding libraries.
 As a result, some declaration files end up being out of date, thus hindering their usability.
-Mismatches between the declaration files and the actual JavaScript Library code have a negative impact on developers.
+Discrepancies between the declaration files and the actual JavaScript Library code have a negative impact on developers.
 Type checking messages and IDE code intelligence features are not accurate, thus increasing developing costs, generating actually more difficulties than before migrating to TypeScript.
@@
 Declaration files are critical for TypeScript projects.
-TypeScript's popularity increase and the continuous growth of packages uploaded to the NPM registry, surpassing \textit{1 million} packages by June 2019, indicate that mismatches between declaration files and JavaScript implementations will likely grow.
-This Thesis tackles the declaration file generation problem and proposes a solution for assiting programmers in creating TypeScript Declaration Files from JavaScript code and keeping them up to date after.
-
-\todo{Agregar una imagen de un modulo simple con codigo JavaScript y que del otro lado salga el declaration file.}
+The increase of TypeScript's popularity and the continuous growth of packages uploaded to the NPM registry, surpassing 1 million packages by June 2019, indicate that mismatches between declaration files and JavaScript implementations will likely grow.
+This Thesis tackles the declaration file generation problem and proposes a solution for assisting programmers in creating TypeScript Declaration Files from JavaScript code and keeping them up to date.
 
@@
 
-Following the goal of easing and automating the generation of TypeScript Declaration Files for existings JavaScript Libraries, the method receives the name of the JavaScript Library published to the NPM Registry as an input argument and returns the corresponding TypeScript declaration file.
+Following the goal of easing and automating the generation of TypeScript Declaration Files for existing JavaScript Libraries, the method receives the name of the JavaScript Library published to the NPM Registry as an input argument and returns the corresponding TypeScript declaration file.
 
@@
 It contains no errors and matches the structure of the JavaScript Library under analysis, so that the JavaScript code generated after compilation runs.
-The conducted experiments included tests that consisted on replacing a specific type definition from DefinitelyTyped \citep{definitely-typed-repository} with the one generated in the experiments: TypeScript compilation was sucessful, the generated JavaScript code ran without errors and code intelligence features performed by IDEs like code completion worked as expected.
+The conducted experiments included tests that consisted on replacing a specific type definition from DefinitelyTyped \citep{definitely-typed-repository} with the one generated in the experiments: TypeScript compilation was successful, the generated JavaScript code ran without errors and code intelligence features performed by IDEs like code completion worked as expected.
 
@@
 It is achieved by retrieving the examples provided in the README files of the repositories of the different libraries.
-This is generally the place where developers explicitely show how to use their code.
+This is generally the place where developers explicitly show how to use their code.
 It showed to be an appropriate and pragmatic way of extracting the developer's intention and providing an useful initial code base with meaningful examples, thus avoiding a possible cold start problem.
@@
 
-\todo{Agregar una imagen mostrando que es un unico script que hace todo. Mostrar imagenes de docker mostrando en qué está escrito cada cosa.}
-
 \section{Initial code base} \label{sec:initial-code-base}
@@
 
-We decided to extract the code examples from the Readme files of the repository associated to the NPM Package.
+It was decided to extract the code examples from the Readme files of the repository associated to the NPM Package.
 Readme files are usually used by developers for briefly describing what the code does, what problem it solves, how to install the application, how to build the code, etc.
@@
 The \mintinline{text}{npm view} command shows data about a package.
-It is possible to retrieve a specific field of the package registry by specifying fhe field name after the package descriptor.
-For example to retrieve information about the \textit{lodash} package, the following command can be used: \mintinline{text}{npm view lodash}.
+It is possible to retrieve a specific field of the package registry by specifying the field name after the package descriptor.
+For example to retrieve information about the \mintinline{text}{lodash} package, the following command can be used: \mintinline{text}{npm view lodash}.
 
 For this particular case, only the url of the repository associated with the package is needed.
-This can be achieved by adding the correponding field name: \mintinline{text}{npm view <PACKAGE> repository.url}.
+This can be achieved by adding the corresponding field name: \mintinline{text}{npm view <PACKAGE> repository.url}.
 
@@
 The implementation wraps variables around Proxy Objects to store tracking data, hereby enabling to associate the observation with a specific argument.
-The effect is reverted using Jalangi's callbacks to modify the behaviour of the affected operators, such as \mintinline{text}{===} or \mintinline{text}{typeof}.
+The effect is reverted using Jalangi's callbacks to modify the behavior of the affected operators, such as \mintinline{text}{===} or \mintinline{text}{typeof}.
 
@@
 No effort was made in trying to fix failing instrumentation.
-Failing JS Libraries were ommited from the analysis by providing a list with the names of the modules that should not be instrumented.
+Failing JS Libraries were omitted from the analysis by providing a list with the names of the modules that should not be instrumented.
 
@@
 The object \mintinline{text}{Interaction} is associated to an \mintinline{text}{ArgumentContainer}.
-If the \mintinline{text}{ArgumentContainer} does not exist when the Jalangi callback is triggerd, the \mintinline{text}{Interaction} is not saved.
+If the \mintinline{text}{ArgumentContainer} does not exist when the Jalangi callback is triggered, the \mintinline{text}{Interaction} is not saved.
 The main goal of gathering run-time information is to describe a function and its arguments.
@@
 \subsubsection{\textit{operator} interaction}
-Identifying if an argument is used with a specific operator is very important for infering the type of that variable.
+Identifying if an argument is used with a specific operator is very important for inferring the type of that variable.
 Relational operators like \mintinline{text}{<, >, <=, >=} are mainly used with type \mintinline{text}{number}.
@@
 Each interaction and return value of a \mintinline{text}{FunctionContainer} stores the corresponding \mintinline{text}{traceId}.
-The \mintinline{text}{traceId} field enables identifying correlations between input types and return values, as explained in \secref{sec:typescript-overloading}.
+The \mintinline{text}{traceId} field enables identifying correlations between input types and return values.
 An example is provided in \figref{fig:run-time-information-gathering-traces}.
@@
 
-Jalangi enables to modify the behaviour of the original code.
+Jalangi enables to modify the behavior of the original code.
 Using \mintinline{text}{invokFunPre} and \mintinline{javascript}{declare} callbacks, a function invocation is detected and the arguments of the function are modified in run-time.
@@
 The implemented solution wraps objects around JavaScript Proxies.
-Primitive values like \mintinline{text}{number} or \mintinline{text}{string} are converted to their native wapper objects, i.e.
+Primitive values like \mintinline{text}{number} or \mintinline{text}{string} are converted to their native wrapper objects, i.e.
 \mintinline{text}{Number} and \mintinline{text}{String}.
@@
 
-The behaviour of the original code is being explicitely modified.
+The behavior of the original code is being explicitly modified.
 Therefore, original values are used instead for affected operators like \mintinline{text}{===, !==, typeof}, for which wrapper objects will not work as expected.
@@
 When instrumenting a JS Library, Jalangi will try to instrument all required modules by resolving and finding the modules.
-Unfortunately, Jalangi's native behaviour is to stop the complete analysis if there is an error in the instrumentation of any of the required modules or in the script itself.
+Unfortunately, Jalangi's native behavior is to stop the complete analysis if there is an error in the instrumentation of any of the required modules or in the script itself.
 
-The workaround introduced by the implementation modifies Jalangi's behaviour for failing instrumentation.
+The workaround introduced by the implementation modifies Jalangi's behavior for failing instrumentation.
 For a given required module, it first tries to instrument it.
-If the instrumentation if unsuccesful, it simply skips the instrumentation and returns the original uninstrumented module.
+If the instrumentation if unsuccessful, it simply skips the instrumentation and returns the original module without instrumenting it.
 
@@
 
-JavaScript construtors where explained in \secref{sec:background-javascript}.
+JavaScript constructors where explained in \secref{sec:background-javascript}.
 There are built-in functions that are not able to be used as a constructors.
@@
 
-\coderef{code:typescript-templates-function-as-constructor} exposes a scenario for which a module could be infered as \mintinline{text}{module-function.d.ts} or \mintinline{text}{module-class.d.ts} depending only on how it was executed.
+\coderef{code:typescript-templates-function-as-constructor} exposes a scenario for which a module could be inferred as \mintinline{text}{module-function.d.ts} or \mintinline{text}{module-class.d.ts} depending only on how it was executed.
 Thus, examples that execute the module under analysis by truly representing the developer's intention are crucial for an accurate generation of declaration files and inference of types and interfaces.
@@
 
-\subsection{Type inference}
-\todo{El tipo se extrae del valor en runtime.}
 \subsection{Implementation}
@@
 \section{Evaluation}
+After generating a declaration file for a published NPM module, it is necessary to evaluate the quality of it.
+It was decided to compare the generated declaration file against the one uploaded to the DefinitelyTyped repository for the same module, as shown in \figref{fig:evaluation-diagram}.
+
+It is arguable, however, to use a baseline for evaluation that it is known to be incorrect or inaccurate for some modules.
+It was exposed in \secref{sec:problem-definition} that mismatches between definition files and JS libraries do exist.
+The presented approach actually tackles that issue.
+
+Nevertheless, it was decided that during a first stage of development comparing against DefinitelyTyped provides an acceptable baseline, guiding the tool to generate declaration files that are \textit{as good as} the ones uploaded to the repository.
+Furthermore, comparing two files is extremely fast, which is useful for creating an automated test suite that aids the development process.
+
+Finally, in a future stage, declaration files should be checked against the JS library itself.
+Approaches based on feedback-directed random testing like TSTest already made considerable progress and new solutions could be built on top of them \citep{DBLP:journals/pacmpl/KristensenM17}.
+
 \input{figures/approach/evaluation/evaluationDiagram.tex}
+
 \subsection{Parsing}
-\subsubsection{TypeScript Compiler API}
-\subsection{Comparison}
-\todo{Number of functions}
-\todo{Number of parameters}
-\todo{Interfaces}
-\todo{methodCall}
-\todo{usedAsArgument}
-\todo{Syntax \& Semantic Errors}
+Before comparison, declaration files need to be parsed so that both of them share the same structure.
+It was achieved by using the TypeScript Compiler API, a library developed by Microsoft that allows to traverse the Abstract Syntax Tree in an easy and intuitive way \citep{ts-compiler-api}.
+
+An simple example for the \mintinline{text}{abs} module described in \coderef{code:dts-generate-example} can be seen in \coderef{code:parsing-module-function}.
+
+\begin{code}
+  \jscode{code/parsing/parsed-abs-definitely-typed.json}
+  \captionsetup{aboveskip=0pt, belowskip=10pt}
+  \caption[Parsed declaration file example]{\textbf{Parsed declaration file example} - Parsed declaration for module \mintinline{text}{abs}.
+The declaration file was generated using the \mintinline{text}{module-function} template.}
+  \label{code:parsing-module-function}
+\end{code}
+
+The parsing consists in creating a structure where declared \mintinline{text}{interfaces}, \mintinline{text}{functions}, \mintinline{text}{classes} and \mintinline{text}{namespaces} are stored separated.
+Function arguments are correctly described, identifying complex types like union types or callbacks.
+Optional parameters are also identified.
+For \mintinline{text}{classes}, a distinction between the constructor and methods is made.
+Finally, syntax and semantic errors are also checked by the TypeScript Compiler API.
+
+The tool is called \mintinline{text}{parse-dts} and is naturally written in TypeScript. It also runs in NodeJS within its corresponding Docker container.
+
+\subsection{Comparator}
+An independent tool will compare two parsed declaration files.
+As described in \figref{fig:evaluation-diagram}, the comparator will compare the generated declaration file against the corresponding file in the DefinitelyTyped repository.
+
+As explained before, in a future stage a generated declaration file should not be compared against an existing one, since there would be no master file to compare against.
+A declaration file should be checked against the JavaScript library itself.
+The comparator explained in this section should assist the development process until the quality of the generated declaration files matches the one of the files uploaded to the DefinitelyTyped repository.
+
+It was discovered that the implementation was easier when focusing on each template independently.
+For this implementation, only the \mintinline{text}{module-function} was considered for the comparison.
+
+The following criteria were applied:
+\begin{itemize}
+  \item Number of declared functions: Checks the number of declared and exported functions for each of both declaration files.
+  \item Name of declared functions: Checks whether both of the declaration files declared a function with the same name.
+  \item Number of parameters: Checks the number of parameters of the declared functions. This is checked for optional and non-optional parameters.
+  \item Interfaces: Checks the number of declared interfaces and the fields within those interfaces.
+  \item Errors: Indicates whether there are errors in the declaration files.
+\end{itemize}
+
+Each of these criteria will add a penalty score. The lower the comparison score the more similar the files according to the presented criteria.
+
+\section{Type Inference}
+Type inference for JavaScript is a topic that was deeply investigated.
+Nevertheless, it was not included in the implementation.
+Types were directly imported from the types at run-time.
+For example, if a function \mintinline{text}{f(input)} was invoked as \mintinline{text}{f("hello")}, argument \mintinline{text}{input} would be declared as \mintinline{text}{string} in the declaration file.
+The refinement of the overall architecture of the generation of TypeScript declaration files was prioritized over type inference accuracy.
+Extracting the code examples from the repositories proved to be a solid and pragmatic way of capturing the developer's intention.
+Using the types imported from those examples allowed to focus on creating an end-to-end solution while still obtaining declaration files with reasonable types.
+It is worth noting that this would not have been the case if randomly generated test cases were used.
+The discussed approach on how to deal with type inference in JavaScript is presented in detail below.
+
+\subsection{Developer's intention}
+The crucial aspect while inferring a type is trying to reproduce the developer's intention as good as possible.
+
+A plausible approach for inferring a type in JavaScript would be binding a type to an operation.
+As a result, for example, everything that is used in an \mintinline{text}{if-else} clause would be inferred as \mintinline{text}{boolean}.
+
+Nevertheless, there are several reasons that may lead a JavaScript developer to use a a value with a type that is not a \mintinline{text}{boolean} in a conditional operator.
+It is a common practice to check for empty strings or zero values.
+As a result, developers end up using a value of type \mintinline{text}{string} or \mintinline{text}{number} in a conditional.
+\coderef{code:type-inference-if-clause-boolean-string} exposes the inconsistencies of binding the usage of an operator with a fix type by providing an example where the argument is inferred as a \mintinline{text}{string} even though it is used in an \mintinline{text}{if} clause.
+
+\begin{code}
+  \jscode{code/type-inference/if_clause_boolean_string.js}
+  \captionsetup{aboveskip=0pt, belowskip=10pt}
+  \caption[String inference]{\textbf{String inference} - Argument \mintinline{text}{a} should be inferred as a \mintinline{text}{string}.
+Usage of methods \mintinline{text}{indexOf()} and \mintinline{text}{toLowerCase()} clearly indicate that \mintinline{text}{a} is a \mintinline{text}{string}, in spite of the \mintinline{text}{if} clause.}
+  \label{code:type-inference-if-clause-boolean-string}
+\end{code}
+
+On the other hand, it can be seen in \coderef{code:type-inference-number-with-plus-operator} that assuming operands in a string concatenation operation to be of type \mintinline{text}{string} is not accurate either.
+As explained in \secref{sec:background-js-type-coercion}, concatenating a \mintinline{text}{string} and a \mintinline{text}{number} without explicitly converting the \mintinline{text}{number} into a \mintinline{text}{string} improves readability by reducing verbosity.
+
+\begin{code}
+  \jscode{code/type-inference/type-inference-number-with-plus-operator.js}
+  \captionsetup{aboveskip=0pt, belowskip=10pt}
+  \caption[Number inference]{\textbf{Number inference} - Even though argument \mintinline{text}{a} is used as part of a string concatenation through the \mintinline{text}{+} operator, the developer's intention of \mintinline{text}{a} being a \mintinline{text}{number} is clear.
+Several classic arithmetic operations support the assumption.}
+  \label{code:type-inference-number-with-plus-operator}
+\end{code}
+
+Both \coderef{code:type-inference-if-clause-boolean-string} and \coderef{code:type-inference-number-with-plus-operator} show that the developer's intention is independent from the input value's type.
+Functions could be invoked with different values but the developer's intention remains unmodified in each invocation, unless new execution paths are explored and new indicators of the developer's intention get known.
+It is useful, though, to consider specific examples provided by the developer as strong indicators of the intention, as explained in \secref{sec:initial-code-base}.
+It is decisive if a developer explicitly manifests her intention in code, for example by creating an example where a specific function is used with a \mintinline{text}{string} type.
+
+Finally, it is possible for JavaScript developers to perform uncommon operations like multiplying an \mintinline{text}{array} with a \mintinline{text}{function}.
+Furthermore, within a particular context unintuitive code fragments may yield a valid development pattern.
+\coderef{code:type-inference-value-of-method-modified} provides an example where a developer might choose to add two arrays because the module she is developing works in a scenario where the \mintinline{text}{valueOf} method of the built-in \mintinline{text}{Array} object is modified.
+
+\begin{code}
+  \jscode{code/type-inference/value-of-method-modified.js}
+  \captionsetup{aboveskip=0pt, belowskip=10pt}
+  \caption[Sum of arrays]{\textbf{Sum of arrays} - Unintuitive usage of operator \mintinline{text}{+} with arrays.
+\mintinline{text}{valueOf()} method of \mintinline{text}{Array} object is modified.
+Operator \mintinline{text}{+} is explained in detail in \secref{sec:background-js-type-coercion}.}
+  \label{code:type-inference-value-of-method-modified}
+\end{code}
+
+It is clear that types can not be inferred only with assumptions based on constraints imposed to the JavaScript language.
+Such constraints could be based on uncommon usages of the language.
+JavaScript allows indeed to perform such unintuitive operations.
+However, type inference is intended to be performed for already existing working JavaScript libraries.
+JavaScript developers may be already applying patterns and using operations that ought to be restricted.
+Developer's intention is exposed through indicators in the code.
+Some indicators are more determinant than other ones in revealing the intention.
+The presented approach elaborates the concept by presenting a map of indicators and a way of determining the relevance.
+
+\subsubsection{Behavior and correctness}
+The definition of incorrect code in JavaScript needs to be broader than a piece of code which throws a run-time error.
+\coderef{code:type-inference-examples-definition-of-incorrect} exposes different examples of legal JavaScript code, i.e.
+code that does not throw any error.
+Even though the code will run flawlessly, accessing a non-existing property of an object, is something that should probably be considered as incorrect, since the result will be \mintinline{text}{undefined}.
+On the other hand, it would probably be incorrect to invoke a function that multiplies two variables with an \mintinline{text}{object} and an \mintinline{text}{array}, since it will \mintinline{text}{NaN}, as shown in \coderef{code:type-inference-multiply-probably-incorrect}.
+
+\begin{code}
+  \jscode{code/type-inference/definition-of-incorrect/examples-definition-of-incorrect.js}
+  \captionsetup{aboveskip=0pt, belowskip=10pt}
+  \caption[Legal JavaScript code]{\textbf{Legal JavaScript code} - The presented examples will not throw any run-time error. These examples expose the problem of considering a code correct or not.}
+  \label{code:type-inference-examples-definition-of-incorrect}
+\end{code}
+
+\begin{code}
+  \jscode{code/type-inference/definition-of-incorrect/multiply-probably-incorrect.js}
+  \captionsetup{aboveskip=0pt, belowskip=10pt}
+  \caption[Number inference]{\textbf{Number inference} - Invoking a function that multiplies both of their arguments with anything different than a \mintinline{text}{number} is probably not expected.}
+  \label{code:type-inference-multiply-probably-incorrect}
+\end{code}
+
+\input{figures/approach/type-inference/behavior_correctness/ecma-script-min-specification.tex}
+
+According to the specification of the native \mintinline{text}{Math.min()} method, each argument will be converted to a number before actually computing the minimum \citep{ecma-script}.
+Therefore, according to the definition shown in \figref{fig:type-inference-ecma-script-min-specification}, it is possible to compute the minimum between, for example, an \mintinline{text}{object} and a \mintinline{text}{function}.
+Although legal, there is no reasonable argument to define the examples in \coderef{code:type-inference-multiply-probably-incorrect} as correct.
+
+\begin{code}
+  \jscode{code/type-inference/definition-of-incorrect/math_min.js}
+  \captionsetup{aboveskip=0pt, belowskip=10pt}
+  \caption[Math.min() possible usage]{\textbf{Math.min() possible usage} - Invoking \mintinline{text}{Math.min()} with values of type different as \mintinline{text}{number} should not be expected.}
+  \label{code:type-inference-multiply-probably-incorrect}
+\end{code}
+
+The key aspect is that the ECMAScript Language Specification defines behavior.
+It perfectly describes what is going to happen if a developer invokes \textit{Math.min()} with an argument that is not a \mintinline{text}{number}: every argument will be coerced to a \mintinline{text}{number}.
+However, such a definition is not related to \textit{correctness}.
+
+Ryan Cavanaugh, development lead for the TypeScript team at Microsoft, stated the following\footnote{https://stackoverflow.com/questions/41750390/what-does-all-legal-javascript-is-legal-typescript-mean}:
+
+\begin{displayquote}
+  \textit{After all, this code does have defined behavior!
+But this misses the forest for the trees - it is plainly implausible that this code is correct for any reasonable definition of "correct".
+TypeScript's existence is predicated on the notion that some JavaScript code is not correct and you'd like to know about it.}
+  \\
+  \\
+  \textit{The fact that the spec describes what happens is not instructive as to whether or not you're writing a correct program.
+A doctor can clearly describe what will happen to you if you eat a rock, but that doesn't mean rocks are food.}
+\end{displayquote}
+
+Probably no developer would argue with the fact that \mintinline{text}{Math.min()} should indeed receive numbers as arguments.
+Actually, TypeScript only accepts numbers as possible arguments for \mintinline{text}{Math.min()}, as shown in
+
+\begin{code}
+  \tscode{code/type-inference/definition-of-incorrect/math_min_typescript_definition.ts}
+  \captionsetup{aboveskip=0pt, belowskip=10pt}
+  \caption[TypeScript definition for Math.min()]{\textbf{TypeScript definition for Math.min()} - TypeScript defines \mintinline{text}{Math.min()} as a function that receives arguments only of type \mintinline{text}{number}.}
+  \label{code:type-inference-multiply-probably-incorrect}
+\end{code}
+
+\todo{What does the developer expect from each operation?
+How do they use them?
+Esa es la clave, en realidad la type inference se trata de adinivar lo que quiere hacer el developer y para eso hace falta responder a esas preguntas.
+Eso se logra con los experiments que se hicieron sobre la info sobre los tipos y los operators.}

### 2019-08-13 figures/approach/run-time-information-gathering/analysis/method-call.tex
Add content to the TypeInference section

@@
-\begin{figure}[h]
+\begin{figure}[tp]
 	\centering
@@
 Method information can be accessed through the reference \mintinline{bash}{functionId}, for which, in this example, the \mintinline{bash}{inputValue} interactions are stored.
-Main entry for \mintinline{bash}{functionId_3} has been ommited.
-Several additional fields and entries of the final output JSON file were also ommited for simplicity.}
+Main entry for \mintinline{bash}{functionId_3} has been omitted.
+Several additional fields and entries of the final output JSON file were also omitted for simplicity.}
 	\label{fig:run-time-information-gathering-method-call}

### 2019-08-13 figures/approach/run-time-information-gathering/analysis/operator.tex
Add usedAsArgument and operator interactions
Add content to the TypeInference section

@@
 The example shows how that value is \mintinline{bash}{left} for one argument and \mintinline{bash}{right} for the other one.
-Several additional fields and entries of the final output JSON file were ommited for simplicity.}
-	\label{fig:run-time-information-gathering-following-interactions}
+Several additional fields and entries of the final output JSON file were omitted for simplicity.}
+	\label{fig:run-time-information-gathering-operator}
 \end{figure}

### 2019-08-13 code/type-inference/definition-of-incorrect/examples-definition-of-incorrect.js

@@ -1,0 +1,17 @@
+let a = "hello world".indexOf(
+	"hello",
+	0,
+	true,
+	null,
+	undefined,
+	{},
+	[]
+); // To many parameters for String.indexOf()
+
+let b = {
+	x: "One",
+	x: "Two"
+}; // Duplicated property "x". ==> b.x === "Two"
+
+let c = "MyString".lenth;
+// "lenth" instead of "length". ===> c === undefined.

### 2019-08-13 code/type-inference/definition-of-incorrect/math_min.js

@@ -1,0 +1,7 @@
+Math.min({}, function() {}); // NaN
+
+Math.min([], "hello world"); // NaN
+
+Math.min(1, false); // 0
+
+Math.min(undefined, null); // NaN

### 2019-08-13 code/type-inference/definition-of-incorrect/math_min_typescript_definition.ts

@@ -1,0 +1,5 @@
+/**
+  * Returns the smaller of a set of supplied numeric expressions.
+  * @param values Numeric expressions to be evaluated.
+  */
+min(...values: number[]): number;

### 2019-08-13 code/type-inference/definition-of-incorrect/multiply-probably-incorrect.js

@@ -1,0 +1,5 @@
+function multiply(a, b) {
+	return a * b;
+}
+
+let x = multiply({}, []); // NaN. Probably incorrect.

### 2019-08-13 code/type-inference/if_clause_boolean_string.js

@@ -1,0 +1,15 @@
+function f(a) {
+	// Some code ...
+
+	if (a) {
+		// ...
+	}
+
+	if (a.indexOf("my string") !== -1) {
+		// Do something ...
+	}
+
+	let b = a.toLowerCase();
+
+	// More code ...
+}

### 2019-08-13 code/type-inference/type-inference-number-with-plus-operator.js

@@ -1,0 +1,19 @@
+function f(a) {
+	log("Input value is: " + a);
+
+	let x = a + 3;
+
+	//...
+
+	let y = a / 10;
+
+	// ...
+
+	let z = a * 1234;
+
+	// ...
+
+	if (a < 100) {
+		// ...
+	}
+}

### 2019-08-13 code/type-inference/value-of-method-modified.js

@@ -1,0 +1,19 @@
+/**
+ * valueOf() method of Array object is modified
+ * and returns the sum of its elements.
+ * 
+ * [1, 2, 3].valueOf() === 6 (1 + 2 + 3)
+ */
+
+Array.prototype.valueOf = function() {
+	return this.reduce(function(previousValue, currentValue) {
+		return previousValue + currentValue;
+	});
+}
+
+let a = [1, 2, 3]; // a.valueOf() = 6
+let b = [4, 5, 6]; // b.valueOf() = 15
+
+let c = a + b;
+
+// c === 21

### 2019-08-14 chapters/4-approach.tex
Finish type inference section

@@
 An example is provided in \figref{fig:run-time-information-gathering-traces}.
+The field can also be helpful in determining common interactions happening on the same execution path.
 
@@
 Probably no developer would argue with the fact that \mintinline{text}{Math.min()} should indeed receive numbers as arguments.
-Actually, TypeScript only accepts numbers as possible arguments for \mintinline{text}{Math.min()}, as shown in
+Actually, TypeScript only accepts numbers as possible arguments for \mintinline{text}{Math.min()}, as shown in \
 
@@
   \caption[TypeScript definition for Math.min()]{\textbf{TypeScript definition for Math.min()} - TypeScript defines \mintinline{text}{Math.min()} as a function that receives arguments only of type \mintinline{text}{number}.}
-  \label{code:type-inference-multiply-probably-incorrect}
+  \label{code:type-inference-math-min-typescript-definition}
 \end{code}
 
-\todo{What does the developer expect from each operation?
-How do they use them?
-Esa es la clave, en realidad la type inference se trata de adinivar lo que quiere hacer el developer y para eso hace falta responder a esas preguntas.
-Eso se logra con los experiments que se hicieron sobre la info sobre los tipos y los operators.}
+\subsection{Case study: inference for \mintinline{text}{+} operator}
+
+The following section exposes the difficulties that may arise when trying to infer the type of a value using the \mintinline{text}{+} operator.
+
+\coderef{code:type-inference-string-concatenation} provides an example where a function can be reasonably used with a \mintinline{text}{string}, a \mintinline{text}{number} or even an \mintinline{text}{object}.
+On the other hand, \coderef{code:type-inference-string-concatenation-only-string} presents an example using the same \mintinline{text}{+} operator where only a \mintinline{text}{string} would be accepted.
+
+Finally, for both cases it would be technically correct to infer the input argument as \mintinline{text}{any}.
+JavaScript will accept a value of any type and will perform the according type coercion without throwing any run-time error.
+It is clear then that the developer's intention is the most relevant aspect for inferring a type.
+
+\begin{code}
+  \jscode{code/type-inference/example-plus-operator/string-concatenation.js}
+  \captionsetup{aboveskip=0pt, belowskip=10pt}
+  \caption[Operator + used with string and number]{\textbf{Operator + used with \mintinline{text}{string} and \mintinline{text}{number}} - The following simple function can be invoked with a \mintinline{text}{string}, a \mintinline{text}{number} or even an \mintinline{text}{object} with a defined \mintinline{text}{valueOf} method.}
+  \label{code:type-inference-string-concatenation}
+\end{code}
+
+\begin{code}
+  \jscode{code/type-inference/example-plus-operator/string-concatenation-only-string.js}
+  \captionsetup{aboveskip=0pt, belowskip=10pt}
+  \caption[Operator + used only with string]{\textbf{Operator + used with \mintinline{text}{string}} - The following function provides an example where only a \mintinline{text}{string} would be accepted.}
+  \label{code:type-inference-string-concatenation-only-string}
+\end{code}
+
+\subsection{Constraints and restrictions}
+The following section presents the restrictions that different type checkers perform on JavaScript code.
+The state of the art is shown through a comparison between \mintinline{text}{TypeScript}{TypeScript}, \mintinline{text}{Flow} and \mintinline{text}{Infernu}.
+All of them introduce the concept of restrictions for JS from different angles and for different use cases.
+The analysis is going to be performed on the \mintinline{text}{+} operator.
+
+\subsubsection{TypeScript}
+Typescript defines the \mintinline{text}{+} operator using the table presented in \figref{fig:type-inference-typescript-definition-plus-operator}.
+It is important to note that TypeScript allows a \mintinline{text}{string} to be added to anything.
+On the other hand, it does not allow, for example, to add a \mintinline{text}{boolean} and an \mintinline{text}{object} or a \mintinline{text}{number} and a \mintinline{text}{boolean}.
+
+\input{figures/approach/type-inference/constraints/typescript.tex}
+
+\subsubsection{Flow}
+Flow is a static type checker for JavaScript developed by Facebook \citep{flow}.
+It is intended to aid developers while writing JS code by detecting different errors.
+It only allows to use the \mintinline{text}{+} operator between \mintinline{text}{string} and \mintinline{text}{string}, \mintinline{text}{number} and \mintinline{text}{number} or \mintinline{text}{string} and \mintinline{text}{number}, as explained in \coderef{code:type-inference-constraints-flow}.
+
+\begin{code}
+  \jscode{code/type-inference/constraints/constraints-flow.js}
+  \captionsetup{aboveskip=0pt, belowskip=10pt}
+  \caption[Restrictions in Flow type checker]{\textbf{Restrictions in Flow type checker} - The \mintinline{text}{+} can be used only with \mintinline{text}{string} or \mintinline{text}{number}.}
+  \label{code:type-inference-constraints-flow}
+\end{code}
+
+\subsubsection{Infernu}
+Infernu is a static type checker for JavaScript that applies the concept of Type Classes to JS \citep{infernu}.
+It is inspired by Haskell's Type Classes.
+According to Infernu's documentation, operator \mintinline{text}{+} can be used only when both operands are either \mintinline{text}{string} or \mintinline{text}{number}, as shown in \coderef{code:type-inference-constraints-infernu}.
+
+\begin{code}
+  \jscode{code/type-inference/constraints/constraints-infernu.js}
+  \captionsetup{aboveskip=0pt, belowskip=10pt}
+  \caption[Restrictions in Infernu type checker]{\textbf{Restrictions in Infernu type checker} - The \mintinline{text}{+} can be used only when both operands are either \mintinline{text}{string} or \mintinline{text}{number}.}
+  \label{code:type-inference-constraints-infernu}
+\end{code}
+
+\subsection{Proposal}
+The challenge in type inference automation is to capture the developer's intention.
+It is fundamental to understand the code written by the developer.
+As explained, JavaScript is a language for which very unintuitive operations do not throw any errors at run-time.
+Developers may be already using this kind of behavior consciously by creating uncommon patterns which are useful in different scenarios.
+Therefore, working with constraints and hard rules that forbid certain operations, like adding an \mintinline{text}{array} and a \mintinline{text}{number}, is not ideal.
+
+The proposal is to use a scoring system that favors coding best practices and disfavors uncommon or unexpected ones.
+In the end, based on this scoring you can infer the type of the variable, since types with higher score will represent the developer's intention more accurately.
+Uncommon operations will not be restricted but they would get lower scores.
+The scoring is generated by registering all interactions performed by the code to a particular variable.
+
+Applying the presented approach would mean that despite uncommon usage of a particular type, if there is enough evidence that supports the developer's intention, that type should be inferred anyway.
+
+\coderef{code:type-inference-unintuitive-developers-intention} presents a code where it is clear that \mintinline{text}{a} is an \mintinline{text}{array}.
+We see in line 9 that the developer is using an array with the \mintinline{text}{+} operator.
+It is known that such an usage is not common.
+However, there is enough evidence to support inferring \mintinline{text}{a} as an \mintinline{text}{array}.
+We are therefore assuming that the developer is consciously adding the number \mintinline{text}{1} to an \mintinline{text}{array}.
+In fact, applying a constraint on using an \mintinline{text}{array} with the \mintinline{text}{+} operator would have made it impossible to actually infer the type \mintinline{text}{array}.
+
+\begin{code}
+  \jscode{code/type-inference/proposal/intention-to-use-array.js}
+  \captionsetup{aboveskip=0pt, belowskip=10pt}
+  \caption[Unintuitive developer's intention]{\textbf{Unintuitive developer's intention} - The developer's intention of using \mintinline{text}{a} as an \mintinline{text}{array} is clear, despite the usage of the \mintinline{text}{+} operator.
+A scoring system would assign a high score to the type \mintinline{text}{array} because of other hints in the code and therefore the type of \mintinline{text}{a} would be inferred as \mintinline{text}{array}.}
+  \label{code:type-inference-unintuitive-developers-intention}
+\end{code}
+
+Code examples provided by the developer should be considered as additional hints for inferring a type.
+In fact, they should be considered as very strong indicators, since the developer is explicitly stating her intention through code.
+
+There is no doubt about the developer's intention, however, when an interface needs to be inferred.
+The developer explicitly reveals the shape of the required interface by accessing an object's property or invoking a method.
+
+Finally, in order to define a scoring system it is crucial to understand what the developers are expecting from each operation and how are they using them.
+For answering those questions, it is necessary to analyze existing code bases and extracting common patterns.
+\secref{sec:experiments-js-operators-usage} presents the results from the analysis performed for this approach, where a heat map was created for each operator by analyzing existing JS libraries uploaded to the NPM registry.
+This way, the most common types for different operators could be identified.
+The following section discusses the application of the obtained results.
+
+\subsubsection{Scoring for operators}
+Results presented in \secref{sec:experiments-js-operators-usage} are briefly commented in the following section.
+Information was retrieved by execution JavaScript libraries through the code examples in the repositories, so the results intent to represent the developer's common usage.
+Obtained values are used to create a scoring for operators and the type of their operands.
+Analyze is limited to operators \mintinline{text}{+} and \mintinline{text}{===} but they can be easily extended.
+Usage distribution for these operators are presented in \figref{fig:type-inference-proposal-scoring-plus} and \figref{fig:type-inference-proposal-scoring-strict-equal}.
+Scores can be built using the results.
+For operator \mintinline{text}{+}, when an operand is known to be a \mintinline{text}{number}, the second operand is extremely likely to be also a \mintinline{text}{number} and should therefore be greatly favored.
+
+\input{figures/approach/type-inference/proposal/scoring/operator-plus.tex}
+
+\input{figures/approach/type-inference/proposal/scoring/operator-strict-equal.tex}

### 2019-08-14 chapters/6-conclusions.tex
Finish type inference section

@@
 \todo{Deberia ser hecho en otro modulo que no sea el runtime information. El runtime information deberia exponer unicamente los datos que le hacen falta al otro para inferir los tipos.}
+
+\todo{Tambien se puede identificar patterns de programacion donde siempre los tipos son los mismos!!! El concepto de inferir el tipo en realidad pasa a detectar la intencion del developer.}
+
 \subsubsection{Return values}
@@
 
-\todo{Tambien se puede identificar patterns de programacion donde siempre los tipos son los mismos.}
-
-\todo{MERGE Declaration Files - Ver si se agrega}
-
-\subsubsection{Operators}
-\todo{Explicar el tema de la type inference con los operators}
-
 \subsection{Uploading Results to Definitely Typed Repository}

### 2019-08-14 figures/approach/type-inference/constraints/typescript.tex
Finish type inference section

\begin{figure}[tp]
	\centering
	\subfloat[Table with allowed types for the \mintinline{text}{+} operator]{
		\includegraphics[width=0.6\textwidth]{figures/approach/type-inference/typescript_plus_operator.png}
	}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.7\textwidth}
			\tscode{code/type-inference/constraints/constraints-typescript.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[Examples of not allowed operations]{\usebox{\mintedbox}}

	\caption[TypeScript definition of the + operator]{\textbf{TypeScript definition of the \mintinline{text}{+} operator} - Rows and columns of the table represent the type of the operands. Values within the table represent the type of the return value. The empty cells specify the invalid operations.}
	\label{fig:type-inference-typescript-definition-plus-operator}
\end{figure}

### 2019-08-14 figures/approach/type-inference/proposal/scoring/operator-plus.tex
Finish type inference section

\begin{figure}[tp]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
			title=Operator + | number,
			width=0.55\textwidth,
			ybar=0pt,
			ymax=100,
			enlargelimits=0.15,
			legend style={at={(0.5,-0.2)}, anchor=north,legend columns=-1},
			symbolic x coords={number,string,undefined,object,null,boolean,function,array},
			xtick=data,
			nodes near coords, 
			nodes near coords align={vertical},
			x tick label style={rotate=90,anchor=east},
		]
		\addplot coordinates {
			(number, 98)
			(string, 2)
			(undefined, 0)
			(object, 0)
			(null, 0)
			(boolean, 0)
			(function, 0)
			(array, 0)
		};
		\end{axis}
	\end{tikzpicture}
	%
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
			title=Operator + | string,
			width=0.55\textwidth,
			ybar=0pt,
			ymax=100,
			enlargelimits=0.15,
			legend style={at={(0.5,-0.2)}, anchor=north,legend columns=-1},
			symbolic x coords={number,string,undefined,object,null,boolean,function,array},
			xtick=data,
			nodes near coords, 
			nodes near coords align={vertical},
			x tick label style={rotate=90,anchor=east},
		]
		\addplot coordinates {
			(number, 7)
			(string, 64)
			(undefined, 29)
			(object, 0)
			(null, 0)
			(boolean, 0)
			(function, 0)
			(array, 0)
		};
		\end{axis}
	\end{tikzpicture}

	\caption[Usage distribution for operator +]{\textbf{Usage distribution for operator \mintinline{text}{+}} - Both graphs were made by fixing the type of one of the operands. If a value is used with the operator \mintinline{text}{+} and the other operand is a \mintinline{text}{number}, the graph on the left should be used. If the second operand is a \mintinline{text}{string}, the graph on the right should be used.
	}

	\label{fig:type-inference-proposal-scoring-plus}
\end{figure}

### 2019-08-14 figures/approach/type-inference/proposal/scoring/operator-strict-equal.tex
Finish type inference section

\begin{figure}[tp]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
			title=Operator ${===}$ | number,
			width=0.55\textwidth,
			ybar=0pt,
			ymax=100,
			enlargelimits=0.15,
			legend style={at={(0.5,-0.2)}, anchor=north,legend columns=-1},
			symbolic x coords={number,string,undefined,object,null,boolean,function,array},
			xtick=data,
			nodes near coords, 
			nodes near coords align={vertical},
			x tick label style={rotate=90,anchor=east},
		]
		\addplot coordinates {
			(number, 97)
			(string, 0)
			(undefined, 2)
			(object, 0)
			(null, 1)
			(boolean, 0)
			(function, 0)
			(array, 0)
		};
		\end{axis}
	\end{tikzpicture}
	%
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
			title=Operator ${===}$ | string,
			width=0.55\textwidth,
			ybar=0pt,
			ymax=100,
			enlargelimits=0.15,
			legend style={at={(0.5,-0.2)}, anchor=north,legend columns=-1},
			symbolic x coords={number,string,undefined,object,null,boolean,function,array},
			xtick=data,
			nodes near coords, 
			nodes near coords align={vertical},
			x tick label style={rotate=90,anchor=east},
		]
		\addplot coordinates {
			(number, 0)
			(string, 96)
			(undefined, 1)
			(object, 0)
			(null, 1)
			(boolean, 1)
			(function, 0)
			(array, 0)
		};
		\end{axis}
	\end{tikzpicture}
	%
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
			title=Operator ${===}$ | undefined,
			width=0.55\textwidth,
			ybar=0pt,
			ymax=100,
			enlargelimits=0.15,
			legend style={at={(0.5,-0.2)}, anchor=north,legend columns=-1},
			symbolic x coords={number,string,undefined,object,null,boolean,function,array},
			xtick=data,
			nodes near coords, 
			nodes near coords align={vertical},
			x tick label style={rotate=90,anchor=east},
		]
		\addplot coordinates {
			(number, 3)
			(string, 6)
			(undefined, 4)
			(object, 69)
			(null, 1)
			(boolean, 3)
			(function, 10)
			(array, 5)
		};
		\end{axis}
	\end{tikzpicture}
	%
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
			title=Operator ${===}$ | null,
			width=0.55\textwidth,
			ybar=0pt,
			ymax=100,
			enlargelimits=0.15,
			legend style={at={(0.5,-0.2)}, anchor=north,legend columns=-1},
			symbolic x coords={number,string,undefined,object,null,boolean,function,array},
			xtick=data,
			nodes near coords, 
			nodes near coords align={vertical},
			x tick label style={rotate=90,anchor=east},
		]
		\addplot coordinates {
			(number, 9)
			(string, 22)
			(undefined, 2)
			(object, 10)
			(null, 6)
			(boolean, 0)
			(function, 17)
			(array, 34)
		};
		\end{axis}
	\end{tikzpicture}

	\caption[Usage distribution for operator ===]{\textbf{Usage distribution for operator \mintinline{text}{===}} - All graphs were made by fixing the type of one of the operands. That means, for example, that if a value is used with the operator \mintinline{text}{===} and the other operand is a \mintinline{text}{number}, the top-left graph should be used.}
	\label{fig:type-inference-proposal-scoring-strict-equal}
\end{figure}

### 2019-08-14 figures/experiments/operators/conditional-operators/conditionalOperators.tex
Add content to the TypeInference section
Finish type inference section

@@
 			title=Conditional operators,
-			width=1\textwidth,
+			width=0.8\textwidth,
 			ybar=0pt,

### 2019-08-14 code/type-inference/constraints/constraints-flow.js

@@ -1,0 +1,17 @@
+// @flow
+
+function f(a, b) { 
+	return a + b;
+}
+
+f(123, 456);
+f("123", "456");
+f(123, "456");
+f("123", 456);
+// f(true, "hello"); -> ERROR
+// f(true, 1); -> ERROR
+// f(12, {}); -> ERROR
+// f({}, {}); -> ERROR
+// f([], []); -> ERROR
+// f([1], [2]); -> ERROR
+// f([1, 2], [3, 4]); -> ERROR

### 2019-08-14 code/type-inference/constraints/constraints-infernu.js

@@ -1,0 +1,32 @@
+function f(a, b) {
+	return a + b;
+}
+
+f(123, 456);
+f("123", "456");
+
+// f(123, "456");
+/* 	Infernu Error:
+ *  --------------
+ *  Failed unifying:
+         ((Number,String) -> ei)
+     With:
+         ((ek,ek) -> ek)
+     Because:
+     Failed unifying:
+         ek
+     With:
+         String
+     Because:
+     Failed unifying:
+         Number
+     With:
+         String 
+*/
+
+// f("123", 456); // -> Error
+// f(12, {}); // -> Error
+// f({}, {}); // -> Error
+// f([], []); // -> Error
+// f([1], [2]); // -> Error
+// f([1, 2], [3, 4]); // -> Error

### 2019-08-14 code/type-inference/constraints/constraints-typescript.ts

@@ -1,0 +1,23 @@
+let a1: number = 1;
+let b1: number = 2;
+let s1 = a1 + b1; // OK. Inferred => s1: number
+
+let a2: string = "hello";
+let b2: string = "hello";
+let s2 = a2 + b2; // OK. Inferred => s2: string
+
+let a3: string = "hello";
+let b3: number = 1234;
+let s3 = a3 + b3; // OK. Inferred => s3: string
+
+let a4: number = 1234;
+let b4: object = {};
+let s4 = a4 + b4; // NOK.
+
+let a5: number = 123;
+let b5: boolean = true;
+let s5 = a5 + b5; // NOK.
+
+let a6: string = "hello";
+let b6: object = {};
+let s6 = a6 + b6; // OK. Inferred => s6: string

### 2019-08-14 code/type-inference/example-plus-operator/string-concatenation-only-string.js

@@ -1,0 +1,5 @@
+function printName(myName) {
+	return "My name is: " + myName;
+}
+
+f("John");

### 2019-08-14 code/type-inference/example-plus-operator/string-concatenation.js

@@ -1,0 +1,16 @@
+function f(a) {
+	return "The total amount is: " + a;
+}
+
+f("unknown"); // The total amount is: unknown
+f(123.45); // The total amount is: 123.45
+
+let o = {
+	val: 123.45
+};
+
+o.valueOf = function() {
+	return this.val;
+};
+
+f(o); // The total amount is: 123.45

### 2019-08-14 code/type-inference/proposal/intention-to-use-array.js

@@ -1,0 +1,13 @@
+function f(a) {
+	a.push(12);
+	a.reverse();
+
+	let otherArray = a.filter(function(v) {
+		// Filter code ...
+	});
+
+	let b = a + 1;
+
+	// ...
+}
+

### 2019-08-16 chapters/5-experiments.tex
Add examples of the obtained results

@@
 \chapter{Experiments}\label{chap:experiments}
+Declaration files were generated for existing modules repository.
+The DefinitelyTyped repository was used as a benchmark.
+Each one of the generated files was compared against the corresponding declaration file already uploaded to the repository.
 
-Experiments often include benchmarks.
-You need to describe you benchmarking setup: parameters of the machine used (architecture, memory, etc), parameters of the software used (OS version, libraries used and their version), the benchmark programs.
-The programs need not be part of the thesis, but they should be described and available separately.
+\figref{fig:experiments-overall-funnel} shows that a declaration file was generated for 244 modules out of 6029.
+The following sections explain each stage of the process.
+Examples of generated declaration files for existing modules are presented in \secref{sec:experiments-declaration-files-generation}.
 
-If you measure run times and evaluate your findings statistically, then there is great potential for errors.
-Consult SIGPLAN's checklist for empirical evaluations \url{http://www.sigplan.org/Resources/EmpiricalEvaluation/} for advice.
+\input{figures/experiments/overall-funnel.tex}
 
-\input{figures/experiments/overallFunnel.tex}
+An analysis on the usage of JavaScript operators can be seen in \secref{sec:experiments-js-operators-usage}.
 
-\section{Definitely Typed Modules}
-\todo{Poner cómo es la distribución de los tipos de declaration files.}
-
-\draft{
-module-class	503	8,27%
-module	4379	71,99%
-module-function	1201	19,74%
-}
+\section{Definitely Typed Benchmark}
 
@@
 
-\section{Code Examples Extraction}
-\subsection{Repositories URL}
-\subsection{Readme Files}
-\subsection{Code Examples}
+\subsection{Code Examples Extraction}
+Extracting the code examples from the repositories of the JavaScript libraries proved to be a pragmatic way of capturing the types.
+However, code examples for only 2260 out of 6029 modules could be retrieved.
 
-\section{Code Execution}
+\subsubsection{Repositories URL}
+The url of the repositories could be retrieved for only 4974 modules.
+More than 1000 modules do not have the repository entry in their corresponding \mintinline{text}{package.json} files.
+Therefore, the \mintinline{text}{npm view <module> repository.url} command returns an empty value.
+This is even happening for important modules like \mintinline{text}{ace}.
 
-\section{Run-time Information Gathering}
+\subsubsection{Readme Files}
+700 modules simply do not have a readme file in their repositories.
+The implementation does contemplate, however, different naming conventions like \mintinline{text}{readme.md} or \mintinline{text}{README.md}.
 
-\section{Declaration Files Generation}
+\subsubsection{Code Examples Extraction}
+The 50\% loss is mainly explained because developers did not tag their code around the \mintinline{text}{javascript} tag in their markdown readme files.
+As explained in \secref{sec:conclusions-future-work}, the process of extracting code examples from a repository can be greatly improved.
+Counting with code examples for 2200 modules was considered to be enough for this work.
 
-\section{Evaluation}
+\subsubsection{Code Examples Execution}
+Only 1448 modules had working code examples.
+The 2260 extracted code examples were executed by installing the required packages and running the code as a \mintinline{text}{node} application.
+812 modules did not run correctly and were discarded from the analysis.
+Some failing examples were analyzed and there were mainly two reasons for the failure: 1) The code example had been properly extracted but the code itself was not working properly and 2) the extracted code example was not correct.
+Improving the retrieval the code examples for a specific module will definitely have a positive impact in this number.
+
+\subsection{Run-time Information Gathering} \label{sec:experiments-run-time-information-gathering}
+Run-time information could only be extracted for 442 modules out of 1448 modules with working code examples.
+As explained in \secref{sec:run-time-wrapper-objects}, the behavior of the code under analysis was explicitly modified by wrapping the arguments around Wrapper Objects.
+This definitely had an influence on the executed code of some modules by causing new run-time errors.
+Furthermore, Jalangi's own limitation in the instrumentation process caused some of the modules to be unable to be instrumented.
+
+A plain Jalangi instrumentation, i.e.
+without user defined analysis modules, was not applied to the 1448 working code examples.
+Therefore, it was not possible to determine which modules are failing only because of Jalangi's own limitations.
+
+\subsection{Declaration Files Generation} \label{sec:experiments-declaration-files-generation}
+Finally, a declaration file was generated for 244 modules.
+Despite the correct gathering of run-time information, some of the 442 modules mentioned in \secref{sec:experiments-run-time-information-gathering} did not contain valid run-time information for the generation of the declaration file.
+The main reason for this behavior is that the extracted code examples did not execute the modules in a way that which gathers valid run-time information.
+
+\subsection{Evaluation}
+As shown in \figref{fig:experiments-typescript-templates-distribution-definitely-typed}, 72\% of the modules uploaded to DefinitelyTyped are written using the \mintinline{text}{modules} template.
+On the other hand, 20\% of them are written using the \mintinline{text}{module-function}.
+The complexity of evaluating declaration files written with the \mintinline{text}{module-function} is considerably lower than for the other templates.
+Templates \mintinline{text}{module-class} and \mintinline{text}{module} were not evaluated.
+20\% was considered enough as a starting point.
+
+\input{figures/experiments/template-distribution-definitely-typed.tex}
+
+\subsection{Results}
+
+\input{figures/experiments/results/module-function-results.tex}
+
+\input{figures/experiments/results/module-class-results-flake-idgen.tex}
+
+\input{figures/experiments/results/module-class-results-route-parser.tex}
+
+\input{figures/experiments/results/module-class-results-timer-machine.tex}
+
+\input{figures/experiments/results/module-results.tex}
+
+\input{figures/experiments/results/datadog-metrics.tex}
 

### 2019-08-16 chapters/6-conclusions.tex
Add examples of the obtained results

@@
 
+\draft{Los code examples son la posta!
+Como estos ejecutan el codigo y la generacion de los declaration files depende de la runtime information que se junta, la calidad de los declaration files es dependiente de los code examples!}
+
+\draft{Compara contra definitely typed puede no estar tan bueno!
+Muchas cosas son elecciones del que escribe el file y hace lo que se le canta.
+Si se sigue comparando con fines de mejorar el desarrollo se puede hcer siempre y cuando sólo se tengan en cuenta los parametros importantes tipo los metodos de una clase, etc etc.}
+
 \draft{Decir que todos los bloques pueden mejorarse independientemente: Extraer mejor los readme files, mejorar la instrumentation usando Babel para quedarse siempre en ES5...
@@
 
-\section{Future Work}
+\section{Future Work} \label{sec:conclusions-future-work}
 

### 2019-08-16 figures/experiments/overall-funnel.tex
Add examples of the obtained results

--- 2019-07-29 figures/experiments/overallFunnel.tex
@@
 
-\begin{figure}[h]
+\begin{figure}[tp]
 	\hspace*{-0.28\textwidth}
@@
 				4.05,
-				9.37,
+				7.33,
 				24.02,
@@
 		\node[left=0.02\textwidth of a1] {Run-time Information};
-		\node[right=0.12\textwidth of a1] {\textbf{565}};
+		\node[right=0.12\textwidth of a1] {\textbf{442}};
 
@@
 It was possible to gather valid run-time information for only 25\% of the modules for which a Code Example was extracted.}
+	\label{fig:experiments-overall-funnel}
 \end{figure}

### 2019-08-16 figures/experiments/results/datadog-metrics.tex
Add examples of the obtained results

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.49\textwidth}
			\tscode{code/experiments/results/module/datadog-metrics/generated.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[datadog-metrics/index.d.ts - Generated]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.49\textwidth}
			\tscode{code/experiments/results/module/datadog-metrics/definitely-typed.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[datadog-metrics/index.d.ts - DefinitelyTyped]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.95\textwidth}
			\tscode{code/experiments/results/module/datadog-metrics/logger-datadog-metrics.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[datadog-metrics/logger.js]{\usebox{\mintedbox} \label{fig:subfloat-datadog-logger} }

	\caption[Missing properties | datadog-metrics]{\textbf{Missing properties | datadog-metrics} - Properties \mintinline{text}{aggregator} and \mintinline{text}{reporter} are not in the DefinitelyTyped version, but they appear in the generated declaration file. However, they are indeed used by the library, as exposed in lines 2 and 3 of the library's source code shown in \protect\subref{fig:subfloat-datadog-logger}.}
	\label{fig:experiments-results-module-datadog-metrics}
\end{figure}

### 2019-08-16 figures/experiments/results/module-class-results-flake-idgen.tex
Add examples of the obtained results

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/experiments/results/module-class/flake-idgen/generated.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[flake-idgen/index.d.ts - Generated]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/experiments/results/module-class/flake-idgen/definitely-typed.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[flake-idgen/index.d.ts - DefinitelyTyped]{\usebox{\mintedbox}}

	\caption[Module-class results | flake-idgen]{\textbf{Module-class results | flake-idgen} - Parameters of interface \mintinline{text}{ConstructorOptions} are correctly detected. Name of interface differs since it is automatically generated based on the name of the argument variable. Optional properties were not implemented, hence the \mintinline{text}{undefined} type for some properties. Analogously, callback \mintinline{text}{cb} is inferred as \mintinline{text}{undefined}.}
	\label{fig:experiments-results-module-class-flake-idgen}
\end{figure}

### 2019-08-16 figures/experiments/results/module-class-results-route-parser.tex
Add examples of the obtained results

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/experiments/results/module-class/route-parser/generated.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[route-parser/index.d.ts - Generated]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/experiments/results/module-class/route-parser/definitely-typed.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[route-parser/index.d.ts - DefinitelyTyped]{\usebox{\mintedbox}}

	\caption[Module-class results | route-parser]{\textbf{Module-class results | route-parser} - Constructor and methods are correctly identified. Specific signature of dictionary \mintinline{text}{params} is not detected. Instead, a generic \mintinline{text}{object} is inferred.}
	\label{fig:experiments-results-module-class-route-parser}
\end{figure}

### 2019-08-16 figures/experiments/results/module-class-results-timer-machine.tex
Add examples of the obtained results

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/experiments/results/module-class/timer-machine/generated.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[timer-machine/index.d.ts - Generated]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/experiments/results/module-class/timer-machine/definitely-typed.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[timer-machine/index.d.ts - DefinitelyTyped]{\usebox{\mintedbox}}

	\caption[Module-class results | timer-machine]{\textbf{Module-class results | timer-machine} - Parameter \mintinline{text}{started} is inferred as \mintinline{text}{undefined} instead of marking it as optional. Methods that were not executed do not appear in the generated declaration file.}
	\label{fig:experiments-results-module-class-timer-machine}
\end{figure}

### 2019-08-16 figures/experiments/results/module-function-results.tex
Add examples of the obtained results

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.49\textwidth}
			\tscode{code/experiments/results/module-function/abs/generated.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[abs/index.d.ts - Generated]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.49\textwidth}
			\tscode{code/experiments/results/module-function/abs/definitely-typed.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[abs/index.d.ts - DefinitelyTyped]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.49\textwidth}
			\tscode{code/experiments/results/module-function/dirname-regex/generated.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[dirname-regex/index.d.ts - Generated]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.49\textwidth}
			\tscode{code/experiments/results/module-function/dirname-regex/definitely-typed.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[dirname-regex/index.d.ts - DefinitelyTyped]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.49\textwidth}
			\tscode{code/experiments/results/module-function/escape-html/generated.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[escape-html/index.d.ts - Generated]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.49\textwidth}
			\tscode{code/experiments/results/module-function/escape-html/definitely-typed.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[escape-html/index.d.ts - DefinitelyTyped]{\usebox{\mintedbox}}

	\caption[Module-function results]{\textbf{Module-function results} - Results are shown for modules \mintinline{text}{abs}, \mintinline{text}{dirname-regex}, \mintinline{text}{escape-html}. On the left side the generated declaration file with \mintinline{text}{dts-generate}. On the right side the corresponding file in the DefinitelyTyped repository. Functions are correctly detected and input types are accurately inferred. Both files are parsed for comparison, as explained in \secref{sec:approach-parsing}. Therefore, subtle differences in the syntax between both files are not important.}
	\label{fig:experiments-results-module-function}
\end{figure}

### 2019-08-16 figures/experiments/results/module-results.tex
Add examples of the obtained results

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/experiments/results/module/is-uuid/generated.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[is-uuid/index.d.ts - Generated]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/experiments/results/module/is-uuid/definitely-typed.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[is-uuid/index.d.ts - DefinitelyTyped]{\usebox{\mintedbox}}

	\caption[Module results | is-uuid]{\textbf{Module results | is-uuid} - Methods that were not executed are not included in the declaration file.}
	\label{fig:experiments-results-module-is-uuid}
\end{figure}

### 2019-08-16 figures/experiments/template-distribution-definitely-typed.tex
Add examples of the obtained results

\begin{figure}[tp]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
			width=0.7\textwidth,
			ybar=0pt,
			ymax=6000,
			enlargelimits=0.15,
			bar width=0.08\textwidth,
			legend style={at={(0.5,-0.2)}, anchor=north,legend columns=-1},
			symbolic x coords={module-class,module-function,module},
			xtick=data,
			nodes near coords, 
			nodes near coords align={vertical},
			x tick label style={rotate=45,anchor=east},
		]
		\addplot coordinates {
			(module-class, 503)
			(module-function, 1201)
			(module, 4379)
		};
		\end{axis}
	\end{tikzpicture}

	\caption[TypeScript templates distribution in DefinitelyTyped]{\textbf{TypeScript templates distribution in DefinitelyTyped} - Out of a total of 6029, 72\% of the modules uploaded to the DefinitelyTyped repository use the \mintinline{text}{module} template.
	}

	\label{fig:experiments-typescript-templates-distribution-definitely-typed}
\end{figure}

### 2019-08-16 code/experiments/results/module-class/flake-idgen/definitely-typed.d.ts

@@ -1,0 +1,16 @@
+interface ConstructorOptions {
+	datacenter?: number;
+	worker?: number;
+	id?: number;
+	epoch?: number;
+	seqMask?: number;
+}
+
+declare namespace FlakeId { }
+
+declare class FlakeId {
+	constructor(options?: ConstructorOptions);
+	next(callback?: (err: Error, id: Buffer) => void): Buffer;
+}
+
+export = FlakeId;

### 2019-08-16 code/experiments/results/module-class/route-parser/definitely-typed.d.ts

@@ -1,0 +1,9 @@
+declare class Route {
+	constructor(spec: string);
+	match(pathname: string): { [i: string]: string } | false;
+
+	reverse(params: { [i: string]: any }): string | false;
+}
+
+declare namespace Route { }
+export = Route;

### 2019-08-16 code/experiments/results/module-class/timer-machine/definitely-typed.d.ts

@@ -1,0 +1,24 @@
+export as namespace Timer;
+export = Timer;
+
+declare namespace Timer {
+	type TimerEvent = "start" | "stop" | "time";
+}
+
+declare class Timer {
+	static get(reference: string): Timer;
+	static destroy(reference: string): Timer;
+
+	constructor(started?: boolean);
+
+	isStarted(): boolean;
+	isStopped(): boolean;
+	start(): void;
+	timeFromStart(): number;
+	stop(): void;
+	time(): number;
+	toggle(): void;
+	emitTime(): void;
+	valueOf(): number;
+	on(event: Timer.TimerEvent, callback?: () => void): void;
+}

### 2019-08-16 code/experiments/results/module-function/escape-html/definitely-typed.d.ts

@@ -1,0 +1,4 @@
+declare function escapeHTML(text: string): string;
+declare namespace escapeHTML { }
+
+export = escapeHTML;

### 2019-08-16 code/experiments/results/module/datadog-metrics/definitely-typed.ts

@@ -1,0 +1,15 @@
+export interface LoggerOptions {
+	apiKey?: string;
+	appKey?: string;
+	defaultTags?: string[];
+	flushIntervalSeconds?: number;
+	host?: string;
+	prefix?: string;
+}
+
+export class BufferedMetricsLogger {
+	constructor(
+		options: LoggerOptions
+	);
+	// ...
+}

### 2019-08-16 code/experiments/results/module/datadog-metrics/generated.d.ts

@@ -1,0 +1,16 @@
+export interface I__opts {
+	'aggregator': undefined;
+	'defaultTags': Array<any>;
+	'reporter': undefined;
+	'apiKey': string;
+	'appKey': undefined;
+	'agent': undefined;
+	'host': string;
+	'prefix': string;
+	'flushIntervalSeconds': number;
+}
+
+export class BufferedMetricsLogger {
+	constructor(opts: I__opts);
+	// ...
+}

### 2019-08-16 code/experiments/results/module/datadog-metrics/logger-datadog-metrics.js

@@ -1,0 +1,10 @@
+function BufferedMetricsLogger(opts) {
+	this.aggregator = opts.aggregator || new Aggregator(opts.defaultTags);
+	this.reporter = opts.reporter || new DataDogReporter(opts.apiKey, opts.appKey, opts.agent);
+	this.host = opts.host;
+	this.prefix = opts.prefix || '';
+	this.flushIntervalSeconds = opts.flushIntervalSeconds;
+
+	// ...
+	// ...
+}

### 2019-08-16 code/experiments/results/module/is-uuid/definitely-typed.d.ts

@@ -1,0 +1,7 @@
+export function v1(value: string): boolean;
+export function v2(value: string): boolean;
+export function v3(value: string): boolean;
+export function v4(value: string): boolean;
+export function v5(value: string): boolean;
+export function nil(value: string): boolean;
+export function anyNonNil(value: string): boolean;

### 2019-08-18 chapters/4-approach.tex
Add examples of the obtained results
Finish experiments section and conclusion

@@
 
-\subsection{Parsing}
+\subsection{Parsing} \label{sec:approach-parsing}
 Before comparison, declaration files need to be parsed so that both of them share the same structure.
@@
 
-\section{Type Inference}
+\section{Type Inference} \label{sec:approach-type-inference}
 Type inference for JavaScript is a topic that was deeply investigated.

### 2019-08-18 chapters/5-experiments.tex
Finish experiments section and conclusion

--- 2019-07-27 chapters/5-experiments.tex
@@
 \chapter{Experiments}\label{chap:experiments}
+Declaration files were generated for existing modules uploaded to the NPM registry.
+The DefinitelyTyped repository was used as a benchmark.
+Each one of the generated files was compared against the corresponding declaration file already uploaded to the repository.
 
-Experiments often include benchmarks.
-You need to describe you benchmarking setup: parameters of the machine used (architecture, memory, etc), parameters of the software used (OS version, libraries used and their version), the benchmark programs.
-The programs need not be part of the thesis, but they should be described and available separately.
+\figref{fig:experiments-overall-funnel} shows that a declaration file was generated for 244 modules out of 6029 modules.
+The following sections elaborate each stage of the process and expose the reasons for generating a declaration file for only 4\% of the modules uploaded to DefinitelyTyped.
+Samples of the generated declaration files for templates \mintinline{text}{module}, \mintinline{text}{module-class} and \mintinline{text}{module-function} are presented in \secref{sec:experiments-declaration-files-generation}.
 
-If you measure run times and evaluate your findings statistically, then there is great potential for errors.
-Consult SIGPLAN's checklist for empirical evaluations \url{http://www.sigplan.org/Resources/EmpiricalEvaluation/} for advice.
+\input{figures/experiments/overall-funnel.tex}
 
-\input{figures/experiments/figExample}
-\input{figures/experiments/tabAccuracy}
+Finally, an analysis on the usage of JavaScript operators can be seen in \secref{sec:experiments-js-operators-usage}.
+The goal is to gather empirical evidence that shows which are the most common types that developers use with every JavaScript operator.
 
-\section{Definitely Typed Modules}
+\section{Definitely Typed Benchmark}
 
-\section{Code Examples Extraction}
-\subsection{Repositories URL}
-\subsection{Readme Files}
 \subsection{Code Examples}
+Retrieving the code examples from the JavaScript libraries' repositories proved to be a pragmatic way of capturing the types.
+However, as shown in \figref{fig:experiments-overall-funnel}, working code examples for only 2260 modules could be retrieved.
+The process of getting a valid code example for a module is divided in 4 blocks:
+\begin{itemize}
+	\item Extracting repositories url.
+	\item Extracting readme files.
+	\item Extracting code examples within readme files.
+	\item Executing code examples and discarding failing ones.
+\end{itemize}
 
-\section{Code Execution}
+The results obtained for each on of them are described in the following sections.
 
-\section{Run-time Information Gathering}
+\subsubsection{Repositories URL}
+The url of the repositories could be retrieved for only 4974 modules.
+More than 1000 modules do not have the repository entry in their corresponding \mintinline{text}{package.json} files.
+Therefore, the \mintinline{text}{npm view <module> repository.url} command returns an empty value.
+This is even happening for important modules like \mintinline{text}{ace}.
 
-\section{Declaration Files Generation}
+\subsubsection{Readme Files}
+700 modules simply do not have a readme file in their repositories.
+The implementation does contemplate, however, different naming conventions like \mintinline{text}{readme.md} or \mintinline{text}{README.md}.
 
-\section{Evaluation}
+\subsubsection{Code Examples Extraction}
+The 50\% loss is mainly explained because developers did not wrap their code around a block using the \mintinline{text}{javascript} or \mintinline{text}{js} tags.
+As explained in \secref{sec:conclusions-future-work}, the process of extracting code examples from a repository can be greatly improved.
+However, counting with code examples for 2200 modules was considered to be enough for evaluating the generation of declaration files.
 
-\section{JS Operators Usage}
-\todo{Agregar que es una matriz simetrica.
-Que era para hacer un analisis mas global.
-Para algunos operators se puede inferir que el orden de los tipos sí afecta.
-Por ejemplo para instance of donde obviamente importa el orden.
-Pero para == por ejemplo generalmente se pone a == null y no null == a}.
+\subsubsection{Code Examples Execution}
+The 2260 extracted code examples were executed by installing the required packages and running the code as a \mintinline{text}{node} application.
+Working and functional code examples could only be extracted for 1448 modules.
+812 modules did not run correctly and were discarded.
+Some failing samples were analyzed and there were mainly two reasons for the failure:
+\begin{enumerate}
+	\item The code example had been properly extracted but the code itself was not working. It was executing the library in an unsupported way, hence the error at run-time.
+	\item The extracted code example was not intended to be executed or it was not even valid JavaScript code.
+\end{enumerate}
+
+Improving the retrieval of code examples for a specific module will definitely have a positive impact in this number.
+
+\subsection{Run-time Information Gathering} \label{sec:experiments-run-time-information-gathering}
+Run-time information could only be extracted for 442 modules out of 1448 modules with working code examples.
+As explained in \secref{sec:run-time-wrapper-objects}, the behavior of the code under analysis was explicitly modified by wrapping the arguments around Wrapper Objects.
+It definitely had an influence on some modules by causing run-time errors when executing the code examples.
+Furthermore, Jalangi's instrumentation itself caused some executions to fail, since the modules contained JavaScript features that are not supported by Jalangi.
+As a result, these modules could not be executed.
+
+An instrumentation without user defined analysis modules was not applied to the 1448 working code examples.
+Therefore, it was not possible to determine which modules were failing only because of Jalangi's own limitations.
+
+\subsection{Declaration Files Generation} \label{sec:experiments-declaration-files-generation}
+Finally, a declaration file was generated for 244 modules.
+Despite the correct execution of instrumented code examples, some of the 442 modules mentioned in \secref{sec:experiments-run-time-information-gathering} did not contain suitable run-time information for generating a declaration file.
+198 of the extracted code examples did not execute the library itself and therefore the collected run-time information was not useful for generating a declaration file.
+
+\subsubsection{Results}
+The following section exhibits some samples of the 244 generated declaration files.
+It shows some results for each of the implemented templates: \mintinline{text}{module}, \mintinline{text}{module-function} and \mintinline{text}{module-class}.
+
+\figref{fig:experiments-results-module-function} shows the generated declaration files for simple modules like \mintinline{text}{abs}, \mintinline{text}{dirname-regex} and \mintinline{text}{escape-html}.
+All of them were generated using the \mintinline{text}{module-function} template.
+There are no differences between the generated files and the corresponding declaration files uploaded to DefinitelyTyped.
+
+Templates of type \mintinline{text}{module-class} are shown for modules \mintinline{text}{flake-idgen}, \mintinline{text}{route-parser} and \mintinline{text}{timer-machine} in \figref{fig:experiments-results-module-class-flake-idgen}, \figref{fig:experiments-results-module-class-route-parser} and \figref{fig:experiments-results-module-class-timer-machine}, respectively.
+Properties of interfaces and class methods are correctly generated.
+Optional parameters are not detected, as it was not considered for the implementation.
+
+Finally, \mintinline{text}{module} template is presented for \mintinline{text}{is-uuid} module in \figref{fig:experiments-results-module-is-uuid}.
+
+It is worth mentioning that for some libraries the declaration file in DefinitelyTyped was not correct.
+For example, for \mintinline{text}{datadog-metrics}, some properties of an interface were included in the generated declaration file but they were not present in the one in DefinitelyTyped.
+However, as shown in \figref{fig:experiments-results-module-datadog-metrics}, the properties are indeed used in the source code and should be included.
+The limitations of using DefinitelyTyped as a baseline for comparison are explained in detail in \chapref{chap:conclusion}.
+
+\input{figures/experiments/results/module-function-results.tex}
+
+\input{figures/experiments/results/module-class-results-flake-idgen.tex}
+
+\input{figures/experiments/results/module-class-results-route-parser.tex}
+
+\input{figures/experiments/results/module-class-results-timer-machine.tex}
+
+\input{figures/experiments/results/module-results.tex}
+
+\input{figures/experiments/results/datadog-metrics.tex}
+
+\subsection{Evaluation}
+As shown in \figref{fig:experiments-typescript-templates-distribution-definitely-typed}, 20\% of the declaration files in DefinitelyTyped are written using the \mintinline{text}{module-function}.
+However, 57\% of the 244 generated declaration files are written with the \mintinline{text}{module-function} template.
+Additionally, the complexity of evaluating declaration files written with the \mintinline{text}{module-function} is considerably lower than for other templates.
+The evaluation for templates \mintinline{text}{module-class} and \mintinline{text}{module} was not implemented.
+
+33 out of 116 evaluated modules have no difference with their corresponding declaration file in DefinitelyTyped.
+
+\input{figures/experiments/template-distribution-definitely-typed.tex}
+
+\section{JavaScript Operators Usage} \label{sec:experiments-js-operators-usage}
+The following section presents a study on the usage of JavaScript operators.
+The types of the operands of each operator were recorded with Jalangi.
+The goal is to better understand how developers are using JavaScript operators and what they expect from them.
+The study was performed on the 442 modules used for gathering the run-time information, as shown in \figref{fig:experiments-overall-funnel}.
+
+The results might have exposed unexpected development patterns like multiplying a \mintinline{text}{function} and an \mintinline{text}{object}.
+Conversely, the obtained results support the idea that developers are using operators in a traditional way, despite the possibility in JavaScript of using any type with any operator.
+
+Table \ref{table:experiments-number-of-events-per-operator} summarizes the distribution of events per operator. In total, 1263866 events were collected.
+
+The most relevant results are presented in the following sections.
+
+\input{figures/experiments/operators/operators-distribution.tex}
+
+\subsection{Relational operators}
+As shown in \figref{fig:experiments-relational-operators}, operators \mintinline{text}{<, >, <=, >=} are only used with both operands of type \mintinline{text}{number}.
+
+Operator \mintinline{text}{instanceof} is mostly used in the expected way. First and second operand are of type \mintinline{text}{object} and \mintinline{text}{function}, respectively.
+
+Operator \mintinline{text}{in} will check if the second operand has a property that matches the name of the first operand.
+Most of the events are \mintinline{text}{[number] in [array]}, which means that developers are using the operator to check whether an array has a specific key.
+The expected usage of this operator is \mintinline{text}{[string] in [object]}, which matches the second most used combination.
+
+\subsection{Additive operators}
+Operator \mintinline{text}{+} is mostly used for adding numbers.
+The combination \mintinline{text}{number-number} represents 77\% of all collected events for this operator.
+Moreover, 16\% of all events corresponds to the pure string concatenation combination \mintinline{text}{string-string}.
+Finally, the concatenation of a string a number represents only 1.5\% of all events.
+
+On the other hand, operator \mintinline{text}{-} is only used with type \mintinline{text}{number} for both operands.
+
+Results for additive operators are summarized in \figref{fig:experiments-additive-operators}.
+
+\subsection{Multiplicative operators}
+It can be seen in \figref{fig:experiments-multiplicative-operators} that operators \mintinline{text}{*, /, %} are only used with type \mintinline{text}{number}.
+
+\subsection{Equality operators}
+\figref{fig:experiments-equality-operators} exhibits the usage of operators \mintinline{text}{==, !=, ===, !==}.
+
+Number and string comparison represent the most used combination for both operators \mintinline{text}{==} and \mintinline{text}{===}.
+
+Operators \mintinline{text}{!=, !==} are mostly used for performing a comparison against \mintinline{text}{null}.
+
+\subsection{Conditional operators}
+Jalangi executes the same callback for all operators that trigger a condition check before branching.
+Therefore, \figref{fig:experiments-conditional-operators} combines operators \mintinline{text}{if-else}, \mintinline{text}{switch-case}, \mintinline{text}{while}, \mintinline{text}{for}, \mintinline{text}{||}, \mintinline{text}{&&} and \mintinline{text}{?
+:}.
+
+Representing an 82.63\% of all collected events, type \mintinline{text}{boolean} is the most used type with these operators.
+Types \mintinline{text}{number} and \mintinline{text}{string} come in second and third place, respectively.
+
 \input{figures/experiments/operators/relational-operators/relationalOperators.tex}
+
 \input{figures/experiments/operators/additive-operators/additiveOperators.tex}
+
 \input{figures/experiments/operators/multiplicative-operators/multiplicativeOperators.tex}
+
 \input{figures/experiments/operators/equality-operators/equalityOperators.tex}
+
 \input{figures/experiments/operators/conditional-operators/conditionalOperators.tex}

### 2019-08-18 chapters/6-conclusions.tex
Finish experiments section and conclusion

--- 2019-08-14 chapters/6-conclusions.tex
@@
 
-Brief wrapup of the achievements of this thesis.
+We have presented \mintinline{text}{dts-generate}, a tool for generating a TypeScript declaration file for a specific JavaScript library.
+The tool downloads code samples written by the developers from the library's repository.
+It uses these samples to execute the library and gather data flow and type information.
+The tool finally generates a TypeScript declaration file based on the information gathered at run-time.
 
-Directions for future work (if applicable).
+We developed an architecture that supports the automatic generation of declaration files for specific JavaScript libraries without additional manual tasks.
+The architecture contemplates a future incorporation of a Symbolic Execution Engine that refines the initial code base enabling the exploration of new execution paths.
+However not implemented in this work, its incorporation would result in small incremental modifications to the presented architecture as it is considered to only expand the existing code base.
 
-\draft{Una de las conclusiones mas importantes es que la type inference al final no es tan importante.
-Lo dificil es generar el declaration file como corresponde, detectar las interfaces, etc.
-El tipo de las variables puede, en principio, inferirse de los ejemplos de los developers.
-Es una aproximación bastante acertada y se atacan los otros problemas.}
+Building and end-to-end solution for the generation of TypeScript declaration files was prioritized over type inference accuracy.
+Consequently, types were taken over from the values at run-time.
+Obtaining the types from the code examples extracted from the repositories proved to be a pragmatic and effective approximation, enabling to work on specific aspects regarding the TypeScript declaration file generation.
 
-\draft{Decir que todos los bloques pueden mejorarse independientemente: Extraer mejor los readme files, mejorar la instrumentation usando Babel para quedarse siempre en ES5...
-mejorar la generation de declaration files, etc.}
+We built a mechanism to automatically create declaration files for every module uploaded to DefinitelyTyped and evaluate the results, creating a TypeScript declaration files parser and a comparator.
 
-\section{Future Work}
+We exposed the fundamental aspect of capturing the developer's intention when inferring types in JavaScript.
+Instead of applying constraints and restrictions for operations with certain types, we presented a proposal where common practices are favored.
+Uncommon usage is not forbidden but greatly disfavored.
+Used in a first place merely as an initial code base for executing the library and approximating types, code examples turned out to be extremely useful for inferring types.
+They expose the developer's intention explicitly through code.
+
+Finally, the architecture is composed of different blocks that interact with each other.
+Each block is independent and has a well defined behavior as well as clear input and output values.
+As a result, each block can be independently and simultaneously improved.
+
+\section{Future Work} \label{sec:conclusions-future-work}
+
+\subsection{Instrumentation}
+\draft{Babel}
+
+\subsection{Declaration files generation}
+\draft{Optional parameters}
+\draft{Functions as paramters}
+\subsubsection{Uploading Results to Definitely Typed Repository}
 
 \subsection{Type Inference}
-\todo{Deberia ser hecho en otro modulo que no sea el runtime information. El runtime information deberia exponer unicamente los datos que le hacen falta al otro para inferir los tipos.}
+% \todo{Deberia ser hecho en otro modulo que no sea el runtime information. El runtime information deberia exponer unicamente los datos que le hacen falta al otro para inferir los tipos.}
 
-\todo{Tambien se puede identificar patterns de programacion donde siempre los tipos son los mismos!!! El concepto de inferir el tipo en realidad pasa a detectar la intencion del developer.}
+% \todo{Tambien se puede identificar patterns de programacion donde siempre los tipos son los mismos!!! El concepto de inferir el tipo en realidad pasa a detectar la intencion del developer.}
 
 \subsubsection{Return values}
-\todo{El return value se podria incluir en el analysis de Jalangi, detectando las interactions que tiene etc etc.
+% \todo{El return value se podria incluir en el analysis de Jalangi, detectando las interactions que tiene etc etc.
 Esto no es tan facil porque no se puede simplemente inspeccionar el return value, sino que hay que saber para qué se usa.
@@
 \subsubsection{Variable Names}
-\todo{Se puede extraer info del nombre de las variables y del nombre de los metodos, etc. La developer's intention.}
+% \todo{Se puede extraer info del nombre de las variables y del nombre de los metodos, etc. La developer's intention.}
 
-\subsection{Uploading Results to Definitely Typed Repository}
+\subsubsection{Variable Content}
+
+\subsubsection{Function Names}

### 2019-08-18 figures/experiments/operators/multiplicative-operators/multiplicativeOperators.tex
Add content to the TypeInference section
Finish experiments section and conclusion

@@
 	\end{tikzpicture}
-	\caption[Relational operators]{\textbf{Type distribution for multiplicative operators} - All multiplicative operators are mainly used with the \textit{number-number} combination}
+	\caption[Relational operators]{\textbf{Type distribution for multiplicative operators} - All multiplicative operators are mainly used with the \textit{number-number} combination.}
+	\label{fig:experiments-multiplicative-operators}
 \end{figure}

### 2019-08-18 figures/experiments/operators/operators-distribution.tex
Finish experiments section and conclusion

\begin{table}[tp]
	\centering

	\begin{tabular}{ |c|c|c| } 
		\hline
		\textbf{Operator} & \multicolumn{2}{|c|}{\textbf{Number of events}} \\
		\hline
		\mintinline{text}{<} & 143666 & 11.37\% \\
		\mintinline{text}{>} & 8914 & 0.71\% \\
		\mintinline{text}{<=} & 2537 & 0.20\% \\
		\mintinline{text}{>=} & 1835 & 0.15\% \\
		\mintinline{text}{instanceof} & 1823 & 0.14\% \\
		\mintinline{text}{in} & 2242 & 0.18\% \\
		\mintinline{text}{+} & 261669 & 20.70\% \\
		\mintinline{text}{-} & 196598 & 15.56\% \\
		\mintinline{text}{*} & 82821 & 6.55\% \\
		\mintinline{text}{/} & 318 & 0.03\% \\
		\mintinline{text}{==} & 4094 & 0.32\% \\
		\mintinline{text}{!=} & 377 & 0.03\% \\
		\mintinline{text}{===} & 28815 & 2.28\% \\
		\mintinline{text}{!==} & 33265 & 2.63\% \\
		\mintinline{text}{conditional} & 366215 & 28.98\% \\
		\mintinline{text}{%} & 939 & 0.07\% \\
		\mintinline{text}{|} & 17815 & 1.41\% \\
		\mintinline{text}{>>>} & 64668 & 5.12\% \\
		\mintinline{text}{delete} & 97 & 0.01\% \\
		\mintinline{text}{^} & 97 & 0.01\% \\
		\mintinline{text}{>>} & 13741 & 1.09\% \\
		\mintinline{text}{&} & 29063 & 2.30\% \\
		\hline
		\textbf{Total} & \textbf{1263866} & \textbf{100\%}\\
		\hline
	\end{tabular}
	\caption{\textbf{Number of events per operator} - More than 1 million events were collected. The \mintinline{text}{conditional} operator includes all operators that trigger a condition check before branching: \mintinline{text}{if-else}, \mintinline{text}{switch-case}, \mintinline{text}{while}, \mintinline{text}{for}, \mintinline{text}{||}, \mintinline{text}{&&} and \mintinline{text}{? :}.}
	\label{table:experiments-number-of-events-per-operator}
\end{table}

### 2019-08-18 figures/experiments/template-distribution-definitely-typed.tex
Finish experiments section and conclusion

@@
 			ybar=0pt,
-			ymax=6000,
+			ymax=100,
 			enlargelimits=0.15,
 			bar width=0.08\textwidth,
-			legend style={at={(0.5,-0.2)}, anchor=north,legend columns=-1},
+			legend columns=1,
+			legend cell align={left},
+			legend style={at={(0.6,0.95)}, anchor=north},
 			symbolic x coords={module-class,module-function,module},
 			xtick=data,
+			ylabel=Percentage,
 			nodes near coords, 
@@
 		\addplot coordinates {
-			(module-class, 503)
-			(module-function, 1201)
-			(module, 4379)
+			(module-class, 8)
+			(module-function, 20)
+			(module, 72)
 		};
+
+		\addplot coordinates {
+			(module-class, 12)
+			(module-function, 57)
+			(module, 31)
+		};
+		\legend{DefinitelyTyped (Total: 6029), Generated (Total: 244)}
 		\end{axis}
@@
 
-	\caption[TypeScript templates distribution in DefinitelyTyped]{\textbf{TypeScript templates distribution in DefinitelyTyped} - Out of a total of 6029, 72\% of the modules uploaded to the DefinitelyTyped repository use the \mintinline{text}{module} template.
+	\caption[TypeScript templates distribution | Generated \& DefinitelyTyped]{\textbf{TypeScript templates distribution | Generated \& DefinitelyTyped} - Out of a total of 6029, 72\% of the modules uploaded to the DefinitelyTyped repository use the \mintinline{text}{module} template and only 20\% use the \mintinline{text}{module-function} one.
+However, 57\% of the 244 generated declaration files use the \mintinline{text}{module-function} template.
 	}

### 2019-08-19 chapters/4-approach.tex
Add introduction

@@
 The following section presents the restrictions that different type checkers perform on JavaScript code.
-The state of the art is shown through a comparison between \mintinline{text}{TypeScript}{TypeScript}, \mintinline{text}{Flow} and \mintinline{text}{Infernu}.
+The state of the art is shown through a comparison between \mintinline{text}{TypeScript}, \mintinline{text}{Flow} and \mintinline{text}{Infernu}.
 All of them introduce the concept of restrictions for JS from different angles and for different use cases.
@@
 \subsubsection{Flow}
-Flow is a static type checker for JavaScript developed by Facebook \citep{flow}.
+Flow is a static type checker for JavaScript developed by Facebook \citep{flow}\citep{FLOW_DBLP:journals/pacmpl/ChaudhuriVGRL17}.
 It is intended to aid developers while writing JS code by detecting different errors.

### 2019-08-19 chapters/6-conclusions.tex
Add introduction

@@
 Uncommon usage is not forbidden but greatly disfavored.
-Used in a first place merely as an initial code base for executing the library and approximating types, code examples turned out to be extremely useful for inferring types.
+Accordingly, we collected evidence regarding the usage of JavaScript operators by analyzing 400 libraries.
+Furthermore, used in a first place merely as an initial code base for executing the library and approximating types, code examples turned out to be extremely useful for inferring types.
 They expose the developer's intention explicitly through code.
@@
 \subsubsection{Function Names}
+
+\draft{JAVERT!!}

### 2019-08-19 code/introduction/javascript-wtfs.js

@@ -1,0 +1,14 @@
+"0" == false; // true
+true == "1.00"; // true
+false == "    \n\r\t     "; // true
+false == []; // true
+0 == []; // true
+ null == undefined; // true 4
+"01" < "00100"; // false
+[1] + 1; //'11'
+[2] == "2"; // true
+null + undefined + [1, 2, 3] // 'NaN1,2,3'
+"hello world".lenth + 1 // NaN | note 'lenth' instead of 'length'
+[1, 15, 20, 100].sort() // [ 1, 100, 15, 20 ]
+typeof null; // object
+null instanceof Object; // false

### 2019-08-19 chapters/3-background.tex
Finish TypeScript Declaration File Generation section
Finish experiments section and conclusion
Finish type coercion background section

--- 2019-07-23 chapters/3-background.tex
@@
 \chapter{Background}\label{chap:background}
-Explain the general context of your work.
-Explain mathematical background required and introduce notation.
-\input{figures/background/algBackpropagation}
-\input{figures/background/figTikz}
+\section{JavaScript} \label{sec:background-javascript}
+% \todo{Agregar img con data de la survey de stackoverflow indicando la popularidad de JS. Agarrar data de encuestas viejas y hacer una comparación de agunas techs, tipo JS, PHP, Python, Java, etc.}
 
-\section{JavaScript}
-\subsection{TypeCoercion}
-\subsection{NPM Packages}
+\input{figures/background/stackoverflowSurvey-ProgrammingLanguages2019.tex}
 
-\section{TypeScript}
-\subsection{Declaration Files}
-\subsection{Definitely Typedd}
+\input{figures/background/stackoverflowSurvey-ProgrammingLanguagesEvolution.tex}
 
-\section{Jalangi}
+% \todo{Agregar tema serverless. Que NodeJS tiene el 75\% de los deploys}
+
+\subsection{TypeCoercion} \label{sec:background-js-type-coercion}
+Type coercion is the process of converting one type into another one.
+In JavaScript, a variable can be converted into different types depending on the operator and the value of the other operands.
+The specific definition can be found in the ECMAScript Language Specification \citep{ecma-script}.
+The following section will cover the basic forms of type coercion but providing a detailed explanation of how JavaScript converts a variable type into another one for each operator is beyond the scope of this section.
+It is however intended to give insight about this kind of behavior to show its importance while performing type inference in JavaScript.
+As a result, specific operator related type coercion is explained only for operator \mintinline{text}{+}.
+
+Finally, there are excellent references regarding JavaScript's Type Coercion, like Kyle Simpson's book `You Don't Know JS: Types \& Grammar' \citep{you-dont-know-js}.
+
+\subsubsection{Types of conversion}
+\label{types_of_conversion}
+There are two types of type coercion: explicit and implicit \citep{you-dont-know-js}.
+Explicit type coercion is simply when the developer chooses to convert one type into another one.
+The developer achieves this by writing a piece of code that shows her intention, as shown in \coderef{code:background-explicit-type-coercion}.
+This type of coercion is also called type casting \citep{you-dont-know-js}.
+
+\begin{code}
+	\jscode{code/background/javascript/type-coercion/explicit-type-coercion.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[Explicit JavaScript Type Coercion]{\textbf{Explicit JavaScript Type Coercion} - The developer explicitly transforms a type into another one.
+The return values of \mintinline{text}{String()}, \mintinline{text}{Number()} and \mintinline{text}{Boolean()} are always of type \mintinline{text}{string}, \mintinline{text}{number} and \mintinline{text}{boolean}, respectively.}
+	\label{code:background-explicit-type-coercion}
+  \end{code}
+
+On the other hand, depending on the operator and the context, JavaScript will perform implicit type transformations automatically.
+If the developer is not aware of this, she may face some unexpected behavior, as shown in \coderef{code:background-implicit-type-coercion}.
+
+\begin{code}
+	\jscode{code/background/javascript/type-coercion/implicit-type-coercion.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[Implicit JavaScript Type Coercion]{\textbf{Implicit JavaScript Type Coercion} - Examples given by Douglas Crockford in his talk `JavaScript: The Good Parts' at Google\footnote{https://www.youtube.com/watch?v=hQVTIJBZook}.}
+	\label{code:background-implicit-type-coercion}
+\end{code}
+
+In any case, JavaScript will always convert any value into a primitive value \citep{ecma-script}, regardless of explicit or implicit type coercion.
+There is no coercion mechanism that results in an \mintinline{text}{object} or a \mintinline{text}{function}.
+
+It is fundamental to understand how a value gets coerced into one of these types.
+JavaScript mainly performs this by applying one of three so called \mintinline{text}{Abstract Operations} \citep{ecma-script}: \mintinline{text}{ToString()}, \mintinline{text}{ToNumber()} and \mintinline{text}{ToBoolean()}.
+
+The operation \mintinline{text}{ToPrimitive()} is also called within the operators.
+However, each operator will call either \mintinline{text}{ToString()}, \mintinline{text}{ToNumber()} or \mintinline{text}{ToBoolean()} after calling \mintinline{text}{ToPrimitive()}.
+This operator will be explained in an independent section.
+
+\subsubsection{ToString}
+JavaScript performs this conversion in a very intuitive way, as explained in \coderef{code:background-to-string-operation}.
+Every primitive gets converted as expected.
+Objects are converted into strings by calling \mintinline{text}{ToPrimtive()} with \mintinline{text}{hint = string}, as presented in \coderef{code:background-to-string-implementation}.
+
+\begin{code}
+	\jscode{code/background/javascript/type-coercion/to-string.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[ToString implementation]{\textbf{ToString implementation}}
+	\label{code:background-to-string-implementation}
+\end{code}
+
+\begin{code}
+	\jscode{code/background/javascript/type-coercion/string-conversion.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[ToString operation]{\textbf{ToString operation}}
+	\label{code:background-to-string-operation}
+\end{code}
+
+\subsubsection{ToNumber}
+\coderef{code:background-to-number-operation} presents this type of conversion.
+Strings and boolean values get converted in the expected way.
+Moreover, \mintinline{text}{undefined} will be converted into \mintinline{text}{NaN} and \mintinline{text}{null} into \mintinline{text}{0}.
+Similarly, objects are converted into numbers by calling \mintinline{text}{ToPrimtive()} with \mintinline{text}{hint = number}, as presented in \coderef{code:background-to-number-implementation}.
+
+\begin{code}
+	\jscode{code/background/javascript/type-coercion/to-number.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[ToNumber implementation]{\textbf{ToNumber implementation}}
+	\label{code:background-to-number-implementation}
+\end{code}
+
+\begin{code}
+	\jscode{code/background/javascript/type-coercion/number-conversion.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[ToNumber operation]{\textbf{ToNumber operation}}
+	\label{code:background-to-number-operation}
+\end{code}
+
+\subsubsection{ToBoolean}
+Only the values presented in \coderef{code:background-to-boolean-operation} are converted into \mintinline{text}{false}.
+Everything else is coerced to \mintinline{text}{true}.
+That means, that every non-empty \mintinline{text}{string}, every non-zero \mintinline{text}{number} and every \mintinline{text}{object} gets coerced to \mintinline{text}{true}.
+
+Values that are converted into \mintinline{text}{false}, like \mintinline{text}{null} or \mintinline{text}{undefined} are called \textit{falsy} values \citep{you-dont-know-js}.
+
+\begin{code}
+	\jscode{code/background/javascript/type-coercion/boolean-conversion.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[ToBoolean operation]{\textbf{ToBoolean operation}}
+	\label{code:background-to-boolean-operation}
+\end{code}
+
+\subsubsection{ToPrimitive}
+The signature of this function is \mintinline{text}{ToPrimitive(value, preferredType)}.
+The allowed values for \mintinline{text}{preferredType} are \mintinline{text}{string} or \mintinline{text}{number}.
+Every JavaScript operator will try to convert an object into a primitive by invoking \mintinline{text}{ToPrimitive} with a different value for \mintinline{text}{preferredType}.
+If no value is passed, \mintinline{text}{number} is considered as the default, except for JavaScript native \mintinline{text}{Date} objects, where \mintinline{text}{string} is used as default.
+
+An implementation of this behavior is shown in \coderef{code:background-to-primitive-operation}.
+
+\begin{code}
+	\jscode{code/background/javascript/type-coercion/to-primitive/to-primitive.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[ToPrimitive operation]{\textbf{ToPrimitive operation}}
+	\label{code:background-to-primitive-operation}
+\end{code}
+
+An important aspect that should be considered, is that the type of the return value of \mintinline{text}{ToPrimitive()} does not necessarily have to match the chosen \mintinline{text}{preferredType}.
+This means concretely that, for example, you can call \mintinline{text}{ToPrimitive(a, number)} and get a \mintinline{text}{string} as a result.
+Of course, when performing an explicit type conversion, JS will eventually convert whatever \mintinline{text}{ToPrimitive} returns into the correct type.
+But some operators will perform different implicit type conversions \textit{after} calling \mintinline{text}{ToPrimitive()}.
+The specific type conversion will then be dependent on the type of the return value of \mintinline{text}{ToPrimitive()} and not on the chosen \mintinline{text}{preferredType}.
+
+All objects that are converted to \mintinline{text}{boolean} will be coerced to \mintinline{text}{true}. The following paragraphs explain how objects get converted into strings or numbers.
+
+\subsubsection{Object to String}
+The object is first converted into a primitive value calling the \mintinline{text}{ToPrimitive()} method, using \mintinline{text}{string} as \mintinline{text}{preferredType}.
+The obtained primitive value is then normally converted into a \mintinline{text}{string}, as explained before in \coderef{code:background-to-string-implementation}.
+
+\coderef{code:background-object-into-string} shows a straightforward example.
+On the other hand, \coderef{code:background-object-into-string-not-string-return-value} provides an example where \mintinline{text}{ToPrimitive(val, String)} does not return a \mintinline{text}{string}, either because simply \mintinline{text}{toString()} does not return a \mintinline{text}{string} or because \mintinline{text}{valueOf()} gets being called instead of \mintinline{text}{toString()}.
+
+\begin{code}
+	\jscode{code/background/javascript/type-coercion/to-primitive/normal-object-to-string.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[Object into string conversion]{\textbf{Object into string conversion} - An object has a \mintinline{text}{toString()} method that returns a string.}
+	\label{code:background-object-into-string}
+\end{code}
+
+\begin{code}
+	\jscode{code/background/javascript/type-coercion/to-primitive/object-to-string-returning-not-a-string.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[Object into string conversion]{\textbf{Object into string conversion} - An object that does not return a \mintinline{text}{string} even though \mintinline{text}{ToPrimitive()} is called with \mintinline{text}{hint = string}.}
+	\label{code:background-object-into-string-not-string-return-value}
+\end{code}
+
+\subsubsection{Object to Number}
+The object gets converted into a primitive by calling \mintinline{text}{ToPrimitive(val, Number)}.
+Normal number conversion is then applied to this primitive value.
+\coderef{code:background-object-into-number} provides a basic example.
+Similarly, \coderef{code:background-object-into-string-not-number-return-value} shows how \mintinline{text}{ToPrimitive()} might return something that is not a \mintinline{text}{number}.
+
+\begin{code}
+	\jscode{code/background/javascript/type-coercion/to-primitive/normal-object-to-number.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[Object into number conversion]{\textbf{Object into number conversion} - An object has a \mintinline{text}{valueOf()} method that returns a \mintinline{text}{number}.}
+	\label{code:background-object-into-number}
+\end{code}
+
+\begin{code}
+	\jscode{code/background/javascript/type-coercion/to-primitive/object-to-number-returning-not-a-number.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[Object into number conversion]{\textbf{Object into number conversion} - An object that does not return a \mintinline{text}{number} even though \mintinline{text}{ToPrimitive()} is called with \mintinline{text}{hint = number}.}
+	\label{code:background-object-into-string-not-number-return-value}
+\end{code}
+
+\subsubsection{The \mintinline{text}{+} operator}
+If any of the operands is an object, it will first convert it into a primitive by passing no \mintinline{text}{preferredType} to the \mintinline{text}{ToPrimitive()} method.
+Afterwards, if any value is a \mintinline{text}{string}, this operator will implicitly convert both values into a \mintinline{text}{string} and perform a normal string concatenation between them.
+Else, it will implicitly convert each value into a \mintinline{text}{number} and perform a normal arithmetic addition.
+An implementation is presented in \coderef{code:background-plus-operator-implementation}.
+
+\coderef{code:background-plus-operator-simple-examples} shows some basic examples.
+As explained before and shown in \coderef{code:background-plus-operator-object-example}, \mintinline{text}{ToPrimitive()} may return a \mintinline{text}{string}, even when it is called with no \mintinline{text}{preferredType}.
+This can happen because:
+\begin{enumerate}
+	\item \mintinline{text}{valueOf()} is returning a String.
+	\item \mintinline{text}{valueOf()} is not returning a primitive, which would make \mintinline{text}{toString()} being called instead.
+	\item One of the operands is the native JavaScript \mintinline{text}{Date} Object.
+\end{enumerate}
+
+\begin{code}
+	\jscode{code/background/javascript/type-coercion/plus-operator-implementation.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[JavaScript + operator implementation]{\textbf{JavaScript \mintinline{text}{+} operator implementation}}
+	\label{code:background-plus-operator-implementation}
+\end{code}
+
+\begin{code}
+	\jscode{code/background/javascript/type-coercion/plus-operator-simple-examples.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[JavaScript + operator examples]{\textbf{JavaScript \mintinline{text}{+} operator examples}}
+	\label{code:background-plus-operator-simple-examples}
+\end{code}
+
+\begin{code}
+	\jscode{code/background/javascript/type-coercion/plus-operator-object.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[JavaScript + operator with object]{\textbf{JavaScript \mintinline{text}{+} operator with object}}
+	\label{code:background-plus-operator-object-example}
+\end{code}
+
+\subsection{NPM Packages} \label{sec:background-npm-packages}
+% \todo{Agregar info de NPM by numbers! Cuantos downloads, cuantos packages, etc.}
+% \todo{https://snyk.io/blog/npm-passes-the-1-millionth-package-milestone-what-can-we-learn/}
+
+% \todo{Agregar info de cómo se resuelve el modulo y dónde se busca, etc -> Mirar link Requiring modules in Node.js: Everything you need to know}
+
+\section{TypeScript} \label{sec:background-typescript}
+\subsection{Declaration Files} \label{sec:declaration-files-background}
+% \todo{Hablar de los tipos de declaration files que hay}
+\subsection{Definitely Typed}
+
+\section{Jalangi} \label{sec:jalangi}

### 2019-08-19 code/background/javascript/type-coercion/boolean-conversion.js

@@ -1,0 +1,14 @@
+// FALSE:
+Boolean(undefined); // false
+Boolean(null); // false
+Boolean(0); // false
+Boolean(NaN); // false
+Boolean(""); // false
+
+// TRUE:
+Boolean(-123.45); // true
+Boolean(" "); // true
+Boolean(" a "); // true
+Boolean([1, 2, 3]); // true
+Boolean(new Date()); // true
+Boolean({}); // true

### 2019-08-19 code/background/javascript/type-coercion/explicit-type-coercion.js

@@ -1,0 +1,10 @@
+let someNumber = 1234;
+let thisIsAString = String(someNumber); // "1234"
+
+let n = 0;
+if (Boolean(n)) {
+	// unreachable code ...
+}
+
+let s = "hello world";
+let r = Number(s) + 1; // NaN

### 2019-08-19 code/background/javascript/type-coercion/implicit-type-coercion.js

@@ -1,0 +1,9 @@
+'' == '0' // false
+0 == '' // true
+0 == '0' // true
+false == 'false' // false
+false == '0' // true
+false == undefined // false
+false == null // false
+null == undefined // true
+"  \t\r\n   " == 0 // true

### 2019-08-19 code/background/javascript/type-coercion/number-conversion.js

@@ -1,0 +1,11 @@
+Number(undefined); // NaN
+
+Number(null); // 0
+
+Number(true); // 1
+Number(false); // 0
+
+Number("123"); // 123
+Number("\n\n     -123.45     \t\t"); // -123.45
+
+Number("123abc"); // NaN

### 2019-08-19 code/background/javascript/type-coercion/plus-operator-implementation.js

@@ -1,0 +1,15 @@
+function AdditionOperator() {
+	this.apply = function (a, b) {
+		let aPrim = ToPrimitive(a); // No preferredType is chosen
+		let bPrim = ToPrimitive(b); // No preferredType is chosen
+
+		if (
+			typeof aPrim === "string" ||
+			typeof bPrim === "string"
+		) {
+			return ToString(aPrim).concat(ToString(bPrim));
+		}
+
+		return ToNumber(a) + ToNumber(b);
+	}
+}

### 2019-08-19 code/background/javascript/type-coercion/plus-operator-object.js

@@ -1,0 +1,17 @@
+let a = {};
+a.valueOf = function () {
+	return "1";
+};
+
+let r = a + 1; // "11"
+
+a = {};
+a.toString = function() {
+	return "2";
+}
+
+r = a + 1 // "21"
+
+let myDate = new Date();
+r = myDate + 123;
+// "Mon mm DD 2019 HH:MM:SS GMT+0200 (Central European Summer Time)123"

### 2019-08-19 code/background/javascript/type-coercion/plus-operator-simple-examples.js

@@ -1,0 +1,19 @@
+1 + 1 // 2
+1 + true // 2
+1 + false // 1
+true + false // 1
+1 + undefined // NaN
+1 + null // 1
+null + undefined // NaN
+"Hello" + "World" // "HelloWorld"
+1 + "1" // "11"
+"1" + true // "1true"
+"1" + undefined // "1undefined"
+"1" + null // "1null"
+
+let a = {};
+a.valueOf = function () {
+	return 1;
+};
+
+let r = a + 1; // 2

### 2019-08-19 code/background/javascript/type-coercion/string-conversion.js

@@ -1,0 +1,14 @@
+String(undefined); // 'undefined'
+
+String(null); // 'null'
+
+String(123); // '123'
+String(-123.45); // '-123.45'
+
+String(NaN); // 'NaN'
+
+String(true); // 'true'
+String(false); // 'false'
+
+String([1, 2, 3]) // '1,2,3'
+String(["hello", "world"]) // 'hello,world'

### 2019-08-19 code/background/javascript/type-coercion/to-number.js

@@ -1,0 +1,12 @@
+function ToNumber(a) {
+	if (isPrimitive(a) === true) {
+		// Native JS Number conversion if
+		// a is a primitive.
+		return Number(a);
+	}
+
+	if (typeof a === "object") {
+		let primValue = ToPrimitive(a, Number);
+		return ToNumber(primValue);
+	}
+}

### 2019-08-19 code/background/javascript/type-coercion/to-primitive/normal-object-to-number.js

@@ -1,0 +1,6 @@
+let a = {};
+a.valueOf = function () {
+	return 12345;
+}
+
+let n = Number(a); // 12345

### 2019-08-19 code/background/javascript/type-coercion/to-primitive/normal-object-to-string.js

@@ -1,0 +1,5 @@
+let a = {};
+a.toString = function () {
+	return "I am a string!";
+}
+let s = String(a); // 'I am a string!'

### 2019-08-19 code/background/javascript/type-coercion/to-primitive/object-to-number-returning-not-a-number.js

@@ -1,0 +1,21 @@
+let a = {};
+a.valueOf = function() {
+	// valueOf() returning undefined, i.e. not a Number.
+	return undefined;
+}
+
+let n = Number(a);
+// NaN. ToPrimitive(a, Number) returns undefined,
+// and isNaN(Number(undefined)) === true.
+
+a.valueOf = function() {
+	// valueOf() is not returning a primitive.
+	// ToPrimitive(a, Number) ends up calling a.toString() instead.
+	return {};
+}
+
+a.toString = function() { return "123"; }
+// toString() returns a String that can
+// be converted into a Number.
+
+n = Number(a); // 123

### 2019-08-19 code/background/javascript/type-coercion/to-primitive/object-to-string-returning-not-a-string.js

@@ -1,0 +1,19 @@
+let a = {};
+a.toString = function () {
+	// toString() returning a Boolean, i.e. not a String.
+	return true;
+}
+
+let s = String(a);
+// 'true'
+// ToPrimitive(a, String) returns the boolean value true.
+
+a.toString = function () {
+	// toString() is not returning a primitive.
+	// ToPrimitive(a, String) ends up calling a.valueOf() instead.
+	return {};
+}
+
+a.valueOf = function () { return undefined; }
+
+s = String(a); // 'undefined'

### 2019-08-19 code/background/javascript/type-coercion/to-primitive/to-primitive.js

@@ -1,0 +1,52 @@
+function ToPrimitive(input, preferredType) {
+	if (typeof input !== "object" || input === null) {
+		return input;
+	}
+
+	if (preferredType === Number) {
+		return primitiveToNumber(input);
+	}
+
+	if (preferredType === String) {
+		return primitiveToString(input);
+	}
+
+	if (input instanceof Date) {
+		return primitiveToString(input);
+	} else {
+		return primitiveToNumber(input);
+	}
+}
+
+function primitiveToString(value) {
+	let toString = value.toString();
+
+	if (isPrimitive(toString)) {
+		return toString;
+	}
+
+	let valueOf = value.valueOf();
+	if (isPrimitive(valueOf)) {
+		return valueOf;
+	}
+
+	throw new TypeError("Cannot convert object to primitive value");
+}
+
+function primitiveToNumber(value) {
+	let valueOf = value.valueOf();
+	if (isPrimitive(valueOf)) {
+		return valueOf;
+	}
+
+	let toString = value.toString();
+	if (isPrimitive(toString)) {
+		return toString;
+	}
+
+	throw new TypeError("Cannot convert object to primitive value");
+}
+
+function isPrimitive(value) {
+	return (value !== Object(value));
+}

### 2019-08-19 code/background/javascript/type-coercion/to-string.js

@@ -1,0 +1,12 @@
+function ToString(a) {
+	if (isPrimitive(a) === true) {
+		// Native JS String conversion if
+		// a is a primitive.
+		return String(a);
+	}
+
+	if (typeof a === "object") {
+		let primValue = ToPrimitive(a, String);
+		return ToString(primValue);
+	}
+}

### 2019-08-19 chapters/3-background.tex
Add background for NPM and Javascript

@@
 \section{JavaScript} \label{sec:background-javascript}
-% \todo{Agregar img con data de la survey de stackoverflow indicando la popularidad de JS. Agarrar data de encuestas viejas y hacer una comparación de agunas techs, tipo JS, PHP, Python, Java, etc.}
+JavaScript is the standard programming language for developing client-side web applications.
+In addition to HTML and CSS, they conform the key technology stack of the Web.
+It is compliant with the ECMAScript Language Specification \citep{ecma-script}.
+
+JavaScript is an interpreted, dynamically typed language that allows to program small scripts very fast.
+Its low learning curve, the possibility of fast prototyping and its easy interaction with the browser's DOM made it the ideal programming languages for introducing dynamic behavior to web applications.
+
+All browsers have built-in interpreters that support running JavaScript code.
+The language is also used to develop backend applications that run in server side runtime environments like NodeJS \citep{nodejs}.
+
+\figref{fig:background-survey-programming-languages} shows that it is the most popular programming languages among professional developers.
+Additionally, 5 of the 12 most popular frameworks used for web developing are JavaScript frameworks.
+Similarly, it can be seen in \figref{fig:background-programming-languages-evolution} that JavaScript's popularity remained constant in the last 3 years while being the most popular programming language for seven consecutive years.
 
@@
 \input{figures/background/stackoverflowSurvey-ProgrammingLanguagesEvolution.tex}
-
-% \todo{Agregar tema serverless. Que NodeJS tiene el 75\% de los deploys}
 
@@
 \subsection{NPM Packages} \label{sec:background-npm-packages}
-% \todo{Agregar info de NPM by numbers! Cuantos downloads, cuantos packages, etc.}
-% \todo{https://snyk.io/blog/npm-passes-the-1-millionth-package-milestone-what-can-we-learn/}
+The Node Package Manager (NPM) is the default package manager for NodeJS, a JavaScript runtime environment built on Google Chrome's V8 JavaScript engine \citep{nodejs}.
+It offers a public registry where JavaScript libraries can be uploaded to be used by the developer's community.
 
-% \todo{Agregar info de cómo se resuelve el modulo y dónde se busca, etc -> Mirar link Requiring modules in Node.js: Everything you need to know}
+Dependencies can be installed either globally or as local dependencies of specific a JavaScript projects.
+NPM installs the dependencies described in the \mintinline{text}{package.json} file and local dependencies are saved in the \mintinline{text}{node_modules} directory in the project's root directory.
+An example of NPM's usage is shown in \coderef{code:background-npm-usage-example}.
+The registry contains currently more than 1 million modules and it has 11000 million downloads per week.
+
+\begin{code}
+	\begin{bashinline}
+$ cat index.js 
+const abs = require('abs');
+console.log(abs('/foo'));
+
+$ node index.js
+Error: Cannot find module 'abs'
+
+$ npm i abs
+$ node index.js
+/foo
+	  \end{bashinline}
+	\caption[NPM usage example]{\textbf{NPM usage example} - Requiring the module before installing the dependencies will fail since the library is not available.
+After installing the dependencies locally under the \mintinline{text}{node_modules} directory, the \mintinline{text}{abs} module can be imported using the \mintinline{text}{require} function.}
+	\label{code:background-npm-usage-example}
+  \end{code}
 

### 2019-08-20 chapters/3-background.tex
Add examples and types description to background

@@
 \section{JavaScript} \label{sec:background-javascript}
-JavaScript is the standard programming language for developing client-side web applications.
-In addition to HTML and CSS, they conform the key technology stack of the Web.
-It is compliant with the ECMAScript Language Specification \citep{ecma-script}.
+JavaScript is the standard programming language for developing client-side web applications. Together with HTML and CSS, they conform the key technology stack of the Web.
 
-JavaScript is an interpreted, dynamically typed language that allows to program small scripts very fast.
+JavaScript is compliant with the ECMAScript Language Specification \citep{ecma-script}.
+It is an interpreted, dynamically typed language that allows to program small scripts very fast.
 Its low learning curve, the possibility of fast prototyping and its easy interaction with the browser's DOM made it the ideal programming languages for introducing dynamic behavior to web applications.
+
+Numerous JavaScript frameworks that aid the development of web applications have been created over the years.
+\mintinline{text}{jQuery} is used by 74.2\% of the 10 million most popular websites, making it the most popular JavaScript framework in the market \citep{jquery}\citep{w3techs-javascript-libraries-statistics}.
+Other frameworks like \mintinline{text}{AngularJS}, \mintinline{text}{React} and \mintinline{text}{Vue.js} appeared as an elegant alternative for building complex and scalable web applications \citep{angularjs}\citep{react}\citep{vuejs}.
+A very simple example of how to use \mintinline{text}{jQuery} for adding dynamic behavior to a web page is shown in \coderef{code:background-jquery-example}.
+
+\begin{code}
+	\htmlcode{code/background/javascript/jquery-example.html}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[Example using JavaScript and HTML]{\textbf{Example using JavaScript and HTML} - Very simple example that uses library jQuery to hide an element from the DOM when it gets clicked.
+The example was extracted from W3Schools website.}
+	\label{code:background-jquery-example}
+\end{code}
 
@@
 \input{figures/background/stackoverflowSurvey-ProgrammingLanguagesEvolution.tex}
+
+\subsection{Types}
+Section 8 of the ECMAScript Language Specification presents the allowed types for a variable \citep{ecma-script}.
+These are: \mintinline{text}{undefined}, \mintinline{text}{null}, \mintinline{text}{boolean}, \mintinline{text}{string}, \mintinline{text}{number}, \mintinline{text}{object} and \mintinline{text}{symbol} for ES6/ECMAScript 2015.
+All of them except for \mintinline{text}{object} are called primitives.
+
+It must be noted that functions and arrays are not built-in types.
+Functions are considered to be `member of the Object type that is an instance of the standard built-in Function constructor and that may be invoked as a subroutine' \citep{ecma-script}.
+The \mintinline{text}{Function} object is a built-in object with a \mintinline{text}{[[Call]]} internal property, which means that it can be invoked.
+Similarly, arrays are instances of the built-in \mintinline{text}{Array} object.
+Both concepts are shown in \coderef{code:background-functions-and-arrays}.
+
+\begin{code}
+	\jscode{code/background/javascript/functions-and-arrays.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[Functions and arrays are built-in objects in JS]{\textbf{Functions and arrays are built-in objects in JS} - A function can be created using the \mintinline{text}{function} keyword or using the \mintinline{text}{Function} built-in constructor.
+Similarly, arrays can be created using brackets or the \mintinline{text}{Array} built-in object.}
+	\label{code:background-functions-and-arrays}
+\end{code}
+
+Finally, the \mintinline{text}{typeof} operator is not an exact match with the built-in types.
+It surprisingly returns \mintinline{text}{'object'} for \mintinline{text}{null}, \mintinline{text}{'function'} for a \mintinline{text}{function} and \mintinline{text}{'object'} for an \mintinline{text}{array}.
+\coderef{code:background-javascript-typeof} gives some examples about this operator.
+
+\begin{code}
+	\jscode{code/background/javascript/javascript-typeof.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[typeof JavaScript operator]{\textbf{\mintinline{text}{typeof} JavaScript operator} - Examples of the \mintinline{text}{typeof} operator for \mintinline{text}{null}, \mintinline{text}{function} and \mintinline{text}{array}.}
+	\label{code:background-javascript-typeof}
+\end{code}
 

### 2019-08-20 chapters/4-approach.tex
Add examples and types description to background

@@
 In spite of the fact that JavaScript was originally intended for simple web-pages scripting, it's popularity has increased over the last years, being the most commonly used programming language among professional developers in 2018.
-Development of robust and high-quality frameworks such as jQuery\footnote{https://jquery.com}, AngularJS\footnote{https://angularjs.org}, React\footnote{https://reactjs.org} or Vue.js\footnote{https://vuejs.org} contributed to JavaScript being the leading programming language for client-side development.
+Development of robust and high-quality frameworks such as \mintinline{text}{jQuery}, \mintinline{text}{AngularJS}, \mintinline{text}{React} or \mintinline{text}{Vue.js} contributed to JavaScript being the leading programming language for client-side development \citep{jquery}\citep{angularjs}\citep{react}\citep{vuejs}.
 

### 2019-08-20 code/background/javascript/functions-and-arrays.js

@@ -1,0 +1,26 @@
+function f(a, b) {
+	console.log("I am a function being called with: " + a + " and " + b);
+}
+
+f(100, 200); // 'I am a function being called with: 100 and 200'
+f instanceof Function; // true
+f instanceof Object; // true
+
+let g = new Function(
+	"a",
+	"console.log('This is the built-in Function Object being called with: ' + a)"
+)
+
+g(456); // 'This is the built-in Function Object being called with: 456'
+g instanceof Function; // true
+g instanceof Object; // true
+
+let a = [1, 2, 3];
+a.push(4); // [1, 2, 3, 4];
+a instanceof Array; // true
+a instanceof Object; // true
+
+let b = new Array(1, 2, 3);
+b.push(4); // [1, 2, 3, 4];
+b instanceof Array; // true
+b instanceof Object; // true

### 2019-08-20 code/background/javascript/javascript-typeof.js

@@ -1,0 +1,13 @@
+let x = null;
+typeof x; // 'object'
+x instanceof Object; // false
+
+let f = function() {};
+typeof f; // 'function'
+f instanceof Function; // true
+f instanceof Object; // true
+
+let a = [1, 2, 3];
+typeof a; // 'object'
+a instanceof Array; // true
+a instanceof Object; // true

### 2019-08-20 chapters/3-background.tex
Add 'Wrapper objects' section to background

@@
 	\label{code:background-javascript-typeof}
+\end{code}
+
+\subsection{Wrapper Objects for Primitives}
+JavaScript also includes the built-in objects \mintinline{text}{Boolean}, \mintinline{text}{String} and \mintinline{text}{Number}.
+Whenever a method or a property of a primitive value of type \mintinline{text}{boolean}, \mintinline{text}{string} and \mintinline{text}{number} is accessed, JavaScript will convert the primitive value into its corresponding built-in object before accessing the property or invoking the method, as illustrated in \coderef{code:background-javascript-wrapper-objects}.
+It is however considered a bad practice to explicitly use the built-in objects instead of the primitive values.
+Several JavaScript linters like JSHint will consider this an error \citep{jshint-wrapper-objects-error}.
+
+\begin{code}
+	\jscode{code/background/javascript/javascript-wrapper-objects.js}
+	\captionsetup{aboveskip=0pt, belowskip=10pt}
+	\caption[Wrapper Objects for primitives]{\textbf{Wrapper Objects for primitives} - When a method is invoked on a primitive value like \mintinline{text}{number} or \mintinline{text}{string}, an intermediate wrapper object is created.}
+	\label{code:background-javascript-wrapper-objects}
 \end{code}

### 2019-08-20 code/background/javascript/javascript-wrapper-objects.js

@@ -1,0 +1,18 @@
+let n = 123;
+typeof n === 'number'; // true
+
+/**
+ * The value of n is converted into the built-in Number object
+ * with 'new Number(n)'. The method 'toExponential()' is invoked
+ * on that hidden object.
+ */
+n.toExponential(3); // '1.230e+2
+
+let s = "hello world";
+typeof s === 'string'; // true
+/**
+ * The value of s is converted into the built-in Number object
+ * with 'new String(s)'. The method 'toUpperCase()' is invoked
+ * on that hidden object.
+ */
+s.toUpperCase(); // 'HELLO WORLD'

### 2019-08-20 chapters/3-background.tex
Add declaration files section

@@
 \section{JavaScript} \label{sec:background-javascript}
-JavaScript is the standard programming language for developing client-side web applications. Together with HTML and CSS, they conform the key technology stack of the Web.
+JavaScript (abbreviated JS) is the standard programming language for developing client-side web applications. Together with HTML and CSS, they conform the key technology stack of the Web.
 
@@
 \subsection{Declaration Files} \label{sec:declaration-files-background}
-% \todo{Hablar de los tipos de declaration files que hay}
+Declaration files contain a typed description of an external JavaScript library's API.
+It allows the IDE to use code intelligence features like auto-completion and avoids misusing the library by declaring specific interfaces and types required by the library.
+They are necessary for integrating new TypeScript projects with existing JavaScript libraries that are still developed in plain JavaScript.
+
+Declaration files have the extension \mintinline{text}{.d.ts}.
+\figref{fig:background-declaration-files-calculator-example} provides an example where a new TypeScript program is using the declaration file of an external JavaScript library called \mintinline{text}{calculator}.
+The example  library \mintinline{text}{calculator} is not part of the TypeScript project, it is imported in run-time after compilation when the JavaScript code is executed.
+
+\input{figures/background/typescript/declaration-files/calculator-example.tex}
+
+It is important to mention that the compiler and the IDE will blindly trust the declaration file.
+There are no run-time checks being performed on the declaration file and the JavaScript code itself.
+TypeScript is not aware if the JavaScript library corresponding to the declaration file even exists.
+\figref{fig:background-declaration-files-calculator-error-example} exposes a scenario where a declaration file that does not match the JavaScript implementation.
+The declaration file states that types \mintinline{text}{number} and \mintinline{text}{string} are allowed for \mintinline{text}{sum} method but the JavaScript code explicitly throws an error when the arguments are not of type \mintinline{text}{number}.
+The IDE, however, will provide suggestions based on the inaccurate declaration file.
+The compiler itself will use the declaration file to perform type checking.
+Therefore, the code is compiled without errors and the error is only encountered at run-time.
+
+\input{figures/background/typescript/declaration-files/calculator-example-error.tex}
+
+It is clear that inaccurate declaration files will cause errors that are tedious and difficult to track and debug.
+Discrepancies with the actual JavaScript implementation is going to frustrate developers, stopping them from using TypeScript as an alternative to JavaScript.
+Hence the importance of accurate declaration files that match the JavaScript implementation.
+
+There are two possibilities for importing declaration files of NPM packages:
+\begin{itemize}
+	\item If the JavaScript library is written in TypeScript, then the declaration file will be generated automatically by the compiler using the \mintinline{text}{--declaration} flag.
+The generated declaration file is bundled with the JavaScript distribution files and uploaded to the NPM registry as a part of the package.
+The \mintinline{text}{types} field in the \mintinline{text}{package.json} file should point to the declaration file.
+Consequently, the declaration file and the JavaScript code will always be synchronized.
+	\item If the library is not written in TypeScript, the declaration files have to be manually created and uploaded to the DefinitelyTyped repository \citep{definitely-typed-repository}.
+After a successful pull request, files will be automatically published to the \mintinline{text}{@types} organization on NPM \citep{types-organization-npm}.
+\end{itemize}
+
 \subsection{Definitely Typed}
+The declaration files for NPM packages that do not already include them in the package are uploaded to the DefinitelyTyped repository \citep{definitely-typed-repository}.
+It currently contains declaration files for more than 6000 modules.
+The developers community adds new declaration files or updates existing ones by creating a pull request with the changes.
+If it is approved, the changes will be automatically pushed to the NPM registry through Microsoft's \mintinline{text}{types-publisher} tool \citep{types-publisher}.
+
+\subsubsection{Consumption}
+As shown in \coderef{code:background-definitely-typed-consumption}, the declaration files can be simply imported by running \mintinline{text}{npm install --save-dev @types/THE-MODULE}.
+NPM will install the files under the \mintinline{text}{node_modules/@types} directory and the compiler will include the declaration files automatically.
+
+\begin{code}
+	\begin{bashinline}
+$ npm install --save-dev @types/abs 
+$ cat node_modules/@types/abs/index.d.ts
+
+// Type definitions for abs 1.3
+// Project: https://github.com/ionicabizau/abs
+// Definitions by: Aya Morisawa <https://github.com/AyaMorisawa>
+// Definitions: https://github.com/DefinitelyTyped/DefinitelyTyped
+
+/**
+ * Compute the absolute path of an input.
+ * @param input The input path.
+ */
+declare function Abs(input: string): string;
+export = Abs;
+	  \end{bashinline}
+	\caption[DefinitelyTyped files consumption]{\textbf{DefinitelyTyped files consumption} - The declaration files are installed as a common NPM package under the \mintinline{text}{@types} directory.}
+	\label{code:background-definitely-typed-consumption}
+  \end{code}
 

### 2019-08-20 chapters/4-approach.tex
Add declaration files section

@@
 Therefore, several industrial developers started using TypeScript in JavaScript projects.
-However, despite TypeScript's popularity, many new and existing JavaScript Libraries are still being written in pure JavaScript.
+However, despite TypeScript's popularity, many new and existing JavaScript libraries are still being written in pure JavaScript.
 
@@
 TypeScript enables it through \textit{declaration files}, a typed description of the library's API.
-A public repository, DefinitelyTyped, contains declaration files for more than 6000 JS Libraries \citep{definitely-typed-repository}.
+A public repository, DefinitelyTyped, contains declaration files for more than 6000 JS libraries \citep{definitely-typed-repository}.
 71\% of the top 1000 depended-upon packages have their corresponding declaration file in DefinitelyTyped, as explained in \chapref{chap:background}.
@@
 Discrepancies between the declaration files and the actual JavaScript Library code have a negative impact on developers.
-Type checking messages and IDE code intelligence features are not accurate, thus increasing developing costs, generating actually more difficulties than before migrating to TypeScript.
+Type checking messages and IDE code intelligence features are not accurate, which increases developing costs, generating actually more difficulties than before migrating to TypeScript.
 
@@
 
-Following the goal of easing and automating the generation of TypeScript Declaration Files for existing JavaScript Libraries, the method receives the name of the JavaScript Library published to the NPM Registry as an input argument and returns the corresponding TypeScript declaration file.
+Following the goal of easing and automating the generation of TypeScript Declaration Files for existing JavaScript libraries, the method receives the name of the JavaScript Library published to the NPM Registry as an input argument and returns the corresponding TypeScript declaration file.
 
@@
 \subsection{Implementation}
-As explained before, the DefinitelyTyped Repository is the official repository for finding declaration files for existing JavaScript Libraries.
+As explained before, the DefinitelyTyped Repository is the official repository for finding declaration files for existing JavaScript libraries.
 The \textit{TypeScript Declaration Files Generation Method} is intended to be used on existing, published npm packages.
@@
 
-Jalangi fails to instrument some JavaScript Libraries.
+Jalangi fails to instrument some JavaScript libraries.
 No effort was made in trying to fix failing instrumentation.
-Failing JS Libraries were omitted from the analysis by providing a list with the names of the modules that should not be instrumented.
+Failing JS libraries were omitted from the analysis by providing a list with the names of the modules that should not be instrumented.
 

### 2019-08-20 figures/background/typescript/declaration-files/calculator-example-error.tex
Add declaration files section

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.48\textwidth}
			\tscode{code/background/typescript/declaration-files/calculator-error-index.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[calculator/index.d.ts]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.48\textwidth}
			\jscode{code/background/typescript/declaration-files/calculator-index.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[calculator/index.js]{\usebox{\mintedbox}}

	\subfloat[Visual Studio Code auto-completion]{\includegraphics[width=0.8\textwidth]{figures/background/typescript/declaration-files/ide-auto-completion-error.png}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.58\textwidth}
			\jscode{code/background/typescript/declaration-files/calculator-error-index.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[index.ts]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.4\textwidth}
			\begin{bashinline}
$ tsc index.ts
$ node index.js
Error: Only numbers are allowed!
			\end{bashinline}
		\end{minipage}
	\end{lrbox}
	\subfloat[Console output]{\usebox{\mintedbox}}

	\caption[TypeScript declaration file with errors]{\textbf{TypeScript declaration file with errors} - An error is explicitly introduced by writing in the declaration file that the \mintinline{text}{sum} method accepts also a \mintinline{text}{string}. IDE and compiler blindly trust the declaration file. An error occurs only at run-time.}
	\label{fig:background-declaration-files-calculator-error-example}
\end{figure}

### 2019-08-20 figures/background/typescript/declaration-files/calculator-example.tex
Add declaration files section

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.48\textwidth}
			\tscode{code/background/typescript/declaration-files/calculator-index.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[calculator/index.d.ts]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.48\textwidth}
			\jscode{code/background/typescript/declaration-files/calculator-index.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[calculator/index.js]{\usebox{\mintedbox}}

	\subfloat[Visual Studio Code auto-completion]{\includegraphics[width=0.7\textwidth]{figures/background/typescript/declaration-files/ide-auto-completion.png}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.58\textwidth}
			\jscode{code/background/typescript/declaration-files/index.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[index.ts]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.4\textwidth}
			\begin{bashinline}
$ tsc index.ts
$ node index.js
Computing [(1+2) * (3+4)] ...
The result is: 21
			\end{bashinline}
		\end{minipage}
	\end{lrbox}
	\subfloat[Console output]{\usebox{\mintedbox}}

	\caption[TypeScript declaration files]{\textbf{TypeScript declaration files} - A TypeScript project using the declaration file of an external JavaScript library \mintinline{text}{calculator}. IDE auto-completion and compiler make use of the declaration file.}
	\label{fig:background-declaration-files-calculator-example}
\end{figure}

### 2019-08-20 code/background/typescript/declaration-files/calculator-error-index.d.ts

@@ -1,0 +1,13 @@
+export = Calculator;
+
+declare class Calculator {
+	constructor();
+	sum(
+		a: number | string,
+		b: number | string): number;
+
+	multiply(a: number, b: number): number;
+}
+
+declare namespace Calculator {
+}

### 2019-08-20 code/background/typescript/declaration-files/calculator-error-index.ts

@@ -1,0 +1,9 @@
+import Calculator from './calculator';
+
+let c = new Calculator();
+
+console.log("Computing [(1+2) * (3+4)] ...")
+
+let r = c.multiply(c.sum("1", "2"), c.sum(3, 4));
+
+console.log("The result is: " + r);

### 2019-08-20 code/background/typescript/declaration-files/calculator-index.d.ts

@@ -1,0 +1,10 @@
+export = Calculator;
+
+declare class Calculator {
+	constructor();
+	sum(a: number, b: number): number;
+	multiply(a: number, b: number): number;
+}
+
+declare namespace Calculator {
+}

### 2019-08-20 code/background/typescript/declaration-files/calculator-index.js

@@ -1,0 +1,16 @@
+module.exports = function () {
+	this.sum = function (a, b) {
+		if (
+			typeof a !== 'number' ||
+			typeof b !== 'number'
+		) {
+			throw new Error("Only numbers are allowed!");
+		}
+
+		return a + b;
+	};
+
+	this.multiply = function (a, b) {
+		return a * b;
+	};
+};

### 2019-08-20 code/background/typescript/declaration-files/index.ts

@@ -1,0 +1,9 @@
+import Calculator from './calculator';
+
+let c = new Calculator();
+
+console.log("Computing [(1+2) * (3+4)] ...")
+
+let r = c.multiply(c.sum(1, 2), c.sum(3, 4));
+
+console.log("The result is: " + r);

### 2019-08-20 chapters/3-background.tex
Complete Jalangi section

@@
 
+\todo{Agregar templates!}
+
 \subsection{Definitely Typed}
@@
 \section{Jalangi} \label{sec:jalangi}
+Jalangi is a framework for performing dynamic analysis in JavaScript with configurable modules \citep{DBLP:conf/sigsoft/SenKBG13}.
+It first instruments the code and then executes the instrumented code adding the analysis module.
+Jalangi offers different callbacks that match virtually every JavaScript operations.
+The relevant callbacks for this work are listed below with short descriptions that have been directly copied from Jalangi's documentation \citep{jalangi-docs}:
+\begin{itemize}
+	\item \textbf{binaryPre()}: `This callback is called before a binary operation.'
+	\item \textbf{declare()}: `This callback is triggered at the beginning of a scope for every local variable declared in the scope, for every formal parameter, for every function defined using a function statement, for arguments variable, and for the formal parameter passed in a catch statement.'
+	\item \textbf{getField()}: `This callback is called after a property of an object is accessed.'
+	\item \textbf{putFieldPre()}: `This callback is called before a property of an object is written.'
+	\item \textbf{functionEnter()}: `This callback is called before the execution of a function body starts.'
+	\item \textbf{functionExit()}: `This callback is called when the execution of a function body completes'
+	\item \textbf{invokeFun()}: `This callback is called after a function, method, or constructor invocation.'
+	\item \textbf{invokeFunPre()}: `This callback is called before a function, method, or constructor invocation.'
+	\item \textbf{unaryPre()}: `This callback is called before a unary operation.'
+	\item \textbf{write()}: `This callback is called before a variable is written.'
+  \end{itemize}
+
+\figref{fig:background-jalangi} provides an example for the usage of Jalangi where invocations of a function are detected and printed by stdout.
+
+\input{figures/background/typescript/jalangi/jalangi-example.tex}

### 2019-08-20 figures/background/typescript/jalangi/jalangi-example.tex
Complete Jalangi section

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.34\textwidth}
			\jscode{code/background/jalangi/sum.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[sum.js]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.64\textwidth}
			\jscode{code/background/jalangi/analysis.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[analysis.js]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\begin{bashinline}
$ node jalangi.js --analysis analysis.js sum.js
Function sum was invoked with params (1,2) and returned 3
Function sum was invoked with params (3,4) and returned 7
Function sum was invoked with params (5,6) and returned 11
			\end{bashinline}
		\end{minipage}
	\end{lrbox}
	\subfloat[Console output]{\usebox{\mintedbox}}

	\caption[Jalangi usage example]{\textbf{Jalangi usage example} - \mintinline{text}{sum.js} file is instrumented and executed alongside the analysis module defined in \mintinline{text}{analysis.js}.}
	\label{fig:background-jalangi}
\end{figure}

### 2019-08-20 code/background/jalangi/analysis.js

@@ -1,0 +1,15 @@
+(function (sandbox) {
+	function MyAnalysis() {
+		this.invokeFun = function (iid, f, base, args, result) {
+			console.log(
+				"Function " + f.name + " " +
+				"was invoked with params (" +
+				Object.values(args).join(',') +
+				") and returned " +
+				result
+			);
+		};
+	}
+
+	sandbox.analysis = new MyAnalysis();
+})(J$);

### 2019-08-20 code/background/jalangi/sum.js

@@ -1,0 +1,7 @@
+function sum(a, b) {
+	return a + b;
+}
+
+sum(1, 2);
+sum(3, 4);
+sum(5, 6);

### 2019-08-21 chapters/3-background.tex
Add declaration files templates

@@
 
-\todo{Agregar templates!}
+\subsubsection{Templates}
+TypeScript provides different templates for writing declaration files \citep{typescript-declaration-files-templates}. The relevant templates are the following:
+\begin{itemize}
+	\item \mintinline{text}{module.d.ts}: It is used when the exported module is a JavaScript object containing several properties and methods.
+
+	\item \mintinline{text}{module-class.d.ts}: This template is used when the exported function is supposed to be used as a constructor with the \mintinline{text}{new} operator.
+
+	\item \mintinline{text}{module-function.d.ts}: It is used when the module exports only a function.
+  \end{itemize}
 

### 2019-08-21 chapters/4-approach.tex
Add declaration files templates

@@
 \begin{enumerate}
-  \item \mintinline{text}{module.d.ts}: It is used when the exported module is a JavaScript object containing several properties and methods.
-It is the most common used template: more than 70\% of the modules uploaded to DefinitelyTyped are based on this template.
-The code checks if the exported module is neither a function nor a constructor, i.e.
+  \item \mintinline{text}{module.d.ts}: The code checks if the exported module is neither a function nor a constructor, i.e.
 if properties \mintinline{text}{isExported} and \mintinline{text}{isConstructor} are \mintinline{text}{false}.
@@
 
-  \item \mintinline{text}{module-class.d.ts}: As shown in \figref{fig:ts-declaration-file-module-class-template}, this template is used when the exported function is supposed to be used as a constructor with the \mintinline{text}{new} operator.
-The code will use this template when both the fields \mintinline{text}{isExported} and \mintinline{text}{isConstructor} are true.
+  \item \mintinline{text}{module-class.d.ts}: The code will use this template when both the fields \mintinline{text}{isExported} and \mintinline{text}{isConstructor} are true.
 It should be noted that the module exports only the constructor function.
@@
 
-  \item \mintinline{text}{module-function.d.ts}: It is used when the module exports only a function.
-The field \mintinline{text}{isExported} field is used to determine whether this templates should be used or not.
+  \item \mintinline{text}{module-function.d.ts}: The field \mintinline{text}{isExported} field is used to determine whether this templates should be used or not.
 An example is provided in \figref{fig:ts-declaration-file-module-function-template}.

### 2019-08-22 chapters/3-background.tex
Add future work

@@
 \section{JavaScript} \label{sec:background-javascript}
-JavaScript (abbreviated JS) is the standard programming language for developing client-side web applications. Together with HTML and CSS, they conform the key technology stack of the Web.
+JavaScript (abbreviated JS) is a programming language that is compliant with the ECMAScript Language Specification \citep{ecma-script}. It is one of the
+key technologies for developing client-side web applications, alongside HTML and CSS.
 
-JavaScript is compliant with the ECMAScript Language Specification \citep{ecma-script}.
 It is an interpreted, dynamically typed language that allows to program small scripts very fast.
-Its low learning curve, the possibility of fast prototyping and its easy interaction with the browser's DOM made it the ideal programming languages for introducing dynamic behavior to web applications.
+Its low learning curve, the possibility of fast prototyping and its simple interaction with the contents of a web page through the DOM (Document Object Model) made it the ideal programming languages for introducing dynamic behavior to web applications.
 
@@
 	\caption[Example using JavaScript and HTML]{\textbf{Example using JavaScript and HTML} - Very simple example that uses library jQuery to hide an element from the DOM when it gets clicked.
-The example was extracted from W3Schools website.}
+The example was extracted from W3Schools website \citep{w3schools}.}
 	\label{code:background-jquery-example}
@@
 Additionally, 5 of the 12 most popular frameworks used for web developing are JavaScript frameworks.
-Similarly, it can be seen in \figref{fig:background-programming-languages-evolution} that JavaScript's popularity remained constant in the last 3 years while being the most popular programming language for seven consecutive years.
+Similarly, it can be seen in \figref{fig:background-programming-languages-evolution} that JavaScript's popularity remained constant in the last 3 years and it remained as the most popular programming language for seven consecutive years.
 
@@
 It must be noted that functions and arrays are not built-in types.
-Functions are considered to be `member of the Object type that is an instance of the standard built-in Function constructor and that may be invoked as a subroutine' \citep{ecma-script}.
+A Function is considered to be `member of the Object type that is an instance of the standard built-in Function constructor and that may be invoked as a subroutine' \citep{ecma-script}.
 The \mintinline{text}{Function} object is a built-in object with a \mintinline{text}{[[Call]]} internal property, which means that it can be invoked.
@@
 JavaScript also includes the built-in objects \mintinline{text}{Boolean}, \mintinline{text}{String} and \mintinline{text}{Number}.
-Whenever a method or a property of a primitive value of type \mintinline{text}{boolean}, \mintinline{text}{string} and \mintinline{text}{number} is accessed, JavaScript will convert the primitive value into its corresponding built-in object before accessing the property or invoking the method, as illustrated in \coderef{code:background-javascript-wrapper-objects}.
+Whenever a method or a property of a primitive value of type \mintinline{text}{boolean}, \mintinline{text}{string} or \mintinline{text}{number} is accessed, JavaScript will convert the primitive value into its corresponding built-in object before accessing the property or invoking the method, as illustrated in \coderef{code:background-javascript-wrapper-objects}.
 It is however considered a bad practice to explicitly use the built-in objects instead of the primitive values.
@@
 The specific definition can be found in the ECMAScript Language Specification \citep{ecma-script}.
-The following section will cover the basic forms of type coercion but providing a detailed explanation of how JavaScript converts a variable type into another one for each operator is beyond the scope of this section.
-It is however intended to give insight about this kind of behavior to show its importance while performing type inference in JavaScript.
-As a result, specific operator related type coercion is explained only for operator \mintinline{text}{+}.
+The following section will cover the basics about JavaScript's type coercion.
+A detailed explanation of how JavaScript converts a variable type into another one for each operator can be seen in the ECMAScript Language Specification \citep{ecma-script}.
+It is however intended to give insight about this kind of behavior to highlight its importance for performing type inference in JavaScript.
+Therefore, specific operator related type coercion is explained only for operator \mintinline{text}{+}.
 
-Finally, there are excellent references regarding JavaScript's Type Coercion, like Kyle Simpson's book `You Don't Know JS: Types \& Grammar' \citep{you-dont-know-js}.
+There are excellent references regarding JavaScript's Type Coercion, like Kyle Simpson's book `You Don't Know JS: Types \& Grammar' \citep{you-dont-know-js}.
 
@@
 Explicit type coercion is simply when the developer chooses to convert one type into another one.
-The developer achieves this by writing a piece of code that shows her intention, as shown in \coderef{code:background-explicit-type-coercion}.
+The developer achieves this by writing a piece of code that explicitly shows her intention, as shown in \coderef{code:background-explicit-type-coercion}.
 This type of coercion is also called type casting \citep{you-dont-know-js}.
@@
 	\captionsetup{aboveskip=0pt, belowskip=10pt}
-	\caption[Implicit JavaScript Type Coercion]{\textbf{Implicit JavaScript Type Coercion} - Examples given by Douglas Crockford in his talk `JavaScript: The Good Parts' at Google\footnote{https://www.youtube.com/watch?v=hQVTIJBZook}.}
+	\caption[Implicit JavaScript Type Coercion]{\textbf{Implicit JavaScript Type Coercion} - Examples given by Douglas Crockford in his talk `JavaScript: The Good Parts' at Google \citep{js-the-good-parts}.}
 	\label{code:background-implicit-type-coercion}
@@
 However, each operator will call either \mintinline{text}{ToString()}, \mintinline{text}{ToNumber()} or \mintinline{text}{ToBoolean()} after calling \mintinline{text}{ToPrimitive()}.
-This operator will be explained in an independent section.
+This operation will be explained in an independent section.
 
@@
 	\captionsetup{aboveskip=0pt, belowskip=10pt}
-	\caption[ToString operation]{\textbf{ToString operation}}
+	\caption[ToString operation examples]{\textbf{ToString operation examples}}
 	\label{code:background-to-string-operation}
@@
 Strings and boolean values get converted in the expected way.
-Moreover, \mintinline{text}{undefined} will be converted into \mintinline{text}{NaN} and \mintinline{text}{null} into \mintinline{text}{0}.
+Curiously, \mintinline{text}{undefined} will be converted into \mintinline{text}{NaN} and \mintinline{text}{null} into \mintinline{text}{0}.
 Similarly, objects are converted into numbers by calling \mintinline{text}{ToPrimtive()} with \mintinline{text}{hint = number}, as presented in \coderef{code:background-to-number-implementation}.
@@
 	\captionsetup{aboveskip=0pt, belowskip=10pt}
-	\caption[ToNumber operation]{\textbf{ToNumber operation}}
+	\caption[ToNumber operation examples]{\textbf{ToNumber operation examples} - Every non-printable character is removed from the string before converting it into a number.}
 	\label{code:background-to-number-operation}
@@
 	\captionsetup{aboveskip=0pt, belowskip=10pt}
-	\caption[ToBoolean operation]{\textbf{ToBoolean operation}}
+	\caption[ToBoolean operation examples]{\textbf{ToBoolean operation examples} - Values that are coerced to false are called \textit{falsy} values.}
 	\label{code:background-to-boolean-operation}
@@
 An important aspect that should be considered, is that the type of the return value of \mintinline{text}{ToPrimitive()} does not necessarily have to match the chosen \mintinline{text}{preferredType}.
-This means concretely that, for example, you can call \mintinline{text}{ToPrimitive(a, number)} and get a \mintinline{text}{string} as a result.
+This means concretely that, for example, \mintinline{text}{ToPrimitive(a, number)} can be called and return a \mintinline{text}{string} as a result.
 Of course, when performing an explicit type conversion, JS will eventually convert whatever \mintinline{text}{ToPrimitive} returns into the correct type.
@@
 \section{TypeScript} \label{sec:background-typescript}
+TypeScript is a programming language developed by Microsoft that compiles to plain JavaScript.
+Its syntax is a superset of JavaScript, which means in practice that `every JavaScript program is also a TypeScript program' \citep{typescript}.
+It provides optional types annotations that enable features like auto-completion and static checking.
+Type annotations are erased and variable names are preserved after compiling, which produces a JavaScript output that is very similar to the TypeScript source code.
+The extension for TypeScript files is \mintinline{text}{.ts}.
+An example about some of the basic TypeScript features is provided in \coderef{code:background-typescript-and-javascript}.
+
+A complete specification of the language can be found in the TypeScript Language Specification document \citep{typescript}. This section will only cover the most relevant aspects of the language.
+
+\begin{code}
+	\begin{bashinline}
+$ cat index.js 
+function f(a, b) {
+	return a + b;
+}
+
+f(1, 2);
+
+$ npm i -g typescript
+$ mv index.js index.ts
+$ tsc index.ts
+$ cat index.js
+function f(a, b) {
+    return a + b;
+}
+f(1, 2);
+
+$ cat index-with-type-annotations.ts
+function f(a: number, b: number) : number {
+	return a + b;
+}
+
+f(1, 2);
+
+$ tsc index-with-type-annotations.ts
+$ cat index-with-type-annotations.js
+function f(a, b) {
+    return a + b;
+}
+f(1, 2);
+	  \end{bashinline}
+	\caption[TypeScript compilation example]{\textbf{TypeScript compilation example} - A plain JavaScript file is renamed into a TypeScript file.
+The TypeScript compiler will compile it into plain JavaScript again.
+Secondly, a \mintinline{text}{.ts} file with typed annotations gets compiled into a common JavaScript file with no annotations.}
+	\label{code:background-typescript-and-javascript}
+  \end{code}
+
+\subsection{Types}
+The type analysis performed by TypeScript will only happen at compile-time, which means that the compiled JavaScript has no additional overhead at run-time.
+
+\subsubsection{Any Type}
+The \mintinline{text}{any} type closely resembles a common value in JavaScript.
+It has been introduced to support values with no type annotations.
+It can be used to represent any JavaScript value.
+Minimal static checking is performed on values of the type \mintinline{text}{any}: properties with any name can be accessed, methods with any name and any argument list can be invoked and the value can be invoked as a function.
+
+\subsubsection{Primitive Types}
+As stated in the specification, `The primitive types are the \mintinline{text}{Number}, \mintinline{text}{Boolean}, \mintinline{text}{String}, \mintinline{text}{Symbol}, \mintinline{text}{Void}, \mintinline{text}{Null} and \mintinline{text}{Undefined} types and all user defined enum types.' \citep{typescript}.
+
+The types \mintinline{text}{Number}, \mintinline{text}{Boolean}, \mintinline{text}{String}, \mintinline{text}{Symbol}, \mintinline{text}{Null} and \mintinline{text}{Undefined} closely match the JavaScript equivalent types.
+The \mintinline{text}{Void} type stands for the absence of a value, for a example when a function returns no value.
+
+\subsubsection{Object Types}
+The following types are considered to be object types: Class and interface type references, array types, tuple types, function types, and constructor types.
+
+Array types can be written as \mintinline{text}{string[]} or as \mintinline{text}{Array<string>}.
+
+The signature of a function can be represented with the \mintinline{text}{Function} type.
+It is useful for the very common case where a callback is an argument of another function.
+For example the following function has two arguments, a \mintinline{text}{string} and a callback that receives a \mintinline{text}{number} and returns a \mintinline{text}{boolean}: \mintinline{text}{function f(a: string, c: (x: number) => boolean) { // ...
+}}.
+
+Interfaces can be used to parametrize object types by defining the types and signatures of properties and methods of a value.
+It is important to mention that interfaces are not exported in any way to the compiled JavaScript.
+They are only useful for static type checking at compile-time and code intelligence features provided by the IDE.
+
+Classes are implemented in the expected way according to the usual Object Oriented programming languages.
+\figref{fig:background-typescript-class-example} shows an example of a TypeScript class implementation and its corresponding JavaScript file after compilation.
+
+\input{figures/background/typescript/classes/class-example.tex}
+
+\subsubsection{Union \& Intersection Types}
+A Union Type represents a value that may have only one of multiple possible types.
+It is represented as \mintinline{text}{let a: string | number}, meaning that \mintinline{text}{a} is either of type \mintinline{text}{string} or \mintinline{text}{number}.
+
+On the other hand, the Intersection Type stands for a value that can simultaneously two types. It can be useful for defining a value that implements two interfaces simultaneously.
+
 \subsection{Declaration Files} \label{sec:declaration-files-background}

### 2019-08-22 chapters/6-conclusions.tex
Add future work

@@
 
-Building and end-to-end solution for the generation of TypeScript declaration files was prioritized over type inference accuracy.
+Building an end-to-end solution for the generation of TypeScript declaration files was prioritized over type inference accuracy.
 Consequently, types were taken over from the values at run-time.
-Obtaining the types from the code examples extracted from the repositories proved to be a pragmatic and effective approximation, enabling to work on specific aspects regarding the TypeScript declaration file generation.
+Since developers expose through code how a library should be used, obtaining the types from the code examples extracted from the repositories proved to be a pragmatic and effective approximation, enabling to work on specific aspects regarding the TypeScript declaration file generation itself.
 
-We built a mechanism to automatically create declaration files for every module uploaded to DefinitelyTyped and evaluate the results, creating a TypeScript declaration files parser and a comparator.
+We built a mechanism to automatically create declaration files for potentially every module uploaded to DefinitelyTyped.
+We managed to generate declaration files for 244 modules.
+We compared the results against the corresponding files uploaded to DefinitelyTyped by creating a TypeScript declaration files parser and a comparator.
 
@@
 Accordingly, we collected evidence regarding the usage of JavaScript operators by analyzing 400 libraries.
-Furthermore, used in a first place merely as an initial code base for executing the library and approximating types, code examples turned out to be extremely useful for inferring types.
-They expose the developer's intention explicitly through code.
 
@@
 \section{Future Work} \label{sec:conclusions-future-work}
+\subsection{Instrumentation}
+The latest commit to Jalangi's Github repository was on April 5, 2017 \citep{jalangi-repository}.
+The are 49 open issues but the last comment on an issue was on January 31, 2018.
+Jalangi plays a key role in the process of generating declaration files.
+In order to increase the number of modules that are able to be instrumented shown in \figref{fig:experiments-overall-funnel}, it is necessary to perform some changes to Jalangi's code.
+Since it is a library that seems to be no longer maintained, the repository might need to be forked.
 
-\subsection{Instrumentation}
-\draft{Babel}
+On of the main drawbacks of Jalangi is that is does not offer full support of ES6, as stated in their repository: `Jalangi2 supports ECMAScript 5.1.
+Some ES6 features may work, but have not been tested.' \citep{jalangi-repository}.
+A possible solution for this would be to use Babel to convert ES6 code into a backwards compatible version like ECMAScript 5.1 before instrumenting the code \citep{babel}.
 
-\subsection{Declaration files generation}
-\draft{Optional parameters}
-\draft{Functions as paramters}
-\subsubsection{Uploading Results to Definitely Typed Repository}
+\subsection{Optional parameters}
+The concept of optional parameters was not implemented.
+The difference between an optional parameter and a parameter of type \mintinline{text}{undefined} has to be further studied.
+In practice, JavaScript does return \mintinline{text}{undefined} for both the cases of accessing a nonexistent property or a property with the value \mintinline{text}{undefined}.
+Therefore, the developer's intention of treating a parameter as optional has to be clearly captured.
+
+\subsection{Function types}
+TypeScript enables the developer to clearly specify the signature of a function that is a parameter of an interface or of another function.
+This is considered a good practice and it greatly enriches the declaration file.
+The current implementation assigns to a parameter that is a function the type \mintinline{text}{Function}, which is naturally very unspecific.
+The gathered run-time information is already sufficient for creating the signatures of function types.
+The code that uses the run-time information for clearly representing the signatures needs to be developed.
+
+\subsection{Uploading Results to DefinitelyTyped}
+Automatic creation of pull requests on the DefinitelyTyped repository could be attached to the pipeline.
+Compliance with the contribution guidelines specified by DefinitelyTyped can be checked before generating the pull requests.
 
 \subsection{Type Inference}
-% \todo{Deberia ser hecho en otro modulo que no sea el runtime information. El runtime information deberia exponer unicamente los datos que le hacen falta al otro para inferir los tipos.}
+Inspecting the semantics of the code can be of great use for inferring types.
+The name that a developer chooses for a variable or a function is very important to determine the purpose of that variable.
+A function that performs the operation \mintinline{text}{firstArgument + secondArgument} has a complete different meaning if its called \mintinline{text}{sum()} or \mintinline{text}{concatenate()}, even though the JavaScript operator is the same in both cases.
 
-% \todo{Tambien se puede identificar patterns de programacion donde siempre los tipos son los mismos!!! El concepto de inferir el tipo en realidad pasa a detectar la intencion del developer.}
+Furthermore, the content of a value can also be useful.
+Some strings may represent, for example, known error codes or RGB colors representations.
+This information can be used by a type inferrer to determine the type more accurately.
 
-\subsubsection{Return values}
-% \todo{El return value se podria incluir en el analysis de Jalangi, detectando las interactions que tiene etc etc.
-Esto no es tan facil porque no se puede simplemente inspeccionar el return value, sino que hay que saber para qué se usa.
-Los ejemplos podrían ser útiles}.
-
-\subsubsection{Variable Names}
-% \todo{Se puede extraer info del nombre de las variables y del nombre de los metodos, etc. La developer's intention.}
-
-\subsubsection{Variable Content}
-
-\subsubsection{Function Names}
-
-\draft{JAVERT!!}
+Finally, detecting common development patterns is important.
+Developers tend to apply the same patterns for solving similar problems.
+For example, a very common pattern that developers use for dealing with optional parameters is \mintinline{text}{let a = firstArgument || {}}.
+Identifying such patterns and relating them with commonly used types can be extremely useful for a better capture of the developer's intention.

### 2019-08-22 figures/background/typescript/classes/class-example.tex
Add future work

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.9\textwidth}
			\tscode{code/background/typescript/classes/index.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[index.ts]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.9\textwidth}
			\jscode{code/background/typescript/classes/index.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[index.js]{\usebox{\mintedbox}}

	\caption[TypeScript classes compilation]{\textbf{TypeScript classes compilation} - A TypeScript class is compiled into the common ES5 construct used for emulating classes in JavaScript. Private and public methods are only checked at compile-time. Both methods are exported in the same way, since JavaScript has no concept of method's visibility.}
	\label{fig:background-typescript-class-example}
\end{figure}

### 2019-08-22 code/background/javascript/type-coercion/boolean-conversion.js

@@ -9,6 +9,6 @@ Boolean(""); // false
 Boolean(-123.45); // true
 Boolean(" "); // true
 Boolean(" a "); // true
-Boolean([1, 2, 3]); // true
+Boolean([]); // true
 Boolean(new Date()); // true
 Boolean({}); // true

### 2019-08-22 code/background/typescript/classes/index.js

@@ -1,0 +1,14 @@
+var Greeter = /** @class */ (function () {
+	function Greeter(name) {
+		this.name = name;
+	}
+	Greeter.prototype.sayHello = function () {
+		console.log("Hello " + this.name + "!");
+	};
+	Greeter.prototype.sayHelloPrivate = function () {
+		console.log("Hello " + this.name + "(private) !");
+	};
+	return Greeter;
+}());
+var g = new Greeter("John");
+g.sayHello();

### 2019-08-22 code/background/typescript/classes/index.ts

@@ -1,0 +1,18 @@
+class Greeter {
+	name: string;
+
+	constructor(name: string) {
+		this.name = name
+	}
+
+	public sayHello() {
+		console.log("Hello " + this.name + "!");
+	}
+
+	private sayHelloPrivate() {
+		console.log("Hello " + this.name + "(private) !");
+	}
+}
+
+let g = new Greeter("John");
+g.sayHello();

### 2019-08-22 chapters/55-related.tex
Finish experiments section and conclusion
Add related work

@@
 \chapter{Related Work}\label{chap:relatedwork}
-Give a brief overview of the work relevant for your thesis.
-Closely related work should be compared in depth: what are the pros and cons of their work compared to yours?
 
-\todo{dts-gen package: https://github.com/microsoft/dts-gen}
+\subsubsection{Microsoft's dts-gen}
+Microsoft developed \mintinline{text}{dts-gen}, a tool that creates starter declaration files for JavaScript libraries \citep{dts-gen}.
+Its documentation states that the result is however intended to be only used a starting point.
+The outcome needs to be refined afterwards by the developers.
+
+They analyze the shape of the objects at runtime after initialization without executing the library.
+This results in many variables being inferred as \mintinline{text}{any}.
+\coderef{code:related-work-dts-gen-example} shows an example for module \mintinline{text}{abs}.
+
+The solution presented in this work, however, is intended to generate declaration files that are ready to be uploaded to DefinitelyTyped without further manual intervention.
+Any amount of manual work that a developer needs to do on a declaration file after updating JavaScript code increases the risk for having discrepancies between the declaration file and the implementation.
+
+Formal aspects like applying the right template and using the correct syntax are perfectly covered by \mintinline{text}{dts-gen}.
+
+\begin{code}
+	\begin{bashinline}
+$ npm i -g dts-gen
+$ npm i -g abs
+$ dts-gen -m abs
+Wrote 5 lines to abs.d.ts.
+
+$ cat abs.d.ts
+/** Declaration file generated by dts-gen */
+
+export = abs;
+
+declare function abs(input: any): any;
+	  \end{bashinline}
+	\caption[Microsoft's dts-gen example]{\textbf{Microsoft's dts-gen example} - A declaration file for module \mintinline{text}{abs} is generated.
+Types are inferred as \mintinline{text}{any}.
+The correct \mintinline{text}{module-function} template is used.}
+	\label{code:related-work-dts-gen-example}
+  \end{code}
+
+\subsubsection{TSInfer \& TSEvolve}
+TSInfer and TSEvolve are presented as part of TSTools \citep{DBLP:conf/fase/KristensenM17}.
+Both tools are the continuation of TSCheck \citep{DBLP:conf/oopsla/FeldthausM14}, a tool for looking for mismatches between a declaration file and an implementation.
+
+TSInfer proceeds in a similar way than TSCheck.
+It initializes the library in a browser and it records a snapshot of the resulting state and then it performs a light weight static analysis on all the functions and objects stored in the snapshot.
+
+The abstraction and the constraints they introduced as part of the static analysis tools for inferring the types have room for improvement.
+A run-time based approach like the one presented in our work will provide more accurate information, thus generating more precise declaration files.
+
+Since they analyze the objects and functions stored in the snapshot, they faced the problem of including in the declaration file internal methods and properties that developers wanted to hide.
+Run-time information would have informed that the developer has no intention of exposing such methods.
+
+Moreover, TSEvolve performs a differential analysis on the changes made to a JavaScript library in order to determine intentional discrepancies between declaration files of two consecutive versions.
+We consider that a differential analysis may not be needed.
+If the developer's intention is accurately extracted and the execution code clearly represents that intention then the generated declaration file would already describe the newer version of a library without the need of a differential analysis.
+
+\subsubsection{TSTest}
+TSTest is a tool that checks for mismatches between a declaration file and a JavaScript implementation \citep{DBLP:journals/pacmpl/KristensenM17}.
+It applies feedback-directed random testing for generating type test scripts.
+These scripts will execute the library in order to check if it behaves the way it is described in the declaration file.
+TSTest also provides concrete executions for mismatches.
+
+We evaluated the generated declaration files comparing them to the declaration files uploaded to DefinitelyTyped.
+The disadvantage of doing this is that since the uploaded files are written manually, they could already contain mismatches with the JavaScript implementation.
+However, it is a suitable choice for a development stage since it is used as a baseline.
+
+In a final stage, declaration files need to be checked against the proper JavaScript implementation and TSTest has to be definitely taken into account.

### 2019-08-22 chapters/0_2-abstract.tex
Add abstract

\chapter*{Abstract}
Developers are starting to write complex and large JavaScript applications in TypeScript, a typed superset of JavaScript. TypeScript applications use untyped JavaScript libraries through typed descriptions of their APIs called declaration files. These files are maintained manually, which is error prone and highly time consuming. Discrepancies between a declaration file and the JavaScript implementation will prevent developers from using them.

This work presents \mintinline{text}{dts-generate}, a tool that generates a declaration file for JavaScript libraries uploaded to the NPM registry. It executes the library, gathers run-time information and generates the declaration file based on it. 244 declaration files were generated and compared against DefinitelyTyped, the standard public repository for declaration files. 33 files out of 244 had no differences.

A study on type inference in JavaScript that leverages the developer's intention is presented along with evidence that describes the usage of JavaScript operators.

### 2019-08-22 chapters/3-background.tex
Final refinements

@@
 It must be noted that functions and arrays are not built-in types.
-A Function is considered to be `member of the Object type that is an instance of the standard built-in Function constructor and that may be invoked as a subroutine' \citep{ecma-script}.
+A Function is considered to be `member of the Object type that is an instance of the standard built-in \mintinline{text}{Function} constructor and that may be invoked as a subroutine' \citep{ecma-script}.
 The \mintinline{text}{Function} object is a built-in object with a \mintinline{text}{[[Call]]} internal property, which means that it can be invoked.
@@
 Every primitive gets converted as expected.
-Objects are converted into strings by calling \mintinline{text}{ToPrimtive()} with \mintinline{text}{hint = string}, as presented in \coderef{code:background-to-string-implementation}.
+Objects are converted into strings by calling \mintinline{text}{ToPrimitive()} with \mintinline{text}{hint = string}, as presented in \coderef{code:background-to-string-implementation}.
 
@@
 	\captionsetup{aboveskip=0pt, belowskip=10pt}
-	\caption[ToString implementation]{\textbf{ToString implementation}}
+	\caption[ToString implementation]{\textbf{ToString implementation} - If the input argument has type \mintinline{text}{object}, it will call \mintinline{text}{ToPrimitive()} with \mintinline{text}{hint = string} and then recursively call itself again.}
 	\label{code:background-to-string-implementation}
@@
 	\captionsetup{aboveskip=0pt, belowskip=10pt}
-	\caption[ToString operation examples]{\textbf{ToString operation examples}}
+	\caption[ToString operation examples]{\textbf{ToString operation examples} - Numbers are converted into string in an expected way.
+The values \mintinline{text}{true, false, null, undefined, NaN}, however, are converted into a \mintinline{text}{string} containing the value's name.
+This is one of the main reasons why it is common to see the words 'undefined', 'null' or 'NaN' in web applications.}
 	\label{code:background-to-string-operation}
@@
 Curiously, \mintinline{text}{undefined} will be converted into \mintinline{text}{NaN} and \mintinline{text}{null} into \mintinline{text}{0}.
-Similarly, objects are converted into numbers by calling \mintinline{text}{ToPrimtive()} with \mintinline{text}{hint = number}, as presented in \coderef{code:background-to-number-implementation}.
+Similarly, objects are converted into numbers by calling \mintinline{text}{ToPrimitive()} with \mintinline{text}{hint = number}, as presented in \coderef{code:background-to-number-implementation}.
 
@@
 The obtained primitive value is then normally converted into a \mintinline{text}{string}, as explained before in \coderef{code:background-to-string-implementation}.
+\coderef{code:background-object-into-string} shows a straightforward example.
 
-\coderef{code:background-object-into-string} shows a straightforward example.
 On the other hand, \coderef{code:background-object-into-string-not-string-return-value} provides an example where \mintinline{text}{ToPrimitive(val, String)} does not return a \mintinline{text}{string}, either because simply \mintinline{text}{toString()} does not return a \mintinline{text}{string} or because \mintinline{text}{valueOf()} gets being called instead of \mintinline{text}{toString()}.
@@
 	\captionsetup{aboveskip=0pt, belowskip=10pt}
-	\caption[JavaScript + operator with object]{\textbf{JavaScript \mintinline{text}{+} operator with object}}
+	\caption[JavaScript + operator with an object]{\textbf{JavaScript \mintinline{text}{+} operator with an object}}
 	\label{code:background-plus-operator-object-example}
@@
 
-Dependencies can be installed either globally or as local dependencies of specific a JavaScript projects.
+Dependencies can be installed either globally or as local dependencies of specific JavaScript projects.
 NPM installs the dependencies described in the \mintinline{text}{package.json} file and local dependencies are saved in the \mintinline{text}{node_modules} directory in the project's root directory.
@@
 f(1, 2);
-
 $ npm i -g typescript
@@
 f(1, 2);
-
 $ cat index-with-type-annotations.ts
@@
 f(1, 2);
-
 $ tsc index-with-type-annotations.ts

### 2019-08-22 chapters/4-approach.tex
Final refinements

@@
 
-Code examples that execute the JavaScript Library are needed to extract the run-time information via code instrumentation.
+Code examples that execute the JavaScript Library are needed in order to extract the run-time information via code instrumentation.
 It is achieved by retrieving the examples provided in the README files of the repositories of the different libraries.
@@
   \jscode{code/run-time-information-gathering/analysis/get-field-changing-at-runtime.js}
-  \caption[Modified accessed property]{\textbf{Modified accessed property} - Variable \mintinline{text}{b} gets declared within the scope of function \mintinline{text}{foo}.
-Therefore, no observation on it gets stored.}
+  \caption[Modified accessed property]{\textbf{Modified accessed property} - Property \mintinline{text}{bar} gets accessed twice. Two interactions are saved.}
   \label{code:get-field-changing-at-runtime}
@@
 However, the implementation is complex and error prone.
-Instead, it was decided to modify the corresponding methods or specific operators so that they are compliant with JavaScript Type Coercion, which is explained in detail in {sec:background-js-type-coercion}.
+Instead, it was decided to modify the corresponding methods or specific operators so that they are compliant with JavaScript Type Coercion, which is explained in detail in \secref{sec:background-js-type-coercion}.
 
@@
   -v PATH_TO_RUN_TIME_INFORMATION_JSON:/tmp/output.json \
-  tsd-generator \
-  --module-name calculator \
-  -i /tmp/output.json
-
+  tsd-generator --module-name calculator -i /tmp/output.json
 $ docker cp generate-declaration-file:/usr/local/app/output/. /tmp/ts-declaration-file
@@
 Before comparison, declaration files need to be parsed so that both of them share the same structure.
-It was achieved by using the TypeScript Compiler API, a library developed by Microsoft that allows to traverse the Abstract Syntax Tree in an easy and intuitive way \citep{ts-compiler-api}.
+It was achieved by using the TypeScript Compiler API, a library developed by Microsoft that allows to traverse the Abstract Syntax Tree in an easy and intuitive way \citep{typescript-compiler-api}.
 

### 2019-08-22 code/introduction/javascript-wtfs.js

@@ -3,7 +3,7 @@ true == "1.00"; // true
 false == "    \n\r\t     "; // true
 false == []; // true
 0 == []; // true
- null == undefined; // true 4
+null == undefined; // true
 "01" < "00100"; // false
 [1] + 1; //'11'
 [2] == "2"; // true

### 2019-08-22 chapters/3-background.tex
Fix code blocks

@@
 An important aspect that should be considered, is that the type of the return value of \mintinline{text}{ToPrimitive()} does not necessarily have to match the chosen \mintinline{text}{preferredType}.
-This means concretely that, for example, \mintinline{text}{ToPrimitive(a, number)} can be called and return a \mintinline{text}{string} as a result.
+This means concretely that \mintinline{text}{ToPrimitive(a, number)} can be called and return a \mintinline{text}{string} as a result, for example.
 Of course, when performing an explicit type conversion, JS will eventually convert whatever \mintinline{text}{ToPrimitive} returns into the correct type.
@@
 
-On the other hand, \coderef{code:background-object-into-string-not-string-return-value} provides an example where \mintinline{text}{ToPrimitive(val, String)} does not return a \mintinline{text}{string}, either because simply \mintinline{text}{toString()} does not return a \mintinline{text}{string} or because \mintinline{text}{valueOf()} gets being called instead of \mintinline{text}{toString()}.
+\coderef{code:background-object-into-string-not-string-return-value} provides an example where \mintinline{text}{ToPrimitive(val, String)} does not return a \mintinline{text}{string}, either because simply \mintinline{text}{toString()} does not return a \mintinline{text}{string} or because \mintinline{text}{valueOf()} gets being called instead of \mintinline{text}{toString()}.
 

### 2019-12-18 ecoop-2020/lipics-v2019-sample-article.tex
Add report for ecoop2020

\documentclass[a4paper,english,cleveref, autoref]{lipics-v2019}
%This is a template for producing LIPIcs articles. 
%See lipics-manual.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling cleveref support, use "autoref"

\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{positioning}

\definecolor{backgroundgray}{rgb}{.94,.94,.94}
\definecolor{lightgray}{rgb}{.35,.35,.5}
\definecolor{darkgray}{rgb}{.2, .2, .2}
\definecolor{purple}{rgb}{0.65, 0.12, 0.82}

\newcommand{\secref}[1]{Section~\ref{#1} - \nameref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\coderef}[1]{Listing~\ref{#1}}

\lstdefinelanguage{JavaScript}{
  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, undefined},
  keywordstyle=\color{black}\bfseries,
  ndkeywords={class, export, boolean, throw, implements, import, this},
  ndkeywordstyle=\color{black}\bfseries,
  identifierstyle=\color{darkgray},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{lightgray}\slshape,
  stringstyle=\color{darkgray}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

\lstdefinelanguage{TypeScript}{
  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, undefined},
  keywordstyle=\color{black}\bfseries,
  ndkeywords={class, export, boolean, throw, implements, import, this, declare, constructor, namespace, interface},
  ndkeywordstyle=\color{black}\bfseries,
  identifierstyle=\color{darkgray},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{lightgray}\slshape,
  stringstyle=\color{darkgray}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

\lstset{
   backgroundcolor=\color{backgroundgray},
   extendedchars=true,
   basicstyle=\footnotesize\ttfamily,
   showstringspaces=false,
   showspaces=false,
   numbers=left,
   numberstyle=\footnotesize,
   numbersep=9pt,
   tabsize=2,
   breaklines=true,
   showtabs=false
}

\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{Generation of TypeScript Declaration Files from JavaScript Code}

% \titlerunning{Dummy short title}%optional, please use if title is longer than one line

\author{John Q. Public}{Dummy University Computing Laboratory, Country \and My second affiliation, Country \and \url{http://www.myhomepage.edu} }{johnqpublic@dummyuni.org}{https://orcid.org/0000-0002-1825-0097}{(Optional) author-specific funding acknowledgements}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional

\authorrunning{J.\,Q. Public}%TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{John Q. Public}%TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[300]{Software and its engineering~Software notations and tools}
\ccsdesc[100]{Software and its engineering~General programming languages}%TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{JavaScript, TypeScript, Dynamic Analysis, Declaration Files}%TODO mandatory; please add comma-separated list of keywords

\category{}%optional, e.g. invited paper

\relatedversion{}%optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversion{A full version of the paper is available at \url{...}.}

\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

% \acknowledgements{I want to thank \dots}%optional

%\nolinenumbers %uncomment to disable line numbering

%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
Developers are starting to write complex and large JavaScript applications in TypeScript, a typed superset of JavaScript. TypeScript applications integrate JavaScript libraries by including typed descriptions of their APIs called declaration files. These files are maintained manually, which is error prone and highly time consuming. Discrepancies between a declaration file and the JavaScript implementation will prevent developers from using them.

This work presents \lstinline[language={}]!dts-generate!, a tool that generates a declaration file for JavaScript libraries uploaded to the NPM registry. It executes the library, gathers run-time information from the code examples written by the developers and generates the declaration file based on it. 244 declaration files were generated and compared against DefinitelyTyped, the standard public repository for declaration files. 33 files out of 244 had no differences.
\end{abstract}

\section{Introduction}
JavaScript has become the most popular language for writing web applications \cite{github-statistics}. It is also gaining popularity for back-end applications running in NodeJS. Its dynamic typing speeds up programming enabling developers to create simple pieces of code very fast, making JavaScript a very appealing programming language.

JavaScript is being used for creating complex and large applications. However, JavaScript was not intended to be more than a scripting language. Maintaining and evolving large JavaScript codebases is notably challenging. Mistakes such as mistyped property names and misunderstood or unexpected type coercion cause developers to spend a significant amount of time in debugging sessions. A JavaScript code blog\footnote{https://wtfjs.com} compiles experiences from developers facing unexpected situations while programming in JavaScript. \coderef{code:introduction-javascript-wtfs} exposes some of these unintuitive JavaScript behaviors.

\begin{lstlisting}[
    caption={\textbf{Unexpected JavaScript behavior} - falsy values, \lstinline!typeof!, \lstinline!null! and \lstinline!undefined! operators and type coercion.},
    language=JavaScript,
	label=code:introduction-javascript-wtfs,
    float,
    abovecaptionskip=-\medskipamount
]
"0" == false; // true
true == "1.00"; // true
false == "    \n\r\t     "; // true
false == []; // true
0 == []; // true
null == undefined; // true
"01" < "00100"; // false
[1] + 1; //'11'
[2] == "2"; // true
null + undefined + [1, 2, 3] // 'NaN1,2,3'
"hello world".lenth + 1 // NaN | note 'lenth' instead of 'length'
[1, 15, 20, 100].sort() // [ 1, 100, 15, 20 ]
typeof null; // object
null instanceof Object; // false
\end{lstlisting}

The overhead produced by such dynamic typing is not present in other languages that use build tools based on type information. The situation motivated the creation of TypeScript, a superset of JavaScript with typed annotations \cite{typescript}. It has become a widely used alternative among JavaScript developers, since it incorporates features that are helpful for developing and maintaining large applications \cite{DBLP:conf/icse/GaoBB17}. TypeScript enables the early detection of run-time errors detection and the integration of code intelligence tools like auto-completion in the IDEs.

Existing JavaScript libraries can be used in a TypeScript project by adding a declaration file that contains a typed description of the library's API. Declaration files are stored in a repository called DefinitelyTyped that contains declaration files for more than 6000 JavaScript libraries \cite{definitely-typed-repository}. Unfortunately, declaration files need to be manually created and maintained, which is error prone and time consuming. TypeScript does not perform any run-time check on these declaration files. A discrepancy between the declaration file and its corresponding JavaScript library would lead to additional frustration and debugging sessions, since type checks and code-intelligence features would be inaccurate.

Some previous work tackled the problem of automatically searching for mismatches between a declaration file and implementation code \cite{DBLP:conf/oopsla/FeldthausM14}. TSTest adapted the feedback-directed random testing technique, mainly used for testing Java libraries, to detect discrepancies between a declaration file and a JavaScript library \cite{DBLP:journals/pacmpl/KristensenM17}. Tools like TSInfer and TSEvolve are designed for assisting the creation of new declaration files and supporting the evolution the declaration file when the corresponding JavaScript library gets modified \cite{DBLP:conf/fase/KristensenM17}, respectively. TypeScript itself developed \lstinline[language={}]{dts-gen}, a tool that generates a declaration file that is meant to be used only as a `starting point for writing a high-quality declaration file' \cite{dts-gen}.

We explore in this work the possibilities for improving the existing tools. We present the tool \lstinline[language={}]{dts-generate}. It is based on an architecture that supports the generation of declaration files for an existing JavaScript library published to the NPM registry. The tool will gather data flow and type information at run-time and generate a declaration file based on that information.

The architecture supports the future incorporation of a Symbolic Execution Engine that expands the initial code base using the signatures in the declaration file. The iterative process of exploring new execution paths will refine the generated declaration file in each iteration.

Finally, we generated declaration files for 244 JavaScript libraries and evaluated them against DefinitelyTyped.

The contributions of this paper are as follows.
\begin{itemize}
    \item We introduce an architecture that supports generating TypeScript Declaration Files for a given JavaScript Library using run-time information. The architecture supports the future incorporation of a Symbolic Execution Engine that expands the initial code base using the signatures in the declaration file. The iterative process of exploring new execution paths will refine the generated declaration file in each iteration.

    \item We present the tool \lstinline{dts-generate}, a command line application that generates a valid TypeScript Declaration File for a specific NPM package using run-time information.

    \item Next, we deliver a test framework that supports generating a TypeScript Declaration File using \lstinline{dts-generate} for all JavaScript libraries in the DefinitelyTyped Repository. We present the results of running \lstinline{dts-generate} for 244 of these libraries.

    \item Finally, we present a TypeScript Declaration Files parser and comparator. Both tools were necessary for evaluating the results against DefinitelyTyped.
\end{itemize}

\section{Motivating Examples}

\section{dts-generate - Generation of TypeScript Declaration Files}
We introduce \lstinline{dts-generate}, a command line tool which generates a valid TypeScript Declaration File for a specific JavaScript Library uploaded to the NPM Registry, as explained in \figref{fig:tsd_generation_method_block_diagram}. The tool is intended to be used on existing, published \lstinline{npm} packages. The generated output TypeScript declaration file is a valid file which can be used for development and uploaded to the DefinitelyTyped Repository.

\begin{figure}[tp]
    \centering
    \includegraphics[width=1\linewidth]{dts-generate-block-diagram.pdf}
    \caption[dts-generate - Architecture overview]{\textbf{dts-generate - Architecture overview} - Initial code base is retrieved from the npm package's repository. A valid TypeScript Declaration File is generated using run-time information. A Symbolic Execution Engine creates test cases based on the generated Declaration File and via a feedback loop enriches the code base until the stopping criteria is reached. The final TypeScript Declaration File gets returned. Feedback loop through the Symbolic Execution Engine was not implemented. It can be added in a future to the existing architecture, without modifying the working blocks.}
    \label{fig:tsd_generation_method_block_diagram}
  \end{figure}

Code examples that execute the JavaScript Library are needed in order to extract the run-time information via code instrumentation. It is achieved by retrieving the examples provided in the README files of the repositories of the different libraries. This is generally the place where developers explicitly show how to use their code. It showed to be an appropriate and pragmatic way of extracting the developer's intention and providing an useful initial code base with meaningful examples, thus avoiding a possible cold start problem.

The examples and the code base of the library are instrumented with Jalangi \cite{DBLP:conf/sigsoft/SenKBG13}\cite{DBLP:conf/sigsoft/SenKBG13a} to gather data flow information and type information at runtime. Jalangi is a dynamic analysis configurable framework that provides several analysis modules that were extended as needed to retrieve the required run-time information, which is then saved to an output JSON file.

A second independent block uses the run-time information to generate a TypeScript Declaration File. It infers the overall structure of the JS Library, the interfaces and the types from the JSON file.

The declaration file returned by the method is valid and fully functional, making it suitable for being used within the development process. It contains no errors and matches the structure of the JavaScript Library under analysis, so that the JavaScript code generated after compiling the TypeScript code runs without errors. The conducted experiments included tests that consisted on replacing a specific type definition from DefinitelyTyped \cite{definitely-typed-repository} with the one generated in the experiments: TypeScript compilation was successful, the generated JavaScript code ran without errors and code intelligence features performed by IDEs like code completion worked as expected.

The command line interface was inspired in the \lstinline{dts-gen} \cite{dts-gen} package. It can be seen in \coderef{code:dts-generate-example} that invoking the package is very simple and the only required argument is the name of the module published to the npm registry.

\begin{lstlisting}[
    caption={\textbf{dts-generate usage} - Example of how to generate a declaration file for module \lstinline{abs}.},
    language=bash,
	label=code:dts-generate-example,
    float,
    abovecaptionskip=-\medskipamount
]
$ ./dts-generate abs
$ cat output/abs/index.d.ts 
export = Abs;

declare function Abs(input: string): string;
\end{lstlisting}

\subsection{Initial Code Base}
To extract run-time information of a JavaScript Library, it is necessary, by definition, to actually execute the code, since the analysis modules provided by Jalangi to gather information are only triggered if the instrumented code gets executed.

It was decided to extract the code examples that execute the JavaScript Library from the Readme files of the repository associated to the NPM Package. Readme files are usually used by developers for briefly describing what the code does, what problem it solves, how to install the application, how to build the code, etc. It is very common that developers provide code examples in the readme files to show how the code works and how to use it. This is specially true for NPM Packages, which are in general created to solve a specific problem of JavaScript development.

Obtaining code examples for a specific NPM Package is achieved in three steps:
\begin{itemize}
  \item Obtain the repository url from the package: The command \lstinline{npm view <PACKAGE> repository.url} can be used for retrieving the url of the package's repository.

  \item Retrieve the README file from the repository.

  \item Extract the code examples from the README file: Readme files are written using Markdown\footnote{https://www.markdownguide.org}, a very common lightweight and simple markup language. It is very common to write code examples within code blocks indicating the programming language, so that it gets highlighted with the specific syntax. The Markdown identifiers for JavaScript are \lstinline{js} or \lstinline{javascript}. The code examples are finally retrieved by filtering the content within the corresponding code blocks.
\end{itemize}

\subsection{Run-time Information Gathering}
The Runtime Information block described in \figref{fig:tsd_generation_method_block_diagram} will gather information such as:

\begin{itemize}
  \item Function \lstinline{f} got invoked with parameters \lstinline{a} and \lstinline{b} with types \lstinline{string} and \lstinline{number}.
  \item Property or method \lstinline{foo} of parameter \lstinline{a} of function \lstinline{f} was accessed within the function.
  \item Parameter \lstinline{a} of function \lstinline{f} was used as operand for operator \lstinline{==}.
\end{itemize}

The dynamic analysis framework used for gathering this kind of information is Jalangi. The configurable analysis modules enable programming custom callbacks that get triggered with virtually any JavaScript event. The events that are observed are:
\begin{itemize}
  \item Binary operations, like \lstinline{==}, \lstinline{+} or \lstinline{===}.
  \item Variable declaration.
  \item Function, method, or constructor invocation.
  \item Access to an object's property.
  \item Unary operations, like \lstinline{!} or \lstinline{typeof}.
\end{itemize}

The implementation stores these observations as entities called \lstinline{interactions}. They are used for translating, modifying and aggregating Jalangi's raw event information in order to get an application specific data representation. The run-time information is finally returned as a JSON file that can be used for later processing. The tool is written in JavaScript and runs in Node.js within a Docker container.

\subsection{TypeScript Declaration File Generation}
\paragraph*{Overwiew}
The actual generation of the declaration file is the next step in the pipeline after gathering the run-time information, as shown in \figref{fig:tsd_generation_method_block_diagram}. It is a lightweight, simple and fast application, which does not interact with the actual JavaScript module at run-time. Instead, it uses the JSON output file containing the run-time information and generates a TypeScript declaration file which is use ready to be used within a TypeScript project. The tool itself is written in TypeScript and runs within a Docker container in NodeJS.

\paragraph*{Templates}
TypeScript provides templates for writing declaration files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/templates.html} and each template corresponds to a different way of exporting a JavaScript module. The tool uses different fields from the run-time information to detect how the module is being used in order to choose the right templates accordingly. The implemented templates are \lstinline{module}, \lstinline{module-class} and \lstinline{module-function}.

\paragraph*{Interfaces}
Finally, interfaces are created by exploring \lstinline{getField} and \lstinline{methodCall} interactions from the run-time information. The code will gather the interactions for a specific argument and build the interface by incrementally adding new properties. Interactions within the \lstinline{followingInteractions} field are recursively traversed, building a new interface in each recursion level.

\subsection{Evaluation}
\label{sec:dts-generate-evaluation}
After generating a declaration file for a published NPM module, it is necessary to evaluate the quality of it. It was decided to compare the generated declaration file against the one uploaded to the DefinitelyTyped repository for the same module, as shown in \figref{fig:evaluation-diagram}.

\begin{figure}[tp]
    \begin{centering}
        {\includegraphics[width=1\textwidth]{evaluation-diagram.pdf}}
        \caption[Evaluation against DefinitelyTyped Repository]{\textbf{Evaluation of generated declaration files against DefinitelyTyped Repository} - A parser transforms the generated declaration file and the equivalent file in the DefinitelyTyped repository into a JSON file using the TypeScript Compiler API \cite{typescript-compiler-api}. Comparison is then performed on the JSON files, i.e. not on the declaration files.}
        \label{fig:evaluation-diagram}
    \end{centering}
\end{figure}

\paragraph*{Parsing}
Before comparison, declaration files need to be parsed so that both of them share the same structure. It was achieved by using the TypeScript Compiler API, a library developed by Microsoft that allows to traverse the Abstract Syntax Tree in an easy and intuitive way \cite{typescript-compiler-api}.

The parsing consists in creating a structure where declared \lstinline{interfaces}, \lstinline{functions}, \lstinline{classes} and \lstinline{namespaces} are stored separated. Function arguments are correctly described, identifying complex types like union types or callbacks. Optional parameters are also identified. For \lstinline{classes}, a distinction between the constructor and methods is made. Finally, syntax and semantic errors are also checked by the TypeScript Compiler API.

The tool is called \lstinline{parse-dts} and is naturally written in TypeScript. It also runs in NodeJS within its corresponding Docker container.

\paragraph*{Comparator}
An independent tool will compare two parsed declaration files. As described in \figref{fig:evaluation-diagram}, the comparator will compare the generated declaration file against the corresponding file in the DefinitelyTyped repository.

It was discovered that the implementation was easier when focusing on each template independently. For this implementation, only the \lstinline{module-function} was considered for the comparison.

The following criteria were applied:
\begin{itemize}
  \item Number of declared functions: Checks the number of declared and exported functions for each of both declaration files.
  \item Name of declared functions: Checks whether both of the declaration files declared a function with the same name.
  \item Number of parameters: Checks the number of parameters of the declared functions. This is checked for optional and non-optional parameters.
  \item Interfaces: Checks the number of declared interfaces and the fields within those interfaces.
  \item Errors: Indicates whether there are errors in the declaration files.
\end{itemize}

\section{Results}
Declaration files were generated for existing modules uploaded to the NPM registry. The DefinitelyTyped repository was used as a benchmark. Each one of the generated files was compared against the corresponding declaration file already uploaded to the repository.

Figure \ref{fig:experiments-overall-funnel} shows that a declaration file was generated for 244 modules out of 6029 modules. Samples of the generated declaration files for templates \lstinline{module}, \lstinline{module-class} and \lstinline{module-function} are presented in \secref{sec:experiments-declaration-files-generation}.

\definecolor{myellow}{RGB}{228,212,0}
\definecolor{mgreen}{RGB}{5,104,57}

\newcommand\funnel[3]{%
\pgfmathsetmacro\mwid{(0.3+\val*0.06)}
\pgfmathsetmacro\mradius{(\val*0.01 + 1)}
\pgfmathsetmacro\mheight{(\val*0.003 + 0.4)}
\pgfmathsetmacro\marc{\mwid-.4}
    \begin{scope}[%
        shift={(0,#1)}, 
        line width=.05pt, 
        %x=5mm, 
        scale=0.7,
        yshift=\xi*0.05
        ]
    \draw[black,bottom color=#2, top color=#2] (-\mwid,0) -- (-\mwid+.4,-\mheight) arc (190:350:\marc cm and \mradius mm) -- (\mwid,0);
    \draw[black,fill=#3] (0,0) ellipse (\mwid cm and \mradius mm);
    \path (-\mwid,0) -- (-\mwid+.4,-\mheight) coordinate[midway] (a\xi);
    \end{scope}
}

\begin{figure}[tp]
	\hspace*{-0.16\textwidth}
	\centering
	\begin{tikzpicture}
		\foreach \val
				[%
				count=\xi starting from 0, 
				evaluate=\xi as \shadecolor using int(25*\xi),
				evaluate=\xi as \coord using int(\xi-12)
				]
			in {
				4.05,
				7.33,
				24.02,
				37.49,
				71.19,
				82.50,
				100.00
			}{
				\funnel{\coord}{mgreen!\shadecolor !myellow}{mgreen!\shadecolor !myellow}
			}

		\node[left=0.02\textwidth of a0] {Generated Declaration Files};
		\node[right=0.04\textwidth of a0] {\textbf{244}};

		\node[left=0.02\textwidth of a1] {Run-time Information};
		\node[right=0.06\textwidth of a1] {\textbf{442}};

		\node[left=0.02\textwidth of a2] {Working Examples};
		\node[right=0.17\textwidth of a2] {\textbf{1448}};

		\node[left=0.02\textwidth of a3] {Code Examples};
		\node[right=0.25\textwidth of a3] {\textbf{2260}};

		\node[left=0.02\textwidth of a4] {README file};
		\node[right=0.45\textwidth of a4] {\textbf{4292}};

		\node[left=0.02\textwidth of a5] {Github Repository};
		\node[right=0.52\textwidth of a5] {\textbf{4974}};

		\node[align=right,left=0.02\textwidth of a6, text width=0.25\textwidth] {Definitely Typed Modules};
		\node[right=0.62\textwidth of a6] {\textbf{6029}};

	\end{tikzpicture}
	\caption[Number of analyzed modules for each stage of the experiment]{\textbf{Number of analyzed modules for each stage of the experiment} - A TypeScript Declaration File was generated for only 244 modules, out of 6029 modules in the DefinitelyTyped Repository. It was possible to gather valid run-time information for only 25\% of the modules for which a Code Example was extracted.}
	\label{fig:experiments-overall-funnel}
\end{figure}

\subsection{Code Examples}
Retrieving the code examples from the JavaScript libraries' repositories proved to be a pragmatic way of capturing the types. However, as shown in \figref{fig:experiments-overall-funnel}, working code examples for only 2260 modules could be retrieved. The process of getting a valid code example for a module is divided in 4 blocks:
\begin{itemize}
	\item Extracting repositories url.
	\item Extracting readme files.
	\item Extracting code examples within readme files.
	\item Executing code examples and discarding failing ones.
\end{itemize}

The results obtained for each on of them are described in the following sections.

\paragraph*{Repositories URL}
The url of the repositories could be retrieved for only 4974 modules. More than 1000 modules do not have the repository entry in their corresponding \lstinline{package.json} files. Therefore, the \lstinline{npm view <module> repository.url} command returns an empty value. This is even happening for important modules like \lstinline{ace}.

\paragraph*{Readme Files}
700 modules simply do not have a readme file in their repositories. The implementation does contemplate, however, different naming conventions like \lstinline{readme.md} or \lstinline{README.md}.

\paragraph*{Code Examples Extraction}
The 50\% loss is mainly explained because developers did not wrap their code around a block using the \lstinline{javascript} or \lstinline{js} tags. Counting with code examples for 2200 modules was considered to be enough for evaluating the generation of declaration files.

\paragraph*{Code Examples Execution}
The 2260 extracted code examples were executed by installing the required packages and running the code as a \lstinline{node} application. Working and functional code examples could only be extracted for 1448 modules. 812 modules did not run correctly and were discarded. Some failing samples were analyzed and there were mainly two reasons for the failure:
\begin{enumerate}
	\item The code example had been properly extracted but the code itself was not working. It was executing the library in an unsupported way, hence the error at run-time.
	\item The extracted code example was not intended to be executed or it was not even valid JavaScript code.
\end{enumerate}

\subsection{Declaration Files Generation} \label{sec:experiments-declaration-files-generation}

The following section exhibits some samples of the 244 generated declaration files. It shows some results for each of the implemented templates: \lstinline{module}, \lstinline{module-function} and \lstinline{module-class}.

\figref{fig:experiments-results-module-function} shows the generated declaration files for simple modules like \lstinline{abs}, \lstinline{dirname-regex} and \lstinline{escape-html}. All of them were generated using the \lstinline{module-function} template. There are no differences between the generated files and the corresponding declaration files uploaded to DefinitelyTyped.

\begin{figure}[tp]
    \centering
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export = Abs;

declare function Abs(input: string): string;
      \end{lstlisting}
      \caption{abs/index.d.ts - Generated}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
declare function Abs(input: string): string;
export = Abs;
      \end{lstlisting}
      \caption{abs/index.d.ts - DefinitelyTyped}
    \end{subfigure}

    \begin{subfigure}{0.48\linewidth}
        \begin{lstlisting}[language=TypeScript]
export = DirnameRegex;

declare function DirnameRegex(): RegExp;
        \end{lstlisting}
        \caption{dirname-regex/index.d.ts - Generated}
      \end{subfigure}
      \hfill
      \begin{subfigure}{0.48\linewidth}
        \begin{lstlisting}[language=TypeScript]
export = dirnameRegex;

declare function dirnameRegex(): RegExp;
        \end{lstlisting}
        \caption{dirname-regex/index.d.ts - DefinitelyTyped}
      \end{subfigure}

      \begin{subfigure}{0.48\linewidth}
        \begin{lstlisting}[language=TypeScript]
export = EscapeHtml;

declare function EscapeHtml(string: string): string;
        \end{lstlisting}
        \caption{escape-html/index.d.ts - Generated}
      \end{subfigure}
      \hfill
      \begin{subfigure}{0.48\linewidth}
        \begin{lstlisting}[language=TypeScript]
declare function escapeHTML(text: string): string;
declare namespace escapeHTML { }

export = escapeHTML;
        \end{lstlisting}
        \caption{escape-html/index.d.ts - DefinitelyTyped}
        \label{fig:experiments-results-module-function}
      \end{subfigure}

    \caption{\textbf{Module-function results} - Results are shown for modules \lstinline{abs}, \lstinline{dirname-regex}, \lstinline{escape-html}. On the left side the generated declaration file with \lstinline{dts-generate}. On the right side the corresponding file in the DefinitelyTyped repository. Functions are correctly detected and input types are accurately inferred. Both files are parsed for comparison, as explained in \secref{sec:dts-generate-evaluation}. Therefore, subtle differences in the syntax between both files are not important.}
\end{figure}

Templates of type \lstinline{module-class} are shown for modules \lstinline{flake-idgen}, \lstinline{route-parser} and \lstinline{timer-machine} in \figref{fig:experiments-results-module-class-flake-idgen}, \figref{fig:experiments-results-module-class-route-parser} and \figref{fig:experiments-results-module-class-timer-machine}, respectively. Properties of interfaces and class methods are correctly generated. Optional parameters are not detected, as it was not considered for the implementation. Finally, \lstinline{module} template is presented for \lstinline{is-uuid} module in \figref{fig:experiments-results-module-is-uuid}.

\begin{figure}[tp]
    \centering
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export = FlakeIdgen;

declare class FlakeIdgen {
    constructor(options: FlakeIdgen.I__options);
    next(cb: undefined): Buffer;
}

declare namespace FlakeIdgen {
    export interface I__options {
        'id': undefined;
        'datacenter': number;
        'worker': number;
        'epoch': undefined;
        'seqMask': undefined;
    }
}
      \end{lstlisting}
      \caption{flake-idgen/index.d.ts - Generated}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
interface ConstructorOptions {
    datacenter?: number;
    worker?: number;
    id?: number;
    epoch?: number;
    seqMask?: number;
}

declare namespace FlakeId { }

declare class FlakeId {
    constructor(options?: ConstructorOptions);
    next(callback?: (err: Error, id: Buffer) => void): Buffer;
}

export = FlakeId;        
      \end{lstlisting}
      \caption{flake-idgen/index.d.ts - DefinitelyTyped}
    \end{subfigure}

    \caption{\textbf{Module-class results | flake-idgen} - Parameters of interface \lstinline{ConstructorOptions} are correctly detected. Name of interface differs since it is automatically generated based on the name of the argument variable. Optional properties were not implemented, hence the \lstinline{undefined} type for some properties. Analogously, callback \lstinline{cb} is inferred as \lstinline{undefined}.}
    \label{fig:experiments-results-module-class-flake-idgen}
\end{figure}

\begin{figure}[tp]
    \centering
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export = Route;

declare class Route {
    constructor(spec: string);
    match(path: string): object;
    reverse(params: object): string;
}

declare namespace Route {
}
      \end{lstlisting}
      \caption{route-parser/index.d.ts - Generated}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
declare class Route {
    constructor(spec: string);
    match(pathname: string): { [i: string]: string } | false;

    reverse(params: { [i: string]: any }): string | false;
}

declare namespace Route { }
export = Route;       
      \end{lstlisting}
      \caption{route-parser/index.d.ts - DefinitelyTyped}
    \end{subfigure}

    \caption{\textbf{Module-class results | route-parser} - Constructor and methods are correctly identified. Specific signature of dictionary \lstinline{params} is not detected. Instead, a generic \lstinline{object} is inferred.}
    \label{fig:experiments-results-module-class-route-parser}
\end{figure}

\begin{figure}[tp]
    \centering
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export = Timer;

declare class Timer {
    constructor(start: undefined);
    start(): boolean;
    isStopped(): boolean;
    emit(): boolean;
    stop(): boolean;
    isStarted(): boolean;
    timeFromStart(): number;
    time(): number;
}

declare namespace Timer {
}
      \end{lstlisting}
      \caption{timer-machine/index.d.ts - Generated}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export as namespace Timer;
export = Timer;

declare namespace Timer {
    type TimerEvent = "start" | "stop" | "time";
}

declare class Timer {
    static get(reference: string): Timer;
    static destroy(reference: string): Timer;

    constructor(started?: boolean);

    isStarted(): boolean;
    isStopped(): boolean;
    start(): void;
    timeFromStart(): number;
    stop(): void;
    time(): number;
    toggle(): void;
    emitTime(): void;
    valueOf(): number;
    on(event: Timer.TimerEvent, callback?: () => void): void;
}
      \end{lstlisting}
      \caption{timer-machine/index.d.ts - DefinitelyTyped}
    \end{subfigure}

    \caption{\textbf{Module-class results | timer-machine} - Parameter \lstinline{started} is inferred as \lstinline{undefined} instead of marking it as optional. Methods that were not executed do not appear in the generated declaration file.}
    \label{fig:experiments-results-module-class-timer-machine}
\end{figure}

\begin{figure}[tp]
    \centering
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export function v1(str: string): boolean;
export function v2(str: string): boolean;
export function v3(str: string): boolean;
export function v4(str: string): boolean;
export function v5(str: string): boolean;
      \end{lstlisting}
      \caption{is-uuid/index.d.ts - Generated}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export function v1(value: string): boolean;
export function v2(value: string): boolean;
export function v3(value: string): boolean;
export function v4(value: string): boolean;
export function v5(value: string): boolean;
export function nil(value: string): boolean;
export function anyNonNil(value: string): boolean;
      \end{lstlisting}
      \caption{is-uuid/index.d.ts - DefinitelyTyped}
    \end{subfigure}

    \caption{\textbf{Module results | is-uuid} - Methods that were not executed are not included in the declaration file.}
    \label{fig:experiments-results-module-is-uuid}
\end{figure}

It is worth mentioning that for some libraries the declaration file in DefinitelyTyped was not correct. For example, for \lstinline{datadog-metrics}, some properties of an interface were included in the generated declaration file but they were not present in the one in DefinitelyTyped. However, as shown in \figref{fig:experiments-results-module-datadog-metrics}, the properties are indeed used in the source code and should be included.

\begin{figure}[tp]
    \centering
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export interface I__opts {
    'aggregator': undefined;
    'defaultTags': Array<any>;
    'reporter': undefined;
    'apiKey': string;
    'appKey': undefined;
    'agent': undefined;
    'host': string;
    'prefix': string;
    'flushIntervalSeconds': number;
}

export class BufferedMetricsLogger {
    constructor(opts: I__opts);
    // ...
}
      \end{lstlisting}
      \caption{datadog-metrics/index.d.ts - Generated}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export interface LoggerOptions {
    apiKey?: string;
    appKey?: string;
    defaultTags?: string[];
    flushIntervalSeconds?: number;
    host?: string;
    prefix?: string;
}

export class BufferedMetricsLogger {
    constructor(
        options: LoggerOptions
    );
    // ...
}
      \end{lstlisting}
      \caption{datadog-metrics/index.d.ts - DefinitelyTyped}
    \end{subfigure}

    \begin{subfigure}{0.80\linewidth}
        \begin{lstlisting}[language=TypeScript]
function BufferedMetricsLogger(opts) {
    this.aggregator = opts.aggregator || new Aggregator(opts.defaultTags);
    this.reporter = opts.reporter || new DataDogReporter(opts.apiKey, opts.appKey, opts.agent);
    this.host = opts.host;
    this.prefix = opts.prefix || '';
    this.flushIntervalSeconds = opts.flushIntervalSeconds;

    // ...
    // ...
}
        \end{lstlisting}
        \caption{datadog-metrics/logger.js}
        \label{fig:subfloat-datadog-logger}
      \end{subfigure}

    \caption{\textbf{Missing properties | datadog-metrics} - Properties \lstinline{aggregator} and \lstinline{reporter} are not in the DefinitelyTyped version, but they appear in the generated declaration file. However, they are indeed used by the library, as exposed in lines 2 and 3 of the library's source code shown in \protect\subref{fig:subfloat-datadog-logger}.}
    \label{fig:experiments-results-module-datadog-metrics}
\end{figure}

\subsection{Evaluation}
As shown in \figref{fig:experiments-typescript-templates-distribution-definitely-typed}, 20\% of the declaration files in DefinitelyTyped are written using the \lstinline{module-function}. However, 57\% of the 244 generated declaration files are written with the \lstinline{module-function} template. Additionally, the complexity of evaluating declaration files written with the \lstinline{module-function} is considerably lower than for other templates. The evaluation for templates \lstinline{module-class} and \lstinline{module} was not implemented.

33 out of 116 evaluated modules have no difference with their corresponding declaration file in DefinitelyTyped.

\begin{figure}[tp]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
            width=0.7\textwidth,
            scale=0.8,
			ybar=0pt,
			ymax=100,
			enlargelimits=0.15,
			bar width=0.04\textwidth,
			legend columns=1,
			legend cell align={left},
			legend style={
                at={(0.6,0.95)},
                anchor=north,
                nodes={scale=0.8, transform shape}
            },
			symbolic x coords={module-class,module-function,module},
			xtick=data,
			ylabel=Percentage,
			nodes near coords, 
			nodes near coords align={vertical}
		]
		\addplot coordinates {
			(module-class, 8)
			(module-function, 20)
			(module, 72)
		};

		\addplot coordinates {
			(module-class, 12)
			(module-function, 57)
			(module, 31)
		};
		\legend{DefinitelyTyped (Total: 6029), Generated (Total: 244)}
		\end{axis}
	\end{tikzpicture}

	\caption[TypeScript templates distribution | Generated \& DefinitelyTyped]{\textbf{TypeScript templates distribution | Generated \& DefinitelyTyped} - Out of a total of 6029, 72\% of the modules uploaded to the DefinitelyTyped repository use the \lstinline{module} template and only 20\% use the \lstinline{module-function} one. However, 57\% of the 244 generated declaration files use the \lstinline{module-function} template.
	}

	\label{fig:experiments-typescript-templates-distribution-definitely-typed}
\end{figure}

\section{Related Work}
\paragraph*{Microsoft's dts-gen}
Microsoft developed \lstinline[language={}]{dts-gen}, a tool that creates starter declaration files for JavaScript libraries \cite{dts-gen}. Its documentation states that the result is however intended to be only used a starting point. The outcome needs to be refined afterwards by the developers.

The tool analyzes the shape of the objects at runtime after initialization without executing the library. This results in many variables being inferred as \lstinline[language={}]{any}. \coderef{code:related-work-dts-gen-example} shows an example for module \lstinline[language={}]{abs}.

The solution presented in this work, however, is intended to generate declaration files that are ready to be uploaded to DefinitelyTyped without further manual intervention. Any amount of manual work that a developer needs to do on a declaration file after updating JavaScript code increases the risk for having discrepancies between the declaration file and the implementation.

Formal aspects like applying the right template and using the correct syntax are perfectly covered by \lstinline[language={}]{dts-gen}.

\begin{lstlisting}[
    language=bash,
    caption={Microsoft's dts-gen example - A declaration file for module \lstinline!abs! is generated. Types are inferred as \lstinline!any!. The correct \lstinline!module-function! template is used.},
	label=code:related-work-dts-gen-example,
    float,
    abovecaptionskip=-\medskipamount
]
$ npm i -g dts-gen
$ npm i -g abs
$ dts-gen -m abs
Wrote 5 lines to abs.d.ts.

$ cat abs.d.ts
/** Declaration file generated by dts-gen */

export = abs;

declare function abs(input: any): any;
\end{lstlisting}

\paragraph*{TSInfer \& TSEvolve}
TSInfer and TSEvolve are presented as part of TSTools \cite{DBLP:conf/fase/KristensenM17}. Both tools are the continuation of TSCheck \cite{DBLP:conf/oopsla/FeldthausM14}, a tool for looking for mismatches between a declaration file and an implementation.

TSInfer proceeds in a similar way than TSCheck. It initializes the library in a browser and it records a snapshot of the resulting state and then it performs a light weight static analysis on all the functions and objects stored in the snapshot.

The abstraction and the constraints they introduced as part of the static analysis tools for inferring the types have room for improvement. A run-time based approach like the one presented in our work will provide more accurate information, thus generating more precise declaration files.

Since they analyze the objects and functions stored in the snapshot, they faced the problem of including in the declaration file internal methods and properties that developers wanted to hide. Run-time information would have informed that the developer has no intention of exposing such methods.

Moreover, TSEvolve performs a differential analysis on the changes made to a JavaScript library in order to determine intentional discrepancies between declaration files of two consecutive versions. We consider that a differential analysis may not be needed. If the developer's intention is accurately extracted and the execution code clearly represents that intention then the generated declaration file would already describe the newer version of a library without the need of a differential analysis.

\paragraph*{TSTest}
TSTest is a tool that checks for mismatches between a declaration file and a JavaScript implementation \cite{DBLP:journals/pacmpl/KristensenM17}. It applies feedback-directed random testing for generating type test scripts. These scripts will execute the library in order to check if it behaves the way it is described in the declaration file. TSTest also provides concrete executions for mismatches.

We evaluated the generated declaration files comparing them to the declaration files uploaded to DefinitelyTyped. The disadvantage of doing this is that since the uploaded files are written manually, they could already contain mismatches with the JavaScript implementation. However, it is a suitable choice for a development stage since it is used as a baseline.

In a final stage, declaration files need to be checked against the proper JavaScript implementation and TSTest has to be definitely taken into account.

\section{Conclusion}
We have presented \lstinline{dts-generate}, a tool for generating a TypeScript declaration file for a specific JavaScript library. The tool downloads code samples written by the developers from the library's repository. It uses these samples to execute the library and gather data flow and type information. The tool finally generates a TypeScript declaration file based on the information gathered at run-time.

We developed an architecture that supports the automatic generation of declaration files for specific JavaScript libraries without additional manual tasks. The architecture contemplates a future incorporation of a Symbolic Execution Engine that refines the initial code base enabling the exploration of new execution paths. However not implemented in this work, its incorporation would result in small incremental modifications to the presented architecture as it is considered to only expand the existing code base.

Building an end-to-end solution for the generation of TypeScript declaration files was prioritized over type inference accuracy. Consequently, types were taken over from the values at run-time. Since developers expose through code how a library should be used, obtaining the types from the code examples extracted from the repositories proved to be a pragmatic and effective approximation, enabling to work on specific aspects regarding the TypeScript declaration file generation itself.

We built a mechanism to automatically create declaration files for potentially every module uploaded to DefinitelyTyped. We managed to generate declaration files for 244 modules. We compared the results against the corresponding files uploaded to DefinitelyTyped by creating a TypeScript declaration files parser and a comparator.

We exposed the fundamental aspect of capturing the developer's intention when inferring types in JavaScript. Instead of applying constraints and restrictions for operations with certain types, we presented a proposal where common practices are favored. Uncommon usage is not forbidden but greatly disfavored. Accordingly, we collected evidence regarding the usage of JavaScript operators by analyzing 400 libraries.

Finally, the architecture is composed of different blocks that interact with each other. Each block is independent and has a well defined behavior as well as clear input and output values. As a result, each block can be independently and simultaneously improved.

\bibliography{lipics-v2019-sample-article}

\end{document}

### 2019-12-18 ecoop-2020/lipics-v2019-sample-article.tex
Add motivating example

@@
 
-\relatedversion{}%optional, e.g. full version hosted on arXiv, HAL, or other respository/website
+\relatedversion{}%optional, e.g. full version hosted on arXiv, HAL, or other repository/website
 %\relatedversion{A full version of the paper is available at \url{...}.}
@@
 
-This work presents \lstinline[language={}]!dts-generate!, a tool that generates a declaration file for JavaScript libraries uploaded to the NPM registry.
+This work presents \lstinline{dts-generate}, a tool that generates a declaration file for JavaScript libraries uploaded to the NPM registry.
 It executes the library, gathers run-time information from the code examples written by the developers and generates the declaration file based on it.
@@
 
-\section{Motivating Examples}
+\section{Motivating Example}
+The NPM package \lstinline{route-parser} is a simple route parsing, matching, and reversing library for Javascript\footnote{https://www.npmjs.com/package/route-parser}.
+It has about 35000 weekly downloads and 221 NPM packages depend on it.
+If a developer is creating or extending a JavaScript library written in TypeScript that depends on \lstinline{route-parser}, the TypeScript compiler and IDEs will need a declaration file for that JS library in order to perform static checking and code completion, respectively.
+We use \lstinline{dts-generate} to automatically generate a TypeScript Declaration File for this library.
+The tool will download the npm package, run it, gather run-time information and generate a valid TypeScript Declaration File.
+\figref{fig:motivating-example-route-parser-vscode} shows the generated declaration file for package \lstinline{route-parser} using \lstinline{dts-generate}.
+The result is a valid declaration file that can be used in a TypeScript project.
+For example, it can be seen in \figref{fig:motivating-example-route-parser-vscode} that Visual Studio Code performs code completion properly.
+Finally, if \lstinline{route-parser} gets modified in a future, a new declaration file could be automatically generated using \lstinline{dts-generate}.
+
+\begin{figure}[tp]
+  \centering
+  \begin{subfigure}{0.70\linewidth}
+    \begin{lstlisting}[language=TypeScript]
+export = Route;
+
+declare class Route {
+  constructor(spec: string);
+  match(path: string): object;
+  reverse(params: object): string;
+}
+
+declare namespace Route {
+}
+    \end{lstlisting}
+  \end{subfigure}
+
+  \hspace{10.\textwidth}
+
+  \begin{subfigure}{1.\linewidth}
+    \centering
+    \includegraphics[width=0.7\linewidth]{motivating-example-route-parser-vscode.png}
+  \end{subfigure}
+
+  \caption{\textbf{Declaration file for \lstinline{route-parser} generated with \lstinline{dts-generate}} - Constructor and methods are correctly identified.
+Declaration file can be correctly used in Visual Studio Code\footnote{https://code.visualstudio.com}.}
+  \label{fig:motivating-example-route-parser-vscode}
+\end{figure}
 
@@
 \subsection{TypeScript Declaration File Generation}
-\paragraph*{Overwiew}
+\paragraph*{Overview}
 The actual generation of the declaration file is the next step in the pipeline after gathering the run-time information, as shown in \figref{fig:tsd_generation_method_block_diagram}.
@@
 
-Templates of type \lstinline{module-class} are shown for modules \lstinline{flake-idgen}, \lstinline{route-parser} and \lstinline{timer-machine} in \figref{fig:experiments-results-module-class-flake-idgen}, \figref{fig:experiments-results-module-class-route-parser} and \figref{fig:experiments-results-module-class-timer-machine}, respectively.
+Templates of type \lstinline{module-class} are shown for modules \lstinline{flake-idgen}, \lstinline{route-parser} and \lstinline{timer-machine} in \figref{fig:experiments-results-module-class-flake-idgen} and \figref{fig:experiments-results-module-class-timer-machine}, respectively.
 Properties of interfaces and class methods are correctly generated.
@@
     \label{fig:experiments-results-module-class-flake-idgen}
-\end{figure}
-
-\begin{figure}[tp]
-    \centering
-    \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
-export = Route;
-
-declare class Route {
-    constructor(spec: string);
-    match(path: string): object;
-    reverse(params: object): string;
-}
-
-declare namespace Route {
-}
-      \end{lstlisting}
-      \caption{route-parser/index.d.ts - Generated}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
-declare class Route {
-    constructor(spec: string);
-    match(pathname: string): { [i: string]: string } | false;
-
-    reverse(params: { [i: string]: any }): string | false;
-}
-
-declare namespace Route { }
-export = Route;       
-      \end{lstlisting}
-      \caption{route-parser/index.d.ts - DefinitelyTyped}
-    \end{subfigure}
-
-    \caption{\textbf{Module-class results | route-parser} - Constructor and methods are correctly identified.
-Specific signature of dictionary \lstinline{params} is not detected.
-Instead, a generic \lstinline{object} is inferred.}
-    \label{fig:experiments-results-module-class-route-parser}
 \end{figure}

### 2019-12-30 ecoop-2020/main.tex
cleanup
abstract

--- 2019-12-18 ecoop-2020/lipics-v2019-sample-article.tex
@@
 \documentclass[a4paper,english,cleveref, autoref]{lipics-v2019}
+% ECOOP 2020: 25 pages excluding references
 %This is a template for producing LIPIcs articles. 
@@
 \begin{abstract}
-Developers are starting to write complex and large JavaScript applications in TypeScript, a typed superset of JavaScript.
-TypeScript applications integrate JavaScript libraries by including typed descriptions of their APIs called declaration files.
-These files are maintained manually, which is error prone and highly time consuming.
-Discrepancies between a declaration file and the JavaScript implementation will prevent developers from using them.
+Developers are starting to write complex and large applications in
+TypeScript, a typed dialect of JavaScript. TypeScript applications
+integrate JavaScript libraries via typed descriptions of their APIs
+called declaration files. These files are available in public
+repository DefinitelyTyped.
+This repository is maintained manually, which
+is error prone and time consuming. Discrepancies between a
+declaration file and the JavaScript implementation lead to
+incorrect feedback from the TypeScript IDE and thus to incorrect uses
+of the underlying JavaScript library.
 
-This work presents \lstinline{dts-generate}, a tool that generates a declaration file for JavaScript libraries uploaded to the NPM registry.
-It executes the library, gathers run-time information from the code examples written by the developers and generates the declaration file based on it.
-244 declaration files were generated and compared against DefinitelyTyped, the standard public repository for declaration files.
-33 files out of 244 had no differences.
+This work presents \lstinline{dts-generate}, a tool that generates
+TypeScript declaration files for JavaScript libraries uploaded to the NPM
+registry. It extracts code examples from the documentation written by
+the developer, executes the library driven by the examples, gathers
+run-time information, and generates a declaration file based on this
+information. To evaluate the tool, 244 declaration files were generated and 
+compared against the declaration file provided on DefinitelyTyped, the
+standard public repository for declaration files. 33 files out of 244 had no differences. 
 \end{abstract}
@@
 
-\bibliography{lipics-v2019-sample-article}
+\bibliography{main}
 

### 2019-12-30 ecoop-2020/main.tex
abstract

@@
 \begin{abstract}
-Developers are starting to write complex and large applications in
+Developers are starting to write large and complex applications in
 TypeScript, a typed dialect of JavaScript. TypeScript applications
 integrate JavaScript libraries via typed descriptions of their APIs
-called declaration files. These files are available in public
-repository DefinitelyTyped.
-This repository is maintained manually, which
+called declaration files. DefinitelyTyped is the standard public
+repository for these files.
+The repository is maintained manually by volunteers, which
 is error prone and time consuming. Discrepancies between a
@@
 information. To evaluate the tool, 244 declaration files were generated and 
-compared against the declaration file provided on DefinitelyTyped, the
-standard public repository for declaration files. 33 files out of 244 had no differences. 
+compared with the declaration file provided on DefinitelyTyped.
+33 files out of 244 had no differences.
+% but what about the remaining generated files
 \end{abstract}

### 2020-01-03 ecoop-2020/main.tex
intro & motivation

@@
 
-This work presents \lstinline{dts-generate}, a tool that generates
+This work presents \texttt{dts-generate}, a tool that generates
 TypeScript declaration files for JavaScript libraries uploaded to the NPM
@@
 \section{Introduction}
-JavaScript has become the most popular language for writing web applications \cite{github-statistics}.
-It is also gaining popularity for back-end applications running in NodeJS.
-Its dynamic typing speeds up programming enabling developers to create simple pieces of code very fast, making JavaScript a very appealing programming language.
-
-JavaScript is being used for creating complex and large applications.
-However, JavaScript was not intended to be more than a scripting language.
-Maintaining and evolving large JavaScript codebases is notably challenging.
-Mistakes such as mistyped property names and misunderstood or unexpected type coercion cause developers to spend a significant amount of time in debugging sessions.
-A JavaScript code blog\footnote{https://wtfjs.com} compiles experiences from developers facing unexpected situations while programming in JavaScript.
-\coderef{code:introduction-javascript-wtfs} exposes some of these unintuitive JavaScript behaviors.
+\label{sec:introduction}
+JavaScript has become the most popular language for writing web
+applications \cite{github-statistics}. It is also gaining popularity
+for back-end applications running in NodeJS, a JavaScript-based
+server-side platform. JavaScript is appealing to developers because
+its forgiving dynamic typing enables 
+them to create simple pieces of code very quickly and proceed on a
+trial-and-error basis.
+
+JavaScript was never intended to be more than a
+scripting language. Hence, it lacks features for maintaining and evolving large
+codebases.
+Presently, however, JavaScript is being used for creating large and complex
+applications. 
+Mistakes such as mistyped property
+names and misunderstood or unexpected type coercions cause developers
+to spend a significant amount of time in debugging. There is ample
+evidence for such mishaps. For example, a 
+JavaScript code blog\footnote{\url{https://wtfjs.com}} collects experiences
+from developers facing unexpected situations while programming in
+JavaScript. \coderef{code:introduction-javascript-wtfs} exposes some
+of these unintuitive JavaScript behaviors.
 
 \begin{lstlisting}[
-    caption={\textbf{Unexpected JavaScript behavior} - falsy values, \lstinline!typeof!, \lstinline!null! and \lstinline!undefined! operators and type coercion.},
+    caption={\textbf{Unexpected JavaScript behavior} - falsy values, \lstinline!typeof!, \lstinline!null! and \lstinline!undefined! operators and type coercion},
     language=JavaScript,
@@
 null == undefined; // true
-"01" < "00100"; // false
 [1] + 1; //'11'
@@
 \end{lstlisting}
-
-The overhead produced by such dynamic typing is not present in other languages that use build tools based on type information.
-The situation motivated the creation of TypeScript, a superset of JavaScript with typed annotations \cite{typescript}.
-It has become a widely used alternative among JavaScript developers, since it incorporates features that are helpful for developing and maintaining large applications \cite{DBLP:conf/icse/GaoBB17}.
-TypeScript enables the early detection of run-time errors detection and the integration of code intelligence tools like auto-completion in the IDEs.
-
-Existing JavaScript libraries can be used in a TypeScript project by adding a declaration file that contains a typed description of the library's API.
-Declaration files are stored in a repository called DefinitelyTyped that contains declaration files for more than 6000 JavaScript libraries \cite{definitely-typed-repository}.
-Unfortunately, declaration files need to be manually created and maintained, which is error prone and time consuming.
-TypeScript does not perform any run-time check on these declaration files.
-A discrepancy between the declaration file and its corresponding JavaScript library would lead to additional frustration and debugging sessions, since type checks and code-intelligence features would be inaccurate.
-
-Some previous work tackled the problem of automatically searching for mismatches between a declaration file and implementation code \cite{DBLP:conf/oopsla/FeldthausM14}.
-TSTest adapted the feedback-directed random testing technique, mainly used for testing Java libraries, to detect discrepancies between a declaration file and a JavaScript library \cite{DBLP:journals/pacmpl/KristensenM17}.
-Tools like TSInfer and TSEvolve are designed for assisting the creation of new declaration files and supporting the evolution the declaration file when the corresponding JavaScript library gets modified \cite{DBLP:conf/fase/KristensenM17}, respectively.
-TypeScript itself developed \lstinline[language={}]{dts-gen}, a tool that generates a declaration file that is meant to be used only as a `starting point for writing a high-quality declaration file' \cite{dts-gen}.
-
-We explore in this work the possibilities for improving the existing tools.
-We present the tool \lstinline[language={}]{dts-generate}.
-It is based on an architecture that supports the generation of declaration files for an existing JavaScript library published to the NPM registry.
-The tool will gather data flow and type information at run-time and generate a declaration file based on that information.
-
-The architecture supports the future incorporation of a Symbolic Execution Engine that expands the initial code base using the signatures in the declaration file.
-The iterative process of exploring new execution paths will refine the generated declaration file in each iteration.
-
-Finally, we generated declaration files for 244 JavaScript libraries and evaluated them against DefinitelyTyped.
+% pjt: why unexpected?
+% "01" < "00100"; // false
+
+The cognitive load produced by such dynamic typing (which includes
+unchecked property names) is not present in other languages that use
+build tools based on type information. This insight motivated the
+creation of TypeScript, a superset of JavaScript with type
+annotations \cite{typescript}. It has become a widely used alternative
+among JavaScript developers, because it incorporates features that are
+helpful for developing and maintaining large applications
+\cite{DBLP:conf/icse/GaoBB17}. TypeScript enables the early detection
+of several kinds of run-time errors and the integration of code intelligence
+tools like auto-completion in an IDE.
+
+Existing JavaScript libraries can be used in a TypeScript project by
+adding a declaration file that contains a description of the library's
+API in terms of types. Lots of declaration files for popular
+JavaScript libraries have been created in a community effort in the
+DefinitelyTyped repository~\cite{definitely-typed-repository}.
+At the time of writing this repository contains declaration files for
+more than 6000 JavaScript libraries. Unfortunately, most declaration
+files in this libraries have been manually created and maintained,
+which is error prone and time consuming. As TypeScript takes a
+declaration file at face value, discrepancies are not detected at
+compile time and the inaccurate code-intelligence features are misleading the
+programmer. Moreover, TypeScript does not perform any run-time 
+checking on types, either, so that a discrepancy  between the declaration file
+and its corresponding JavaScript library can lead to unexpected 
+behavior and crashes. Such behavior can lead to developer
+frustration, longish debugging sessions, and decreasing confidence in
+the tool chain.
+
+Some previous work tackled the problem of improving the quality of
+declaration files. Feldthaus et al
+\cite{DBLP:conf/oopsla/FeldthausM14} search automatically for
+mismatches between a declaration file and implementation code. TSTest
+\cite{DBLP:journals/pacmpl/KristensenM17}
+adapted feedback-directed random testing to detect discrepancies between a
+declaration file and a JavaScript library. Tools like TSInfer and
+TSEvolve~\cite{DBLP:conf/fase/KristensenM17} are designed for
+assisting the creation of new declaration 
+files and supporting the evolution the declaration file when the
+corresponding JavaScript library gets modified, respectively. They
+rely on an existing static analyzer for JavaScript \cite{TAJS}.
+TypeScript itself
+developed \texttt{dts-gen}, a tool that generates a
+declaration file that is meant to be used as a \emph{starting point
+for writing a high-quality declaration file} \cite{dts-gen}.
+
+In this work we present the tool \texttt{dts-generate}
+as a first step to explore the possibilities for
+generating useful declaration files without the heavy lifting of
+static analysis. \texttt{dts-generate} comes with a framework that
+supports the generation of declaration files for an existing
+JavaScript library published to the NPM registry. The tool gathers
+data flow and type information at run-time to generate a declaration
+file based on that information.
+
+The novelty of our tool is twofold:
+\begin{enumerate}
+\item 
+  We do not rely on static analysis, which is hard to implement
+  soundly and precisely and which is prone to maintenance problems
+  when keeping up with JavaScript's yearly language updates.
+\item
+  Instead we extract example code from the programmer's library
+  documentation and rely on dynamic analysis to extract typed usage
+  patterns for the library from the example runs.
+\end{enumerate}
+% The architecture supports the future incorporation of a Symbolic
+% Execution Engine that expands the initial code base using the
+% signatures in the declaration file. The iterative process of exploring
+% new execution paths will refine the generated declaration file in each
+% iteration.
+
+% Finally, we generated declaration files for 244 JavaScript libraries
+% and evaluated them against the DefinitelyTyped repository.
 
@@
 \begin{itemize}
-    \item We introduce an architecture that supports generating TypeScript Declaration Files for a given JavaScript Library using run-time information.
-The architecture supports the future incorporation of a Symbolic Execution Engine that expands the initial code base using the signatures in the declaration file.
-The iterative process of exploring new execution paths will refine the generated declaration file in each iteration.
-
-    \item We present the tool \lstinline{dts-generate}, a command line application that generates a valid TypeScript Declaration File for a specific NPM package using run-time information.
-
-    \item Next, we deliver a test framework that supports generating a TypeScript Declaration File using \lstinline{dts-generate} for all JavaScript libraries in the DefinitelyTyped Repository.
-We present the results of running \lstinline{dts-generate} for 244 of these libraries.
-
-    \item Finally, we present a TypeScript Declaration Files parser and comparator. Both tools were necessary for evaluating the results against DefinitelyTyped.
+% \item We introduce an architecture that supports generating TypeScript
+%   Declaration Files for a given JavaScript Library using run-time
+%   information. The architecture supports the future incorporation of a
+%   Symbolic Execution Engine that expands the initial code base using
+%   the signatures in the declaration file. The iterative process of
+%   exploring new execution paths will refine the generated declaration
+%   file in each iteration.
+\item A framework that extracts code examples from the
+  documentation of an NPM modules and collects run-time type
+  information from running these examples.
+
+\item The tool \texttt{dts-generate}, a command line
+  application that generates a valid TypeScript declaration file for a
+  specific NPM package using run-time information.
+
+\item A comparator for TypeScript declaration files. This tool is
+  necessary for evaluating our framework and also useful to detect
+  incompatibilities when evolving JavaScript modules.
+\item An evaluation of our framework. We examined all 6029 entries in
+  the DefinitelyTyped repository and found 244 sufficiently
+  well-documented NPM libraries, on which we ran \texttt{dts-generate}
+  and compared the outcome with the respective declaration file from the
+  DefinitelyTyped repository.
+
 \end{itemize}
@@
 \section{Motivating Example}
-The NPM package \lstinline{route-parser} is a simple route parsing, matching, and reversing library for Javascript\footnote{https://www.npmjs.com/package/route-parser}.
-It has about 35000 weekly downloads and 221 NPM packages depend on it.
-If a developer is creating or extending a JavaScript library written in TypeScript that depends on \lstinline{route-parser}, the TypeScript compiler and IDEs will need a declaration file for that JS library in order to perform static checking and code completion, respectively.
-We use \lstinline{dts-generate} to automatically generate a TypeScript Declaration File for this library.
-The tool will download the npm package, run it, gather run-time information and generate a valid TypeScript Declaration File.
-\figref{fig:motivating-example-route-parser-vscode} shows the generated declaration file for package \lstinline{route-parser} using \lstinline{dts-generate}.
-The result is a valid declaration file that can be used in a TypeScript project.
-For example, it can be seen in \figref{fig:motivating-example-route-parser-vscode} that Visual Studio Code performs code completion properly.
-Finally, if \lstinline{route-parser} gets modified in a future, a new declaration file could be automatically generated using \lstinline{dts-generate}.
+The NPM package \texttt{route-parser} is a simple route parsing,
+matching, and reversing library for
+Javascript\footnote{https://www.npmjs.com/package/route-parser}. It
+has about 35000 weekly downloads and 221 NPM packages depend on it. If
+a developer creates or extends TypeScript code that depends on the
+\texttt{route-parser} library, the TypeScript
+compiler and IDE requires a declaration file for that library to
+perform static checking and code completion, respectively. We 
+use \texttt{dts-generate} to automatically generate a TypeScript
+declaration file for \texttt{route-parser}. The tool downloads the NPM 
+package, runs the examples extracted from its documentation, gathers
+run-time information, and generates a TypeScript declaration
+file. The result is a valid declaration file
+shown in \figref{fig:motivating-example-route-parser-vscode}  that 
+can be used in a TypeScript project. For example,  Visual Studio
+Code performs code completion properly
+(\figref{fig:motivating-example-route-parser-vscode}). If the
+\texttt{route-parser} package gets modified in a future, a new declaration
+file can be generated automatically using
+\texttt{dts-generate}. Our comparator tool can compare the new file
+for incompatibilities with the previous declaration file.
 
@@
 
-  \caption{\textbf{Declaration file for \lstinline{route-parser} generated with \lstinline{dts-generate}} - Constructor and methods are correctly identified.
+  \caption{\textbf{Declaration file for \texttt{route-parser} generated with \texttt{dts-generate}} - Constructor and methods are correctly identified.
 Declaration file can be correctly used in Visual Studio Code\footnote{https://code.visualstudio.com}.}
@@
 \section{dts-generate - Generation of TypeScript Declaration Files}
-We introduce \lstinline{dts-generate}, a command line tool which generates a valid TypeScript Declaration File for a specific JavaScript Library uploaded to the NPM Registry, as explained in \figref{fig:tsd_generation_method_block_diagram}.
-The tool is intended to be used on existing, published \lstinline{npm} packages.
+We introduce \texttt{dts-generate}, a command line tool which generates a valid TypeScript Declaration File for a specific JavaScript Library uploaded to the NPM Registry, as explained in \figref{fig:tsd_generation_method_block_diagram}.
+The tool is intended to be used on existing, published \texttt{npm} packages.
 The generated output TypeScript declaration file is a valid file which can be used for development and uploaded to the DefinitelyTyped Repository.
@@
 
-The command line interface was inspired in the \lstinline{dts-gen} \cite{dts-gen} package.
+The command line interface was inspired in the \texttt{dts-gen} \cite{dts-gen} package.
 It can be seen in \coderef{code:dts-generate-example} that invoking the package is very simple and the only required argument is the name of the module published to the npm registry.
@@
 \begin{lstlisting}[
-    caption={\textbf{dts-generate usage} - Example of how to generate a declaration file for module \lstinline{abs}.},
+    caption={\textbf{dts-generate usage} - Example of how to generate a declaration file for module \texttt{abs}.},
     language=bash,
@@
 \begin{itemize}
-  \item Obtain the repository url from the package: The command \lstinline{npm view <PACKAGE> repository.url} can be used for retrieving the url of the package's repository.
+  \item Obtain the repository url from the package: The command \texttt{npm view <PACKAGE> repository.url} can be used for retrieving the url of the package's repository.
 
@@
 It is very common to write code examples within code blocks indicating the programming language, so that it gets highlighted with the specific syntax.
-The Markdown identifiers for JavaScript are \lstinline{js} or \lstinline{javascript}.
+The Markdown identifiers for JavaScript are \texttt{js} or \texttt{javascript}.
 The code examples are finally retrieved by filtering the content within the corresponding code blocks.
@@
 
-The implementation stores these observations as entities called \lstinline{interactions}.
+The implementation stores these observations as entities called \texttt{interactions}.
 They are used for translating, modifying and aggregating Jalangi's raw event information in order to get an application specific data representation.
@@
 
-The tool is called \lstinline{parse-dts} and is naturally written in TypeScript. It also runs in NodeJS within its corresponding Docker container.
+The tool is called \texttt{parse-dts} and is naturally written in TypeScript. It also runs in NodeJS within its corresponding Docker container.
 
@@
 The url of the repositories could be retrieved for only 4974 modules.
-More than 1000 modules do not have the repository entry in their corresponding \lstinline{package.json} files.
-Therefore, the \lstinline{npm view <module> repository.url} command returns an empty value.
-This is even happening for important modules like \lstinline{ace}.
+More than 1000 modules do not have the repository entry in their corresponding \texttt{package.json} files.
+Therefore, the \texttt{npm view <module> repository.url} command returns an empty value.
+This is even happening for important modules like \texttt{ace}.
 
 \paragraph*{Readme Files}
-700 modules simply do not have a readme file in their repositories. The implementation does contemplate, however, different naming conventions like \lstinline{readme.md} or \lstinline{README.md}.
+700 modules simply do not have a readme file in their repositories. The implementation does contemplate, however, different naming conventions like \texttt{readme.md} or \texttt{README.md}.
 
 \paragraph*{Code Examples Extraction}
-The 50\% loss is mainly explained because developers did not wrap their code around a block using the \lstinline{javascript} or \lstinline{js} tags.
+The 50\% loss is mainly explained because developers did not wrap their code around a block using the \texttt{javascript} or \texttt{js} tags.
 Counting with code examples for 2200 modules was considered to be enough for evaluating the generation of declaration files.
@@
 \paragraph*{Code Examples Execution}
-The 2260 extracted code examples were executed by installing the required packages and running the code as a \lstinline{node} application.
+The 2260 extracted code examples were executed by installing the required packages and running the code as a \texttt{node} application.
 Working and functional code examples could only be extracted for 1448 modules.
@@
 
-\figref{fig:experiments-results-module-function} shows the generated declaration files for simple modules like \lstinline{abs}, \lstinline{dirname-regex} and \lstinline{escape-html}.
-All of them were generated using the \lstinline{module-function} template.
+\figref{fig:experiments-results-module-function} shows the generated declaration files for simple modules like \texttt{abs}, \texttt{dirname-regex} and \texttt{escape-html}.
+All of them were generated using the \texttt{module-function} template.
 There are no differences between the generated files and the corresponding declaration files uploaded to DefinitelyTyped.
@@
 
-    \caption{\textbf{Module-function results} - Results are shown for modules \lstinline{abs}, \lstinline{dirname-regex}, \lstinline{escape-html}.
+    \caption{\textbf{Module-function results} - Results are shown for modules \texttt{abs}, \texttt{dirname-regex}, \texttt{escape-html}.
 On the left side the generated declaration file with \lstinline{dts-generate}.
@@
 
-Templates of type \lstinline{module-class} are shown for modules \lstinline{flake-idgen}, \lstinline{route-parser} and \lstinline{timer-machine} in \figref{fig:experiments-results-module-class-flake-idgen} and \figref{fig:experiments-results-module-class-timer-machine}, respectively.
+Templates of type \lstinline{module-class} are shown for modules \texttt{flake-idgen}, \texttt{route-parser} and \texttt{timer-machine} in \figref{fig:experiments-results-module-class-flake-idgen} and \figref{fig:experiments-results-module-class-timer-machine}, respectively.
 Properties of interfaces and class methods are correctly generated.
 Optional parameters are not detected, as it was not considered for the implementation.
-Finally, \lstinline{module} template is presented for \lstinline{is-uuid} module in \figref{fig:experiments-results-module-is-uuid}.
+Finally, \lstinline{module} template is presented for \texttt{is-uuid} module in \figref{fig:experiments-results-module-is-uuid}.
 
@@
 
-    \caption{\textbf{Module-class results | flake-idgen} - Parameters of interface \lstinline{ConstructorOptions} are correctly detected.
+    \caption{\textbf{Module-class results | flake-idgen} - Parameters of interface \texttt{ConstructorOptions} are correctly detected.
 Name of interface differs since it is automatically generated based on the name of the argument variable.
 Optional properties were not implemented, hence the \lstinline{undefined} type for some properties.
-Analogously, callback \lstinline{cb} is inferred as \lstinline{undefined}.}
+Analogously, callback \texttt{cb} is inferred as \lstinline{undefined}.}
     \label{fig:experiments-results-module-class-flake-idgen}
@@
 It is worth mentioning that for some libraries the declaration file in DefinitelyTyped was not correct.
-For example, for \lstinline{datadog-metrics}, some properties of an interface were included in the generated declaration file but they were not present in the one in DefinitelyTyped.
+For example, for \texttt{datadog-metrics}, some properties of an interface were included in the generated declaration file but they were not present in the one in DefinitelyTyped.
 However, as shown in \figref{fig:experiments-results-module-datadog-metrics}, the properties are indeed used in the source code and should be included.
@@
 
-    \caption{\textbf{Missing properties | datadog-metrics} - Properties \lstinline{aggregator} and \lstinline{reporter} are not in the DefinitelyTyped version, but they appear in the generated declaration file.
+    \caption{\textbf{Missing properties | datadog-metrics} - Properties \texttt{aggregator} and \texttt{reporter} are not in the DefinitelyTyped version, but they appear in the generated declaration file.
 However, they are indeed used by the library, as exposed in lines 2 and 3 of the library's source code shown in \protect\subref{fig:subfloat-datadog-logger}.}
@@
 \paragraph*{Microsoft's dts-gen}
-Microsoft developed \lstinline[language={}]{dts-gen}, a tool that creates starter declaration files for JavaScript libraries \cite{dts-gen}.
+Microsoft developed \texttt{dts-gen}, a tool that creates starter declaration files for JavaScript libraries \cite{dts-gen}.
 Its documentation states that the result is however intended to be only used a starting point.
@@
 
-Formal aspects like applying the right template and using the correct syntax are perfectly covered by \lstinline[language={}]{dts-gen}.
+Formal aspects like applying the right template and using the correct syntax are perfectly covered by \texttt{dts-gen}.
 
@@
 \section{Conclusion}
-We have presented \lstinline{dts-generate}, a tool for generating a TypeScript declaration file for a specific JavaScript library.
+We have presented \texttt{dts-generate}, a tool for generating a TypeScript declaration file for a specific JavaScript library.
 The tool downloads code samples written by the developers from the library's repository.

### 2020-01-04 ecoop-2020/main.tex
cosmetics and further tweaking

@@
 \usetikzlibrary{positioning}
+
+\input{macros}
 
@@
 This work presents \texttt{dts-generate}, a tool that generates
-TypeScript declaration files for JavaScript libraries uploaded to the NPM
+TypeScript declaration files for JavaScript libraries uploaded to the \NPM{}
 registry. It extracts code examples from the documentation written by
@@
 applications \cite{github-statistics}. It is also gaining popularity
-for back-end applications running in NodeJS, a JavaScript-based
+for back-end applications running in \NodeJS{}, a JavaScript-based
 server-side platform. JavaScript is appealing to developers because
@@
 supports the generation of declaration files for an existing
-JavaScript library published to the NPM registry. The tool gathers
+JavaScript library published to the \NPM{} registry. The tool gathers
 data flow and type information at run-time to generate a declaration
@@
 \item A framework that extracts code examples from the
-  documentation of an NPM modules and collects run-time type
+  documentation of an \NPM{} package and collects run-time type
   information from running these examples.
@@
   application that generates a valid TypeScript declaration file for a
-  specific NPM package using run-time information.
+  specific \NPM{} package using run-time information.
 
@@
   the DefinitelyTyped repository and found 244 sufficiently
-  well-documented NPM libraries, on which we ran \texttt{dts-generate}
+  well-documented \NPM{} packages, on which we ran \texttt{dts-generate}
   and compared the outcome with the respective declaration file from the
@@
 \section{Motivating Example}
-The NPM package \texttt{route-parser} is a simple route parsing,
+\label{sec:motivating-example}
+The \NPM{} package \texttt{route-parser} is a simple route parsing,
 matching, and reversing library for
 Javascript\footnote{https://www.npmjs.com/package/route-parser}. It
-has about 35000 weekly downloads and 221 NPM packages depend on it. If
+has about 35,000 weekly downloads and 221 \NPM{} packages depend on it. If
 a developer creates or extends TypeScript code that depends on the
@@
 use \texttt{dts-generate} to automatically generate a TypeScript
-declaration file for \texttt{route-parser}. The tool downloads the NPM 
+declaration file for \texttt{route-parser}. The tool downloads the \NPM{} 
 package, runs the examples extracted from its documentation, gathers
 run-time information, and generates a TypeScript declaration
-file. The result is a valid declaration file
-shown in \figref{fig:motivating-example-route-parser-vscode}  that 
-can be used in a TypeScript project. For example,  Visual Studio
-Code performs code completion properly
+file. The result is 
+shown in \figref{fig:motivating-example-route-parser-vscode} and it is
+ready for use in a TypeScript project. For example,  Visual Studio's
+code completion runs properly with it
 (\figref{fig:motivating-example-route-parser-vscode}). If the
-\texttt{route-parser} package gets modified in a future, a new declaration
+\texttt{route-parser} package gets modified in the future, a new declaration
 file can be generated automatically using
@@
 
-\section{dts-generate - Generation of TypeScript Declaration Files}
-We introduce \texttt{dts-generate}, a command line tool which generates a valid TypeScript Declaration File for a specific JavaScript Library uploaded to the NPM Registry, as explained in \figref{fig:tsd_generation_method_block_diagram}.
-The tool is intended to be used on existing, published \texttt{npm} packages.
-The generated output TypeScript declaration file is a valid file which can be used for development and uploaded to the DefinitelyTyped Repository.
+In the next section, we examine each step in the generation process in
+detail and refer back to this example for concreteness.
+
+\section{The Generation of TypeScript Declaration Files}
+\label{sec:gener-typescr-decl}
+This section gives an overview of our approach to generating
+TypeScript declaration files from JavaScript libraries packaged in
+\NPM. \figref{fig:tsd_generation_method_block_diagram} gives a rough
+picture of the inner working of our tool \texttt{dts-generate}. The
+input is an \NPM{} package and the output is a TypeScript declaration
+file for the package if it is ``sufficiently documented'', which we
+substantiate in the next subsection.
+
+% We introduce \texttt{dts-generate}, a command line tool which
+% generates a valid TypeScript Declaration File for a specific
+% JavaScript Library uploaded to the \NPM{} Registry, as explained in
+% \figref{fig:tsd_generation_method_block_diagram}. The tool is intended
+% to be used on existing, published \NPM{} packages. The generated
+% output TypeScript declaration file is a valid file which can be used
+% for development and uploaded to the DefinitelyTyped Repository.
 
@@
     \includegraphics[width=1\linewidth]{dts-generate-block-diagram.pdf}
-    \caption[dts-generate - Architecture overview]{\textbf{dts-generate - Architecture overview} - Initial code base is retrieved from the npm package's repository.
+    \caption[dts-generate - Architecture overview]{\textbf{dts-generate - Architecture overview} - Initial code base is retrieved from the \NPM{} package's repository.
 A valid TypeScript Declaration File is generated using run-time information.
@@
 
-Code examples that execute the JavaScript Library are needed in order to extract the run-time information via code instrumentation.
-It is achieved by retrieving the examples provided in the README files of the repositories of the different libraries.
-This is generally the place where developers explicitly show how to use their code.
-It showed to be an appropriate and pragmatic way of extracting the developer's intention and providing an useful initial code base with meaningful examples, thus avoiding a possible cold start problem.
-
-The examples and the code base of the library are instrumented with Jalangi \cite{DBLP:conf/sigsoft/SenKBG13}\cite{DBLP:conf/sigsoft/SenKBG13a} to gather data flow information and type information at runtime.
-Jalangi is a dynamic analysis configurable framework that provides several analysis modules that were extended as needed to retrieve the required run-time information, which is then saved to an output JSON file.
-
-A second independent block uses the run-time information to generate a TypeScript Declaration File. It infers the overall structure of the JS Library, the interfaces and the types from the JSON file.
-
-The declaration file returned by the method is valid and fully functional, making it suitable for being used within the development process.
-It contains no errors and matches the structure of the JavaScript Library under analysis, so that the JavaScript code generated after compiling the TypeScript code runs without errors.
-The conducted experiments included tests that consisted on replacing a specific type definition from DefinitelyTyped \cite{definitely-typed-repository} with the one generated in the experiments: TypeScript compilation was successful, the generated JavaScript code ran without errors and code intelligence features performed by IDEs like code completion worked as expected.
-
-The command line interface was inspired in the \texttt{dts-gen} \cite{dts-gen} package.
-It can be seen in \coderef{code:dts-generate-example} that invoking the package is very simple and the only required argument is the name of the module published to the npm registry.
+As \texttt{dts-generate} is based on run-time information, 
+exemplary code fragments that execute the JavaScript library are
+needed to obtain 
+significant run-time information from running the instrumented library
+code.
+
+The examples and the code base of the library are instrumented with
+Jalangi \cite{DBLP:conf/sigsoft/SenKBG13} to gather data flow
+information and type information at runtime. Jalangi is a configurable
+framework for dynamic analysis of JavaScript. It provides several
+analysis modules that were extended as needed to retrieve the required
+run-time information, which is then saved in a JSON file.
+
+A second independent block uses the run-time information to generate a
+TypeScript declaration file. It infers the overall structure of the JavaScript
+library, its interfaces, and the types from the run-time information. 
+The resulting declaration file is ready for use in the development
+process. Its content mimics the usage of the library in the example
+code fragments and matches the structure of the
+JavaScript library under analysis, so that the JavaScript code
+generated after compiling the TypeScript code runs without
+interface-related errors.
+
+The command line interface is inspired by the \texttt{dts-gen} package
+\cite{dts-gen}. \coderef{code:dts-generate-example} shows
+that invoking the package is very simple and the only required
+argument is the name of the module published to the \NPM{} registry.
 
@@
 \subsection{Initial Code Base}
-To extract run-time information of a JavaScript Library, it is necessary, by definition, to actually execute the code, since the analysis modules provided by Jalangi to gather information are only triggered if the instrumented code gets executed.
-
-It was decided to extract the code examples that execute the JavaScript Library from the Readme files of the repository associated to the NPM Package.
-Readme files are usually used by developers for briefly describing what the code does, what problem it solves, how to install the application, how to build the code, etc.
-It is very common that developers provide code examples in the readme files to show how the code works and how to use it.
-This is specially true for NPM Packages, which are in general created to solve a specific problem of JavaScript development.
-
-Obtaining code examples for a specific NPM Package is achieved in three steps:
+\label{sec:initial-code-base}
+
+To extract run-time information from a JavaScript library, it is
+necessary, by definition, to actually execute the code, because the
+analysis modules provided by Jalangi to gather information are only
+triggered if the instrumented code gets executed.
+
+There are several options to obtain code fragments that drive the
+library code:
+\begin{enumerate}
+\item\label{item:1} execute code that imports the library;
+\item\label{item:2} execute the test cases that come with the library;
+\item\label{item:3} execute code fragments extracted from the library documentation.
+\end{enumerate}
+
+Option~\ref{item:1} does not solve our problem, it just delegates it
+to the importing library, which also needs code to drive it. Moreover,
+it is costly to download and instrument another package.
+
+We considered option~\ref{item:2} under the assumption that most
+libraries would come with test cases. However, there is no standard
+for testing JavaScript code so that test cases were difficult to reap
+from the \NPM{} packages: they used different directory structures,
+different (or no) testing tools, or some packages did not have tests
+at all.
+
+In the end, option~\ref{item:3} was the most viable even though there
+is no standard for documentation, either. However, almost all repositories
+contain README files where the library authors briefly describe what
+the code does, which problem it solves, how to install the
+application, how to build the code, etc. It is very common that
+developers provide code examples in the readme files to show how the
+library works and how to use it. This observation holds in particular
+for \NPM{} packages, which are generally created to solve a specific
+problem of JavaScript development.
+
+Obtaining code examples for a specific \NPM{} package is done in three steps.
 \begin{itemize}
-  \item Obtain the repository url from the package: The command \texttt{npm view <PACKAGE> repository.url} can be used for retrieving the url of the package's repository.
+\item Obtain the repository url from the package: The command
+  \texttt{npm view <PACKAGE> repository.url} can be used for
+  retrieving the url of the package's repository.
 
@@
 
-  \item Extract the code examples from the README file: Readme files are written using Markdown\footnote{https://www.markdownguide.org}, a very common lightweight and simple markup language.
-It is very common to write code examples within code blocks indicating the programming language, so that it gets highlighted with the specific syntax.
-The Markdown identifiers for JavaScript are \texttt{js} or \texttt{javascript}.
-The code examples are finally retrieved by filtering the content within the corresponding code blocks.
+\item Extract the code examples from the README file: README files are
+  written using Markdown\footnote{\url{https://www.markdownguide.org}}, a
+  popular markup language. In a README file, it is customary to write
+  code examples within code blocks labeled with the programming
+  language, so that syntax highlighting is done correctly. Hence, we
+  retrieve the code examples from code blocks labeled \texttt{js} or
+  \texttt{javascript}, which both stand for JavaScript in
+  Markdown.
 \end{itemize}
 
+For example, in the case of \texttt{route-parser} ...
+
+Obtaining the code fragments from the examples provided in the
+README files of the repository proved to be an
+appropriate and pragmatic way of extracting the developer's
+intention and providing an useful initial code base with meaningful
+examples, thus avoiding possible cold start problems.
+
 \subsection{Run-time Information Gathering}
-The Runtime Information block described in \figref{fig:tsd_generation_method_block_diagram} will gather information such as:
+The Runtime Information block described in
+\figref{fig:tsd_generation_method_block_diagram} gathers
+information such as:
 
 \begin{itemize}
-  \item Function \lstinline{f} got invoked with parameters \lstinline{a} and \lstinline{b} with types \lstinline{string} and \lstinline{number}.
-  \item Property or method \lstinline{foo} of parameter \lstinline{a} of function \lstinline{f} was accessed within the function.
-  \item Parameter \lstinline{a} of function \lstinline{f} was used as operand for operator \lstinline{==}.
+  \item Function \lstinline{f} was invoked where parameter
+    \lstinline{a} held a value of type \lstinline{string} and
+    \lstinline{b} a value of type \lstinline{number}. 
+  \item Property \lstinline{foo} of parameter \lstinline{a} of
+    function \lstinline{f} was accessed within the function. 
+  \item Parameter \lstinline{a} of function \lstinline{f} was used as
+    operand for operator \lstinline{==}. 
 \end{itemize}
 
-The dynamic analysis framework used for gathering this kind of information is Jalangi.
-The configurable analysis modules enable programming custom callbacks that get triggered with virtually any JavaScript event.
-The events that are observed are:
+The dynamic analysis framework used for gathering this kind of
+information is Jalangi. The configurable analysis modules enable
+programming custom callbacks that get triggered with virtually any
+JavaScript event. Our instrumentation observes the following events: 
 \begin{itemize}
-  \item Binary operations, like \lstinline{==}, \lstinline{+} or \lstinline{===}.
+  \item Binary operations, like \lstinline{==}, \lstinline{+} or
+    \lstinline{===}.  
   \item Variable declaration.
@@
 
-The implementation stores these observations as entities called \texttt{interactions}.
-They are used for translating, modifying and aggregating Jalangi's raw event information in order to get an application specific data representation.
-The run-time information is finally returned as a JSON file that can be used for later processing.
-The tool is written in JavaScript and runs in Node.js within a Docker container.
+The implementation stores these observations as entities called
+\texttt{interactions}. They are used for translating, modifying, and
+aggregating Jalangi's raw event information to get an application
+specific data representation. The run-time information is saved as a
+JSON file that can be used for later processing. The tool is written
+in JavaScript and runs in \NodeJS{} in a Docker container.
 
 \subsection{TypeScript Declaration File Generation}
+\label{sec:typescr-decl-file}
 \paragraph*{Overview}
-The actual generation of the declaration file is the next step in the pipeline after gathering the run-time information, as shown in \figref{fig:tsd_generation_method_block_diagram}.
-It is a lightweight, simple and fast application, which does not interact with the actual JavaScript module at run-time.
-Instead, it uses the JSON output file containing the run-time information and generates a TypeScript declaration file which is use ready to be used within a TypeScript project.
-The tool itself is written in TypeScript and runs within a Docker container in NodeJS.
+The actual generation of the declaration file is the next step in the
+pipeline after gathering the run-time information, as shown in
+\figref{fig:tsd_generation_method_block_diagram}. It is a lightweight,
+simple and fast application, which does not interact with the actual
+JavaScript module at run-time. Instead, it uses the JSON output file
+containing the run-time information and generates a TypeScript
+declaration file which is use ready to be used in a TypeScript
+project. The tool itself is written in TypeScript and runs in a
+Docker container in \NodeJS.
 
@@
 
-\subsection{Evaluation}
+\section{Evaluation}
 \label{sec:dts-generate-evaluation}
-After generating a declaration file for a published NPM module, it is necessary to evaluate the quality of it.
+After generating a declaration file for a published \NPM{} package, it is necessary to evaluate the quality of it.
 It was decided to compare the generated declaration file against the one uploaded to the DefinitelyTyped repository for the same module, as shown in \figref{fig:evaluation-diagram}.
@@
 
-The parsing consists in creating a structure where declared \lstinline{interfaces}, \lstinline{functions}, \lstinline{classes} and \lstinline{namespaces} are stored separated.
+The parsing consists in creating a structure where declared \lstinline{interfaces}, \lstinline{functions}, \lstinline{classes} and \lstinline{namespaces} are stored separately.
 Function arguments are correctly described, identifying complex types like union types or callbacks.
@@
 
-The tool is called \texttt{parse-dts} and is naturally written in TypeScript. It also runs in NodeJS within its corresponding Docker container.
+The tool is called \texttt{parse-dts} and is naturally written in TypeScript. It also runs in \NodeJS{} in its corresponding Docker container.
 
@@
 \section{Results}
-Declaration files were generated for existing modules uploaded to the NPM registry.
+\label{sec:results}
+
+\begin{quotation}
+  The conducted experiments included tests that consisted of replacing
+  a specific type definition from DefinitelyTyped
+  \cite{definitely-typed-repository} with the one generated in the
+  experiments: TypeScript compilation was successful, the generated
+  JavaScript code ran without errors and code intelligence features
+  performed by IDEs like code completion worked as expected.
+\end{quotation}
+
+Declaration files were generated for existing modules uploaded to the \NPM{} registry.
 The DefinitelyTyped repository was used as a benchmark.
@@
 	\item Extracting readme files.
-	\item Extracting code examples within readme files.
+	\item Extracting code examples from README files.
 	\item Executing code examples and discarding failing ones.
@@
 \paragraph*{Code Examples Execution}
-The 2260 extracted code examples were executed by installing the required packages and running the code as a \texttt{node} application.
+The 2260 extracted code examples were executed by installing the required packages and running the code as a \NodeJS{} application.
 Working and functional code examples could only be extracted for 1448 modules.
@@
 \section{Related Work}
+\label{sec:related-work}
 \paragraph*{Microsoft's dts-gen}
@@
 \section{Conclusion}
+\label{sec:conclusion}
 We have presented \texttt{dts-generate}, a tool for generating a TypeScript declaration file for a specific JavaScript library.

### 2020-01-06 ecoop-2020/main.tex
rewrites and questions

@@
 
-In the next section, we examine each step in the generation process in
-detail and refer back to this example for concreteness.
+After filling in some background information on declaration files,
+Section~\ref{sec:gener-typescr-decl} examines each step in the
+generation process in detail and refer back to this example for
+concreteness.
+
+\section{TypeScript Declaration Files}
+\label{sec:typescr-decl-files}
+
+The declaration file shown in
+\figref{fig:motivating-example-route-parser-vscode} describes a
+package with a single exported class \lstinline{Route}. The
+description of the class comprises the type signature of the
+constructor (line~4) and the methods \lstinline{match} and
+\lstinline{reverse} (lines~5 and~6).
+
+!!!TODO: explain meaning of namespace.
+
+This file is an instance of one of the standard templates for writing
+declaration
+files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/templates.html}:
+\textbf{module}, \textbf{module-class} and
+\textbf{module-function}. 
+Each template corresponds to a different way of describing the exports
+of a JavaScript library. Choosing the template depends on the
+particular organization of the underlying JavaScript library:
+\begin{description}
+\item[module] several exported functions,
+\item[module-class] a class-like structure,
+\item[module-function] exactly one exported functions.
+\end{description}
+
+% $ ./dts-generate abs
+% $ cat output/abs/index.d.ts 
+\begin{lstlisting}[
+    caption={Example for template \textbf{module-function}},
+    language=TypeScript,
+	label=code:dts-generate-example,
+    float,
+    abovecaptionskip=-\medskipamount
+]
+export = Abs;
+
+declare function Abs(input: string): string;
+\end{lstlisting}
+!!!TODO: example for simple \textbf{module}
+
+The \texttt{route-parser} library is an instance of a
+\textbf{module-class}. The \texttt{abs} library provides a single
+function \lstinline{Abs} so it uses template \textbf{module-function}
+as shown in \ref{code:dts-generate-example}.
 
@@
 framework for dynamic analysis of JavaScript. It provides several
-analysis modules that were extended as needed to retrieve the required
+analysis modules that we extended as needed to retrieve the required
 run-time information, which is then saved in a JSON file.
@@
 
-\begin{lstlisting}[
-    caption={\textbf{dts-generate usage} - Example of how to generate a declaration file for module \texttt{abs}.},
-    language=bash,
-	label=code:dts-generate-example,
-    float,
-    abovecaptionskip=-\medskipamount
-]
-$ ./dts-generate abs
-$ cat output/abs/index.d.ts 
-export = Abs;
-
-declare function Abs(input: string): string;
-\end{lstlisting}
-
 \subsection{Initial Code Base}
@@
 for testing JavaScript code so that test cases were difficult to reap
-from the \NPM{} packages: they used different directory structures,
-different (or no) testing tools, or some packages did not have tests
+from the \NPM{} packages: they use different directory structures,
+employ differing (or no) testing tools, or do not have tests
 at all.
@@
 is no standard for documentation, either. However, almost all repositories
-contain README files where the library authors briefly describe what
+contain README files where the library authors briefly describe in prose what
 the code does, which problem it solves, how to install the
 application, how to build the code, etc. It is very common that
-developers provide code examples in the readme files to show how the
+developers provide code examples in the README files to show how the
 library works and how to use it. This observation holds in particular
@@
 \begin{itemize}
-\item Obtain the repository url from the package: The command
-  \texttt{npm view <PACKAGE> repository.url} can be used for
-  retrieving the url of the package's repository.
-
-\item Retrieve the README file from the repository.
-
-\item Extract the code examples from the README file: README files are
+\item Obtain the  repository's URL with the command
+  \texttt{npm view <PACKAGE> repository.url}
+
+\item Retrieve the README file from the top-level directory of the repository.
+
+\item Extract the code examples from the README file. To this end,
+  observe that README files are
   written using Markdown\footnote{\url{https://www.markdownguide.org}}, a
-  popular markup language. In a README file, it is customary to write
-  code examples within code blocks labeled with the programming
+  popular markup language. In such a file, it is customary to write
+  code examples in code blocks labeled with the programming
   language, so that syntax highlighting is done correctly. Hence, we
@@
 appropriate and pragmatic way of extracting the developer's
-intention and providing an useful initial code base with meaningful
+intention and providing a useful initial code base with meaningful
 examples, thus avoiding possible cold start problems.
@@
 
-The dynamic analysis framework used for gathering this kind of
-information is Jalangi. The configurable analysis modules enable
+The dynamic analysis framework Jalangi is used for gathering this kind of
+information. The configurable analysis modules enable
 programming custom callbacks that get triggered with virtually any
@@
 \label{sec:typescr-decl-file}
-\paragraph*{Overview}
-The actual generation of the declaration file is the next step in the
-pipeline after gathering the run-time information, as shown in
-\figref{fig:tsd_generation_method_block_diagram}. It is a lightweight,
-simple and fast application, which does not interact with the actual
-JavaScript module at run-time. Instead, it uses the JSON output file
-containing the run-time information and generates a TypeScript
-declaration file which is use ready to be used in a TypeScript
-project. The tool itself is written in TypeScript and runs in a
-Docker container in \NodeJS.
-
-\paragraph*{Templates}
-TypeScript provides templates for writing declaration files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/templates.html} and each template corresponds to a different way of exporting a JavaScript module.
-The tool uses different fields from the run-time information to detect how the module is being used in order to choose the right templates accordingly.
-The implemented templates are \lstinline{module}, \lstinline{module-class} and \lstinline{module-function}.
-
-\paragraph*{Interfaces}
-Finally, interfaces are created by exploring \lstinline{getField} and \lstinline{methodCall} interactions from the run-time information.
-The code will gather the interactions for a specific argument and build the interface by incrementally adding new properties.
-Interactions within the \lstinline{followingInteractions} field are recursively traversed, building a new interface in each recursion level.
+%\paragraph*{Overview}
+
+The next step in the pipeline after gathering the run-time information is
+the actual generation of the declaration file (cf.\
+\figref{fig:tsd_generation_method_block_diagram}). It is a lightweight,
+simple and fast application, which solely relies on the run-time
+information gathered in the previous step.
+% The tool itself is written in TypeScript and runs in a
+% Docker container in \NodeJS.
+
+!!!TODO too vague: more information --- how do you choose between the templates?
+
+We use different fields (which?) from the run-time information to detect
+how the module is used to choose the most suitable template. The implemented
+templates are \textbf{module}, \textbf{module-class} and
+\textbf{module-function}.
+
+% \paragraph*{Interfaces}
+Finally, interfaces are created by exploring \texttt{getField} and
+\texttt{methodCall} interactions from the run-time information. We
+gather the interactions for a specific argument and build the
+interface by incrementally adding new properties. Interactions within
+the \texttt{followingInteractions} field are recursively traversed,
+building a new interface in each recursion level.
+
+!!!TODO: insufficient explanation of \texttt{followingInteractions}
+--- needs to be mentioned in the previous step
 
@@
 \label{sec:dts-generate-evaluation}
-After generating a declaration file for a published \NPM{} package, it is necessary to evaluate the quality of it.
-It was decided to compare the generated declaration file against the one uploaded to the DefinitelyTyped repository for the same module, as shown in \figref{fig:evaluation-diagram}.
+After generating a declaration file for an \NPM{} package, we need to
+evaluate its quality. To this end, we compare the generated
+declaration file against the one uploaded to the DefinitelyTyped
+repository for the same module, as shown in
+\figref{fig:evaluation-diagram}. While this approach does not provide
+an absolute measure of quality, it gives us at least an indication of
+the pragmatics of \texttt{dts-generate}: The files on DefinitelyTyped
+are perceived to be useful by the community. If the accuracy of the
+generated files is comparable with the accuracy of the files on
+DefinitelyTyped, then \texttt{dts-generate} is a viable alternative to
+manually creating a definition file.
+
+!!!TODO: check how many files on DT are just generated by
+\texttt{dts-gen} (e.g., the TypeScript tool)
 
@@
 
-\paragraph*{Parsing}
-Before comparison, declaration files need to be parsed so that both of them share the same structure.
-It was achieved by using the TypeScript Compiler API, a library developed by Microsoft that allows to traverse the Abstract Syntax Tree in an easy and intuitive way \cite{typescript-compiler-api}.
-
-The parsing consists in creating a structure where declared \lstinline{interfaces}, \lstinline{functions}, \lstinline{classes} and \lstinline{namespaces} are stored separately.
-Function arguments are correctly described, identifying complex types like union types or callbacks.
-Optional parameters are also identified.
-For \lstinline{classes}, a distinction between the constructor and methods is made.
-Finally, syntax and semantic errors are also checked by the TypeScript Compiler API.
-
-The tool is called \texttt{parse-dts} and is naturally written in TypeScript. It also runs in \NodeJS{} in its corresponding Docker container.
-
-\paragraph*{Comparator}
-An independent tool will compare two parsed declaration files.
-As described in \figref{fig:evaluation-diagram}, the comparator will compare the generated declaration file against the corresponding file in the DefinitelyTyped repository.
-
-It was discovered that the implementation was easier when focusing on each template independently. For this implementation, only the \lstinline{module-function} was considered for the comparison.
+To this end, we need a means of comparing two declaration files. Of
+course, we are not interested in a textual comparison, but in a
+comparison of the structures described by the files. Technically, our
+comparison of two structures A and B yields one of five outcomes:
+\begin{itemize}
+\item A and B are equivalent;
+\item A is more general than B;
+\item B is more general than A;
+\item A and B have a common generalization (i.e., they are
+  compatible);
+\item none of the above holds (i.e., A and B are incompatible).
+\end{itemize}
+
+The first step is to parse declaration files, which can be
+achieved by using the TypeScript Compiler API, a library traverse the
+Abstract Syntax Tree of a TypeScript program
+\cite{typescript-compiler-api}. 
+The step also performs a sanity check of the generated declaration
+files as it rejects files with syntactic or semantics errors.
+The output of the parser is a structure where declared
+{interfaces}, {functions}, {classes} and
+{namespaces} are stored separately. Function arguments are
+correctly described, identifying complex types like union types or
+callbacks. Optional parameters are also identified. For
+{classes}, a distinction between the constructor and methods
+is made.
+
+% The tool is called \texttt{parse-dts} and is naturally written in
+% TypeScript. It also runs in \NodeJS{} in its corresponding Docker
+% container.
+
+It remains to compare the parsed representations of two declaration
+files.
+% It was discovered that the implementation was easier when
+% focusing on each template independently.
+As an example, we discuss the implementation of comparison for
+the \textbf{module-function} template.
 
@@
 \begin{itemize}
-  \item Number of declared functions: Checks the number of declared and exported functions for each of both declaration files.
-  \item Name of declared functions: Checks whether both of the declaration files declared a function with the same name.
-  \item Number of parameters: Checks the number of parameters of the declared functions. This is checked for optional and non-optional parameters.
-  \item Interfaces: Checks the number of declared interfaces and the fields within those interfaces.
+\item Number of declared functions: Checks the number of declared
+  and exported functions for each of both declaration files. 
+\item Name of declared functions: Checks whether both of the
+  declaration files declared a function with the same name. 
+\item Number of parameters: Checks the number of parameters of the
+  declared functions. This is checked for optional and non-optional
+  parameters.
+
+  !!!TODO: what about parameter names? what about object
+  types/interface types of parameters?
+\item Interfaces: Checks the number of declared interfaces and the
+  fields within those interfaces. 
 \item Errors: Indicates whether there are errors in the declaration files.
@@
 
+!!!TODO: too informal. we need to define an ordering on interfaces etc.
+
 \section{Results}
 \label{sec:results}
+
+!!!TODO: first we need to say, what we want to achieve with the
+experimental evaluation
 
@@
 
-Declaration files were generated for existing modules uploaded to the \NPM{} registry.
-The DefinitelyTyped repository was used as a benchmark.
-Each one of the generated files was compared against the corresponding declaration file already uploaded to the repository.
-
-Figure \ref{fig:experiments-overall-funnel} shows that a declaration file was generated for 244 modules out of 6029 modules.
-Samples of the generated declaration files for templates \lstinline{module}, \lstinline{module-class} and \lstinline{module-function} are presented in \secref{sec:experiments-declaration-files-generation}.
+Declaration files were generated for existing modules uploaded to the
+\NPM{} registry. The DefinitelyTyped repository was used as a
+benchmark. Each one of the generated files was compared against the
+corresponding declaration file already uploaded to the repository.
+
+Figure \ref{fig:experiments-overall-funnel} shows that a declaration
+file was generated for 244 modules out of 6029 modules. Samples of the
+generated declaration files for templates \textbf{module},
+\textbf{module-class} and \textbf{module-function} are presented in
+\secref{sec:experiments-declaration-files-generation}.
 
@@
 \subsection{Code Examples}
-Retrieving the code examples from the JavaScript libraries' repositories proved to be a pragmatic way of capturing the types.
-However, as shown in \figref{fig:experiments-overall-funnel}, working code examples for only 2260 modules could be retrieved.
-The process of getting a valid code example for a module is divided in 4 blocks:
+Retrieving the code examples for the JavaScript libraries proved to be
+a pragmatic way of driving the type gathering at run time. However, as
+shown in \figref{fig:experiments-overall-funnel}, it was only possible
+to obtain working code examples for 2260 packages. The
+process of getting a valid code example for a module is divided in four
+stages: 
 \begin{itemize}
-	\item Extracting repositories url.
-	\item Extracting readme files.
+\item Extracting repository URL.
+\item Extracting README file.
 \item Extracting code examples from README files.
@@
 
-The results obtained for each on of them are described in the following sections.
-
-\paragraph*{Repositories URL}
-The url of the repositories could be retrieved for only 4974 modules.
-More than 1000 modules do not have the repository entry in their corresponding \texttt{package.json} files.
-Therefore, the \texttt{npm view <module> repository.url} command returns an empty value.
-This is even happening for important modules like \texttt{ace}.
-
-\paragraph*{Readme Files}
-700 modules simply do not have a readme file in their repositories. The implementation does contemplate, however, different naming conventions like \texttt{readme.md} or \texttt{README.md}.
-
-\paragraph*{Code Examples Extraction}
-The 50\% loss is mainly explained because developers did not wrap their code around a block using the \texttt{javascript} or \texttt{js} tags.
-Counting with code examples for 2200 modules was considered to be enough for evaluating the generation of declaration files.
-
-\paragraph*{Code Examples Execution}
-The 2260 extracted code examples were executed by installing the required packages and running the code as a \NodeJS{} application.
-Working and functional code examples could only be extracted for 1448 modules.
-812 modules did not run correctly and were discarded.
-Some failing samples were analyzed and there were mainly two reasons for the failure:
-\begin{enumerate}
-	\item The code example had been properly extracted but the code itself was not working. It was executing the library in an unsupported way, hence the error at run-time.
-	\item The extracted code example was not intended to be executed or it was not even valid JavaScript code.
-\end{enumerate}
-
-\subsection{Declaration Files Generation} \label{sec:experiments-declaration-files-generation}
-
-The following section exhibits some samples of the 244 generated declaration files.
-It shows some results for each of the implemented templates: \lstinline{module}, \lstinline{module-function} and \lstinline{module-class}.
-
-\figref{fig:experiments-results-module-function} shows the generated declaration files for simple modules like \texttt{abs}, \texttt{dirname-regex} and \texttt{escape-html}.
-All of them were generated using the \texttt{module-function} template.
-There are no differences between the generated files and the corresponding declaration files uploaded to DefinitelyTyped.
+The results obtained for each on of them are described in the
+following sections.
+
+\paragraph*{Repository URL}
+The URL of the source repository could be retrieved for only 4974
+packages. More than 1000 packages on \NPM{} do not have a repository
+entry in their corresponding \texttt{package.json} file. Therefore, the
+\texttt{npm view <module> repository.url} command returns no
+value. Even important modules like \texttt{ace} provide no repository URL.
+
+\paragraph*{README Files}
+700 packages do not have a README file in their repository, although
+the implementation checks for several naming conventions like
+\texttt{readme.md} or \texttt{README.md}.
+
+\paragraph*{Extraction of Code Examples}
+In this step, we loose another 50\% of modules! This loss is mainly
+explained because some developers do not wrap their code in a block
+using the \texttt{javascript} or \texttt{js} tags. However, as we are
+still left with code examples for 2200 modules, we did not further
+look into code extraction as this number was considered sufficient for
+evaluating the generation of declaration files.
+
+\paragraph*{Execution of Code Examples}
+We executed the remaining 2260 extracted code by installing the
+required packages and running the code as a \NodeJS{}
+application. Unfortunately, the code examples only worked for 1448
+modules. 812 modules did not run correctly and had to be
+discarded. Some failing samples were analyzed and there were mainly
+two reasons for the failure: 
+\begin{itemize}
+\item The code fragment had been properly extracted but the code was
+  not faulty. It was executing the library in an unsupported
+  (obsolete?) way, which lead to a run-time error.
+\item The extracted code fragment was not intended to be executed
+  and/or it was not even valid JavaScript code. 
+\end{itemize}
+
+!!!TODO: the remaining two stages ``run-time information'' and
+``generated declaration files'' need to be explained
+
+\subsection{Declaration Files Generation}
+\label{sec:experiments-declaration-files-generation}
+
+This section exhibits some samples of the 244 generated
+declaration files. It shows some results for each of the implemented
+templates: \textbf{module}, \textbf{module-function} and
+\textbf{module-class}.
+
+\figref{fig:experiments-results-module-function} shows the generated
+declaration files for simple modules like \texttt{abs},
+\texttt{dirname-regex}, and \texttt{escape-html}. All of them were
+generated using the \textbf{module-function} template. There are no
+differences between the generated files and the corresponding
+declaration files on DefinitelyTyped.
+
+The left side of the figure shows the generated declaration file with
+\lstinline{dts-generate}, the right side shows the corresponding file
+in the DefinitelyTyped repository. Functions are correctly detected; input
+and output types are accurately inferred.
+
+!!!TODO: explain, what about
+\begin{itemize}
+\item uppercase/lowercase in dirname-regex?
+\item extra namespace in escape-html?
+\end{itemize}
 
@@
         \caption{escape-html/index.d.ts - DefinitelyTyped}
+      \end{subfigure}
+
+    \caption{Results for \textbf{module-function} template}
     \label{fig:experiments-results-module-function}
-      \end{subfigure}
-
-    \caption{\textbf{Module-function results} - Results are shown for modules \texttt{abs}, \texttt{dirname-regex}, \texttt{escape-html}.
-On the left side the generated declaration file with \lstinline{dts-generate}.
-On the right side the corresponding file in the DefinitelyTyped repository.
-Functions are correctly detected and input types are accurately inferred.
-Both files are parsed for comparison, as explained in \secref{sec:dts-generate-evaluation}.
-Therefore, subtle differences in the syntax between both files are not important.}
   \end{figure}
 
-Templates of type \lstinline{module-class} are shown for modules \texttt{flake-idgen}, \texttt{route-parser} and \texttt{timer-machine} in \figref{fig:experiments-results-module-class-flake-idgen} and \figref{fig:experiments-results-module-class-timer-machine}, respectively.
-Properties of interfaces and class methods are correctly generated.
-Optional parameters are not detected, as it was not considered for the implementation.
-Finally, \lstinline{module} template is presented for \texttt{is-uuid} module in \figref{fig:experiments-results-module-is-uuid}.
+Templates of type \textbf{module-class} are shown for modules
+\texttt{flake-idgen}, \texttt{route-parser}, and \texttt{timer-machine}
+in \figref{fig:experiments-results-module-class-flake-idgen} and
+\figref{fig:experiments-results-module-class-timer-machine},
+respectively. Properties of interfaces and class methods are correctly
+generated. Optional parameters are not detected, as it was not
+considered for the implementation.
+
+For \texttt{flake-idgen},  the parameters of interface
+\texttt{ConstructorOptions} are correctly detected. The name of the
+interface differs because it is automatically generated based on the
+name of the argument variable. Optional properties were not
+implemented, which explains the type \lstinline{undefined} for some
+properties. Analogously, the callback \texttt{cb} is inferred as
+\lstinline{undefined}.
+
+!!!TODO:
+\begin{itemize}
+\item why is the callback undefined? is it due to the example? (what does it
+  look like?) why is it called \texttt{cb}?
+\item the export / class is named differently; doesn't that matter?
+\end{itemize}
+
+For \texttt{timer-machine},  the parameter \lstinline{started} is
+inferred as \lstinline{undefined} instead of marking it as
+optional. Methods that were not executed cannot appear in the
+generated declaration file. For that reason, a type corresponding to
+\lstinline{TimerEvent} is not generated, either.
 
@@
 
-    \caption{\textbf{Module-class results | flake-idgen} - Parameters of interface \texttt{ConstructorOptions} are correctly detected.
-Name of interface differs since it is automatically generated based on the name of the argument variable.
-Optional properties were not implemented, hence the \lstinline{undefined} type for some properties.
-Analogously, callback \texttt{cb} is inferred as \lstinline{undefined}.}
+    \caption{Results for \textbf{module-class} module \texttt{flake-idgen}}
     \label{fig:experiments-results-module-class-flake-idgen}
@@
 
-    \caption{\textbf{Module-class results | timer-machine} - Parameter \lstinline{started} is inferred as \lstinline{undefined} instead of marking it as optional.
-Methods that were not executed do not appear in the generated declaration file.}
+    \caption{Results for \textbf{module-class} module \texttt{timer-machine}}
     \label{fig:experiments-results-module-class-timer-machine}
@@
 
-    \caption{\textbf{Module results | is-uuid} - Methods that were not executed are not included in the declaration file.}
+    \caption{Results for \textbf{module}  module \texttt{is-uuid}}
     \label{fig:experiments-results-module-is-uuid}
@@
 
-It is worth mentioning that for some libraries the declaration file in DefinitelyTyped was not correct.
-For example, for \texttt{datadog-metrics}, some properties of an interface were included in the generated declaration file but they were not present in the one in DefinitelyTyped.
-However, as shown in \figref{fig:experiments-results-module-datadog-metrics}, the properties are indeed used in the source code and should be included.
-
+Finally, a sample for the \textbf{module} template
+for the \texttt{is-uuid} module is shown in
+\figref{fig:experiments-results-module-is-uuid}.
+Again, methods that were not executed by the extracted examples are
+not included in the declaration file.
+
+It is worth mentioning that for some libraries the declaration file in
+DefinitelyTyped was not correct. For example, for the module
+\texttt{datadog-metrics}, some properties of an interface were
+included in the generated declaration file but they were not present
+in the declaration presented in DefinitelyTyped. However, as shown in
+\figref{fig:experiments-results-module-datadog-metrics}, the
+properties \texttt{aggregator} and \texttt{reporter} are not present
+in the DefinitelyTyped version, but they appear in the generated
+declaration file. However, they are indeed used by the library, as
+exposed in lines~2 and~3 of the library's source code shown in
+\figref{fig:subfloat-datadog-logger}. 
 \begin{figure}[tp]
@@
 
-    \caption{\textbf{Missing properties | datadog-metrics} - Properties \texttt{aggregator} and \texttt{reporter} are not in the DefinitelyTyped version, but they appear in the generated declaration file.
-However, they are indeed used by the library, as exposed in lines 2 and 3 of the library's source code shown in \protect\subref{fig:subfloat-datadog-logger}.}
+    \caption{Missing properties for module \texttt{datadog-metrics}}
     \label{fig:experiments-results-module-datadog-metrics}
@@
 \subsection{Evaluation}
-As shown in \figref{fig:experiments-typescript-templates-distribution-definitely-typed}, 20\% of the declaration files in DefinitelyTyped are written using the \lstinline{module-function}.
-However, 57\% of the 244 generated declaration files are written with the \lstinline{module-function} template.
-Additionally, the complexity of evaluating declaration files written with the \lstinline{module-function} is considerably lower than for other templates.
-The evaluation for templates \lstinline{module-class} and \lstinline{module} was not implemented.
-
-33 out of 116 evaluated modules have no difference with their corresponding declaration file in DefinitelyTyped.
+
+!!!TODO: explain what's going on in this subsection
+
+!!!TODO: what's the baseline for the percentages in the paragraph, the
+244 modules considered or all of DT? Ah, you say that in the figure;
+but then the figure is comparing apples with pears (percentages for
+different baselines!).
+
+As shown in
+\figref{fig:experiments-typescript-templates-distribution-definitely-typed},
+20\% of the declaration files in DefinitelyTyped are written using the
+\textbf{module-function}. However, 57\% of the 244 generated
+declaration files are written with the \textbf{module-function}
+template. Additionally, the complexity of evaluating declaration files
+written with the \textbf{module-function} is considerably lower than
+for other templates. The evaluation for templates
+\textbf{module-class} and \textbf{module} was not implemented.
+
+33 out of 116 evaluated modules have no difference with their
+corresponding declaration file in DefinitelyTyped.
+
+!!!TODO: that's a bit lame as a result. See above for a more
+meaningful output.
 
@@
 
-	\caption[TypeScript templates distribution | Generated \& DefinitelyTyped]{\textbf{TypeScript templates distribution | Generated \& DefinitelyTyped} - Out of a total of 6029, 72\% of the modules uploaded to the DefinitelyTyped repository use the \lstinline{module} template and only 20\% use the \lstinline{module-function} one.
-However, 57\% of the 244 generated declaration files use the \lstinline{module-function} template.
+	\caption[TypeScript templates distribution | Generated \& DefinitelyTyped]{\textbf{TypeScript templates distribution | Generated \& DefinitelyTyped} - Out of a total of 6029, 72\% of the modules uploaded to the DefinitelyTyped repository use the \textbf{module} template and only 20\% use the \textbf{module-function} one.
+However, 57\% of the 244 generated declaration files use the \textbf{module-function} template.
 	}

### 2020-06-26 ecoop-2020/main.tex
Add new evaluation concept and fix different todos

@@
 run-time information, and generates a declaration file based on this
-information. To evaluate the tool, 244 declaration files were generated and 
-compared with the declaration file provided on DefinitelyTyped.
-33 files out of 244 had no differences.
-% but what about the remaining generated files
+information. To evaluate the tool, 244 declaration files were generated and 82 were
+compared with the corresponding declaration file provided on DefinitelyTyped.
+72 files presented positive results, having no differences at all or differences that can be solved by modifying the code examples accordingly.
+
 \end{abstract}
@@
 \begin{lstlisting}[
-    caption={\textbf{Unexpected JavaScript behavior} - falsy values, \lstinline!typeof!, \lstinline!null! and \lstinline!undefined! operators and type coercion},
+    caption={\textbf{Unintuitive JavaScript behavior} - falsy values, \lstinline!typeof!, \lstinline!null! and \lstinline!undefined! operators and type coercion},
     language=JavaScript,
@@
 null instanceof Object; // false
+"01" < "00100"; // false
 \end{lstlisting}
 % pjt: why unexpected?
-% "01" < "00100"; // false
 
@@
 \label{sec:motivating-example}
-The \NPM{} package \texttt{route-parser} is a simple route parsing,
-matching, and reversing library for
-Javascript\footnote{https://www.npmjs.com/package/route-parser}. It
-has about 35,000 weekly downloads and 221 \NPM{} packages depend on it. If
+The \NPM{} package \texttt{glob-to-regexp} is a simple JavaScript library, which turns a glob expression into a regular expression\footnote{https://www.npmjs.com/package/glob-to-regexp}. It
+has about 6,500,000 weekly downloads and 181 \NPM{} packages depend on it. If
 a developer creates or extends TypeScript code that depends on the
-\texttt{route-parser} library, the TypeScript
+\texttt{glob-to-regexp} library, the TypeScript
 compiler and IDE requires a declaration file for that library to
@@
 use \texttt{dts-generate} to automatically generate a TypeScript
-declaration file for \texttt{route-parser}. The tool downloads the \NPM{} 
+declaration file for \texttt{glob-to-regexp}. The tool downloads the \NPM{} 
 package, runs the examples extracted from its documentation, gathers
@@
 file. The result is 
-shown in \figref{fig:motivating-example-route-parser-vscode} and it is
+shown in \figref{fig:motivating-example-glob-to-regexp-vscode} and it is
 ready for use in a TypeScript project. For example,  Visual Studio's
 code completion runs properly with it
-(\figref{fig:motivating-example-route-parser-vscode}). If the
-\texttt{route-parser} package gets modified in the future, a new declaration
+(\figref{fig:motivating-example-glob-to-regexp-vscode}). If the
+\texttt{glob-to-regexp} package gets modified in the future, a new declaration
 file can be generated automatically using
@@
     \begin{lstlisting}[language=TypeScript]
-export = Route;
-
-declare class Route {
-  constructor(spec: string);
-  match(path: string): object;
-  reverse(params: object): string;
-}
-
-declare namespace Route {
+export = GlobToRegexp;
+
+declare function GlobToRegexp(glob: string, opts: undefined): RegExp;
+declare function GlobToRegexp(glob: string, opts: GlobToRegexp.I__opts): RegExp;
+declare namespace GlobToRegexp {
+  export interface I__opts {
+    'extended': undefined | boolean;
+    'globstar': undefined | boolean;
+    'flags': undefined | string;
+  }
+
 }
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{motivating-example-route-parser-vscode.png}
-  \end{subfigure}
-
-  \caption{\textbf{Declaration file for \texttt{route-parser} generated with \texttt{dts-generate}} - Constructor and methods are correctly identified.
+    \includegraphics[width=0.8\linewidth]{motivating-example-glob-to-regexp-vscode.png}
+  \end{subfigure}
+
+  \caption{\textbf{Declaration file for \texttt{glob-to-regexp} generated with \texttt{dts-generate}} - Interfaces is correctly detected.
+Optional parameters are explicitly declared as \texttt{undefined}.
 Declaration file can be correctly used in Visual Studio Code\footnote{https://code.visualstudio.com}.}
-  \label{fig:motivating-example-route-parser-vscode}
+  \label{fig:motivating-example-glob-to-regexp-vscode}
 \end{figure}
@@
 The declaration file shown in
-\figref{fig:motivating-example-route-parser-vscode} describes a
-package with a single exported class \lstinline{Route}. The
-description of the class comprises the type signature of the
-constructor (line~4) and the methods \lstinline{match} and
-\lstinline{reverse} (lines~5 and~6).
-
-!!!TODO: explain meaning of namespace.
+\figref{fig:motivating-example-glob-to-regexp-vscode} describes a
+package with a single exported function.
+The first parameter is of type \texttt{string} and the second one is an optional interface.
+TypeScript namespaces are used for organizing types declared in a declaration file and avoiding name collisions with other types \cite{typescript-namespaces}.
+For example, the interface \texttt{I\_\_opts} declared in \figref{fig:motivating-example-glob-to-regexp-vscode} is used as \lstinline[language=TypeScript]{globToRegexp.I__opts}, since it is declared within a namespace.
 
@@
 
-% $ ./dts-generate abs
-% $ cat output/abs/index.d.ts 
 \begin{lstlisting}[
     caption={Example for template \textbf{module-function}},
-    language=TypeScript,
+    language=bash,
 	label=code:dts-generate-example,
@@
 ]
+$ ./dts-generate abs
+$ cat output/abs/index.d.ts 
 export = Abs;
@@
 \end{lstlisting}
-!!!TODO: example for simple \textbf{module}
-
-The \texttt{route-parser} library is an instance of a
-\textbf{module-class}. The \texttt{abs} library provides a single
-function \lstinline{Abs} so it uses template \textbf{module-function}
-as shown in \ref{code:dts-generate-example}.
+
+TypeScript provides a guide on how to write high quality declaration files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/by-example.html}.
+They explain the main concepts through examples.
+We extracted one example for each template and generated used \texttt{dts-generate} to generate a corresponding declaration file, as presented in \figref{fig:typescript-templates-by-example}.
+The \texttt{glob-to-regexp} library is an instance of a \textbf{module-function}.
+It is worth mentioning that templates are a way of describing how a JavaScript library is intended to be used.
+Although highly unlikely, it is possible to create a library that can be represented with a different template depending on how it is used, as shown in \figref{fig:multi-purpose-module-example}.
+Run-time analysis proves useful in this case, since it is not sufficient to statically analyze the exported module.
+
+\begin{figure}[tp]
+  \centering
+  \begin{subfigure}{0.48\linewidth}
+    \begin{lstlisting}[language=JavaScript]
+var greet = require("./greet-settings-module");
+
+greet({
+  greeting: "hello world",
+  duration: 4000
+});
+
+greet({
+  greeting: "hello world",
+  color: "#00ff00"
+});
+    \end{lstlisting}
+    \caption{TypeScript example for \texttt{module-function} template}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{0.48\linewidth}
+    \begin{lstlisting}[language=TypeScript]
+export = GreetSettingsModule;
+
+declare function GreetSettingsModule(settings: GreetSettingsModule.I__settings): void;
+declare namespace GreetSettingsModule {
+  export interface I__settings {
+    'greeting': string;
+    'duration': number | undefined;
+    'color': undefined | string;
+  }
+
+}
+    \end{lstlisting}
+    \caption{Declaration file for \texttt{module-function} template}
+  \end{subfigure}
+
+  \begin{subfigure}{0.48\linewidth}
+      \begin{lstlisting}[language=JavaScript]
+var myLib = require("./greet-module");
+
+var result = myLib.makeGreeting("hello, world");
+console.log("The computed greeting is: " + result);
+
+var goodbye = myLib.makeGoodBye();
+console.log("The computed goodbye is: " + goodbye);    
+      \end{lstlisting}
+      \caption{Example for \texttt{module}}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.48\linewidth}
+      \begin{lstlisting}[language=TypeScript]
+export function makeGreeting(str: string): string;
+export function makeGoodBye(): string;        
+      \end{lstlisting}
+      \caption{Declaration file for \texttt{module} template}
+    \end{subfigure}
+
+    \begin{subfigure}{0.48\linewidth}
+      \begin{lstlisting}[language=JavaScript]
+var Greeter = require("./greet-classes-module.js");
+
+var myGreeter = new Greeter("hello, world");
+myGreeter.greeting = "howdy";
+myGreeter.showGreeting();
+      \end{lstlisting}
+      \caption{Example for \texttt{module-class}}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.48\linewidth}
+      \begin{lstlisting}[language=TypeScript]
+export = Greeter;
+
+declare class Greeter {
+  constructor(message: string);
+  showGreeting(): void;
+}
+
+declare namespace Greeter {
+}
+      \end{lstlisting}
+      \caption{Declaration file for \texttt{module-class} template}
+    \end{subfigure}
+
+  \caption{Results for \textbf{module-function} template}
+  \label{fig:typescript-templates-by-example}
+\end{figure}
+
+\begin{figure}[tp]
+  \centering
+  \begin{subfigure}{0.80\linewidth}
+      \begin{lstlisting}[language=JavaScript]
+var multiPurposeModule = require('./multi-purpose-module');
+
+// module-function
+console.log(multiPurposeModule("John", "Doe"));
+// John Doe
+
+// module-class
+var m = new multiPurposeModule("Jane", "Doe");
+console.log(m.firstName + " - " + m.lastName);
+// Jane - Doe
+
+console.log(m.sayHello());
+// Hello, my name is Jane Doe
+
+// module
+console.log(multiPurposeModule.anotherFunction());
+// I am another function! 
+      \end{lstlisting}
+      \caption{index.js}
+    \end{subfigure}
+
+    \begin{subfigure}{0.80\linewidth}
+      \begin{lstlisting}[language=TypeScript]
+var multiPurposeModule = function(firstName, lastName) {
+  this.firstName = firstName;
+  this.lastName = lastName;
+
+  return firstName + " " + lastName;
+};
+
+multiPurposeModule.prototype.sayHello = function() {
+  return "Hello, my name is " + this.firstName + " " + this.lastName;
+};
+
+multiPurposeModule.anotherFunction = function() {
+  return "I am another function!";
+};
+
+module.exports = multiPurposeModule;   
+      \end{lstlisting}
+      \caption{multi-purpose-module/index.js}
+    \end{subfigure}
+  \caption{Example module that implements all three templates simultaneously}
+  \label{fig:multi-purpose-module-example}
+\end{figure}
 
@@
 at all.
+Moreover, developers try to cover edge cases or even not allowed values in their tests.
+Such tests are not really useful for describing a library from the consumer's point of view.
+For example, a developer might write a test invoking a function with \texttt{null} just to validate that it throws an error in such scenario.
 
@@
 for \NPM{} packages, which are generally created to solve a specific
-problem of JavaScript development.
+problem of JavaScript development. These code fragments do not present the problem of option~\ref{item:2} since they are meant to be used by developers using the package.
 
@@
 
-For example, in the case of \texttt{route-parser} ...
+The extracted example for the case of \texttt{glob-to-regexp} is shown in \coderef{code:code-example-extracted-motivating-example}.
+
+\begin{lstlisting}[
+  caption={Extracted example for \texttt{glob-to-regexp}},
+  language=JavaScript,
+label=code:code-example-extracted-motivating-example,
+  float,
+  abovecaptionskip=-\medskipamount
+]
+var globToRegExp = require('glob-to-regexp');
+var re = globToRegExp("p*uck");
+re.test("pot luck"); // true
+re.test("pluck"); // true
+re.test("puck"); // true
+
+re = globToRegExp("*.min.js");
+re.test("http://example.com/jquery.min.js"); // true
+re.test("http://example.com/jquery.min.js.map"); // false
+
+re = globToRegExp("*/www/*.js");
+re.test("http://example.com/www/app.js"); // true
+re.test("http://example.com/www/lib/factory-proxy-model-observer.js"); // true
+
+// Extended globs
+re = globToRegExp("*/www/{*.js,*.html}", { extended: true });
+re.test("http://example.com/www/app.js"); // true
+re.test("http://example.com/www/index.html"); // true
+\end{lstlisting}
 
@@
 aggregating Jalangi's raw event information to get an application
-specific data representation. The run-time information is saved as a
-JSON file that can be used for later processing. The tool is written
-in JavaScript and runs in \NodeJS{} in a Docker container.
+specific data representation.
+Each function invocation is stored as a \texttt{FunctionContainer} which contains an \texttt{ArgumentContainer} for each argument.
+Each \texttt{ArgumentContainer} contains the name and index of the argument and a collection of \texttt{interactions}.
+The interaction \texttt{getField} gets triggered whenever a property of an object gets accessed and it tracks the name of the accessed field.
+The invocation of an object's method is tracked with the \texttt{methodCall} interaction.
+The \texttt{usedAsArgument} interaction gets triggered when an argument is used in the invocation of a function.
+The \texttt{followingInteractions} property is used for inferring nested interfaces.
+It is an array that recursively records \texttt{interactions} on the return value of \texttt{getField} or \texttt{methodCall}.
+
+We wrap each function's argument around a wrapper object, which enables to store meta-information and greatly simplifies the mapping of an observation with the corresponding \texttt{ArgumentContainer} or \texttt{interaction}.
+We then use the original values for critical operators such as \texttt{===} or \texttt{typeof}, for which wrapper objects will definitely modify the behavior of the code.
+
+The property \texttt{requiredModule} stores the name of the module that declared the invoked function.
+If a function is explicitly exported by the module, the property \texttt{isExported} will be set to \texttt{true} when it is invoked.
+If a function of an exported object is invoked, \texttt{isExported} will be \texttt{false} and \texttt{requiredModule} will contain the name of the required module.
+
+The run-time information is saved as a JSON file that can be used for later processing. The tool is written in JavaScript and runs in \NodeJS{} in a Docker container.
 
@@
 
-!!!TODO too vague: more information --- how do you choose between the templates?
-
-We use different fields (which?) from the run-time information to detect
-how the module is used to choose the most suitable template. The implemented
-templates are \textbf{module}, \textbf{module-class} and
-\textbf{module-function}.
+Instead of analyzing the shape of the exported module, we inspect how the module is used to choose between the templates.
+We analyze the properties \texttt{isExported}, \texttt{requiredModule} and \texttt{isConstructor} from the run-time JSON to distinguish between \texttt{module-class} and \texttt{module-function}.
+If a function is invoked and it is imported from the module we are analyzing we infer the template \texttt{module-class} or \texttt{module-function}.
+We choose \texttt{module-class} if the function is used as a constructor.
+If not, we choose \texttt{module-function}.
+Otherwise, we infer the \texttt{module} template.
+\figref{fig:dts-generate-choose-templates} exposes this logic using the example provided in \figref{fig:multi-purpose-module-example}.
+
+\begin{figure}[tp]
+  \centering
+  \begin{subfigure}{0.48\linewidth}
+    \begin{lstlisting}[language=JavaScript]
+var multiPurposeModule = require('./multi-purpose-module');
+
+multiPurposeModule("John", "Doe");
+// John Doe
+    \end{lstlisting}
+    \caption{Example module \texttt{multi-purpose-module} used as \texttt{module-function}}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{0.48\linewidth}
+    \begin{lstlisting}[language=TypeScript]
+export = MultiPurposeModule;
+
+declare function MultiPurposeModule(firstName: string, lastName: string): string;
+    \end{lstlisting}
+    \caption{Generated declaration file with \texttt{module-function} template}
+  \end{subfigure}
+
+  \begin{subfigure}{0.48\linewidth}
+    \begin{lstlisting}[language=JavaScript]
+var multiPurposeModule = require('./multi-purpose-module');
+
+var m = new multiPurposeModule("Jane", "Doe");
+var fullName = m.firstName + " - " + m.lastName;
+// Jane - Doe
+
+m.sayHello();
+// Hello, my name is Jane Doe
+    \end{lstlisting}
+    \caption{Example module \texttt{multi-purpose-module} used as \texttt{module-class}}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.48\linewidth}
+      \begin{lstlisting}[language=TypeScript]
+export = MultiPurposeModule;
+
+declare class MultiPurposeModule {
+  constructor(firstName: string, lastName: string);
+  sayHello(): string;
+}
+
+declare namespace MultiPurposeModule {
+}
+      \end{lstlisting}
+      \caption{Generated declaration file with \texttt{module-class} template}
+    \end{subfigure}
+
+    \begin{subfigure}{0.48\linewidth}
+      \begin{lstlisting}[language=JavaScript]
+var multiPurposeModule = require('./multi-purpose-module');
+
+multiPurposeModule.anotherFunction();
+// I am another function!
+      \end{lstlisting}
+      \caption{Example module \texttt{multi-purpose-module} used as \texttt{module}}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.48\linewidth}
+      \begin{lstlisting}[language=TypeScript]
+export function anotherFunction(): string;
+      \end{lstlisting}
+      \caption{Generated declaration file with \texttt{module} template}
+    \end{subfigure}
+
+  \caption{Different ways of using example module presented in \figref{fig:multi-purpose-module-example}.
+\texttt{dts-generate} generates infers the correct template for each case.
+The JavaScript implementation of the module remained unmodified for each example.}
+  \label{fig:dts-generate-choose-templates}
+\end{figure}
+
+\begin{figure}[tp]
+  \centering
+  \begin{subfigure}{0.48\linewidth}
+    \begin{lstlisting}[language=TypeScript]
+export function v1(str: string): boolean;
+export function v2(str: string): boolean;
+export function v3(str: string): boolean;
+export function v4(str: string): boolean;
+export function v5(str: string): boolean;
+    \end{lstlisting}
+    \caption{is-uuid/index.d.ts - Generated}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{0.48\linewidth}
+    \begin{lstlisting}[language=TypeScript]
+export function v1(value: string): boolean;
+export function v2(value: string): boolean;
+export function v3(value: string): boolean;
+export function v4(value: string): boolean;
+export function v5(value: string): boolean;
+export function nil(value: string): boolean;
+export function anyNonNil(value: string): boolean;
+    \end{lstlisting}
+    \caption{is-uuid/index.d.ts - DefinitelyTyped}
+  \end{subfigure}
+
+  \caption{Results for \textbf{module}  module \texttt{is-uuid}}
+  \label{fig:experiments-results-module-is-uuid}
+\end{figure}
+
+We implemented only basic TypeScript features and we focused our effort in building and end-to-end solution that generates valid and useful declaration files.
+Only the basic types \texttt{string}, \texttt{number}, \texttt{boolean} and its union were considered both for function parameters and interface properties.
+Common types such as \texttt{any}, arrays, function callbacks, tuples, intersections and features such as generics, index signatures were not implemented.
+Each declaration file was tagged with \texttt{dts-parse} and those containing unimplemented features were ignored.
+We add the \texttt{undefined} type for describing an optional parameter instead of adding the corresponding token \texttt{?}, which is also done by TypeScript in strict null checking mode\cite{typescript-optional-parameters-and-properties}.
 
@@
 
-!!!TODO: insufficient explanation of \texttt{followingInteractions}
---- needs to be mentioned in the previous step
-
 \section{Evaluation}
@@
 After generating a declaration file for an \NPM{} package, we need to
-evaluate its quality. To this end, we compare the generated
-declaration file against the one uploaded to the DefinitelyTyped
-repository for the same module, as shown in
-\figref{fig:evaluation-diagram}. While this approach does not provide
-an absolute measure of quality, it gives us at least an indication of
-the pragmatics of \texttt{dts-generate}: The files on DefinitelyTyped
+evaluate its quality. To this end, we created two tools:
+\begin{itemize}
+  \item \texttt{dts-parse}: A TypeScript declaration files parser, which transforms the file into a JSON file with a defined structure.
+  \item \texttt{dts-compare}: A tool that uses \texttt{dts-parse} and compares two TypeScript declaration files computing differences by applying several criteria.
+\end{itemize}
+
+We applied \texttt{dts-compare} to compare the generated declaration file against the one uploaded to the DefinitelyTyped repository for the same module, as shown in \figref{fig:evaluation-diagram}.
+While this approach does not provide an absolute measure of quality, it gives us at least an indication of the pragmatics of \texttt{dts-generate}: The files on DefinitelyTyped
 are perceived to be useful by the community. If the accuracy of the
@@
 
-!!!TODO: check how many files on DT are just generated by
-\texttt{dts-gen} (e.g., the TypeScript tool)
+%!!!TODO: check how many files on DT are just generated by \texttt{dts-gen} (e.g., the TypeScript tool)
 
@@
 
-To this end, we need a means of comparing two declaration files. Of
+\subsection{dts-parse}
+We need a means of comparing two declaration files. Of
 course, we are not interested in a textual comparison, but in a
-comparison of the structures described by the files. Technically, our
-comparison of two structures A and B yields one of five outcomes:
-\begin{itemize}
-\item A and B are equivalent;
-\item A is more general than B;
-\item B is more general than A;
-\item A and B have a common generalization (i.e., they are
-  compatible);
-\item none of the above holds (i.e., A and B are incompatible).
-\end{itemize}
-
-The first step is to parse declaration files, which can be
-achieved by using the TypeScript Compiler API, a library traverse the
+comparison of the structures described by the files.
+
+The first step is to parse the declaration files, which can be
+achieved by using the TypeScript Compiler API, a library which traverse the
 Abstract Syntax Tree of a TypeScript program
@@
 files as it rejects files with syntactic or semantics errors.
-The output of the parser is a structure where declared
+The output of \texttt{dts-parse} is a structure where declared
 {interfaces}, {functions}, {classes} and
@@
 {classes}, a distinction between the constructor and methods
-is made.
-
-% The tool is called \texttt{parse-dts} and is naturally written in
-% TypeScript. It also runs in \NodeJS{} in its corresponding Docker
-% container.
-
-It remains to compare the parsed representations of two declaration
-files.
-% It was discovered that the implementation was easier when
-% focusing on each template independently.
-As an example, we discuss the implementation of comparison for
-the \textbf{module-function} template.
-
-The following criteria were applied:
+is made. Declaration files are tagged according to their features, which is useful for identifying and filtering out declaration files containing unimplemented features.
+
+\texttt{dts-parse} is itself written in TypeScript. It receives the file as input parameter and returns a JSON, as shown in \coderef{code:dts-parse-example}.
+
+\begin{lstlisting}[
+  caption={Example for \texttt{dts-parse} executed on module \texttt{glob-to-regexp} and combined with \texttt{jq} to browse through the returned JSON},
+  language=bash,
+label=code:dts-parse-example,
+  float,
+  abovecaptionskip=-\medskipamount
+]
+$ dts-parse -i glob-to-regexp/index.d.ts | jq '.parsing.functions[0].name'
+"GlobToRegexp"
+\end{lstlisting}
+
+\subsection{dts-compare}
+\label{sec:dts-compare}
+The tool is written in TypeScript and receives 2 declaration files as an input and returns a JSON containing the result of the comparison, as shown in \coderef{code:dts-compare-example}.
+Following the naming convention of any standard Unit Testing framework, the input files are flagged as \lstinline{expected} or \lstinline{actual}.
+We assigned the \lstinline{expected} flag to the DefinitelyTyped declaration file.
+
+\begin{lstlisting}[
+  caption={Example for \texttt{dts-compare} executed on declaration files without differences},
+  language=bash,
+label=code:dts-compare-example,
+  float,
+  abovecaptionskip=-\medskipamount
+]
+$ dts-compare -e expected.d.ts -a actual.d.ts --module-name "my-module"
+{
+    "module": "my-module",
+    "template": "module",
+    "differences": [],
+    "tags": []
+}
+\end{lstlisting}
+
+\texttt{dts-compare} applies \texttt{dts-parse} internally to both files to get a common structure eligible for comparison.
+The result of the comparison is an array of an abstract entity called \texttt{Difference}.
+We extracted code examples written by developers to execute the code and gather run-time information.
+The quality of the generated declaration file is heavily dependent on the code examples.
+For example, if a property of an interface does not get accessed due to a code example not exploring a particular execution path, that property will not exist in the generated declaration file.
+Consequently, in order to isolate the impact of incomplete code examples on our quality analysis we identified each difference as follows:
 \begin{itemize}
-\item Number of declared functions: Checks the number of declared
-  and exported functions for each of both declaration files. 
-\item Name of declared functions: Checks whether both of the
-  declaration files declared a function with the same name. 
-\item Number of parameters: Checks the number of parameters of the
-  declared functions. This is checked for optional and non-optional
-  parameters.
-
-  !!!TODO: what about parameter names? what about object
-  types/interface types of parameters?
-\item Interfaces: Checks the number of declared interfaces and the
-  fields within those interfaces. 
-\item Errors: Indicates whether there are errors in the declaration files.
+  \item \texttt{solvable}: Differences that can be solved by expanding the code examples. For example, invoking a function with different parameters so that a missing interface property gets accessed.
+  \item \texttt{unsolvable}: Differences that can not be solved are marked as \texttt{unsolvable}. These differences are implementation errors of \texttt{dts-generate}.
 \end{itemize}
 
-!!!TODO: too informal. we need to define an ordering on interfaces etc.
+The concrete evaluated differences are as follows:
+\begin{itemize}
+  \item \texttt{TemplateDifference}: A difference between the formal TypeScript templates.
+For example, if the file in DefinitelyTyped is written with \texttt{module} and we generated a file using \texttt{module-function}.
+The comparison stops if the templates differ.
+  \item \texttt{ExportAssignmentDifference}: An equality check between the export assignments of both files, that is in the expression \lstinline{export = XXX}.
+  \item \texttt{FunctionMissingDifference}: A function is present in the DefinitelyTyped file but not in the generated one. This applies to functions and methods of classes and interfaces.
+  \item \texttt{FunctionMissingDifference}: A function is present in the generated declaration file but not in DefinitelyTyped.
+  \item \texttt{FunctionOverloadingDifference}: The number of declarations for the same functions is different.
+This is particularly common when developers declare a function multiple times with different single types instead of using union types.
+  \item \texttt{ParameterMissingDifference}: A parameter of a function or a property of an interface is not present in the generated file.
+  \item \texttt{ParameterExtraDifference}: A parameter of a function or a property of an interface is present in the generated file but not in the DefinitelyTyped file.
+  \item \texttt{ParameterTypeDifference}: A parameter of a function or a property of an interface is generated with a different type than in the DefinitelyTyped file.
+Here we differentiate between \texttt{SolvableDifference} and \texttt{UnsolvableDifference}.
+  \begin{itemize}
+    \item \texttt{SolvableDifference}: A type difference that can be solved by writing better code examples.
+This includes a basic type that is converted to a union type, a function overloading or a parameter or property that is marked as optional.
+    \item \texttt{UnsolvableDifference}: Anything not considered as \texttt{SolvableDifference}.
+  \end{itemize}
+\end{itemize}
+
+Type aliases are invisible to \texttt{dts-compare} as only final types were considered: the declaration \lstinline{type T = string | number; declare function F(a: T);} is equivalent to \lstinline{declare function F(a: string | number);}.
+The same concept applies for literal interfaces.
+\texttt{dts-compare} does not contemplate differences in function return types, since interface inference of function return types is not covered in this version of \texttt{dts-generate}.
 
@@
 \label{sec:results}
-
-!!!TODO: first we need to say, what we want to achieve with the
-experimental evaluation
+We analyzed the obtained results focusing on the following aspects:
+\begin{itemize}
+  \item The quality of the inferred types, interfaces and module structure using runtime information.
+  \item The benefits of using code examples provided by developers as a first approximation to execute the libraries and avoid a cold start problem.
+  \item The usability of the generated declaration file and whether \texttt{dts-generate} can be used in a proper development environment.
+\end{itemize}
 
@@
 Figure \ref{fig:experiments-overall-funnel} shows that a declaration
-file was generated for 244 modules out of 6029 modules. Samples of the
+file was generated for 244 modules out of 6029 modules and we identified 80 modules that have only the features implemented by \texttt{dts-generate}.
+We obtained positive results for 69 of them.
+A detailed explanation of the overall quality of the generated files is provided in \secref{sec:experiments-evaluation}.
+Samples of the
 generated declaration files for templates \textbf{module},
@@
 
-!!!TODO: the remaining two stages ``run-time information'' and
-``generated declaration files'' need to be explained
+\paragraph*{Run-time Information}
+Run-time information was extracted for only 442 out of 1448 modules with working code examples.
+As explained, in order to extract the run-time information, the behavior of the code under analysis was explicitly modified by wrapping the arguments around Wrapper Objects.
+Furthermore, Jalangi’s instrumentation itself caused some executions to fail, since the modules contained JavaScript features that are not supported by Jalangi.
+As a result, run-time information could not be extracted for 1006 modules.
+An instrumentation without user defined analysis modules was not applied, so it was not possible to determine which modules were failing only because of Jalangi’s own limitations.
+
+\paragraph*{Generated Declaration Files}
+A declaration file was generated for 244 out of 442 modules.
+Despite the correct execution of instrumented code examples, 198 modules did not contain suitable run-time information for generating a declaration file.
+The extracted code examples for these modules did not execute the library itself and therefore the collected run-time information was not useful for generating a declaration file.
 
@@
 \texttt{dirname-regex}, and \texttt{escape-html}. All of them were
-generated using the \textbf{module-function} template. There are no
-differences between the generated files and the corresponding
-declaration files on DefinitelyTyped.
-
-The left side of the figure shows the generated declaration file with
+generated using the \textbf{module-function} template. Left side of the figure shows the generated declaration file with
 \lstinline{dts-generate}, the right side shows the corresponding file
-in the DefinitelyTyped repository. Functions are correctly detected; input
-and output types are accurately inferred.
-
-!!!TODO: explain, what about
-\begin{itemize}
-\item uppercase/lowercase in dirname-regex?
-\item extra namespace in escape-html?
-\end{itemize}
+in the DefinitelyTyped repository. There are no relevant
+differences between the files. Functions are correctly detected; input and output types are accurately inferred.
+
+Differences in the export assignment by modules \texttt{dirname-regex} and \texttt{escape-html} is not problematic since the exported module can receive a custom name given by the developer.
+\texttt{dts-generate} automatically generates the export assignment by transforming the module name into camel case form, following TypeScript guidelines.
+The parameter names are extracted from the variable names of the JavaScript code.
+The JavaScript concrete implementation of module \texttt{escape-html} declares the first parameter of the exported function with the name \texttt{string} and not \texttt{text}, as specified in DefinitelyTyped.
+A difference in the names of the function parameters will anyway not affect the code compilation in any way.
+Furthermore, the omitted namespace in module \texttt{escape-html} is irrelevant.
+It is actually not required, since there is nothing declared within the namespace.
 
@@
     \label{fig:experiments-results-module-function}
-  \end{figure}
-
-Templates of type \textbf{module-class} are shown for modules
-\texttt{flake-idgen}, \texttt{route-parser}, and \texttt{timer-machine}
-in \figref{fig:experiments-results-module-class-flake-idgen} and
-\figref{fig:experiments-results-module-class-timer-machine},
-respectively. Properties of interfaces and class methods are correctly
-generated. Optional parameters are not detected, as it was not
-considered for the implementation.
-
-For \texttt{flake-idgen},  the parameters of interface
-\texttt{ConstructorOptions} are correctly detected. The name of the
-interface differs because it is automatically generated based on the
-name of the argument variable. Optional properties were not
-implemented, which explains the type \lstinline{undefined} for some
-properties. Analogously, the callback \texttt{cb} is inferred as
-\lstinline{undefined}.
-
-!!!TODO:
-\begin{itemize}
-\item why is the callback undefined? is it due to the example? (what does it
-  look like?) why is it called \texttt{cb}?
-\item the export / class is named differently; doesn't that matter?
-\end{itemize}
-
-For \texttt{timer-machine},  the parameter \lstinline{started} is
-inferred as \lstinline{undefined} instead of marking it as
-optional. Methods that were not executed cannot appear in the
-generated declaration file. For that reason, a type corresponding to
-\lstinline{TimerEvent} is not generated, either.
-
-\begin{figure}[tp]
-    \centering
-    \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
-export = FlakeIdgen;
-
-declare class FlakeIdgen {
-    constructor(options: FlakeIdgen.I__options);
-    next(cb: undefined): Buffer;
-}
-
-declare namespace FlakeIdgen {
-    export interface I__options {
-        'id': undefined;
-        'datacenter': number;
-        'worker': number;
-        'epoch': undefined;
-        'seqMask': undefined;
-    }
-}
-      \end{lstlisting}
-      \caption{flake-idgen/index.d.ts - Generated}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
-interface ConstructorOptions {
-    datacenter?: number;
-    worker?: number;
-    id?: number;
-    epoch?: number;
-    seqMask?: number;
-}
-
-declare namespace FlakeId { }
-
-declare class FlakeId {
-    constructor(options?: ConstructorOptions);
-    next(callback?: (err: Error, id: Buffer) => void): Buffer;
-}
-
-export = FlakeId;        
-      \end{lstlisting}
-      \caption{flake-idgen/index.d.ts - DefinitelyTyped}
-    \end{subfigure}
-
-    \caption{Results for \textbf{module-class} module \texttt{flake-idgen}}
-    \label{fig:experiments-results-module-class-flake-idgen}
-\end{figure}
-
-\begin{figure}[tp]
-    \centering
-    \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
-export = Timer;
-
-declare class Timer {
-    constructor(start: undefined);
-    start(): boolean;
-    isStopped(): boolean;
-    emit(): boolean;
-    stop(): boolean;
-    isStarted(): boolean;
-    timeFromStart(): number;
-    time(): number;
-}
-
-declare namespace Timer {
-}
-      \end{lstlisting}
-      \caption{timer-machine/index.d.ts - Generated}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
-export as namespace Timer;
-export = Timer;
-
-declare namespace Timer {
-    type TimerEvent = "start" | "stop" | "time";
-}
-
-declare class Timer {
-    static get(reference: string): Timer;
-    static destroy(reference: string): Timer;
-
-    constructor(started?: boolean);
-
-    isStarted(): boolean;
-    isStopped(): boolean;
-    start(): void;
-    timeFromStart(): number;
-    stop(): void;
-    time(): number;
-    toggle(): void;
-    emitTime(): void;
-    valueOf(): number;
-    on(event: Timer.TimerEvent, callback?: () => void): void;
-}
-      \end{lstlisting}
-      \caption{timer-machine/index.d.ts - DefinitelyTyped}
-    \end{subfigure}
-
-    \caption{Results for \textbf{module-class} module \texttt{timer-machine}}
-    \label{fig:experiments-results-module-class-timer-machine}
 \end{figure}
@@
 
-Finally, a sample for the \textbf{module} template
+A sample for the \textbf{module} template
 for the \texttt{is-uuid} module is shown in
@@
 
+Finally, we could not generate a correct declaration file with the template \texttt{module-class}.
+All the modules corresponding to this template had at least one of the unimplemented TypeScript features.
+
 It is worth mentioning that for some libraries the declaration file in
 DefinitelyTyped was not correct. For example, for the module
-\texttt{datadog-metrics}, some properties of an interface were
-included in the generated declaration file but they were not present
-in the declaration presented in DefinitelyTyped. However, as shown in
-\figref{fig:experiments-results-module-datadog-metrics}, the
-properties \texttt{aggregator} and \texttt{reporter} are not present
-in the DefinitelyTyped version, but they appear in the generated
-declaration file. However, they are indeed used by the library, as
-exposed in lines~2 and~3 of the library's source code shown in
-\figref{fig:subfloat-datadog-logger}. 
+\texttt{glob-base}, the parameter \texttt{basePath} is declared as optional in DefinitelyTyped.
+However, when invoking the function without that parameter an Error is thrown at runtime.
+The file generated by \texttt{dts-generate} will never mark the parameter as optional, as shown in \figref{fig:experiments-results-module-glob-base}.
+Function return types interfaces are not inferred by \texttt{dts-generate}.
+Parameter name is also inferred as \texttt{pattern} instead of \texttt{basePath}, since the JavaScript implementation declares the parameter as \texttt{pattern}, as shown in \figref{fig:subfloat-globbase-js-implementation}.
+We even discovered errors in the selected template in DefinitelyTyped.
+The module \texttt{smart-truncate} in DefinitelyTyped uses the \texttt{module} template, but \texttt{dts-generate} generates a file using the \texttt{module-function} template.
+We see in \coderef{code:experiments-results-module-smart-truncate} that the module is indeed exported as a function, as inferred by \texttt{dts-generate}.
+
 \begin{figure}[tp]
@@
       \begin{lstlisting}[language=TypeScript]
-export interface I__opts {
-    'aggregator': undefined;
-    'defaultTags': Array<any>;
-    'reporter': undefined;
-    'apiKey': string;
-    'appKey': undefined;
-    'agent': undefined;
-    'host': string;
-    'prefix': string;
-    'flushIntervalSeconds': number;
-}
-
-export class BufferedMetricsLogger {
-    constructor(opts: I__opts);
+export = GlobBase;
+
+declare function GlobBase(pattern: string): object;
+      \end{lstlisting}
+      \caption{glob-base/index.d.ts - Generated}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.48\linewidth}
+      \begin{lstlisting}[language=TypeScript]
+declare function globbase(basePath?: string): globbase.GlobBaseResult;
+
+declare namespace globbase {
+    interface GlobBaseResult {
+        base: string;
+        isGlob: boolean;
+        glob: string;
+    }
+}
+
+export = globbase;
+      \end{lstlisting}
+      \caption{glob-base/index.d.ts - DefinitelyTyped}
+    \end{subfigure}
+
+    \begin{subfigure}{0.48\linewidth}
+      \begin{lstlisting}[language=JavaScript]
+var globBase = require('glob-base');
+
+globBase();
+
+// Throws: TypeError: glob-base expects a string.
+      \end{lstlisting}
+      \caption{Code example invoking the exported function without the first parameter, as specified in DefinitelyTyped}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.48\linewidth}
+      \begin{lstlisting}[language=TypeScript]
+'use strict';
+
 // ...
-}
-      \end{lstlisting}
-      \caption{datadog-metrics/index.d.ts - Generated}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
-export interface LoggerOptions {
-    apiKey?: string;
-    appKey?: string;
-    defaultTags?: string[];
-    flushIntervalSeconds?: number;
-    host?: string;
-    prefix?: string;
-}
-
-export class BufferedMetricsLogger {
-    constructor(
-        options: LoggerOptions
-    );
+
+module.exports = function globBase(pattern) {
+  if (typeof pattern !== 'string') {
+    throw new TypeError('glob-base expects a string.');
+  }
+
   // ...
@@
       \end{lstlisting}
-      \caption{datadog-metrics/index.d.ts - DefinitelyTyped}
-    \end{subfigure}
-
-    \begin{subfigure}{0.80\linewidth}
-        \begin{lstlisting}[language=TypeScript]
-function BufferedMetricsLogger(opts) {
-    this.aggregator = opts.aggregator || new Aggregator(opts.defaultTags);
-    this.reporter = opts.reporter || new DataDogReporter(opts.apiKey, opts.appKey, opts.agent);
-    this.host = opts.host;
-    this.prefix = opts.prefix || '';
-    this.flushIntervalSeconds = opts.flushIntervalSeconds;
-
-    // ...
-    // ...
-}
-        \end{lstlisting}
-        \caption{datadog-metrics/logger.js}
-        \label{fig:subfloat-datadog-logger}
-      \end{subfigure}
-
-    \caption{Missing properties for module \texttt{datadog-metrics}}
-    \label{fig:experiments-results-module-datadog-metrics}
+      \caption{glob-base/index.js - Fraction of JS implementation that throws an Error when input parameter is not a string}
+      \label{fig:subfloat-globbase-js-implementation}
+    \end{subfigure}
+
+    \caption{Incorrect optional parameter for module \texttt{glob-base}}
+    \label{fig:experiments-results-module-glob-base}
 \end{figure}
 
+\begin{lstlisting}[
+  caption={Code example for \texttt{smart-truncate} module exporting a function},
+label=code:experiments-results-module-smart-truncate,
+  language=JavaScript
+]
+var smartTruncate = require('smart-truncate');
+
+var string = 'To iterate is human, to recurse divine.';
+
+// Append an ellipsis at the end of the truncated string.
+var truncated = smartTruncate(string, 15);
+\end{lstlisting}
+
 \subsection{Evaluation}
-
-!!!TODO: explain what's going on in this subsection
-
-!!!TODO: what's the baseline for the percentages in the paragraph, the
-244 modules considered or all of DT? Ah, you say that in the figure;
-but then the figure is comparing apples with pears (percentages for
-different baselines!).
-
-As shown in
-\figref{fig:experiments-typescript-templates-distribution-definitely-typed},
-20\% of the declaration files in DefinitelyTyped are written using the
-\textbf{module-function}. However, 57\% of the 244 generated
-declaration files are written with the \textbf{module-function}
-template. Additionally, the complexity of evaluating declaration files
-written with the \textbf{module-function} is considerably lower than
-for other templates. The evaluation for templates
-\textbf{module-class} and \textbf{module} was not implemented.
-
-33 out of 116 evaluated modules have no difference with their
-corresponding declaration file in DefinitelyTyped.
-
-!!!TODO: that's a bit lame as a result. See above for a more
-meaningful output.
+\label{sec:experiments-evaluation}
+We analyzed only 82 of the 244 generated files.
+The remaining 162 files contained at least one of the unimplemented TypeScript features and were therefore not considered.
+These features were identified by \texttt{dts-parse} as tags.
+
+As shown in \figref{fig:experiments-results-module-is-uuid}, the code examples have a direct influence on the quality of the generated declaration file.
+We introduced the concept of \texttt{solvable} and \texttt{unsolvable} differences in \secref{sec:dts-compare}.
+Modules containing only \texttt{solvable} differences were considered as positive results.
+We identified 72 such modules.
+We completed the examples manually for 10 of those modules, making them equivalent to the DefinitelyTyped files, as shown in \figref{fig:experiments-results-manually-completed-examples}.
+Expected differences are present: optional parameters are declared both with the actual and \texttt{undefined} type and the union type is handled through multiple overloads.
 
@@
   \centering
-	\begin{tikzpicture}
-		\begin{axis}[
-			ybar,
-            width=0.7\textwidth,
-            scale=0.8,
-			ybar=0pt,
-			ymax=100,
-			enlargelimits=0.15,
-			bar width=0.04\textwidth,
-			legend columns=1,
-			legend cell align={left},
-			legend style={
-                at={(0.6,0.95)},
-                anchor=north,
-                nodes={scale=0.8, transform shape}
-            },
-			symbolic x coords={module-class,module-function,module},
-			xtick=data,
-			ylabel=Percentage,
-			nodes near coords, 
-			nodes near coords align={vertical}
-		]
-		\addplot coordinates {
-			(module-class, 8)
-			(module-function, 20)
-			(module, 72)
-		};
-
-		\addplot coordinates {
-			(module-class, 12)
-			(module-function, 57)
-			(module, 31)
-		};
-		\legend{DefinitelyTyped (Total: 6029), Generated (Total: 244)}
-		\end{axis}
-	\end{tikzpicture}
-
-	\caption[TypeScript templates distribution | Generated \& DefinitelyTyped]{\textbf{TypeScript templates distribution | Generated \& DefinitelyTyped} - Out of a total of 6029, 72\% of the modules uploaded to the DefinitelyTyped repository use the \textbf{module} template and only 20\% use the \textbf{module-function} one.
-However, 57\% of the 244 generated declaration files use the \textbf{module-function} template.
-	}
-
-	\label{fig:experiments-typescript-templates-distribution-definitely-typed}
+  \begin{subfigure}{0.80\linewidth}
+    \begin{lstlisting}[language=TypeScript]
+declare namespace gh {
+  interface Options {
+      enterprise?: boolean;
+  }
+  interface Result {
+      user: string;
+      repo: string;
+      branch: string;
+      https_url: string;
+      tarball_url: string;
+      clone_url: string;
+      travis_url: string;
+      api_url: string;
+      zip_url: string;
+  }
+}
+
+declare function gh(url: string | {url: string}, options?: gh.Options): gh.Result | null;
+
+export = gh;    
+    \end{lstlisting}
+    \caption{github-url-to-object/index.d.ts - DefinitelyTyped}
+  \end{subfigure}
+
+  \begin{subfigure}{0.80\linewidth}
+    \begin{lstlisting}[language=TypeScript]
+export = GithubUrlToObject;
+
+declare function GithubUrlToObject(repoUrl: string|GithubUrlToObject.I__repoUrl, opts: undefined): object;
+declare function GithubUrlToObject(repoUrl: string|GithubUrlToObject.I__repoUrl, opts: GithubUrlToObject.I__opts): null|object;
+declare function GithubUrlToObject(repoUrl: GithubUrlToObject.I__repoUrl, opts: undefined): object;
+declare namespace GithubUrlToObject {
+  export interface I__repoUrl {
+    'url': undefined | string;
+  }
+
+  export interface I__opts {
+    'enterprise': undefined | boolean;
+  }
+}
+    \end{lstlisting}
+    \caption{github-url-to-object/index.d.ts - Generated with manually expanded code example}
+  \end{subfigure}
+
+  \begin{subfigure}{0.80\linewidth}
+    \begin{lstlisting}[language=JavaScript]
+var gh = require('github-url-to-object')
+
+gh('github:monkey/business');
+gh('https://github.com/monkey/business');
+gh('https://github.com/monkey/business/tree/master');
+gh('https://github.com/monkey/business/tree/master/nested/file.js');
+gh('https://github.com/monkey/business.git');
+gh('http://github.com/monkey/business');
+gh('git://github.com/monkey/business.git');
+
+// Manually added:
+gh('git+https://githuuub.com/monkey/business.git', {});
+gh('git+https://githuuub.com/monkey/business.git', {enterprise: true});
+gh({url: 'git://github.com/monkey/business.git'});
+
+    \end{lstlisting}
+    \caption{Code example for module \texttt{github-url-to-object}.}
+    \end{subfigure}
+  \caption{Manually expanded code example for \texttt{github-url-to-object} to match DefinitelyTyped declaration file.
+Literal interface in DefinitelyTyped gets replace by a declared interface.
+The \texttt{undefined} type is used instead of the optional type.
+Return types are not considered.}
+  \label{fig:experiments-results-manually-completed-examples}
 \end{figure}
+
+We identified only 10 modules with \texttt{unsolvable} differences.
+Most of them were caused for the same error: 8 modules had an incorrectly declared interface for the \texttt{string} type.
+This interface contained standard properties or methods from the JavaScript \texttt{String} object, such as \texttt{length} or \texttt{indexOf()}.
+The remaining modules are \texttt{duplexer2} and \texttt{duplexer3}.
+Both of them used native NodeJS types within the declaration file and \texttt{dts-generate} could not identify those properly.
 
@@
 We managed to generate declaration files for 244 modules.
-We compared the results against the corresponding files uploaded to DefinitelyTyped by creating a TypeScript declaration files parser and a comparator.
+We compared the results against the corresponding files uploaded to DefinitelyTyped by creating \texttt{dts-parse}, a TypeScript declaration files parser and \texttt{dts-compare}, a comparator.
 

### 2020-07-30 programming-2021/main.tex
ported to programming class

--- 2020-06-26 ecoop-2020/main.tex
@@
-\documentclass[a4paper,english,cleveref, autoref]{lipics-v2019}
-% ECOOP 2020: 25 pages excluding references
-%This is a template for producing LIPIcs articles. 
-%See lipics-manual.pdf for further information.
-%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
-%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
-%for section-numbered lemmas etc., use "numberwithinsect"
-%for enabling cleveref support, use "cleveref"
-%for enabling cleveref support, use "autoref"
+% -*- coding: utf-8; -*-
+\documentclass[english,cleveref,autoref,submission]{programming}
+% Programming:
+% The main part of the paper should not exceed 22 pages (in the
+% provided style of the journal), but there is no limit for
+% bibliography and appendices. The page limit for the main part of the
+% paper is in place in order to keep the paper on focus and to avoid
+% overloading the reviewers. Authors are encouraged to move important
+% details to appendices, which may be consulted by the reviewers. In
+% some cases, if authors feel that the main part requires
+% substantially more pages, they should explain the reasons why in the
+% additional comments field of the submission form; examples of these
+% cases may include papers with substantial source code listings, and
+% essays. Papers whose length is incommensurate with their
+% contribution will be rejected.
+
+\usepackage[backend=biber]{biblatex}
+\addbibresource{main.bib}
 
@@
 
-\bibliographystyle{plainurl}% the mandatory bibstyle
+%%\bibliographystyle{plainurl}% the mandatory bibstyle
 
@@
 
-\author{John Q.
-Public}{Dummy University Computing Laboratory, Country \and My second affiliation, Country \and \url{http://www.myhomepage.edu} }{johnqpublic@dummyuni.org}{https://orcid.org/0000-0002-1825-0097}{(Optional) author-specific funding acknowledgements}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty.
-Please provide at least the name of the affiliation and the country.
-The full address is optional
-
-\authorrunning{J.\,Q. Public}%TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'
-
-\Copyright{John Q. Public}%TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
+\author{Fernando Cristiani}
+\authorinfo{contact \email{fc@example.org}}
+\affiliation{Hochschule Karlsruhe, Germany }
+
+\author{Peter Thiemann}
+\authorinfo{\email{thiemann@informatik.uni-freiburg.de}}
+\affiliation{Universität Freiburg, Germany}
+
+\keywords{JavaScript, TypeScript, Dynamic Analysis, Declaration
+  Files}%TODO mandatory; please add comma-separated list of 1--5 keywords
+
+%%%%%%%%%%%%%%%%%%
+%% These data MUST be filled for your submission. (see 5.3)
+\paperdetails{
+  %% perspective options are: art, sciencetheoretical, scienceempirical, engineering.
+  %% Choose exactly the one that best describes this work. (see 2.1)
+  perspective=science-empirical,
+  %% State one or more areas, separated by a comma. (see 2.2)
+  %% Please see list of areas in http://programming-journal.org/cfp/
+  %% The list is open-ended, so use other areas if yours is/are not listed.
+  area={General-purpose programming, Web programming},
+  %% You may choose the license for your paper (see 3.)
+  %% License options include: cc-by (default), cc-by-nc
+  % license=cc-by,
+}
+%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%
+%% These data are provided by the editors. May be left out on submission.
+%\paperdetails{
+%  submitted=2016-08-10,
+%  published=2016-10-11,
+%  year=2016,
+%  volume=1,
+%  issue=1,
+%  articlenumber=1,
+%}
+%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Please go to https://dl.acm.org/ccs/ccs.cfm and generate your Classification
+% System [view CCS TeX Code] stanz and copy _all of it_ to this place.
+%% From HERE
+%%% !!! REVISE
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10002944.10011122.10003459</concept_id>
+<concept_desc>General and reference~Computing standards, RFCs and guidelines</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+<concept>
+<concept_id>10010405.10010476.10010477</concept_id>
+<concept_desc>Applied computing~Publishing</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
 
 \ccsdesc[300]{Software and its engineering~Software notations and tools}
-\ccsdesc[100]{Software and its engineering~General programming languages}%TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm
-
-\keywords{JavaScript, TypeScript, Dynamic Analysis, Declaration Files}%TODO mandatory; please add comma-separated list of keywords
-
-\category{}%optional, e.g. invited paper
-
-\relatedversion{}%optional, e.g. full version hosted on arXiv, HAL, or other repository/website
-%\relatedversion{A full version of the paper is available at \url{...}.}
-
-\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
-
-%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors.
-Please enter author specific funding statements as fifth argument of the \author macro.
-
-% \acknowledgements{I want to thank \dots}%optional
-
-%\nolinenumbers %uncomment to disable line numbering
-
-%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
-
-%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\EventEditors{John Q. Open and Joan R. Access}
-\EventNoEds{2}
-\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
-\EventShortTitle{CVIT 2016}
-\EventAcronym{CVIT}
-\EventYear{2016}
-\EventDate{December 24--27, 2016}
-\EventLocation{Little Whinging, United Kingdom}
-\EventLogo{}
-\SeriesVolume{42}
-\ArticleNo{23}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\ccsdesc[100]{Software and its engineering~General programming languages}
+
+% To HERE
+%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 corresponding JavaScript library gets modified, respectively. They
-rely on an existing static analyzer for JavaScript \cite{TAJS}.
+rely on an existing static analyzer for JavaScript \cite{DBLP:conf/sas/JensenMT09}.
 TypeScript itself
@@
 
-\bibliography{main}
+\printbibliography
+%%\bibliography{main}
 

### 2020-07-30 programming-2021/main.tex
intro

@@
 \label{sec:introduction}
-JavaScript has become the most popular language for writing web
+JavaScript is the most popular language for writing web
 applications \cite{github-statistics}. It is also gaining popularity
@@
 JavaScript was never intended to be more than a
-scripting language. Hence, it lacks features for maintaining and evolving large
-codebases.
-Presently, however, JavaScript is being used for creating large and complex
+scripting language and thus lacks features for maintaining and evolving large
+codebases. However, JavaScript is increasingly used for creating large and complex
 applications. 
@@
 
-The cognitive load produced by such dynamic typing (which includes
-unchecked property names) is not present in other languages that use
+The cognitive load produced by these behaviors is mitigated in languages that use
 build tools based on type information. This insight motivated the
-creation of TypeScript, a superset of JavaScript with type
+creation of TypeScript, a superset of JavaScript with expressive type
 annotations \cite{typescript}. It has become a widely used alternative
@@
 of several kinds of run-time errors and the integration of code intelligence
-tools like auto-completion in an IDE.
-
-Existing JavaScript libraries can be used in a TypeScript project by
+tooling like auto-completion in an IDE.
+
+Despite the advantages, it is unrealistic to expect the world to
+switch to TypeScript in a day. Therefore, 
+existing JavaScript libraries can be used in a TypeScript project by
 adding a declaration file that contains a description of the library's
@@
 more than 6000 JavaScript libraries. Unfortunately, most declaration
-files in this libraries have been manually created and maintained,
+files in this repository have been manually created and maintained,
 which is error prone and time consuming. As TypeScript takes a
 declaration file at face value, discrepancies are not detected at
-compile time and the inaccurate code-intelligence features are misleading the
+compile time and the resulting inaccurate code-intelligence features
+are confusing for the
 programmer. Moreover, TypeScript does not perform any run-time 
-checking on types, either, so that a discrepancy  between the declaration file
+checking of types, either, so that a discrepancy  between the declaration file
 and its corresponding JavaScript library can lead to unexpected 
@@
 
-Some previous work tackled the problem of improving the quality of
-declaration files. Feldthaus et al
+We aim to improve on this situation by automatically generating
+sound TypeScript declaration files. To this end, we rely on the examples provided
+by programmers as part of the documentation of their library. If these
+examples are sufficiently rich, then our toolchain can generate the
+desired declaration files from them. This approach takes advantage of
+best practices in the dynamic languages community which favors
+examples over writing out type signatues.
+
+\begin{description}
+\item[PJT: How about a short example at this point?] 
+\item[PJT: maybe later] Some previous work tackled the problem of improving
+  the quality of declaration files. Feldthaus et al
   \cite{DBLP:conf/oopsla/FeldthausM14} search automatically for
-mismatches between a declaration file and implementation code. TSTest
-\cite{DBLP:journals/pacmpl/KristensenM17}
-adapted feedback-directed random testing to detect discrepancies between a
+  mismatches between a declaration file and implementation
+  code. TSTest \cite{DBLP:journals/pacmpl/KristensenM17} adapted
+  feedback-directed random testing to detect discrepancies between a
   declaration file and a JavaScript library. Tools like TSInfer and
   TSEvolve~\cite{DBLP:conf/fase/KristensenM17} are designed for
-assisting the creation of new declaration 
-files and supporting the evolution the declaration file when the
-corresponding JavaScript library gets modified, respectively. They
-rely on an existing static analyzer for JavaScript \cite{DBLP:conf/sas/JensenMT09}.
-TypeScript itself
-developed \texttt{dts-gen}, a tool that generates a
+  assisting the creation of new declaration files and supporting the
+  evolution the declaration file when the corresponding JavaScript
+  library gets modified, respectively. They rely on an existing static
+  analyzer for JavaScript \cite{DBLP:conf/sas/JensenMT09}.  TypeScript
+  itself developed \texttt{dts-gen}, a tool that generates a
   declaration file that is meant to be used as a \emph{starting point
     for writing a high-quality declaration file} \cite{dts-gen}.
-
+\end{description}
 In this work we present the tool \texttt{dts-generate}

### 2020-08-07 programming-2021/main.tex
section 2, 3 examples

@@
 
+\usepackage{cleveref}
 \usepackage{pgfplots}
@@
 \lstdefinelanguage{JavaScript}{
-  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, undefined},
+  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, undefined,const},
   keywordstyle=\color{black}\bfseries,
@@
 \lstdefinelanguage{TypeScript}{
-  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, undefined},
+  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, undefined,const},
   keywordstyle=\color{black}\bfseries,
@@
 JavaScript is the most popular language for writing web
-applications \cite{github-statistics}. It is also gaining popularity
+applications \cite{github-statistics}. It is also increasingly used
 for back-end applications running in \NodeJS{}, a JavaScript-based
@@
 scripting language and thus lacks features for maintaining and evolving large
-codebases. However, JavaScript is increasingly used for creating large and complex
-applications. 
+codebases. However, nowadays developers create large and complex
+applications in JavaScript. 
 Mistakes such as mistyped property
@@
 \begin{lstlisting}[
-    caption={\textbf{Unintuitive JavaScript behavior} - falsy values, \lstinline!typeof!, \lstinline!null! and \lstinline!undefined! operators and type coercion},
+    caption={\textbf{Unintuitive JavaScript behavior} Falsy values, \lstinline!typeof!, \lstinline!null! and \lstinline!undefined! operators and type coercion},
     language=JavaScript,
 	label=code:introduction-javascript-wtfs,
-    float,
+    float=bp,
     abovecaptionskip=-\medskipamount
@@
 of several kinds of run-time errors and the integration of code intelligence
-tooling like auto-completion in an IDE.
+tools like autocompletion in an IDE.
 
@@
 adding a declaration file that contains a description of the library's
-API in terms of types. Lots of declaration files for popular
-JavaScript libraries have been created in a community effort in the
-DefinitelyTyped repository~\cite{definitely-typed-repository}.
-At the time of writing this repository contains declaration files for
-more than 6000 JavaScript libraries. Unfortunately, most declaration
-files in this repository have been manually created and maintained,
+API in terms of types. 
+The DefinitelyTyped repository~\cite{definitely-typed-repository} has
+been created as a community effort to collect declaration files for
+popular JavaScript libraries. At the time of writing it contains
+declaration files for more than 6000 libraries.
+
+Unfortunately, creation and maintenance for most declaration
+files in DefinitelyTyped is conducted manually,
 which is error prone and time consuming. As TypeScript takes a
 declaration file at face value, discrepancies are not detected at
-compile time and the resulting inaccurate code-intelligence features
-are confusing for the
-programmer. Moreover, TypeScript does not perform any run-time 
-checking of types, either, so that a discrepancy  between the declaration file
+compile time and the resulting inaccuracies in autocompletion, say,
+are confusing for the programmer.
+As TypeScript does not perform any run-time 
+checking of types, either,  discrepancies  between the declaration file
 and its corresponding JavaScript library can lead to unexpected 
-behavior and crashes. Such behavior can lead to developer
+behavior and crashes. Such an experience can lead to developer
 frustration, longish debugging sessions, and decreasing confidence in
@@
 
-We aim to improve on this situation by automatically generating
-sound TypeScript declaration files. To this end, we rely on the examples provided
-by programmers as part of the documentation of their library. If these
-examples are sufficiently rich, then our toolchain can generate the
-desired declaration files from them. This approach takes advantage of
+\subsection{Approach}
+\label{sec:approach}
+
+We aim to improve on this situation by providing a tool that generates
+sound TypeScript declaration files automatically. To do so, we rely on the examples provided
+as part of the documentation of a library. If these
+examples are sufficiently rich, then our toolchain can generate
+high-quality declaration files from them. This approach takes
+advantage of 
 best practices in the dynamic languages community which favors
-examples over writing out type signatues.
-
-\begin{description}
-\item[PJT: How about a short example at this point?] 
-\item[PJT: maybe later] Some previous work tackled the problem of improving
-  the quality of declaration files. Feldthaus et al
-  \cite{DBLP:conf/oopsla/FeldthausM14} search automatically for
-  mismatches between a declaration file and implementation
-  code. TSTest \cite{DBLP:journals/pacmpl/KristensenM17} adapted
-  feedback-directed random testing to detect discrepancies between a
-  declaration file and a JavaScript library. Tools like TSInfer and
-  TSEvolve~\cite{DBLP:conf/fase/KristensenM17} are designed for
-  assisting the creation of new declaration files and supporting the
-  evolution the declaration file when the corresponding JavaScript
-  library gets modified, respectively. They rely on an existing static
-  analyzer for JavaScript \cite{DBLP:conf/sas/JensenMT09}.  TypeScript
-  itself developed \texttt{dts-gen}, a tool that generates a
-  declaration file that is meant to be used as a \emph{starting point
-    for writing a high-quality declaration file} \cite{dts-gen}.
-\end{description}
-In this work we present the tool \texttt{dts-generate}
-as a first step to explore the possibilities for
-generating useful declaration files without the heavy lifting of
-static analysis. \texttt{dts-generate} comes with a framework that
+examples and tests over writing out type signatures.
+
+As an example, consider the NPM module
+\texttt{abs} that
+``computes the absolute path of an input''. Its documentation consists
+of the following three examples\footnote{\url{https://www.npmjs.com/package/abs}}.
+\begin{lstlisting}[language=JavaScript,numbers=none]
+const abs = require("abs");
+
+console.log(abs("/foo"));
+// => "/foo"
+
+console.log(abs("foo"));
+// => "/path/to/where/you/are/foo"
+
+console.log(abs("~/foo"));
+// => "/home/username/foo"
+\end{lstlisting}
+From the examples, our tool generates a TypeScript declaration file,
+which is equivalent to the declaration file provided in
+DefinitelyTyped\footnote{\url{https://github.com/DefinitelyTyped/DefinitelyTyped/blob/master/types/abs/index.d.ts}}: 
+\begin{lstlisting}[language=TypeScript,numbers=none]
+export = Abs;
+declare function Abs(input: string): string;
+\end{lstlisting}
+% \begin{description}
+% \item[PJT: maybe later] Some previous work tackled the problem of improving
+%   the quality of declaration files. Feldthaus et al
+%   \cite{DBLP:conf/oopsla/FeldthausM14} search automatically for
+%   mismatches between a declaration file and implementation
+%   code. TSTest \cite{DBLP:journals/pacmpl/KristensenM17} adapted
+%   feedback-directed random testing to detect discrepancies between a
+%   declaration file and a JavaScript library. Tools like TSInfer and
+%   TSEvolve~\cite{DBLP:conf/fase/KristensenM17} are designed for
+%   assisting the creation of new declaration files and supporting the
+%   evolution the declaration file when the corresponding JavaScript
+%   library gets modified, respectively. They rely on an existing static
+%   analyzer for JavaScript \cite{DBLP:conf/sas/JensenMT09}.  TypeScript
+%   itself developed \texttt{dts-gen}, a tool that generates a
+%   declaration file that is meant to be used as a \emph{starting point
+%     for writing a high-quality declaration file} \cite{dts-gen}.
+% \end{description}
+
+Our tool \texttt{dts-generate} is a first step to explore the possibilities for
+generating useful declaration files from code
+examples. \texttt{dts-generate} comes with a framework that 
 supports the generation of declaration files for an existing
 JavaScript library published to the \NPM{} registry. The tool gathers
-data flow and type information at run-time to generate a declaration
+data flow and type information at run time to generate a declaration
 file based on that information.
@@
 \item
+  %% without the heavy lifting of static analysis
   We do not rely on static analysis, which is hard to implement
   soundly and precisely and which is prone to maintenance problems
-  when keeping up with JavaScript's yearly language updates.
+  when keeping up with JavaScript's frequent language updates.
 \item
@@
 
-The contributions of this paper are as follows.
+\subsection{Contributions}
+\label{sec:contributions}
+
 \begin{itemize}
@@
 
-\item The tool \texttt{dts-generate}, a command line
+\item Design and implementation of the tool \texttt{dts-generate}, a command line
   application that generates a valid TypeScript declaration file for a
@@
 The \NPM{} package \texttt{glob-to-regexp} is a simple JavaScript library, which turns a glob expression into a regular expression\footnote{https://www.npmjs.com/package/glob-to-regexp}. It
-has about 6,500,000 weekly downloads and 181 \NPM{} packages depend on it. If
+has about 9,900,000 weekly downloads and 188 \NPM{} packages depend on it. If
 a developer creates or extends TypeScript code that depends on the
@@
 compiler and IDE requires a declaration file for that library to
-perform static checking and code completion, respectively. We 
-use \texttt{dts-generate} to automatically generate a TypeScript
+perform static checking and code completion, respectively. With
+\texttt{dts-generate} we automatically generate a TypeScript
 declaration file for \texttt{glob-to-regexp}. The tool downloads the \NPM{} 
@@
 ready for use in a TypeScript project. For example,  Visual Studio's
-code completion runs properly with it
+code completion runs properly with the generated file
 (\figref{fig:motivating-example-glob-to-regexp-vscode}). If the
@@
   \centering
-  \begin{subfigure}{0.70\linewidth}
-    \begin{lstlisting}[language=TypeScript]
+%  \begin{subfigure}{0.70\linewidth}
+    \begin{lstlisting}[language=TypeScript,numbers=none]
 export = GlobToRegexp;
 
-declare function GlobToRegexp(glob: string, opts: undefined): RegExp;
-declare function GlobToRegexp(glob: string, opts: GlobToRegexp.I__opts): RegExp;
+declare function GlobToRegexp(glob: string, opts ?: GlobToRegexp.I__opts): RegExp;
 declare namespace GlobToRegexp {
   export interface I__opts {
-    'extended': undefined | boolean;
-    'globstar': undefined | boolean;
-    'flags': undefined | string;
-  }
-
-}
-    \end{lstlisting}
-  \end{subfigure}
-
-  \hspace{10.\textwidth}
-
-  \begin{subfigure}{1.\linewidth}
-    \centering
+    'extended' ?: boolean;
+    'globstar' ?: boolean;
+    'flags'    ?: string;
+  }
+}
+    \end{lstlisting}
+  % \end{subfigure}
+  \begin{center}
     \includegraphics[width=0.8\linewidth]{motivating-example-glob-to-regexp-vscode.png}
-  \end{subfigure}
-
-  \caption{\textbf{Declaration file for \texttt{glob-to-regexp} generated with \texttt{dts-generate}} - Interfaces is correctly detected.
-Optional parameters are explicitly declared as \texttt{undefined}.
-Declaration file can be correctly used in Visual Studio Code\footnote{https://code.visualstudio.com}.}
+  \end{center}
+
+  \caption{\textbf{Declaration file for \texttt{glob-to-regexp}
+      generated with \texttt{dts-generate}} - The interface is
+    detected correctly. Optional parameters are detected. The
+    declaration file is usable in Visual Studio
+    Code\footnote{https://code.visualstudio.com}.} 
   \label{fig:motivating-example-glob-to-regexp-vscode}
@@
 
-After filling in some background information on declaration files,
-Section~\ref{sec:gener-typescr-decl} examines each step in the
-generation process in detail and refer back to this example for
-concreteness.
+% After filling in some background information on declaration files,
+% Section~\ref{sec:gener-typescr-decl} examines each step in the
+% generation process in detail and refer back to this example for
+% concreteness.
 
@@
 
-The declaration file shown in
-\figref{fig:motivating-example-glob-to-regexp-vscode} describes a
-package with a single exported function.
-The first parameter is of type \texttt{string} and the second one is an optional interface.
-TypeScript namespaces are used for organizing types declared in a declaration file and avoiding name collisions with other types \cite{typescript-namespaces}.
-For example, the interface \texttt{I\_\_opts} declared in \figref{fig:motivating-example-glob-to-regexp-vscode} is used as \lstinline[language=TypeScript]{globToRegexp.I__opts}, since it is declared within a namespace.
+The declaration file shown in \figref{fig:motivating-example-glob-to-regexp-vscode}
+describes a package with a single exported function. The first parameter is of type
+\texttt{string} and the second one is an optional interface. TypeScript namespaces
+organize types declared in a declaration file and avoid name clashes with other types
+\cite{typescript-namespaces}. For example, the interface \texttt{I\_\_opts} declared in
+\figref{fig:motivating-example-glob-to-regexp-vscode} is used as
+\lstinline[language=TypeScript]{globToRegexp.I__opts} because it is declared within a
+namespace.
 
@@
 files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/templates.html}:
-\textbf{module}, \textbf{module-class} and
+\textbf{module}, \textbf{module-class}, and
 \textbf{module-function}. 
 Each template corresponds to a different way of describing the exports
-of a JavaScript library. Choosing the template depends on the
-particular organization of the underlying JavaScript library:
+of a JavaScript library. The choice of the template depends on the
+organization of the underlying JavaScript library:
 \begin{description}
@@
 \item[module-class] a class-like structure,
-\item[module-function] exactly one exported functions.
+\item[module-function] exactly one exported function.
 \end{description}
@@
 
-TypeScript provides a guide on how to write high quality declaration files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/by-example.html}.
-They explain the main concepts through examples.
-We extracted one example for each template and generated used \texttt{dts-generate} to generate a corresponding declaration file, as presented in \figref{fig:typescript-templates-by-example}.
-The \texttt{glob-to-regexp} library is an instance of a \textbf{module-function}.
-It is worth mentioning that templates are a way of describing how a JavaScript library is intended to be used.
-Although highly unlikely, it is possible to create a library that can be represented with a different template depending on how it is used, as shown in \figref{fig:multi-purpose-module-example}.
-Run-time analysis proves useful in this case, since it is not sufficient to statically analyze the exported module.
-
 \begin{figure}[tp]
@@
   \begin{subfigure}{0.48\linewidth}
-    \begin{lstlisting}[language=JavaScript]
+    \begin{lstlisting}[language=JavaScript,numbers=none]
 var greet = require("./greet-settings-module");
@@
     \end{lstlisting}
-    \caption{TypeScript example for \texttt{module-function} template}
+    \caption{Example for \texttt{module-function} template}
+    \label{fig:example-module-function}
   \end{subfigure}
@@
   \begin{subfigure}{0.48\linewidth}
-    \begin{lstlisting}[language=TypeScript]
+    \begin{lstlisting}[language=TypeScript,numbers=none]
 export = GreetSettingsModule;
@@
     'greeting': string;
-    'duration': number | undefined;
-    'color': undefined | string;
+    'duration' ?: number;
+    'color' ?: string;
   }
@@
   \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=JavaScript]
+      \begin{lstlisting}[language=JavaScript,numbers=none]
 var myLib = require("./greet-module");
@@
       \end{lstlisting}
-      \caption{Example for \texttt{module}}
+    \caption{Example for \texttt{module} template}
+    \label{fig:example-module}
     \end{subfigure}
@@
     \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
+      \begin{lstlisting}[language=TypeScript,numbers=none]
 export function makeGreeting(str: string): string;
@@
     \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=JavaScript]
+      \begin{lstlisting}[language=JavaScript,numbers=none]
 var Greeter = require("./greet-classes-module.js");
@@
       \end{lstlisting}
-      \caption{Example for \texttt{module-class}}
+      \caption{Example for \texttt{module-class} template}
+      \label{fig:example-class}
     \end{subfigure}
@@
     \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
+      \begin{lstlisting}[language=TypeScript,numbers=none]
 export = Greeter;
@@
 
-  \caption{Results for \textbf{module-function} template}
+  \caption{Example uses and declaration files for different templates}
   \label{fig:typescript-templates-by-example}
@@
 
-\begin{figure}[tp]
-  \centering
-  \begin{subfigure}{0.80\linewidth}
-      \begin{lstlisting}[language=JavaScript]
-var multiPurposeModule = require('./multi-purpose-module');
-
-// module-function
-console.log(multiPurposeModule("John", "Doe"));
-// John Doe
-
-// module-class
-var m = new multiPurposeModule("Jane", "Doe");
-console.log(m.firstName + " - " + m.lastName);
-// Jane - Doe
-
-console.log(m.sayHello());
-// Hello, my name is Jane Doe
-
-// module
-console.log(multiPurposeModule.anotherFunction());
-// I am another function! 
-      \end{lstlisting}
-      \caption{index.js}
-    \end{subfigure}
-
-    \begin{subfigure}{0.80\linewidth}
-      \begin{lstlisting}[language=TypeScript]
-var multiPurposeModule = function(firstName, lastName) {
-  this.firstName = firstName;
-  this.lastName = lastName;
-
-  return firstName + " " + lastName;
-};
-
-multiPurposeModule.prototype.sayHello = function() {
-  return "Hello, my name is " + this.firstName + " " + this.lastName;
-};
-
-multiPurposeModule.anotherFunction = function() {
-  return "I am another function!";
-};
-
-module.exports = multiPurposeModule;   
-      \end{lstlisting}
-      \caption{multi-purpose-module/index.js}
-    \end{subfigure}
-  \caption{Example module that implements all three templates simultaneously}
-  \label{fig:multi-purpose-module-example}
-\end{figure}
+TypeScript provides a guide on how to write high-quality declaration
+files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/by-example.html}. The guide
+explains the main concepts through examples. We selected one example for each template
+from the guide and
+used \texttt{dts-generate} to generate a corresponding declaration file, as shown in
+\figref{fig:typescript-templates-by-example}.
+
+Like \texttt{dts-gen}, a TypeScript package that generates a template for a declaration
+file, we determine the kind of declaration file by examining the provided example. If the
+imported entity is used solely as a function as in Figure~\ref{fig:example-module-function}, the
+generated declaration file follows the module-function template.
+If the example only accesses properties of the imported entity as in
+Figure~\ref{fig:example-module}, we generate according to the module template.
+If the imported entity is used with \lstinline/new/ to create new instances as in
+Figure~\ref{fig:example-class}, we generate a module-class template.
+
+The libraries \texttt{glob-to-regexp} as well as \texttt{abs} are both instances of a
+\textbf{module-function}.
+Most libraries fall in one of these three categories. 
+While it is possible to create a library that can be represented with a different template
+depending on how it is used, our selection is driven by the examples in the documentation and
+hence do reflect the developers intent.
+% Run-time analysis proves useful in this case,
+% since it is not sufficient to statically analyze the exported module.
+
+% \begin{figure}[tp]
+%   \centering
+%   \begin{subfigure}{0.80\linewidth}
+%       \begin{lstlisting}[language=JavaScript]
+% var multiPurposeModule = require('./multi-purpose-module');
+
+% // module-function
+% console.log(multiPurposeModule("John", "Doe"));
+% // John Doe
+
+% // module-class
+% var m = new multiPurposeModule("Jane", "Doe");
+% console.log(m.firstName + " - " + m.lastName);
+% // Jane - Doe
+
+% console.log(m.sayHello());
+% // Hello, my name is Jane Doe
+
+% // module
+% console.log(multiPurposeModule.anotherFunction());
+% // I am another function! 
+%       \end{lstlisting}
+%       \caption{index.js}
+%     \end{subfigure}
+
+%     \begin{subfigure}{0.80\linewidth}
+%       \begin{lstlisting}[language=TypeScript]
+% var multiPurposeModule = function(firstName, lastName) {
+%   this.firstName = firstName;
+%   this.lastName = lastName;
+
+%   return firstName + " " + lastName;
+% };
+
+% multiPurposeModule.prototype.sayHello = function() {
+%   return "Hello, my name is " + this.firstName + " " + this.lastName;
+% };
+
+% multiPurposeModule.anotherFunction = function() {
+%   return "I am another function!";
+% };
+
+% module.exports = multiPurposeModule;   
+%       \end{lstlisting}
+%       \caption{multi-purpose-module/index.js}
+%     \end{subfigure}
+%   \caption{Example module that implements all three templates simultaneously}
+%   \label{fig:multi-purpose-module-example}
+% \end{figure}
 
@@
 
-\begin{figure}[tp]
-  \centering
-  \begin{subfigure}{0.48\linewidth}
-    \begin{lstlisting}[language=JavaScript]
-var multiPurposeModule = require('./multi-purpose-module');
-
-multiPurposeModule("John", "Doe");
-// John Doe
-    \end{lstlisting}
-    \caption{Example module \texttt{multi-purpose-module} used as \texttt{module-function}}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{0.48\linewidth}
-    \begin{lstlisting}[language=TypeScript]
-export = MultiPurposeModule;
-
-declare function MultiPurposeModule(firstName: string, lastName: string): string;
-    \end{lstlisting}
-    \caption{Generated declaration file with \texttt{module-function} template}
-  \end{subfigure}
-
-  \begin{subfigure}{0.48\linewidth}
-    \begin{lstlisting}[language=JavaScript]
-var multiPurposeModule = require('./multi-purpose-module');
-
-var m = new multiPurposeModule("Jane", "Doe");
-var fullName = m.firstName + " - " + m.lastName;
-// Jane - Doe
-
-m.sayHello();
-// Hello, my name is Jane Doe
-    \end{lstlisting}
-    \caption{Example module \texttt{multi-purpose-module} used as \texttt{module-class}}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
-export = MultiPurposeModule;
-
-declare class MultiPurposeModule {
-  constructor(firstName: string, lastName: string);
-  sayHello(): string;
-}
-
-declare namespace MultiPurposeModule {
-}
-      \end{lstlisting}
-      \caption{Generated declaration file with \texttt{module-class} template}
-    \end{subfigure}
-
-    \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=JavaScript]
-var multiPurposeModule = require('./multi-purpose-module');
-
-multiPurposeModule.anotherFunction();
-// I am another function!
-      \end{lstlisting}
-      \caption{Example module \texttt{multi-purpose-module} used as \texttt{module}}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
-export function anotherFunction(): string;
-      \end{lstlisting}
-      \caption{Generated declaration file with \texttt{module} template}
-    \end{subfigure}
-
-  \caption{Different ways of using example module presented in \figref{fig:multi-purpose-module-example}.
+% \begin{figure}[tp]
+%   \centering
+%   \begin{subfigure}{0.48\linewidth}
+%     \begin{lstlisting}[language=JavaScript]
+% var multiPurposeModule = require('./multi-purpose-module');
+
+% multiPurposeModule("John", "Doe");
+% // John Doe
+%     \end{lstlisting}
+%     \caption{Example module \texttt{multi-purpose-module} used as \texttt{module-function}}
+%   \end{subfigure}
+%   \hfill
+%   \begin{subfigure}{0.48\linewidth}
+%     \begin{lstlisting}[language=TypeScript]
+% export = MultiPurposeModule;
+
+% declare function MultiPurposeModule(firstName: string, lastName: string): string;
+%     \end{lstlisting}
+%     \caption{Generated declaration file with \texttt{module-function} template}
+%   \end{subfigure}
+
+%   \begin{subfigure}{0.48\linewidth}
+%     \begin{lstlisting}[language=JavaScript]
+% var multiPurposeModule = require('./multi-purpose-module');
+
+% var m = new multiPurposeModule("Jane", "Doe");
+% var fullName = m.firstName + " - " + m.lastName;
+% // Jane - Doe
+
+% m.sayHello();
+% // Hello, my name is Jane Doe
+%     \end{lstlisting}
+%     \caption{Example module \texttt{multi-purpose-module} used as \texttt{module-class}}
+%     \end{subfigure}
+%     \hfill
+%     \begin{subfigure}{0.48\linewidth}
+%       \begin{lstlisting}[language=TypeScript]
+% export = MultiPurposeModule;
+
+% declare class MultiPurposeModule {
+%   constructor(firstName: string, lastName: string);
+%   sayHello(): string;
+% }
+
+% declare namespace MultiPurposeModule {
+% }
+%       \end{lstlisting}
+%       \caption{Generated declaration file with \texttt{module-class} template}
+%     \end{subfigure}
+
+%     \begin{subfigure}{0.48\linewidth}
+%       \begin{lstlisting}[language=JavaScript]
+% var multiPurposeModule = require('./multi-purpose-module');
+
+% multiPurposeModule.anotherFunction();
+% // I am another function!
+%       \end{lstlisting}
+%       \caption{Example module \texttt{multi-purpose-module} used as \texttt{module}}
+%     \end{subfigure}
+%     \hfill
+%     \begin{subfigure}{0.48\linewidth}
+%       \begin{lstlisting}[language=TypeScript]
+% export function anotherFunction(): string;
+%       \end{lstlisting}
+%       \caption{Generated declaration file with \texttt{module} template}
+%     \end{subfigure}
+
+%   \caption{Different ways of using example module presented in \figref{fig:multi-purpose-module-example}.
 \texttt{dts-generate} generates infers the correct template for each case.
 The JavaScript implementation of the module remained unmodified for each example.}
-  \label{fig:dts-generate-choose-templates}
-\end{figure}
+%   \label{fig:dts-generate-choose-templates}
+% \end{figure}
 

### 2020-08-07 programming-2021/main.tex
up to section 6

@@
 \usetikzlibrary{positioning}
+
+\usepackage{xcolor}
+\usepackage{tcolorbox}
+\tcbuselibrary{breakable}
+
+\newenvironment{changethis}{%
+  \begin{tcolorbox}[breakable,notitle,boxrule=0pt,colback=blue!20,colframe=blue!20]}{%
+  \end{tcolorbox}}
 
@@
   documentation of an \NPM{} package and collects run-time type
-  information from running these examples.
+  information from running these examples (Sections~\ref{sec:initial-code-base}
+  and~\ref{sec:run-time-information}).
 
@@
   application that generates a valid TypeScript declaration file for a
-  specific \NPM{} package using run-time information.
-
-\item A comparator for TypeScript declaration files. This tool is
+  specific \NPM{} package using run-time information (Section~\ref{sec:typescr-decl-file}).
+
+\item A comparator for TypeScript declaration files (Section~\ref{sec:dts-compare}). This tool is
   necessary for evaluating our framework and also useful to detect
   incompatibilities when evolving JavaScript modules.
-\item An evaluation of our framework. We examined all 6029 entries in
+\item An evaluation of our framework (Sections~\ref{sec:dts-generate-evaluation}
+  and~\ref{sec:results}). We examined all 6029 entries in 
   the DefinitelyTyped repository and found 244 sufficiently
@@
   DefinitelyTyped repository. 
-
 \end{itemize}
@@
 \label{sec:motivating-example}
-The \NPM{} package \texttt{glob-to-regexp} is a simple JavaScript library, which turns a glob expression into a regular expression\footnote{https://www.npmjs.com/package/glob-to-regexp}. It
+The \NPM{} package \texttt{glob-to-regexp} is a simple JavaScript library, which turns a
+glob expression for matching filenames in the shell into a regular
+expression\footnote{https://www.npmjs.com/package/glob-to-regexp}. It 
 has about 9,900,000 weekly downloads and 188 \NPM{} packages depend on it. If
@@
     \caption{Declaration file for \texttt{module-function} template}
+    \label{fig:template-module-function}
   \end{subfigure}
@@
       \caption{Declaration file for \texttt{module} template}
+      \label{fig:template-module}
     \end{subfigure}
@@
       \caption{Declaration file for \texttt{module-class} template}
+      \label{fig:template-class}
     \end{subfigure}
@@
 imported entity is used solely as a function as in Figure~\ref{fig:example-module-function}, the
-generated declaration file follows the module-function template.
+generated declaration file follows the module-function template as shown in Figure~\ref{fig:template-module-function}.
 If the example only accesses properties of the imported entity as in
-Figure~\ref{fig:example-module}, we generate according to the module template.
+Figure~\ref{fig:example-module}, we generate according to the module template (Figure~\ref{fig:template-module}).
 If the imported entity is used with \lstinline/new/ to create new instances as in
-Figure~\ref{fig:example-class}, we generate a module-class template.
-
-The libraries \texttt{glob-to-regexp} as well as \texttt{abs} are both instances of a
-\textbf{module-function}.
+Figure~\ref{fig:example-class}, we generate a module-class template (Figure~\ref{fig:template-class}).
+
+The libraries \texttt{glob-to-regexp} as well as \texttt{abs} are both instances of the
+\textbf{module-function} template.
 Most libraries fall in one of these three categories. 
-While it is possible to create a library that can be represented with a different template
-depending on how it is used, our selection is driven by the examples in the documentation and
-hence do reflect the developers intent.
+While it is possible to create a library of undetermined category, our selection is driven
+by the examples in the documentation and 
+hence reflects the developers intent.
 % Run-time analysis proves useful in this case,
@@
 This section gives an overview of our approach to generating
-TypeScript declaration files from JavaScript libraries packaged in
+TypeScript declaration files from a JavaScript library packaged in
 \NPM. \figref{fig:tsd_generation_method_block_diagram} gives a rough
@@
 Jalangi \cite{DBLP:conf/sigsoft/SenKBG13} to gather data flow
-information and type information at runtime. Jalangi is a configurable
+information and type information at run time. Jalangi is a configurable
 framework for dynamic analysis of JavaScript. It provides several
@@
 The resulting declaration file is ready for use in the development
-process. Its content mimics the usage of the library in the example
+process. Its contents mimic the usage of the library in the example
 code fragments and matches the structure of the
@@
 
-The command line interface is inspired by the \texttt{dts-gen} package
+The command-line interface is inspired by the \texttt{dts-gen} tool
 \cite{dts-gen}. \coderef{code:dts-generate-example} shows
-that invoking the package is very simple and the only required
+that invoking the package is very simple. Its only required
 argument is the name of the module published to the \NPM{} registry.
@@
 employ differing (or no) testing tools, or do not have tests
-at all.
-Moreover, developers try to cover edge cases or even not allowed values in their tests.
-Such tests are not really useful for describing a library from the consumer's point of view.
-For example, a developer might write a test invoking a function with \texttt{null} just to validate that it throws an error in such scenario.
+at all. Moreover, developers try to cover edge cases or to trigger errors by using illegal
+input values in their tests. Such tests are not really useful for describing a library
+from the consumer's point of view. For example, a developer might write a test invoking a
+function with \texttt{null} just to validate that it throws an error in such scenario.
 
@@
 developers provide code examples in the README files to show how the
-library works and how to use it. This observation holds in particular
-for \NPM{} packages, which are generally created to solve a specific
-problem of JavaScript development. These code fragments do not present the problem of option~\ref{item:2} since they are meant to be used by developers using the package.
+library works and how to use it. These code fragments represent common usage cases rather
+than stress-testing the implementation (the problem of
+option~\ref{item:2}) because they are meant to be instructive to users of the package.
 
@@
   written using Markdown\footnote{\url{https://www.markdownguide.org}}, a
-  popular markup language. In such a file, it is customary to write
-  code examples in code blocks labeled with the programming
+  popular markup language, where
+  code examples are presented in code blocks labeled with the programming
   language, so that syntax highlighting is done correctly. Hence, we
@@
 
-The extracted example for the case of \texttt{glob-to-regexp} is shown in \coderef{code:code-example-extracted-motivating-example}.
+\coderef{code:code-example-extracted-motivating-example} shows
+the examples extracted from  the README file of the \texttt{glob-to-regexp} package in
+that way.
 
 \begin{lstlisting}[
-  caption={Extracted example for \texttt{glob-to-regexp}},
+  caption={Extracted example code for the \texttt{glob-to-regexp} package},
   language=JavaScript,
@@
 appropriate and pragmatic way of extracting the developer's
-intention and providing a useful initial code base with meaningful
+intention and provides a useful initial code base with meaningful
 examples, thus avoiding possible cold start problems.
@@
 \subsection{Run-time Information Gathering}
+\label{sec:run-time-information}
 The Runtime Information block described in
 \figref{fig:tsd_generation_method_block_diagram} gathers
-information such as:
+type-related information as well as usage-related information for each function parameter.
+For example,
 
@@
 information. The configurable analysis modules enable
-programming custom callbacks that get triggered with virtually any
+programming custom callbacks that can get triggered with virtually any
 JavaScript event. Our instrumentation observes the following events: 
 \begin{itemize}
-  \item Binary operations, like \lstinline{==}, \lstinline{+} or
-    \lstinline{===}.  
-  \item Variable declaration.
-  \item Function, method, or constructor invocation.
-  \item Access to an object's property.
-  \item Unary operations, like \lstinline{!} or \lstinline{typeof}.
+  \item binary operations, like \lstinline{==}, \lstinline{+}, or
+    \lstinline{===};
+  \item variable declarations;
+  \item function, method, or constructor invocations;
+  \item access to an object's property;
+  \item unary operations, like \lstinline{!} or \lstinline{typeof}.
 \end{itemize}
@@
 \texttt{interactions}. They are used for translating, modifying, and
-aggregating Jalangi's raw event information to get an application
-specific data representation.
+aggregating Jalangi's raw event information to get an application-specific data representation.
 Each function invocation is stored as a \texttt{FunctionContainer} which contains an \texttt{ArgumentContainer} for each argument.
@@
 
-We wrap each function's argument around a wrapper object, which enables to store meta-information and greatly simplifies the mapping of an observation with the corresponding \texttt{ArgumentContainer} or \texttt{interaction}.
-We then use the original values for critical operators such as \texttt{===} or \texttt{typeof}, for which wrapper objects will definitely modify the behavior of the code.
+We wrap each function's argument in a wrapper object, which enables to store meta-information and greatly simplifies the mapping of an observation with the corresponding \texttt{ArgumentContainer} or \texttt{interaction}.
+We use the original values for critical operators such as \texttt{===} or \texttt{typeof}, for which wrapper objects would modify the behavior of the code.
 
@@
 
-The run-time information is saved as a JSON file that can be used for later processing. The tool is written in JavaScript and runs in \NodeJS{} in a Docker container.
+The run-time information is saved as a JSON file that can be used for later
+processing. The instrumentation tool is written in JavaScript and runs in \NodeJS{} in a
+Docker container.
 
@@
 The next step in the pipeline after gathering the run-time information is
-the actual generation of the declaration file (cf.\
+the generation of the declaration file (cf.\
 \figref{fig:tsd_generation_method_block_diagram}). It is a lightweight,
-simple and fast application, which solely relies on the run-time
+simple, and fast application, which solely relies on the run-time
 information gathered in the previous step.
@@
 
-Instead of analyzing the shape of the exported module, we inspect how the module is used to choose between the templates.
-We analyze the properties \texttt{isExported}, \texttt{requiredModule} and \texttt{isConstructor} from the run-time JSON to distinguish between \texttt{module-class} and \texttt{module-function}.
-If a function is invoked and it is imported from the module we are analyzing we infer the template \texttt{module-class} or \texttt{module-function}.
-We choose \texttt{module-class} if the function is used as a constructor.
-If not, we choose \texttt{module-function}.
-Otherwise, we infer the \texttt{module} template.
-\figref{fig:dts-generate-choose-templates} exposes this logic using the example provided in \figref{fig:multi-purpose-module-example}.
+Instead of analyzing the shape of the exported module, we inspect how the module is used
+to choose between the templates. We analyze the properties \texttt{isExported},
+\texttt{requiredModule} and \texttt{isConstructor} from the run-time JSON to distinguish
+between \texttt{module-class} and \texttt{module-function}. If a function is invoked and
+it is imported from the module we are analyzing we infer the template
+\texttt{module-class} or \texttt{module-function}. We choose \texttt{module-class} if the
+function is used as a constructor. Otherwise, we choose
+\texttt{module-function}.
+If no function is invoked, we use the \texttt{module}
+template.
+% \figref{fig:dts-generate-choose-templates} exposes this logic using the example
+% provided in \figref{fig:multi-purpose-module-example}.
 
@@
 
-We implemented only basic TypeScript features and we focused our effort in building and end-to-end solution that generates valid and useful declaration files.
-Only the basic types \texttt{string}, \texttt{number}, \texttt{boolean} and its union were considered both for function parameters and interface properties.
-Common types such as \texttt{any}, arrays, function callbacks, tuples, intersections and features such as generics, index signatures were not implemented.
-Each declaration file was tagged with \texttt{dts-parse} and those containing unimplemented features were ignored.
-We add the \texttt{undefined} type for describing an optional parameter instead of adding the corresponding token \texttt{?}, which is also done by TypeScript in strict null checking mode\cite{typescript-optional-parameters-and-properties}.
+\begin{changethis}
+  We implemented only basic TypeScript features and we focused our effort in building an
+  end-to-end solution that generates valid and useful declaration files. Only the basic
+  types \texttt{string}, \texttt{number}, \texttt{boolean} and its union were considered
+  both for function parameters and interface properties. Common types such as
+  \texttt{any}, arrays, function callbacks, tuples, intersections and features such as
+  generics, index signatures were not implemented. Each declaration file was tagged with
+  \texttt{dts-parse} and those containing unimplemented features were ignored. We add the
+  \texttt{undefined} type for describing an optional parameter instead of adding the
+  corresponding token \texttt{?}, which is also done by TypeScript in strict null checking
+  mode\cite{typescript-optional-parameters-and-properties}.
+\end{changethis}
 
@@
 the \texttt{followingInteractions} field are recursively traversed,
-building a new interface in each recursion level.
+building a new interface at each level up to a predefined depth. Interfaces that turn out
+to be equivalent are merged.
 
@@
 After generating a declaration file for an \NPM{} package, we need to
-evaluate its quality. To this end, we created two tools:
-\begin{itemize}
-  \item \texttt{dts-parse}: A TypeScript declaration files parser, which transforms the file into a JSON file with a defined structure.
-  \item \texttt{dts-compare}: A tool that uses \texttt{dts-parse} and compares two TypeScript declaration files computing differences by applying several criteria.
-\end{itemize}
-
-We applied \texttt{dts-compare} to compare the generated declaration file against the one uploaded to the DefinitelyTyped repository for the same module, as shown in \figref{fig:evaluation-diagram}.
-While this approach does not provide an absolute measure of quality, it gives us at least an indication of the pragmatics of \texttt{dts-generate}: The files on DefinitelyTyped
-are perceived to be useful by the community. If the accuracy of the
-generated files is comparable with the accuracy of the files on
-DefinitelyTyped, then \texttt{dts-generate} is a viable alternative to
-manually creating a definition file.
-
-%!!!TODO: check how many files on DT are just generated by \texttt{dts-gen} (e.g., the TypeScript tool)
+evaluate its quality. To this end, we created \texttt{dts-compare}, a tool to compute the
+differences between two TypeScript declaration files.
+% \begin{itemize}
+%   \item \texttt{dts-parse}: A TypeScript declaration files parser, which transforms the file into a JSON file with a defined structure.
+%   \item \texttt{dts-compare}: A tool that uses \texttt{dts-parse} and compares two TypeScript declaration files computing differences by applying several criteria.
+% \end{itemize}
+
+For each module, we applied \texttt{dts-compare} to the generated declaration file and the
+one uploaded to the DefinitelyTyped repository for the same module, as shown in
+\figref{fig:evaluation-diagram}. While this approach does not provide an absolute measure
+of quality, it gives us at least an indication of the pragmatics of \texttt{dts-generate}:
+The files on DefinitelyTyped are perceived to be useful by the community. If the accuracy
+of the generated files is comparable with the accuracy of the files on DefinitelyTyped,
+then \texttt{dts-generate} is a viable alternative to manually created definition files.
+
+\begin{changethis}
+  Question: check how many files on DT are just generated by \texttt{dts-gen} (e.g., the TypeScript tool)
+\end{changethis}
 
@@
         {\includegraphics[width=1\textwidth]{evaluation-diagram.pdf}}
-        \caption[Evaluation against DefinitelyTyped Repository]{\textbf{Evaluation of generated declaration files against DefinitelyTyped Repository} - A parser transforms the generated declaration file and the equivalent file in the DefinitelyTyped repository into a JSON file using the TypeScript Compiler API \cite{typescript-compiler-api}.
-Comparison is then performed on the JSON files, i.e.
-not on the declaration files.}
+        \caption[Evaluation against DefinitelyTyped Repository]{\textbf{Evaluation of
+            generated declaration files against DefinitelyTyped Repository} - The parser
+          uses the TypeScript Compiler           API \cite{typescript-compiler-api} to
+          transform declaration files into an abstract syntax which is serialized in
+          JSON. The comparison is performed on the JSON format.} 
         \label{fig:evaluation-diagram}
@@
 \subsection{dts-parse}
+\label{sec:dts-parse}
 We need a means of comparing two declaration files. Of
@@
 
-\texttt{dts-parse} is itself written in TypeScript. It receives the file as input parameter and returns a JSON, as shown in \coderef{code:dts-parse-example}.
+\texttt{dts-parse} is itself written in TypeScript. It receives the file as input
+parameter and returns an AST serialized in JSON, as shown in \coderef{code:dts-parse-example}.
 
 \begin{lstlisting}[
-  caption={Example for \texttt{dts-parse} executed on module \texttt{glob-to-regexp} and combined with \texttt{jq} to browse through the returned JSON},
+  caption={Example for \texttt{dts-parse} executed on module \texttt{glob-to-regexp} and combined with \texttt{jq} to browse the returned JSON},
   language=bash,
@@
 "GlobToRegexp"
-\end{lstlisting}
+\end{lstlisting} %$
 
@@
 \label{sec:dts-compare}
-The tool is written in TypeScript and receives 2 declaration files as an input and returns a JSON containing the result of the comparison, as shown in \coderef{code:dts-compare-example}.
-Following the naming convention of any standard Unit Testing framework, the input files are flagged as \lstinline{expected} or \lstinline{actual}.
-We assigned the \lstinline{expected} flag to the DefinitelyTyped declaration file.
+The tool is written in TypeScript. It takes two declaration files as input and returns a
+JSON file containing the result of the comparison, as shown in
+\coderef{code:dts-compare-example}. Following the naming convention of any standard Unit
+Testing framework, the input files are marked as \lstinline{expected} or
+\lstinline{actual}. We use \lstinline{expected} for the DefinitelyTyped
+declaration files.
 
@@
 
-\texttt{dts-compare} applies \texttt{dts-parse} internally to both files to get a common structure eligible for comparison.
-The result of the comparison is an array of an abstract entity called \texttt{Difference}.
-We extracted code examples written by developers to execute the code and gather run-time information.
-The quality of the generated declaration file is heavily dependent on the code examples.
-For example, if a property of an interface does not get accessed due to a code example not exploring a particular execution path, that property will not exist in the generated declaration file.
-Consequently, in order to isolate the impact of incomplete code examples on our quality analysis we identified each difference as follows:
+\texttt{dts-compare} applies \texttt{dts-parse} internally to both files to get 
+normalized structures that are easy to compare. The result of the comparison is an array
+of an abstract entity \texttt{Difference}.
+
+We extracted code examples written by
+developers to execute the code and gather run-time information. The quality of the
+generated declaration file is heavily dependent on the code examples. For example, if a
+property of an interface does not get accessed due to a code example not exploring a
+particular execution path, that property will not exist in the generated declaration
+file.
+
+To measure the impact of incomplete code examples on the
+quality of generated declaration files we categorized each difference as follows: 
 \begin{itemize}
-  \item \texttt{solvable}: Differences that can be solved by expanding the code examples. For example, invoking a function with different parameters so that a missing interface property gets accessed.
-  \item \texttt{unsolvable}: Differences that can not be solved are marked as \texttt{unsolvable}. These differences are implementation errors of \texttt{dts-generate}.
+  \item \texttt{solvable}: Differences that can be resolved by adding more code
+    examples. For example, invoking a function with different parameters so that a missing
+    interface property gets accessed. 
+  \item \texttt{unsolvable}: Differences that cannot be resolved with further examples.
+    % are marked as \texttt{unsolvable}. These differences are implementation errors of \texttt{dts-generate}.
 \end{itemize}
@@
 \begin{itemize}
-  \item \texttt{TemplateDifference}: A difference between the formal TypeScript templates.
-For example, if the file in DefinitelyTyped is written with \texttt{module} and we generated a file using \texttt{module-function}.
-The comparison stops if the templates differ.
-  \item \texttt{ExportAssignmentDifference}: An equality check between the export assignments of both files, that is in the expression \lstinline{export = XXX}.
-  \item \texttt{FunctionMissingDifference}: A function is present in the DefinitelyTyped file but not in the generated one. This applies to functions and methods of classes and interfaces.
-  \item \texttt{FunctionMissingDifference}: A function is present in the generated declaration file but not in DefinitelyTyped.
-  \item \texttt{FunctionOverloadingDifference}: The number of declarations for the same functions is different.
-This is particularly common when developers declare a function multiple times with different single types instead of using union types.
+  \item \texttt{TemplateDifference}: A difference between the choice of TypeScript
+    templates. For example, if the file in DefinitelyTyped is written as a \texttt{module}
+    but we generated the file using the \texttt{module-function} template. The comparison
+    stops if the templates differ. 
+  \item \texttt{ExportAssignmentDifference}: An equality check between the export
+    assignments of both files, that is in the expression \lstinline{export = XXX}. 
+  \item \texttt{FunctionMissingDifference}: A function declaration is present in the expected
+    file but not in the generated one. This difference applies to functions as well as methods of classes
+    and interfaces. 
+  \item \texttt{FunctionExtraDifference}: A function is present in the generated declaration file but not in DefinitelyTyped.
+  \item \texttt{FunctionOverloadingDifference}: The number of declarations for the same
+    function is different. This difference can be due to inexperience of the author of a
+    declaration file, who uses, e.g., multiple declarations where a union typed argument
+    would be appropriate.
   \item \texttt{ParameterMissingDifference}: A parameter of a function or a property of an interface is not present in the generated file.
-  \item \texttt{ParameterExtraDifference}: A parameter of a function or a property of an interface is present in the generated file but not in the DefinitelyTyped file.
+  \item \texttt{ParameterExtraDifference}: A parameter of a function or a property of an
+    interface is present in the generated file but not in the DefinitelyTyped file. 
   \item \texttt{ParameterTypeDifference}: A parameter of a function or a property of an interface is generated with a different type than in the DefinitelyTyped file.
@@
   \begin{itemize}
-    \item \texttt{SolvableDifference}: A type difference that can be solved by writing better code examples.
-This includes a basic type that is converted to a union type, a function overloading or a parameter or property that is marked as optional.
-    \item \texttt{UnsolvableDifference}: Anything not considered as \texttt{SolvableDifference}.
+    \item \texttt{SolvableDifference}: A type difference that can be solved by writing
+      further code examples. For example, a basic type that is converted to a union type,
+      a function overloading, or a parameter or property that is marked as optional. 
+    \item \texttt{UnsolvableDifference}: Any difference not considered as \texttt{SolvableDifference}.
   \end{itemize}
@@
 
-Type aliases are invisible to \texttt{dts-compare} as only final types were considered: the declaration \lstinline{type T = string | number; declare function F(a: T);} is equivalent to \lstinline{declare function F(a: string | number);}.
-The same concept applies for literal interfaces.
-\texttt{dts-compare} does not contemplate differences in function return types, since interface inference of function return types is not covered in this version of \texttt{dts-generate}.
+Type aliases are expanded before the comparison and thus invisible to
+\texttt{dts-compare}: the declaration
+\lstinline{type T = string | number; declare function F(a: T);} is equivalent to
+\lstinline{declare function F(a: string | number);}. The same approach applies to literal
+interfaces. \texttt{dts-compare} does 
+not contemplate differences in function return types, because interface inference of
+function return types is not covered in the current version of \texttt{dts-generate}.
 
@@
 \NPM{} registry. The DefinitelyTyped repository was used as a
-benchmark. Each one of the generated files was compared against the
-corresponding declaration file already uploaded to the repository.
+benchmark. Each of the generated files was compared against the
+corresponding declaration file uploaded to the repository.
 
@@
 generated declaration files for templates \textbf{module},
-\textbf{module-class} and \textbf{module-function} are presented in
+\textbf{module-class}, and \textbf{module-function} are presented in
 \secref{sec:experiments-declaration-files-generation}.
@@
 \paragraph*{Execution of Code Examples}
-We executed the remaining 2260 extracted code by installing the
+We executed the remaining 2260 extracted code examples by installing the
 required packages and running the code as a \NodeJS{}
@@
 \item The code fragment had been properly extracted but the code was
-  not faulty. It was executing the library in an unsupported
+  faulty. It invoked the library in an unsupported
   (obsolete?) way, which lead to a run-time error.
@@
 \paragraph*{Run-time Information}
-Run-time information was extracted for only 442 out of 1448 modules with working code examples.
-As explained, in order to extract the run-time information, the behavior of the code under analysis was explicitly modified by wrapping the arguments around Wrapper Objects.
-Furthermore, Jalangi’s instrumentation itself caused some executions to fail, since the modules contained JavaScript features that are not supported by Jalangi.
-As a result, run-time information could not be extracted for 1006 modules.
-An instrumentation without user defined analysis modules was not applied, so it was not possible to determine which modules were failing only because of Jalangi’s own limitations.
+Run-time information was extracted for only 442 out of 1448 modules with working code
+examples. As explained, to extract the run-time information, the behavior of the code
+under analysis was explicitly modified by wrapping the arguments. Furthermore, Jalangi’s
+instrumentation itself caused some executions to fail, because the modules contained
+JavaScript features that are not supported by Jalangi. As a result, run-time information
+could not be extracted for 1006 modules. An instrumentation without user-defined analysis
+modules was not applied, so it was not possible to determine which modules were failing
+only because of Jalangi’s own limitations.
 
 \paragraph*{Generated Declaration Files}
-A declaration file was generated for 244 out of 442 modules.
-Despite the correct execution of instrumented code examples, 198 modules did not contain suitable run-time information for generating a declaration file.
-The extracted code examples for these modules did not execute the library itself and therefore the collected run-time information was not useful for generating a declaration file.
+A declaration file was generated for 244 out of 442 modules. Despite the correct execution
+of the instrumented code examples, 198 modules did not yield suitable run-time information
+for generating a declaration file. The extracted code examples for these modules did not
+execute the library. Hence, the collected run-time information was not useful
+for generating a declaration file.
 
@@
 
-This section exhibits some samples of the 244 generated
+%%% !!! TBC !!!
+
+This section exhibits some examples of the 244 generated
 declaration files. It shows some results for each of the implemented
-templates: \textbf{module}, \textbf{module-function} and
+templates: \textbf{module}, \textbf{module-function}, and
 \textbf{module-class}.
@@
 \texttt{dirname-regex}, and \texttt{escape-html}. All of them were
-generated using the \textbf{module-function} template. Left side of the figure shows the generated declaration file with
+generated using the \textbf{module-function} template. The left side of the figure shows
+the generated declaration file with 
 \lstinline{dts-generate}, the right side shows the corresponding file
@@
 
-Differences in the export assignment by modules \texttt{dirname-regex} and \texttt{escape-html} is not problematic since the exported module can receive a custom name given by the developer.
-\texttt{dts-generate} automatically generates the export assignment by transforming the module name into camel case form, following TypeScript guidelines.
-The parameter names are extracted from the variable names of the JavaScript code.
-The JavaScript concrete implementation of module \texttt{escape-html} declares the first parameter of the exported function with the name \texttt{string} and not \texttt{text}, as specified in DefinitelyTyped.
-A difference in the names of the function parameters will anyway not affect the code compilation in any way.
-Furthermore, the omitted namespace in module \texttt{escape-html} is irrelevant.
-It is actually not required, since there is nothing declared within the namespace.
+Differences in the export assignment by modules \texttt{dirname-regex} and
+\texttt{escape-html} is not problematic since the exported module can receive a custom
+name given by the developer. \texttt{dts-generate} automatically generates the export
+assignment by transforming the module name into camel case form, following TypeScript
+guidelines. The parameter names are extracted from the variable names of the JavaScript
+code. The JavaScript concrete implementation of module \texttt{escape-html} declares the
+first parameter of the exported function with the name \texttt{string} and not
+\texttt{text}, as specified in DefinitelyTyped. A difference in the names of the function
+parameters will anyway not affect the code compilation in any way. Furthermore, the
+omitted namespace in module \texttt{escape-html} is irrelevant. It is actually not
+required, since there is nothing declared within the namespace.
 

### 2020-08-08 programming-2021/main.tex
first-pass-done

@@
    showspaces=false,
-   numbers=left,
+   numbers=none,
    numberstyle=\footnotesize,
@@
     float=bp,
-    abovecaptionskip=-\medskipamount
+    captionpos=b
 ]
@@
 	label=code:dts-generate-example,
-    float,
-    abovecaptionskip=-\medskipamount
+    float,captionpos=b
 ]
@@
 
-TypeScript provides a guide on how to write high-quality declaration
+The TypeScript project provides a guide on how to write high-quality declaration
 files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/by-example.html}. The guide
@@
 %     \begin{subfigure}{0.80\linewidth}
-%       \begin{lstlisting}[language=TypeScript]
+%       \begin{lstlisting}[language=TypeScript,numbers=none]
 % var multiPurposeModule = function(firstName, lastName) {
@@
     \includegraphics[width=1\linewidth]{dts-generate-block-diagram.pdf}
-    \caption[dts-generate - Architecture overview]{\textbf{dts-generate - Architecture overview} - Initial code base is retrieved from the \NPM{} package's repository.
-A valid TypeScript Declaration File is generated using run-time information.
-A Symbolic Execution Engine creates test cases based on the generated Declaration File and via a feedback loop enriches the code base until the stopping criteria is reached.
-The final TypeScript Declaration File gets returned.
-Feedback loop through the Symbolic Execution Engine was not implemented.
-It can be added in a future to the existing architecture, without modifying the working blocks.}
+    \caption[dts-generate - Architecture overview]{\textbf{dts-generate - Architecture
+        overview} - The initial code base is retrieved from the \NPM{} package's
+      repository.
+      Run-time information is gathered driven by example code.
+      A TypeScript Declaration File is generated using run-time
+      information.
+      % A Symbolic Execution Engine creates test cases based on the generated
+      % Declaration File and via a feedback loop enriches the code base until the stopping
+      % criteria is reached.
+      % The final TypeScript Declaration File gets returned.
+      % Feedback
+      % loop through the Symbolic Execution Engine was not implemented. It can be added in a
+      % future to the existing architecture, without modifying the working blocks.
+    } 
     \label{fig:tsd_generation_method_block_diagram}
@@
   label=code:code-example-extracted-motivating-example,
-  float,
-  abovecaptionskip=-\medskipamount
+  float,captionpos=b
 ]
@@
 %   \begin{subfigure}{0.48\linewidth}
-%     \begin{lstlisting}[language=TypeScript]
+%     \begin{lstlisting}[language=TypeScript,numbers=none]
 % export = MultiPurposeModule;
@@
 %     \begin{subfigure}{0.48\linewidth}
-%       \begin{lstlisting}[language=TypeScript]
+%       \begin{lstlisting}[language=TypeScript,numbers=none]
 % export = MultiPurposeModule;
@@
 %     \begin{subfigure}{0.48\linewidth}
-%       \begin{lstlisting}[language=TypeScript]
+%       \begin{lstlisting}[language=TypeScript,numbers=none]
 % export function anotherFunction(): string;
@@
   \begin{subfigure}{0.48\linewidth}
-    \begin{lstlisting}[language=TypeScript]
+    \begin{lstlisting}[language=TypeScript,numbers=none]
 export function v1(str: string): boolean;
@@
   \begin{subfigure}{0.48\linewidth}
-    \begin{lstlisting}[language=TypeScript]
+    \begin{lstlisting}[language=TypeScript,numbers=none]
 export function v1(value: string): boolean;
@@
 label=code:dts-parse-example,
-  float,
-  abovecaptionskip=-\medskipamount
+  float=tp, captionpos=b
 ]
@@
 label=code:dts-compare-example,
-  float,
-  abovecaptionskip=-\medskipamount
+  float=tp,captionpos=b
 ]
@@
 \begin{itemize}
-  \item The quality of the inferred types, interfaces and module structure using runtime information.
+  \item The quality of the inferred types, interfaces, and module structure using run-time information.
   \item The benefits of using code examples provided by developers as a first approximation to execute the libraries and avoid a cold start problem.
@@
 
-Figure \ref{fig:experiments-overall-funnel} shows that a declaration
-file was generated for 244 modules out of 6029 modules and we identified 80 modules that have only the features implemented by \texttt{dts-generate}.
-We obtained positive results for 69 of them.
-A detailed explanation of the overall quality of the generated files is provided in \secref{sec:experiments-evaluation}.
-Samples of the
-generated declaration files for templates \textbf{module},
-\textbf{module-class}, and \textbf{module-function} are presented in
-\secref{sec:experiments-declaration-files-generation}.
+Figure \ref{fig:experiments-overall-funnel} shows that a declaration file was generated
+for 244 modules out of 6029 modules and we identified 80 modules that have only the
+features implemented by \texttt{dts-generate}. We obtained positive results for 69 of
+them. Section~\ref{sec:experiments-evaluation} provides a detailed explanation of the overall
+quality of the generated files.
+Section~\ref{sec:experiments-declaration-files-generation} presents examples of the generated declaration files for
+templates \textbf{module}, \textbf{module-class}, and \textbf{module-function}.
 
@@
 \begin{itemize}
-\item Extracting repository URL.
-\item Extracting README file.
-\item Extracting code examples from README files.
-\item Executing code examples and discarding failing ones.
+\item extracting the repository URL;
+\item extracting a README file;
+\item extracting code examples from README files;
+\item executing code examples and discarding failing ones.
 \end{itemize}
 
-The results obtained for each on of them are described in the
+The results obtained for each step are described in the
 following sections.
@@
 
-Differences in the export assignment by modules \texttt{dirname-regex} and
-\texttt{escape-html} is not problematic since the exported module can receive a custom
-name given by the developer. \texttt{dts-generate} automatically generates the export
+The differences in the export assignment by modules \texttt{dirname-regex} and
+\texttt{escape-html} are not problematic because the module can receive an arbitrary 
+name when it is imported. \texttt{dts-generate} automatically generates the export
 assignment by transforming the module name into camel case form, following TypeScript
-guidelines. The parameter names are extracted from the variable names of the JavaScript
-code. The JavaScript concrete implementation of module \texttt{escape-html} declares the
-first parameter of the exported function with the name \texttt{string} and not
-\texttt{text}, as specified in DefinitelyTyped. A difference in the names of the function
-parameters will anyway not affect the code compilation in any way. Furthermore, the
-omitted namespace in module \texttt{escape-html} is irrelevant. It is actually not
-required, since there is nothing declared within the namespace.
+guidelines.
+
+The parameter names are extracted from the variable names of the JavaScript
+code.
+% The JavaScript implementation of module \texttt{escape-html} declares the
+% first parameter of the exported function with the name \texttt{string} and not
+% \texttt{text}, as specified in DefinitelyTyped. 
+A difference in the names of the function parameters does not affect the correctness of
+the declaration file.
+Furthermore, the module \texttt{escape-html} from DefinitelyTyped declares an empty
+namespace, which is not generated by our tool.
 
@@
     \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
+      \begin{lstlisting}[language=TypeScript,numbers=none]
 export = Abs;
@@
     \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
+      \begin{lstlisting}[language=TypeScript,numbers=none]
 declare function Abs(input: string): string;
@@
     \begin{subfigure}{0.48\linewidth}
-        \begin{lstlisting}[language=TypeScript]
+        \begin{lstlisting}[language=TypeScript,numbers=none]
 export = DirnameRegex;
@@
       \begin{subfigure}{0.48\linewidth}
-        \begin{lstlisting}[language=TypeScript]
+        \begin{lstlisting}[language=TypeScript,numbers=none]
 export = dirnameRegex;
@@
       \begin{subfigure}{0.48\linewidth}
-        \begin{lstlisting}[language=TypeScript]
+        \begin{lstlisting}[language=TypeScript,numbers=none]
 export = EscapeHtml;
@@
       \begin{subfigure}{0.48\linewidth}
-        \begin{lstlisting}[language=TypeScript]
+        \begin{lstlisting}[language=TypeScript,numbers=none]
 declare function escapeHTML(text: string): string;
@@
     \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
+      \begin{lstlisting}[language=TypeScript,numbers=none]
 export function v1(str: string): boolean;
@@
     \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
+      \begin{lstlisting}[language=TypeScript,numbers=none]
 export function v1(value: string): boolean;
@@
 
-    \caption{Results for \textbf{module}  module \texttt{is-uuid}}
+    \caption{Results for \textbf{module} \texttt{is-uuid}}
     \label{fig:experiments-results-module-is-uuid}
@@
 
-A sample for the \textbf{module} template
-for the \texttt{is-uuid} module is shown in
+An example for the \textbf{module} template
+is shown for the \texttt{is-uuid} module in
 \figref{fig:experiments-results-module-is-uuid}.
-Again, methods that were not executed by the extracted examples are
-not included in the declaration file.
-
-Finally, we could not generate a correct declaration file with the template \texttt{module-class}.
-All the modules corresponding to this template had at least one of the unimplemented TypeScript features.
-
-It is worth mentioning that for some libraries the declaration file in
-DefinitelyTyped was not correct. For example, for the module
-\texttt{glob-base}, the parameter \texttt{basePath} is declared as optional in DefinitelyTyped.
-However, when invoking the function without that parameter an Error is thrown at runtime.
-The file generated by \texttt{dts-generate} will never mark the parameter as optional, as shown in \figref{fig:experiments-results-module-glob-base}.
-Function return types interfaces are not inferred by \texttt{dts-generate}.
-Parameter name is also inferred as \texttt{pattern} instead of \texttt{basePath}, since the JavaScript implementation declares the parameter as \texttt{pattern}, as shown in \figref{fig:subfloat-globbase-js-implementation}.
-We even discovered errors in the selected template in DefinitelyTyped.
-The module \texttt{smart-truncate} in DefinitelyTyped uses the \texttt{module} template, but \texttt{dts-generate} generates a file using the \texttt{module-function} template.
-We see in \coderef{code:experiments-results-module-smart-truncate} that the module is indeed exported as a function, as inferred by \texttt{dts-generate}.
+The generated file does not contain methods that were not executed by the extracted
+examples. The invoked function are correctly detected.
+
+Finally, we were unable to  generate a declaration file based on the
+\texttt{module-class} template. All modules corresponding to this template exhibited at least one of
+the unimplemented TypeScript features.
+
+It is worth mentioning that for some libraries the declaration file in DefinitelyTyped was
+not correct. For example, for the module \texttt{glob-base}, the parameter
+\texttt{basePath} is declared as optional in DefinitelyTyped. However, when invoking the
+function without the \texttt{basePath} parameter, an error is thrown at run time. The file generated by
+\texttt{dts-generate} does not mark this parameter as optional, as shown in
+\figref{fig:experiments-results-module-glob-base}. Function return type interfaces are
+not inferred by \texttt{dts-generate}. The parameter name is \texttt{pattern}
+instead of \texttt{basePath}, because the JavaScript implementation declares the parameter
+as \texttt{pattern}, as shown in \figref{fig:subfloat-globbase-js-implementation}. We also
+discovered errors in the selected template in DefinitelyTyped. The module
+\texttt{smart-truncate} in DefinitelyTyped uses the \texttt{module} template, but
+\texttt{dts-generate} generates a file using the \texttt{module-function}
+template. \coderef{code:experiments-results-module-smart-truncate} shows that the module is indeed
+exported as a function, as inferred by \texttt{dts-generate}.
 
@@
     \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
+      \begin{lstlisting}[language=TypeScript,numbers=none]
 export = GlobBase;
@@
     \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
+      \begin{lstlisting}[language=TypeScript,numbers=none]
 declare function globbase(basePath?: string): globbase.GlobBaseResult;
@@
     \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=JavaScript]
+      \begin{lstlisting}[language=JavaScript,numbers=none]
 var globBase = require('glob-base');
@@
     \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript]
+      \begin{lstlisting}[language=TypeScript,numbers=none]
 'use strict';
@@
   label=code:experiments-results-module-smart-truncate,
-  language=JavaScript
+  language=JavaScript,captionpos=b,float=tp,numbers=none
 ]
@@
 \label{sec:experiments-evaluation}
-We analyzed only 82 of the 244 generated files.
-The remaining 162 files contained at least one of the unimplemented TypeScript features and were therefore not considered.
-These features were identified by \texttt{dts-parse} as tags.
-
-As shown in \figref{fig:experiments-results-module-is-uuid}, the code examples have a direct influence on the quality of the generated declaration file.
-We introduced the concept of \texttt{solvable} and \texttt{unsolvable} differences in \secref{sec:dts-compare}.
-Modules containing only \texttt{solvable} differences were considered as positive results.
-We identified 72 such modules.
-We completed the examples manually for 10 of those modules, making them equivalent to the DefinitelyTyped files, as shown in \figref{fig:experiments-results-manually-completed-examples}.
-Expected differences are present: optional parameters are declared both with the actual and \texttt{undefined} type and the union type is handled through multiple overloads.
-
+We analyzed 82 of the 244 generated files. The remaining 162 files contained at least one
+of the unimplemented TypeScript features so that they could not be considered. These
+features were identified by \texttt{dts-parse} as tags.
+
+As shown in \figref{fig:experiments-results-module-is-uuid}, the code examples have a
+direct influence on the quality of the generated declaration file. We introduced the
+concept of \texttt{solvable} and \texttt{unsolvable} differences in
+\secref{sec:dts-compare}. Modules containing only \texttt{solvable} differences were
+considered as positive results. We identified 72 such modules. We manually completed the
+examples for 10 of those modules. Applying \texttt{dts-generate} with the completed
+examples generates declaration files, which are equivalent to the files from DefinitelyTyped, as
+shown in \figref{fig:experiments-results-manually-completed-examples}.
+\begin{changethis}
+  Expected differences are present: optional parameters are declared both with the actual
+  and \texttt{undefined} type and the union type is handled through multiple overloads.
+\end{changethis}
 \begin{figure}[tp]
@@
 
-We identified only 10 modules with \texttt{unsolvable} differences.
-Most of them were caused for the same error: 8 modules had an incorrectly declared interface for the \texttt{string} type.
-This interface contained standard properties or methods from the JavaScript \texttt{String} object, such as \texttt{length} or \texttt{indexOf()}.
-The remaining modules are \texttt{duplexer2} and \texttt{duplexer3}.
-Both of them used native NodeJS types within the declaration file and \texttt{dts-generate} could not identify those properly.
+We discovered only 10 modules with \texttt{unsolvable} differences. Most of them were
+caused by the same error: 8 modules incorrectly declared an interface for the
+\texttt{string} type. This interface contains standard properties and methods from the
+JavaScript \texttt{String} object, such as \texttt{length} or \texttt{indexOf()}. The
+remaining modules are \texttt{duplexer2} and \texttt{duplexer3}. Both of them rely on native
+NodeJS types within the declaration file, which \texttt{dts-generate} could not identify
+properly.
 
@@
 	label=code:related-work-dts-gen-example,
-    float,
-    abovecaptionskip=-\medskipamount
+    float=tp,captionpos=b
 ]
@@
 
-\section{Conclusion}
+\section{Conclusions and Future Work}
 \label{sec:conclusion}
-We have presented \texttt{dts-generate}, a tool for generating a TypeScript declaration file for a specific JavaScript library.
-The tool downloads code samples written by the developers from the library's repository.
-It uses these samples to execute the library and gather data flow and type information.
-The tool finally generates a TypeScript declaration file based on the information gathered at run-time.
-
-We developed an architecture that supports the automatic generation of declaration files for specific JavaScript libraries without additional manual tasks.
-The architecture contemplates a future incorporation of a Symbolic Execution Engine that refines the initial code base enabling the exploration of new execution paths.
-However not implemented in this work, its incorporation would result in small incremental modifications to the presented architecture as it is considered to only expand the existing code base.
-
-Building an end-to-end solution for the generation of TypeScript declaration files was prioritized over type inference accuracy.
-Consequently, types were taken over from the values at run-time.
-Since developers expose through code how a library should be used, obtaining the types from the code examples extracted from the repositories proved to be a pragmatic and effective approximation, enabling to work on specific aspects regarding the TypeScript declaration file generation itself.
-
-We built a mechanism to automatically create declaration files for potentially every module uploaded to DefinitelyTyped.
-We managed to generate declaration files for 244 modules.
-We compared the results against the corresponding files uploaded to DefinitelyTyped by creating \texttt{dts-parse}, a TypeScript declaration files parser and \texttt{dts-compare}, a comparator.
-
-We exposed the fundamental aspect of capturing the developer's intention when inferring types in JavaScript.
-Instead of applying constraints and restrictions for operations with certain types, we presented a proposal where common practices are favored.
-Uncommon usage is not forbidden but greatly disfavored.
-Accordingly, we collected evidence regarding the usage of JavaScript operators by analyzing 400 libraries.
-
-Finally, the architecture is composed of different blocks that interact with each other.
-Each block is independent and has a well defined behavior as well as clear input and output values.
-As a result, each block can be independently and simultaneously improved.
+We have presented \texttt{dts-generate}, a tool for generating a TypeScript declaration
+file for a specific JavaScript library. The tool downloads code examples written by the
+developers from the library's repository. It uses these examples to execute the library and
+gather data flow and type information. The tool finally generates a TypeScript declaration
+file based on the information gathered at run time.
+
+We developed an architecture that supports the automatic generation of declaration files
+for specific JavaScript libraries without additional manual tasks. The architecture
+contemplates future incorporation of a Symbolic Execution Engine that refines the
+initial code base enabling the exploration of new execution paths. Although it is not
+implemented 
+in this work, its incorporation would result in small incremental modifications to the
+presented architecture as it is considered to only expand the existing code base.
+
+Building an end-to-end solution for the generation of TypeScript declaration files was
+prioritized over type inference accuracy. Consequently, types were taken over from the
+values at run time. Developers express their intent how a library should be used with
+example code in the documentation. Hence,
+obtaining the types from the code examples extracted from the repositories proved to be a
+pragmatic and effective approximation, enabling to work on specific aspects regarding the
+TypeScript declaration file generation itself.
+
+We built a mechanism to automatically create declaration files for potentially every
+module uploaded to DefinitelyTyped. We managed to generate declaration files for 244
+modules. We compared the results against the corresponding files uploaded to
+DefinitelyTyped by creating \texttt{dts-parse}, a TypeScript declaration files parser and
+\texttt{dts-compare}, a comparator.
+
+\begin{changethis}
+  We exposed the fundamental aspect of capturing the developer's intention when inferring
+  types in JavaScript. Instead of applying constraints and restrictions for operations
+  with certain types, we presented a proposal that favors best practice. Uncommon usage is
+  not forbidden but greatly discouraged. Accordingly, we collected evidence regarding the
+  usage of JavaScript operators by analyzing 400 libraries.
+\end{changethis}
+% Finally, the architecture is composed of different blocks that interact with each
+% other. Each block is independent and has a well defined behavior as well as clear input
+% and output values. As a result, each block can be independently and simultaneously
+% improved.
 

### 2020-08-09 programming-2021/main.tex
related work

@@
 
+\section{Discussion}
+\label{sec:discussion}
+
+How can we improve further? Given that JavaScript developers are more incentivized to develop good
+examples than writing ``boring'' type declaration, how can we capitalize on that attitude?
+The main issue is that any example-driven algorithm can only come up with an
+under-approximation of the intended meaning, so some form of help is required when
+generalization is desired.
+
+\paragraph*{Literal types vs string}
+JavaScript libraries often (?) use literal strings to select
+options and configurations. For example, \dots
+
+For our framework, it is hard to distinguish a string that is meant literally from a
+string that just serves as an example. For an automatic solution, one might collect
+typical values of literal strings and generate the corresponding literal types (or
+unions thereof) when the arguments used in the example are all typical. Alternatively, one
+might always generate literal types unless the number of different examples provided is
+greater than some threshold, in which case the generator generalizes to type
+\lstinline/string/. A further alternative would be to communicate to the developer a set
+of typical example strings, \emph{marker strings}, that the generator always resolves to the \lstinline/string/
+type.
+
+\paragraph*{Any}
+Suppose the programmer supplies examples that use \lstinline/number/, \lstinline/bool/, and
+\lstinline/string/ at the same argument position. The obvious approximation to the type of
+this argument is the union type \lstinline/number|bool|string/. But what if the programmer
+wants to advertise the type \lstinline/any/ for an argument? This intent is hard to
+advertise with a finite number of examples.
+
+We propose that the programmer relies on marker strings like \lstinline/"any"/ to indicate
+the \lstinline/any/ type for an argument position.
+
+\paragraph*{Further types}
+Similar approaches could be conceived for indexed types, polymorphic types, dependent
+types, etc. However, at present we believe incorporating these features in a tool to be
+overblown because they are not widely used in DefinitelyTyped (see
+Table~\ref{table-of-occurrences}).
+
 \section{Related Work}
@@
 \paragraph*{Microsoft's dts-gen}
-Microsoft developed \texttt{dts-gen}, a tool that creates starter declaration files for JavaScript libraries \cite{dts-gen}.
-Its documentation states that the result is however intended to be only used a starting point.
-The outcome needs to be refined afterwards by the developers.
-
-The tool analyzes the shape of the objects at runtime after initialization without executing the library.
-This results in many variables being inferred as \lstinline[language={}]{any}.
-\coderef{code:related-work-dts-gen-example} shows an example for module \lstinline[language={}]{abs}.
-
-The solution presented in this work, however, is intended to generate declaration files that are ready to be uploaded to DefinitelyTyped without further manual intervention.
-Any amount of manual work that a developer needs to do on a declaration file after updating JavaScript code increases the risk for having discrepancies between the declaration file and the implementation.
-
-Formal aspects like applying the right template and using the correct syntax are perfectly covered by \texttt{dts-gen}.
+Microsoft developed \texttt{dts-gen}, a tool that creates declaration files for
+JavaScript libraries \cite{dts-gen}. Its documentation states that the result is 
+intended to be used a starting point for development of a declaration file. The outcome
+needs to be refined afterwards by the developer.
+
+The tool analyzes the shape of the objects at runtime after initialization without
+executing the library. This results in many variables being inferred as
+\lstinline[language={}]{any}. \coderef{code:related-work-dts-gen-example} shows an example
+for module \lstinline[language={}]{abs}.
+
+In contrast, our tool \texttt{dts-generate} is intended to generate declaration files
+that are ready to be uploaded to DefinitelyTyped without further manual intervention. Any
+amount of manual work that a developer needs to do on a declaration file after updating
+JavaScript code increases the risk for having discrepancies between the declaration file
+and the implementation.
+
+% Formal aspects like applying the right template and using the correct syntax are perfectly covered by \texttt{dts-gen}.
 
@@
 \paragraph*{TSInfer \& TSEvolve}
-TSInfer and TSEvolve are presented as part of TSTools \cite{DBLP:conf/fase/KristensenM17}.
-Both tools are the continuation of TSCheck \cite{DBLP:conf/oopsla/FeldthausM14}, a tool for looking for mismatches between a declaration file and an implementation.
-
-TSInfer proceeds in a similar way than TSCheck.
-It initializes the library in a browser and it records a snapshot of the resulting state and then it performs a light weight static analysis on all the functions and objects stored in the snapshot.
-
-The abstraction and the constraints they introduced as part of the static analysis tools for inferring the types have room for improvement.
-A run-time based approach like the one presented in our work will provide more accurate information, thus generating more precise declaration files.
-
-Since they analyze the objects and functions stored in the snapshot, they faced the problem of including in the declaration file internal methods and properties that developers wanted to hide.
-Run-time information would have informed that the developer has no intention of exposing such methods.
-
-Moreover, TSEvolve performs a differential analysis on the changes made to a JavaScript library in order to determine intentional discrepancies between declaration files of two consecutive versions.
-We consider that a differential analysis may not be needed.
-If the developer's intention is accurately extracted and the execution code clearly represents that intention then the generated declaration file would already describe the newer version of a library without the need of a differential analysis.
+TSInfer and TSEvolve are presented as part of TSTools
+\cite{DBLP:conf/fase/KristensenM17}. Both tools are the continuation of TSCheck
+\cite{DBLP:conf/oopsla/FeldthausM14}, a tool for detecting mismatches between a
+declaration file and the implementation of the module.
+
+TSInfer proceeds in a similar way than TSCheck. It initializes the library in a browser
+and records a snapshot of the resulting state.  Then it performs a lightweight static
+analysis on all the functions and objects stored in the snapshot.
+
+The abstraction and the constraints introduced as part of the static analysis tools
+for inferring the types have room for improvement. A run-time based approach like the one
+presented in our work will provide more accurate information, thus generating more precise
+declaration files.
+
+As TSInfer analyzes the objects and functions stored in the snapshot, it faces the problem
+of including internal methods and private properties in the declaration file. Run-time
+information would have informed that the developer has no intention of exposing these
+methods.
+
+Moreover, TSEvolve performs a differential analysis on the changes made to a JavaScript
+library to determine intentional discrepancies between the declaration files of two
+consecutive versions. However, a differential analysis may not be needed. If the
+developer's intention were accurately represented by the extracted code, then the generated
+declaration file would already describe the newer version of a library without the need of
+a differential analysis.
 
 \paragraph*{TSTest}
-TSTest is a tool that checks for mismatches between a declaration file and a JavaScript implementation \cite{DBLP:journals/pacmpl/KristensenM17}.
-It applies feedback-directed random testing for generating type test scripts.
-These scripts will execute the library in order to check if it behaves the way it is described in the declaration file.
-TSTest also provides concrete executions for mismatches.
-
-We evaluated the generated declaration files comparing them to the declaration files uploaded to DefinitelyTyped.
-The disadvantage of doing this is that since the uploaded files are written manually, they could already contain mismatches with the JavaScript implementation.
-However, it is a suitable choice for a development stage since it is used as a baseline.
-
-In a final stage, declaration files need to be checked against the proper JavaScript implementation and TSTest has to be definitely taken into account.
+TSTest is a tool that checks for mismatches between a declaration file and the JavaScript
+implementation of the module \cite{DBLP:journals/pacmpl/KristensenM17}. It applies feedback-directed
+random testing for generating type test scripts. These scripts execute the library with
+random arguments with the typings from the declarations file and check whether the output
+matches the prescribed type. TSTest thus provides concrete counterexamples if it detects mismatches.
+
+% We evaluated the generated declaration files comparing them to the declaration files
+% uploaded to DefinitelyTyped. The disadvantage of doing this is that since the uploaded
+% files are written manually, they could already contain mismatches with the JavaScript
+% implementation. However, it is a suitable choice for a development stage since it is used
+% as a baseline.
+
+% In a final stage, declaration files need to be checked against the proper JavaScript
+% implementation and TSTest has to be definitely taken into account.
+
+TSTest could be used to extend our tool with a feedback loop. If TSTest detects a problem
+with a declaration file generated by \texttt{dts-generate}, then we would add the
+resulting counterexamples to the example code and restart the generation process.
 

### 2020-08-12 programming-2021/main.tex
Update numbers

@@
 				7.33,
-				24.02,
+				18.02,
 				37.49,
@@
 		\node[left=0.02\textwidth of a1] {Run-time Information};
-		\node[right=0.06\textwidth of a1] {\textbf{442}};
+		\node[right=0.06\textwidth of a1] {\textbf{436}};
 
 		\node[left=0.02\textwidth of a2] {Working Examples};
-		\node[right=0.17\textwidth of a2] {\textbf{1448}};
+		\node[right=0.17\textwidth of a2] {\textbf{946}};
 
@@
 required packages and running the code as a \NodeJS{}
-application. Unfortunately, the code examples only worked for 1448
-modules. 812 modules did not run correctly and had to be
+application. Unfortunately, the code examples only worked for 946
+modules. 1314 modules did not run correctly and had to be
 discarded. Some failing samples were analyzed and there were mainly
@@
 \paragraph*{Run-time Information}
-Run-time information was extracted for only 442 out of 1448 modules with working code
+Run-time information was extracted for only 436 out of 946 modules with working code
 examples. As explained, to extract the run-time information, the behavior of the code
@@
 JavaScript features that are not supported by Jalangi. As a result, run-time information
-could not be extracted for 1006 modules. An instrumentation without user-defined analysis
+could not be extracted for 510 modules. An instrumentation without user-defined analysis
 modules was not applied, so it was not possible to determine which modules were failing
@@
 \paragraph*{Generated Declaration Files}
-A declaration file was generated for 244 out of 442 modules. Despite the correct execution
-of the instrumented code examples, 198 modules did not yield suitable run-time information
+A declaration file was generated for 244 out of 436 modules. Despite the correct execution
+of the instrumented code examples, 192 modules did not yield suitable run-time information
 for generating a declaration file. The extracted code examples for these modules did not

### 2020-08-17 programming-2021/main.tex
Update results numbers and glob-to-regexp code example

@@
 
-declare function GlobToRegexp(glob: string, opts ?: GlobToRegexp.I__opts): RegExp;
+declare function GlobToRegexp(glob: string, opts?: GlobToRegexp.I__opts): RegExp;
 declare namespace GlobToRegexp {
   export interface I__opts {
-    'extended' ?: boolean;
-    'globstar' ?: boolean;
-    'flags'    ?: string;
-  }
+    'extended'?: boolean;
+    'globstar'?: boolean;
+    'flags'?: string;
+  }
+
 }
@@
   generics, index signatures were not implemented. Each declaration file was tagged with
-  \texttt{dts-parse} and those containing unimplemented features were ignored. We add the
-  \texttt{undefined} type for describing an optional parameter instead of adding the
-  corresponding token \texttt{?}, which is also done by TypeScript in strict null checking
-  mode\cite{typescript-optional-parameters-and-properties}.
+  \texttt{dts-parse} and those containing unimplemented features were ignored.
 \end{changethis}
@@
 Figure \ref{fig:experiments-overall-funnel} shows that a declaration file was generated
-for 244 modules out of 6029 modules and we identified 80 modules that have only the
-features implemented by \texttt{dts-generate}. We obtained positive results for 69 of
+for 244 modules out of 6029 modules and we identified 60 modules that have only the
+features implemented by \texttt{dts-generate}. We obtained positive results for 52 of
 them. Section~\ref{sec:experiments-evaluation} provides a detailed explanation of the overall
@@
 \label{sec:experiments-evaluation}
-We analyzed 82 of the 244 generated files. The remaining 162 files contained at least one
+We analyzed 60 of the 244 generated files. The remaining 184 files contained at least one
 of the unimplemented TypeScript features so that they could not be considered. These
@@
 \secref{sec:dts-compare}. Modules containing only \texttt{solvable} differences were
-considered as positive results. We identified 72 such modules. We manually completed the
+considered as positive results. We identified 52 such modules. We manually completed the
 examples for 10 of those modules. Applying \texttt{dts-generate} with the completed
@@
 shown in \figref{fig:experiments-results-manually-completed-examples}.
-\begin{changethis}
-  Expected differences are present: optional parameters are declared both with the actual
-  and \texttt{undefined} type and the union type is handled through multiple overloads.
-\end{changethis}
+
 \begin{figure}[tp]
@@
 
-We discovered only 10 modules with \texttt{unsolvable} differences. Most of them were
-caused by the same error: 8 modules incorrectly declared an interface for the
+We discovered only 8 modules with \texttt{unsolvable} differences. Most of them were
+caused by the same error: 7 modules incorrectly declared an interface for the
 \texttt{string} type. This interface contains standard properties and methods from the
-JavaScript \texttt{String} object, such as \texttt{length} or \texttt{indexOf()}. The
+JavaScript \texttt{String} object, such as \texttt{length} or \texttt{indexOf()}. The remaining module is \texttt{stream}
 remaining modules are \texttt{duplexer2} and \texttt{duplexer3}. Both of them rely on native

### 2020-08-17 programming-2021/code/greet-interfaces/typescript-declaration-file/generated/greet-settings-module/index.d.ts
Update code examples

@@ -4,8 +4,8 @@ declare function GreetSettingsModule(settings: GreetSettingsModule.I__settings):
 declare namespace GreetSettingsModule {
 	export interface I__settings {
 		'greeting': string;
-		'duration': number | undefined;
-		'color': undefined | string;
+		'duration'?: number;
+		'color'?: string;
 	}
 
 }

### 2020-08-17 programming-2021/code/greet-organizing-types/typescript-declaration-file/generated/greet-organizing-module/index.d.ts

@@ -2,8 +2,8 @@ export = Greeter;
 
 declare class Greeter {
 	constructor(message: string);
-	log(logSetting: I__logSetting): void;
-	alert(alertSetting: I__alertSetting): void;
+	log(logSetting: Greeter.I__logSetting): void;
+	alert(alertSetting: Greeter.I__alertSetting): void;
 }
 
 declare namespace Greeter {

### 2020-08-17 programming-2021/code/greet-type-aliases/typescript-declaration-file/generated/greet-alias-module/index.d.ts

@@ -9,7 +9,5 @@ declare namespace Greet {
 		'greeting': string;
 	}
 
-export function greetAliasModule(init: string): void;
-export function greetAliasModule(init: Function): void;
-export function greetAliasModule(init: Greet|Greet.I__init): void;
+export function greetAliasModule(init: string | Function | Greet | Greet.I__init): void;
 }

### 2020-08-17 programming-2021/main.tex
Update highlighted box removing comments regarding undefined and question mark token

@@
     'greeting': string;
-    'duration' ?: number;
-    'color' ?: string;
+    'duration'?: number;
+    'color'?: string;
   }
@@
 
-\begin{changethis}
 We implemented only basic TypeScript features and we focused our effort in building an
@@
 \texttt{dts-parse} and those containing unimplemented features were ignored.
-\end{changethis}
 

### 2020-08-19 programming-2021/main.tex
Update example for github-url-to-object

@@
 
-declare function GithubUrlToObject(repoUrl: string|GithubUrlToObject.I__repoUrl, opts: undefined): object;
-declare function GithubUrlToObject(repoUrl: string|GithubUrlToObject.I__repoUrl, opts: GithubUrlToObject.I__opts): null|object;
-declare function GithubUrlToObject(repoUrl: GithubUrlToObject.I__repoUrl, opts: undefined): object;
+declare function GithubUrlToObject(repoUrl: string | GithubUrlToObject.I__repoUrl, opts?: GithubUrlToObject.I__opts): object | null;
 declare namespace GithubUrlToObject {
   export interface I__repoUrl {
-    'url': undefined | string;
+    'url'?: string;
   }
@@
   export interface I__opts {
-    'enterprise': undefined | boolean;
-  }
+    'enterprise'?: boolean;
+  }
+
 }
@@
 gh({url: 'git://github.com/monkey/business.git'});
-
+gh('this is not a proper url');
+gh({url: 'this is not a proper url'});
     \end{lstlisting}
@@
 \texttt{string} type. This interface contains standard properties and methods from the
-JavaScript \texttt{String} object, such as \texttt{length} or \texttt{indexOf()}. The remaining module is \texttt{stream}
-remaining modules are \texttt{duplexer2} and \texttt{duplexer3}. Both of them rely on native
-NodeJS types within the declaration file, which \texttt{dts-generate} could not identify
-properly.
+JavaScript \texttt{String} object, such as \texttt{length} or \texttt{indexOf()}.
 

### 2020-08-21 programming-2021/main.tex
Update number in abstract. Remove limiations regarding undefined type

@@
 compared with the corresponding declaration file provided on DefinitelyTyped.
-72 files presented positive results, having no differences at all or differences that can be solved by modifying the code examples accordingly.
+52 files presented positive results, having no differences at all or differences that can be solved by modifying the code examples accordingly.
 
@@
 Literal interface in DefinitelyTyped gets replace by a declared interface.
-The \texttt{undefined} type is used instead of the optional type.
 Return types are not considered.}

### 2020-08-26 programming-2021/main.tex
Add table with dts-parse statistics

@@
 end-to-end solution that generates valid and useful declaration files. Only the basic
-types \texttt{string}, \texttt{number}, \texttt{boolean} and its union were considered
+types \texttt{string}, \texttt{number}, \texttt{boolean}, its union and the optional type were considered
 both for function parameters and interface properties. Common types such as
 \texttt{any}, arrays, function callbacks, tuples, intersections and features such as
-generics, index signatures were not implemented. Each declaration file was tagged with
-\texttt{dts-parse} and those containing unimplemented features were ignored.
+generics, index signatures were not implemented.
+
+Each declaration file in the DT repository was tagged with
+\texttt{dts-parse}. We summarized the results in Table~\ref{tab:dts-parse-stats}. We based our decisions of prioritizing particular features on these results.
+
+\begin{table}[tp]
+  \begin{center}
+    \begin{tabular}{ |c|c|c| } 
+      \hline
+      \textbf{TypeScript feature} & \textbf{Count Declaration Files} & \textbf{\%} \\ 
+      \hline
+      string & 5086 & 76.50\% \\
+      optional-parameter & 4915 & 73.93\% \\
+      boolean & 3891 & 58.53\% \\
+      number & 3699 & 55.64\% \\
+      void & 3548 & 53.37\% \\
+      type-union & 3456 & 51.99\% \\
+      type-function & 3286 & 49.43\% \\
+      type-array & 3264 & 49.10\% \\
+      type-any & 3127 & 47.04\% \\
+      literals & 1925 & 28.96\% \\
+      alias-type & 1899 & 28.56\% \\
+      index-signature & 1271 & 19.12\% \\
+      generics-function & 1145 & 17.22\% \\
+      dot-dot-dot-token & 882 & 13.27\% \\
+      call-signature & 817 & 12.29\% \\
+      generics-interface & 718 & 10.80\% \\
+      object & 661 & 9.94\% \\
+      undefined & 577 & 8.68\% \\
+      type-intersection & 431 & 6.48\% \\
+      readonly & 373 & 5.61\% \\
+      type-tuple & 355 & 5.34\% \\
+      generics-class & 265 & 3.99\% \\
+      static & 251 & 3.78\% \\
+      private & 82 & 1.23\% \\
+      public & 43 & 0.65\% \\
+      protected & 19 & 0.29\% \\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{Distribution of TypeScript features in declaration files uploaded to DefinitelyTyped.
+The 2nd column shows how many files in DT make use of a specific TypeScript feature.
+The 3rd column shows the corresponding percentage for a total of 6648 declaration files in DT.
+Numbers were obtained from commit \texttt{da93d7b13094bacb170ead3f4a289f3b8687e4e5} (April 4, 2020).}
+  \label{tab:dts-parse-stats}
+\end{table}
 
 % \paragraph*{Interfaces}
-Finally, interfaces are created by exploring \texttt{getField} and
-\texttt{methodCall} interactions from the run-time information. We
+Finally, interfaces are created by exploring \texttt{getField} and \texttt{methodCall} interactions from the run-time information. We
 gather the interactions for a specific argument and build the
@@
 overblown because they are not widely used in DefinitelyTyped (see
-Table~\ref{table-of-occurrences}).
+Table~\ref{tab:dts-parse-stats}).
 

### 2020-08-26 programming-2021/main.tex
Correct email address. Simplified code example

@@
 \author{Fernando Cristiani}
-\authorinfo{contact \email{fc@example.org}}
-\affiliation{Hochschule Karlsruhe, Germany }
+\authorinfo{\email{fernando.cristiani@hs-karlsruhe.de}}
+\affiliation{Hochschule Karlsruhe, Germany}
 
@@
 \lstinline{type T = string | number; declare function F(a: T);} is equivalent to
-\lstinline{declare function F(a: string | number);}. The same approach applies to literal
+\lstinline{declare function} \lstinline{F(a: string | number);}. The same approach applies to literal
 interfaces. \texttt{dts-compare} does 
@@
   \centering
-  \begin{subfigure}{0.80\linewidth}
+  \begin{subfigure}{0.45\linewidth}
     \begin{lstlisting}[language=TypeScript]
@@
   interface Result {
-      user: string;
-      repo: string;
-      branch: string;
-      https_url: string;
-      tarball_url: string;
-      clone_url: string;
-      travis_url: string;
-      api_url: string;
-      zip_url: string;
+      ...
   }
@@
     \end{lstlisting}
-    \caption{github-url-to-object/index.d.ts - DefinitelyTyped}
-  \end{subfigure}
-
-  \begin{subfigure}{0.80\linewidth}
+    \caption{github-url-to-object/index.d.ts - DefinitelyTyped. Properties of interface \texttt{Result} were ignored for readability, since return types were not considered.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{0.45\linewidth}
     \begin{lstlisting}[language=TypeScript]
@@
 
-\begin{changethis}
-  We exposed the fundamental aspect of capturing the developer's intention when inferring
-  types in JavaScript. Instead of applying constraints and restrictions for operations
-  with certain types, we presented a proposal that favors best practice. Uncommon usage is
-  not forbidden but greatly discouraged. Accordingly, we collected evidence regarding the
-  usage of JavaScript operators by analyzing 400 libraries.
-\end{changethis}
+% \begin{changethis}
+%   We exposed the fundamental aspect of capturing the developer's intention when inferring
+%   types in JavaScript. Instead of applying constraints and restrictions for operations
+%   with certain types, we presented a proposal that favors best practice. Uncommon usage is
+%   not forbidden but greatly discouraged. Accordingly, we collected evidence regarding the
+%   usage of JavaScript operators by analyzing 400 libraries.
+% \end{changethis}
+
 % Finally, the architecture is composed of different blocks that interact with each

### 2020-08-27 programming-2021/main.tex

@@
 % and evaluated them against the DefinitelyTyped repository.
+
+The implementation and the results are available under the following repositories:
+
+\begin{itemize}
+  \item \url{https://github.com/proglang/run-time-information-gathering/tree/v1.1}
+  \item \url{https://github.com/proglang/ts-declaration-file-generator/tree/v1.1}
+  \item \url{https://github.com/proglang/ts-declaration-file-generator-service/tree/v1.1}
+  \item \url{https://github.com/proglang/dts-generate-method/tree/v1.1}
+  \item \url{https://github.com/proglang/dts-generate-results/tree/v1.1}
+\end{itemize}
 

### 2020-08-27 programming-2021/main.tex
section 0-4

@@
 \title{Generation of TypeScript Declaration Files from JavaScript Code}
-
+\subtitle{Catching the Developer's Intent from Examples}
 % \titlerunning{Dummy short title}%optional, please use if title is longer than one line
@@
 \authorinfo{\email{thiemann@informatik.uni-freiburg.de}}
-\affiliation{Universität Freiburg, Germany}
+\affiliation{Albert-Ludwigs-Universität Freiburg, Germany}
 
@@
 repository for these files.
-The repository is maintained manually by volunteers, which
+The repository is populated and maintained manually by volunteers, which
 is error prone and time consuming. Discrepancies between a
@@
 run-time information, and generates a declaration file based on this
-information. To evaluate the tool, 244 declaration files were generated and 82 were
-compared with the corresponding declaration file provided on DefinitelyTyped.
-52 files presented positive results, having no differences at all or differences that can be solved by modifying the code examples accordingly.
-
+information. To evaluate the tool, 244 declaration files were
+generated directly from an \NPM{} module and 82 of these were
+compared with the corresponding declaration file provided on
+DefinitelyTyped. We obtained positive results for 52 files, which
+exhibited no differences at all or differences that can be resolved by
+modifying the developer-provided examples. 
 \end{abstract}
@@
 \begin{lstlisting}[
-    caption={\textbf{Unintuitive JavaScript behavior} Falsy values, \lstinline!typeof!, \lstinline!null! and \lstinline!undefined! operators and type coercion},
+    caption={\textbf{Unintuitive JavaScript behavior:} Falsy values,
+    \lstinline!null! vs. \lstinline!undefined!,
+ \lstinline!typeof!, and type coercion},
     language=JavaScript,
@@
 null == undefined; // true
+null + undefined + [1, 2, 3] // 'NaN1,2,3' 
+typeof null; // object
+null instanceof Object; // false
 [1] + 1; //'11'
 [2] == "2"; // true
-null + undefined + [1, 2, 3] // 'NaN1,2,3'
 "hello world".lenth + 1 // NaN | note 'lenth' instead of 'length'
 [1, 15, 20, 100].sort() // [ 1, 100, 15, 20 ]
-typeof null; // object
-null instanceof Object; // false
-"01" < "00100"; // false
-\end{lstlisting}
+\end{lstlisting}
+% "01" < "00100"; // false 
 % pjt: why unexpected?
@@
 
-Unfortunately, creation and maintenance for most declaration
+Unfortunately, creation and maintenance of most declaration
 files in DefinitelyTyped is conducted manually,
-which is error prone and time consuming. As TypeScript takes a
-declaration file at face value, discrepancies are not detected at
-compile time and the resulting inaccuracies in autocompletion, say,
-are confusing for the programmer.
+which is time consuming and  error prone. Errors are aggravated because TypeScript takes a
+declaration file at face value. Discrepancies between declaration and
+implementing JavaScript library are not detected at 
+compile time and result in incorrect behavior (e.g., autocompletion
+hints) of the IDE. 
 As TypeScript does not perform any run-time 
-checking of types, either,  discrepancies  between the declaration file
-and its corresponding JavaScript library can lead to unexpected 
+checking of types, these  discrepancies can lead to unexpected 
 behavior and crashes. Such an experience can lead to developer
@@
 sound TypeScript declaration files automatically. To do so, we rely on the examples provided
-as part of the documentation of a library. If these
+by the developer as part of the documentation of a library. If these
 examples are sufficiently rich, then our toolchain can generate
-high-quality declaration files from them. This approach takes
+high-quality declaration files for the library. This approach takes
 advantage of 
@@
 
-As an example, consider the NPM module
+As an example, consider the NPM\footnote{NPM is the \href{https://www.npmjs.com/}{Node package
+  manager}, an online repository for JavaScript modules.} module
 \texttt{abs} that
@@
   the DefinitelyTyped repository and found 244 sufficiently
-  well-documented \NPM{} packages, on which we ran \texttt{dts-generate}
+  well documented \NPM{} packages, on which we ran \texttt{dts-generate}
   and compared the outcome with the respective declaration file from the
@@
 glob expression for matching filenames in the shell into a regular
-expression\footnote{https://www.npmjs.com/package/glob-to-regexp}. It 
+expression\footnote{\url{https://www.npmjs.com/package/glob-to-regexp}}. It 
 has about 9,900,000 weekly downloads and 188 \NPM{} packages depend on it. If
@@
 file can be generated automatically using
-\texttt{dts-generate}. Our comparator tool can compare the new file
+\texttt{dts-generate}. Our comparator tool (Section~\ref{sec:dts-compare}) can compare the new file
 for incompatibilities with the previous declaration file.
@@
 describes a package with a single exported function. The first parameter is of type
-\texttt{string} and the second one is an optional interface. TypeScript namespaces
+\texttt{string} and the second one is an optional object described by an interface. TypeScript namespaces
 organize types declared in a declaration file and avoid name clashes with other types
-\cite{typescript-namespaces}. For example, the interface \texttt{I\_\_opts} declared in
-\figref{fig:motivating-example-glob-to-regexp-vscode} is used as
-\lstinline[language=TypeScript]{globToRegexp.I__opts} because it is declared within a
-namespace.
+\cite{typescript-namespaces}. For example, the interface
+\lstinline[language=TypeScript]{I__opts} is declared in namespace
+\lstinline[language=TypeScript]{globToRegexp} in
+\figref{fig:motivating-example-glob-to-regexp-vscode}. Hence its uses
+must be qualified by the namespace as in
+\lstinline[language=TypeScript]{globToRegexp.I__opts}.
 
@@
 declaration
-files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/templates.html}:
+files\footnote{\url{https://www.typescriptlang.org/docs/handbook/declaration-files/templates.html}}:
 \textbf{module}, \textbf{module-class}, and
 \textbf{module-function}. 
-Each template corresponds to a different way of describing the exports
+Each template corresponds to a different way of organizing the exports
 of a JavaScript library. The choice of the template depends on the
-organization of the underlying JavaScript library:
+structure of the underlying JavaScript library:
 \begin{description}
@@
 \end{description}
+There are further templates, but most libraries fall in one of these
+three categories.
+Both libraries \texttt{glob-to-regexp} and \texttt{abs} are instances of the
+\textbf{module-function} template.
 
 \begin{lstlisting}[
-    caption={Example for template \textbf{module-function}},
+    caption={Example for \textbf{module-function} template},
     language=bash,
@@
     \end{lstlisting}
-    \caption{Example for \texttt{module-function} template}
+    \caption{Example for \textbf{module-function} template}
     \label{fig:example-module-function}
@@
     \end{lstlisting}
-    \caption{Declaration file for \texttt{module-function} template}
+    \caption{Declaration file for \textbf{module-function} template}
     \label{fig:template-module-function}
@@
       \end{lstlisting}
-    \caption{Example for \texttt{module} template}
+    \caption{Example for \textbf{module} template}
     \label{fig:example-module}
@@
       \end{lstlisting}
-      \caption{Declaration file for \texttt{module} template}
+      \caption{Declaration file for \textbf{module} template}
       \label{fig:template-module}
@@
       \end{lstlisting}
-      \caption{Example for \texttt{module-class} template}
+      \caption{Example for \textbf{module-class} template}
       \label{fig:example-class}
@@
       \end{lstlisting}
-      \caption{Declaration file for \texttt{module-class} template}
+      \caption{Declaration file for \textbf{module-class} template}
       \label{fig:template-class}
@@
 Like \texttt{dts-gen}, a TypeScript package that generates a template for a declaration
-file, we determine the kind of declaration file by examining the provided example. If the
+file, we determine the kind of declaration file by examining the usage
+of the module. If the
 imported entity is used solely as a function as in Figure~\ref{fig:example-module-function}, the
-generated declaration file follows the module-function template as shown in Figure~\ref{fig:template-module-function}.
+generated declaration file follows the \textbf{module-function} template as shown in Figure~\ref{fig:template-module-function}.
 If the example only accesses properties of the imported entity as in
-Figure~\ref{fig:example-module}, we generate according to the module template (Figure~\ref{fig:template-module}).
+Figure~\ref{fig:example-module}, we generate according to the \textbf{module} template (Figure~\ref{fig:template-module}).
 If the imported entity is used with \lstinline/new/ to create new instances as in
-Figure~\ref{fig:example-class}, we generate a module-class template (Figure~\ref{fig:template-class}).
-
-The libraries \texttt{glob-to-regexp} as well as \texttt{abs} are both instances of the
-\textbf{module-function} template.
-Most libraries fall in one of these three categories. 
+Figure~\ref{fig:example-class}, we generate a \textbf{module-class} template (Figure~\ref{fig:template-class}).
 While it is possible to create a library of undetermined category, our selection is driven
@@
 analysis modules that we extended as needed to retrieve the required
-run-time information, which is then saved in a JSON file.
+run-time information.
+%, which is then saved in a JSON file.
 
@@
 process. Its contents mimic the usage of the library in the example
-code fragments and matches the structure of the
+code fragments and match the structure of the
 JavaScript library under analysis, so that the JavaScript code
@@
 
-The command-line interface is inspired by the \texttt{dts-gen} tool
-\cite{dts-gen}. \coderef{code:dts-generate-example} shows
-that invoking the package is very simple. Its only required
+The command-line interface is very simple and inspired by the \texttt{dts-gen} tool
+\cite{dts-gen} (see \coderef{code:dts-generate-example}). The only required
 argument is the name of the module published to the \NPM{} registry.
@@
 To extract run-time information from a JavaScript library, it is
-necessary, by definition, to actually execute the code, because the
-analysis modules provided by Jalangi to gather information are only
-triggered if the instrumented code gets executed.
+necessary, by definition, to execute the code, because the
+analysis modules provided by Jalangi can only gather information if the instrumented code gets executed.
 
@@
 to the importing library, which also needs code to drive it. Moreover,
-it is costly to download and instrument another package.
+we need to identify a dependent package, which is costly to download and instrument.
 
@@
 from the consumer's point of view. For example, a developer might write a test invoking a
-function with \texttt{null} just to validate that it throws an error in such scenario.
-
-In the end, option~\ref{item:3} was the most viable even though there
+function with \texttt{null} just to validate that it throws an error
+in such a scenario.
+
+In the end, option~\ref{item:3} was the most viable option even though there
 is no standard for documentation, either. However, almost all repositories
@@
 developers provide code examples in the README files to show how the
-library works and how to use it. These code fragments represent common usage cases rather
+library works and how to use it. These code fragments showcase common use cases rather
 than stress-testing the implementation (the problem of
@@
 \label{sec:run-time-information}
-The Runtime Information block described in
+The Runtime Information block in
 \figref{fig:tsd_generation_method_block_diagram} gathers
@@
 
-Instead of analyzing the shape of the exported module, we inspect how the module is used
-to choose between the templates. We analyze the properties \texttt{isExported},
-\texttt{requiredModule} and \texttt{isConstructor} from the run-time JSON to distinguish
-between \texttt{module-class} and \texttt{module-function}. If a function is invoked and
+Instead of analyzing the shape of the exported module, we choose the
+template based on how the module is used. We analyze the properties \texttt{isExported},
+\texttt{requiredModule}, and \texttt{isConstructor} from the
+information gathered at run time to distinguish
+between \textbf{module-class} and \textbf{module-function}. If a function is invoked and
 it is imported from the module we are analyzing we infer the template
-\texttt{module-class} or \texttt{module-function}. We choose \texttt{module-class} if the
+\textbf{module-class} or \textbf{module-function}. We choose \textbf{module-class} if the
 function is used as a constructor. Otherwise, we choose
-\texttt{module-function}.
-If no function is invoked, we use the \texttt{module}
+\textbf{module-function}.
+If no function is invoked, we use the \textbf{module}
 template.
@@
 end-to-end solution that generates valid and useful declaration files. Only the basic
-types \texttt{string}, \texttt{number}, \texttt{boolean}, its union and the optional type were considered
-both for function parameters and interface properties. Common types such as
-\texttt{any}, arrays, function callbacks, tuples, intersections and features such as
+types \texttt{string}, \texttt{number}, \texttt{boolean}, unions
+thereof, and optional elements were considered
+both for function parameters and interface properties. We also
+considered overloaded functions.
+Common types such as
+\texttt{any}, arrays, function callbacks, tuples, intersections (that
+is, types formed using the intersection operator \lstinline[language=TypeScript]/&/) and features such as
 generics, index signatures were not implemented.
 
-Each declaration file in the DT repository was tagged with
-\texttt{dts-parse}. We summarized the results in Table~\ref{tab:dts-parse-stats}. We based our decisions of prioritizing particular features on these results.
+Our selection of type features covered in the tool is driven by
+examination of the files in
+the DefinitelyTyped repository. We parsed each declaration file with
+\texttt{dts-parse} and tagged it according to the features used in the file.
+Table~\ref{tab:dts-parse-stats} summarizes the results.
+We decided to prioritize the features with the highest number of
+occurrences and support them first in our framework.
 
@@
   \begin{center}
-    \begin{tabular}{ |c|c|c| } 
+    \begin{tabular}{ |l|c|r| } 
       \hline
-      \textbf{TypeScript feature} & \textbf{Count Declaration Files} & \textbf{\%} \\ 
+      \textbf{TypeScript feature} & \textbf{Count} & \multicolumn{1}{|c|}{\textbf{\%}} \\ 
       \hline
-      string & 5086 & 76.50\% \\
+      type-string & 5086 & 76.50\% \\
       optional-parameter & 4915 & 73.93\% \\
-      boolean & 3891 & 58.53\% \\
-      number & 3699 & 55.64\% \\
-      void & 3548 & 53.37\% \\
+      type-boolean & 3891 & 58.53\% \\
+      type-number & 3699 & 55.64\% \\
+      type-void & 3548 & 53.37\% \\
       type-union & 3456 & 51.99\% \\
@@
       type-any & 3127 & 47.04\% \\
-      literals & 1925 & 28.96\% \\
+      type-literals & 1925 & 28.96\% \\
       alias-type & 1899 & 28.56\% \\
@@
       generics-interface & 718 & 10.80\% \\
-      object & 661 & 9.94\% \\
-      undefined & 577 & 8.68\% \\
+      type-object & 661 & 9.94\% \\
+      type-undefined & 577 & 8.68\% \\
       type-intersection & 431 & 6.48\% \\
@@
   \end{center}
-  \caption{Distribution of TypeScript features in declaration files uploaded to DefinitelyTyped.
+  \caption{Distribution of TypeScript features in declaration files uploaded to DefinitelyTyped (DT).
 The 2nd column shows how many files in DT make use of a specific TypeScript feature.

### 2020-08-27 programming-2021/main.tex
in section 6

@@
 
-Our selection of type features covered in the tool is driven by
+The selection of typing features covered in the tool is driven by
 examination of the files in
@@
 For each module, we applied \texttt{dts-compare} to the generated declaration file and the
-one uploaded to the DefinitelyTyped repository for the same module, as shown in
+one uploaded to the DefinitelyTyped repository for the same module as shown in
 \figref{fig:evaluation-diagram}. While this approach does not provide an absolute measure
@@
 The first step is to parse the declaration files, which can be
-achieved by using the TypeScript Compiler API, a library which traverse the
+achieved by using the TypeScript Compiler API, a library which traverses the
 Abstract Syntax Tree of a TypeScript program
@@
 The output of \texttt{dts-parse} is a structure where declared
-{interfaces}, {functions}, {classes} and
+{interfaces}, {functions}, {classes}, and
 {namespaces} are stored separately. Function arguments are
@@
 callbacks. Optional parameters are also identified. For
-{classes}, a distinction between the constructor and methods
-is made. Declaration files are tagged according to their features, which is useful for identifying and filtering out declaration files containing unimplemented features.
+{classes}, a distinction between constructors and methods
+is made. Declaration files are tagged according to their features,
+which is useful for identifying and filtering out declaration files
+that contain unimplemented features.
 
 \texttt{dts-parse} is itself written in TypeScript. It receives the file as input
-parameter and returns an AST serialized in JSON, as shown in \coderef{code:dts-parse-example}.
-
-\begin{lstlisting}[
-  caption={Example for \texttt{dts-parse} executed on module \texttt{glob-to-regexp} and combined with \texttt{jq} to browse the returned JSON},
-  language=bash,
-label=code:dts-parse-example,
-  float=tp, captionpos=b
-]
-$ dts-parse -i glob-to-regexp/index.d.ts | jq '.parsing.functions[0].name'
-"GlobToRegexp"
-\end{lstlisting} %$
+parameter and returns an AST serialized in JSON.
+
+% , as shown in \coderef{code:dts-parse-example}.
+% \begin{lstlisting}[
+%   caption={Example for \texttt{dts-parse} executed on module \texttt{glob-to-regexp} and combined with \texttt{jq} to browse the returned JSON},
+%   language=bash,
+% label=code:dts-parse-example,
+%   float=tp, captionpos=b
+% ]
+% $ dts-parse -i glob-to-regexp/index.d.ts | jq '.parsing.functions[0].name'
+% "GlobToRegexp"
+% \end{lstlisting} %$
 
@@
 \label{sec:dts-compare}
-The tool is written in TypeScript. It takes two declaration files as input and returns a
-JSON file containing the result of the comparison, as shown in
-\coderef{code:dts-compare-example}. Following the naming convention of any standard Unit
-Testing framework, the input files are marked as \lstinline{expected} or
+
+% The tool is written in TypeScript.
+The compare tool takes two declaration files as input and returns a
+JSON file containing the result of the comparison as shown in
+\coderef{code:dts-compare-example}. Following the standard naming convention of unit
+testing frameworks, the input files are marked as \lstinline{expected} or
 \lstinline{actual}. We use \lstinline{expected} for the DefinitelyTyped
@@
 of an abstract entity \texttt{Difference}. 
-
-We extracted code examples written by
-developers to execute the code and gather run-time information. The quality of the
-generated declaration file is heavily dependent on the code examples. For example, if a
-property of an interface does not get accessed due to a code example not exploring a
-particular execution path, that property will not exist in the generated declaration
-file.
-
-To measure the impact of incomplete code examples on the
-quality of generated declaration files we categorized each difference as follows: 
-\begin{itemize}
-  \item \texttt{solvable}: Differences that can be resolved by adding more code
+A difference is \texttt{solvable} if it can be resolved by adding more code
 examples. For example, invoking a function with different parameters so that a missing
-    interface property gets accessed. 
-  \item \texttt{unsolvable}: Differences that cannot be resolved with further examples.
-    % are marked as \texttt{unsolvable}. These differences are implementation errors of \texttt{dts-generate}.
-\end{itemize}
+interface property gets accessed. Otherwise it is  \texttt{unsolvable}.
 
@@
   \item \texttt{TemplateDifference}: A difference between the choice of TypeScript
-    templates. For example, if the file in DefinitelyTyped is written as a \texttt{module}
-    but we generated the file using the \texttt{module-function} template. The comparison
+    templates. For example, if the file in DefinitelyTyped is written as a \textbf{module}
+    but we generated the file using the \textbf{module-function} template. The comparison
     stops if the templates differ. 
   \item \texttt{ExportAssignmentDifference}: An equality check between the export
-    assignments of both files, that is in the expression \lstinline{export = XXX}. 
+    assignments of both files, that is in the expression
+    \lstinline{export = XXX}.
+    \begin{changethis}
+      Q: how severe is that? That's not an issue, right?
+    \end{changethis}
   \item \texttt{FunctionMissingDifference}: A function declaration is present in the expected
     file but not in the generated one. This difference applies to functions as well as methods of classes
-    and interfaces. 
+    and properties of interfaces. 
   \item \texttt{FunctionExtraDifference}: A function is present in the generated declaration file but not in DefinitelyTyped.
@@
 function return types is not covered in the current version of \texttt{dts-generate}.
+
+\begin{changethis}
+
+To measure the impact of incomplete code examples on the
+quality of generated declaration files we categorized each difference as follows:
+
+  We extracted code examples written by developers to execute the code
+  and gather run-time information. The quality of the generated
+  declaration file is heavily dependent on the code examples. For
+  example, if a property of an interface does not get accessed due to
+  a code example not exploring a particular execution path, that
+  property will not exist in the generated declaration file.
+\end{changethis}
 
@@
 using the \texttt{javascript} or \texttt{js} tags. However, as we are
-still left with code examples for 2200 modules, we did not further
+still left with code examples for 2260 modules, we did not further
 look into code extraction as this number was considered sufficient for

### 2020-08-28 programming-2021/main.tex
up to section 6.3

@@
 \lstinline{dts-generate}, the right side shows the corresponding file
-in the DefinitelyTyped repository. There are no relevant
-differences between the files. Functions are correctly detected; input and output types are accurately inferred.
-
-The differences in the export assignment by modules \texttt{dirname-regex} and
-\texttt{escape-html} are not problematic because the module can receive an arbitrary 
-name when it is imported. \texttt{dts-generate} automatically generates the export
+in the DefinitelyTyped repository. 
+There are no relevant differences between the files. 
+The different names in the export assignment by modules \texttt{dirname-regex} and
+\texttt{escape-html} do not matter because the module can receive an arbitrary 
+name when it is imported: in
+\lstinline[language=TypeScript]/import Abs = require('abs');/ the
+identifier \lstinline[language=TypeScript]/Abs/ can be chosen by the
+programmer. We generate the name in the export 
 assignment by transforming the module name into camel case form, following TypeScript
@@
 
-The parameter names are extracted from the variable names of the JavaScript
+Functions are correctly detected; input and output types are
+accurately inferred. 
+We extract the parameter names from the variable names of the JavaScript
 code.
-% The JavaScript implementation of module \texttt{escape-html} declares the
-% first parameter of the exported function with the name \texttt{string} and not
-% \texttt{text}, as specified in DefinitelyTyped. 
 A difference in the names of the function parameters does not affect the correctness of
 the declaration file.
-Furthermore, the module \texttt{escape-html} from DefinitelyTyped declares an empty
-namespace, which is not generated by our tool.
+Furthermore, neither the order of declarations nor unused namespaces
+(as in module \texttt{escape-html}) matter.
 
@@
     \centering
-    \begin{subfigure}{0.48\linewidth}
+    \begin{subfigure}[t]{0.48\linewidth}
       \begin{lstlisting}[language=TypeScript,numbers=none]
@@
     \hfill
-    \begin{subfigure}{0.48\linewidth}
+    \begin{subfigure}[t]{0.48\linewidth}
       \begin{lstlisting}[language=TypeScript,numbers=none]
@@
       \end{lstlisting}
-      \caption{abs/index.d.ts - DefinitelyTyped}
-    \end{subfigure}
-
-    \begin{subfigure}{0.48\linewidth}
+      \caption{abs/index.d.ts - DT}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{0.48\linewidth}
         \begin{lstlisting}[language=TypeScript,numbers=none]
@@
       \hfill
-      \begin{subfigure}{0.48\linewidth}
+      \begin{subfigure}[t]{0.48\linewidth}
         \begin{lstlisting}[language=TypeScript,numbers=none]
@@
         \end{lstlisting}
-        \caption{dirname-regex/index.d.ts - DefinitelyTyped}
-      \end{subfigure}
-
-      \begin{subfigure}{0.48\linewidth}
+        \caption{dirname-regex/index.d.ts - DT}
+      \end{subfigure}
+
+      \begin{subfigure}[t]{0.48\linewidth}
         \begin{lstlisting}[language=TypeScript,numbers=none]
@@
 declare function EscapeHtml(string: string): string;
+
         \end{lstlisting}
@@
         \end{lstlisting}
-        \caption{escape-html/index.d.ts - DefinitelyTyped}
+        \caption{escape-html/index.d.ts - DT}
       \end{subfigure}
@@
       \end{lstlisting}
-      \caption{is-uuid/index.d.ts - DefinitelyTyped}
+      \caption{is-uuid/index.d.ts - DT}
     \end{subfigure}
@@
 
-An example for the \textbf{module} template
-is shown for the \texttt{is-uuid} module in
-\figref{fig:experiments-results-module-is-uuid}.
-The generated file does not contain methods that were not executed by the extracted
-examples. The invoked function are correctly detected.
-
-Finally, we were unable to  generate a declaration file based on the
-\texttt{module-class} template. All modules corresponding to this template exhibited at least one of
+\figref{fig:experiments-results-module-is-uuid} contains an example
+for the \textbf{module} template, the declaration for the \texttt{is-uuid} module.
+The generated file only contain methods that were executed by the extracted
+examples. All invoked functions are correctly detected; the additional
+functions declared on DefinitelyTyped are not used in the example
+code.
+It would be easy to generate a perfectly matching declaration file by
+adding two examples using functions \texttt{nil} and
+\texttt{anyNonNil}.
+
+We were unable to  generate a declaration file based on the
+\textbf{module-class} template. All modules corresponding to this template exhibited at least one of
 the unimplemented TypeScript features.
@@
 \figref{fig:experiments-results-module-glob-base}. Function return type interfaces are
-not inferred by \texttt{dts-generate}. The parameter name is \texttt{pattern}
-instead of \texttt{basePath}, because the JavaScript implementation declares the parameter
-as \texttt{pattern}, as shown in \figref{fig:subfloat-globbase-js-implementation}. We also
+not inferred by \texttt{dts-generate}.
+% The parameter name is \texttt{pattern}
+% instead of \texttt{basePath}, because the JavaScript implementation declares the parameter
+% as \texttt{pattern}, as shown in \figref{fig:subfloat-globbase-js-implementation}. 
+We also
 discovered errors in the selected template in DefinitelyTyped. The module
-\texttt{smart-truncate} in DefinitelyTyped uses the \texttt{module} template, but
-\texttt{dts-generate} generates a file using the \texttt{module-function}
+\texttt{smart-truncate} in DefinitelyTyped uses the \textbf{module} template, but
+\texttt{dts-generate} generates a file using the \textbf{module-function}
 template. \coderef{code:experiments-results-module-smart-truncate} shows that the module is indeed
@@
 
-    \begin{subfigure}{0.48\linewidth}
-      \begin{lstlisting}[language=JavaScript,numbers=none]
-var globBase = require('glob-base');
-
-globBase();
-
-// Throws: TypeError: glob-base expects a string.
-      \end{lstlisting}
-      \caption{Code example invoking the exported function without the first parameter, as specified in DefinitelyTyped}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.48\linewidth}
+    \begin{subfigure}[t]{0.48\linewidth}
       \begin{lstlisting}[language=TypeScript,numbers=none]
@@
       \end{lstlisting}
-      \caption{glob-base/index.js - Fraction of JS implementation that throws an Error when input parameter is not a string}
+      \caption{glob-base/index.js - Excerpt from JS implementation
+        that throws an Error when input parameter is not a string} 
       \label{fig:subfloat-globbase-js-implementation}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.48\linewidth}
+      \begin{lstlisting}[language=JavaScript,numbers=none]
+var globBase = require('glob-base');
+
+globBase();
+
+// Throws: TypeError: glob-base expects a string.
+      \end{lstlisting}
+      \caption{Code example invoking the exported function without the first parameter, as specified in DefinitelyTyped}
     \end{subfigure}

### 2020-08-28 programming-2021/main.tex
all sections processed

@@
 
-\begin{figure}[tp]
+\begin{figure}[t]
   \centering
-  \begin{subfigure}{0.48\linewidth}
+  \begin{subfigure}[t]{0.48\linewidth}
     \begin{lstlisting}[language=JavaScript,numbers=none]
@@
   \hfill
-  \begin{subfigure}{0.48\linewidth}
+  \begin{subfigure}[t]{0.48\linewidth}
     \begin{lstlisting}[language=TypeScript,numbers=none]
@@
 
-  \begin{subfigure}{0.48\linewidth}
+  \begin{subfigure}[t]{0.48\linewidth}
       \begin{lstlisting}[language=JavaScript,numbers=none]
@@
 
-    \begin{subfigure}{0.48\linewidth}
+    \begin{subfigure}[t]{0.48\linewidth}
       \begin{lstlisting}[language=JavaScript,numbers=none]
@@
       \begin{lstlisting}[language=TypeScript,numbers=none]
-export = GlobBase;
-
-declare function GlobBase(pattern: string): object;
-      \end{lstlisting}
-      \caption{glob-base/index.d.ts - Generated}
+'use strict';
+
+// ...
+
+module.exports = function globBase(pattern) {
+  if (typeof pattern !== 'string') {
+    throw new TypeError('glob-base expects a string.');
+  }
+
+  // ...
+} 
+      \end{lstlisting}
+      \caption{glob-base/index.js - Excerpt of JS implementation
+        that throws an Error when input parameter is not a string} 
+      \label{fig:subfloat-globbase-js-implementation}
     \end{subfigure}
@@
 \begin{subfigure}[t]{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript,numbers=none]
-'use strict';
-
-// ...
-
-module.exports = function globBase(pattern) {
-  if (typeof pattern !== 'string') {
-    throw new TypeError('glob-base expects a string.');
-  }
-
-  // ...
-} 
-      \end{lstlisting}
-      \caption{glob-base/index.js - Excerpt from JS implementation
-        that throws an Error when input parameter is not a string} 
-      \label{fig:subfloat-globbase-js-implementation}
+      \begin{lstlisting}[language=JavaScript,numbers=none]
+var globBase = require('glob-base');
+
+globBase();
+
+// Throws: TypeError: glob-base expects a string.
+      \end{lstlisting}
+      \caption{Code example invoking the exported function without the first parameter, as specified in DefinitelyTyped}
     \end{subfigure}
@@
     \begin{subfigure}[t]{0.48\linewidth}
-      \begin{lstlisting}[language=JavaScript,numbers=none]
-var globBase = require('glob-base');
-
-globBase();
-
-// Throws: TypeError: glob-base expects a string.
-      \end{lstlisting}
-      \caption{Code example invoking the exported function without the first parameter, as specified in DefinitelyTyped}
+      \begin{lstlisting}[language=TypeScript,numbers=none]
+export = GlobBase;
+
+declare function GlobBase(pattern: string): object;
+      \end{lstlisting}
+      \caption{glob-base/index.d.ts - Generated}
     \end{subfigure}
@@
 
-\begin{figure}[tp]
+\begin{figure}[t]
   \centering
-  \begin{subfigure}{0.45\linewidth}
+  \begin{subfigure}[t]{0.45\linewidth}
     \begin{lstlisting}[language=TypeScript]
@@
   \hfill
-  \begin{subfigure}{0.45\linewidth}
+  \begin{subfigure}[t]{0.45\linewidth}
     \begin{lstlisting}[language=TypeScript]
@@
     \end{subfigure}
-  \caption{Manually expanded code example for \texttt{github-url-to-object} to match DefinitelyTyped declaration file.
-Literal interface in DefinitelyTyped gets replace by a declared interface.
-Return types are not considered.}
+  \caption{Manually expanded code example for
+    \texttt{github-url-to-object} to match DefinitelyTyped declaration
+    file. The literal interface in DefinitelyTyped is replaced by a
+    declared interface. Return types are not considered.} 
   \label{fig:experiments-results-manually-completed-examples}
@@
 caused by the same error: 7 modules incorrectly declared an interface for the
-\texttt{string} type. This interface contains standard properties and methods from the
+\texttt{string} type. This interface contains standard properties and methods of the
 JavaScript \texttt{String} object, such as \texttt{length} or \texttt{indexOf()}.
@@
 
-How can we improve further? Given that JavaScript developers are more incentivized to develop good
-examples than writing ``boring'' type declaration, how can we capitalize on that attitude?
+How can we improve further? Given that JavaScript developers are incentivized to develop good
+examples rather than writing ``boring'' type declarations, how can we capitalize on that attitude?
 The main issue is that any example-driven algorithm can only come up with an
@@
 greater than some threshold, in which case the generator generalizes to type
-\lstinline/string/. A further alternative would be to communicate to the developer a set
-of typical example strings, \emph{marker strings}, that the generator always resolves to the \lstinline/string/
-type.
+\lstinline/string/. Both alternatives come with the problem that they
+create a local over-approximation, so that it would become harder to
+qualify the output of the generator. A further alternative that does
+not suffer from this drawback would be to communicate to the developer a set
+of typical example strings, \emph{marker strings}, that the generator
+always generalizes to the \lstinline/string/ type.
 
@@
 wants to advertise the type \lstinline/any/ for an argument? This intent is hard to
-advertise with a finite number of examples.
+communicate with a finite number of examples.
 
@@
 \label{sec:related-work}
-\paragraph*{Microsoft's dts-gen}
-Microsoft developed \texttt{dts-gen}, a tool that creates declaration files for
+\paragraph*{dts-gen}
+TypeScript comes with \texttt{dts-gen}, a tool that creates declaration files for
 JavaScript libraries \cite{dts-gen}. Its documentation states that the result is 
-intended to be used a starting point for development of a declaration file. The outcome
+intended to be used as a starting point for the development of a declaration file. The outcome
 needs to be refined afterwards by the developer.
@@
 The tool analyzes the shape of the objects at runtime after initialization without
-executing the library. This results in many variables being inferred as
-\lstinline[language={}]{any}. \coderef{code:related-work-dts-gen-example} shows an example
-for module \lstinline[language={}]{abs}.
+executing the library. This results in most parameters and results being inferred as
+\lstinline[language={}]{any}.
+% \coderef{code:related-work-dts-gen-example} shows an example
+% for module \lstinline[language={}]{abs}.
 
@@
 amount of manual work that a developer needs to do on a declaration file after updating
-JavaScript code increases the risk for having discrepancies between the declaration file
+JavaScript code increases the risk of discrepancies between the declaration file
 and the implementation.
@@
 
-\begin{lstlisting}[
-    language=bash,
-    caption={Microsoft's dts-gen example - A declaration file for module \lstinline!abs!
-is generated.
-Types are inferred as \lstinline!any!.
-The correct \lstinline!module-function!
-template is used.},
-	label=code:related-work-dts-gen-example,
-    float=tp,captionpos=b
-]
-$ npm i -g dts-gen
-$ npm i -g abs
-$ dts-gen -m abs
-Wrote 5 lines to abs.d.ts.
-
-$ cat abs.d.ts
-/** Declaration file generated by dts-gen */
-
-export = abs;
-
-declare function abs(input: any): any;
-\end{lstlisting}
+%$ npm i -g dts-gen
+% $ npm i -g abs
+% $ dts-gen -m abs
+% Wrote 5 lines to abs.d.ts.
+
+% $ cat abs.d.ts
+
+% \begin{lstlisting}[
+%     language=TypeScript,
+%     caption={Microsoft's dts-gen example - A declaration file for
+% module \lstinline!abs! is generated. Types are inferred as
+% \lstinline!any!. The correct \textbf{module-function} template is
+% used.}, 
+% 	label=code:related-work-dts-gen-example,
+%     float=tp,captionpos=b
+% ]
+% /** Declaration file generated by dts-gen */
+
+% export = abs;
+
+% declare function abs(input: any): any;
+% \end{lstlisting}
 

### 2020-08-31 programming-2021/main.tex
example packages for literal types

@@
 \paragraph*{Literal types vs string}
-JavaScript libraries often (?) use literal strings to select
-options and configurations. For example, \dots
+Many JavaScript libraries use literal strings to select
+options and configurations. To be concrete, 1925 DT packages rely on
+literal types (see Table~\ref{tab:dts-parse-stats}). For example, the bonjour
+package\footnote{\url{https://www.npmjs.com/package/@types/bonjour}}
+relies on strings to control events and indicate protocols in (different) interfaces:
+\begin{lstlisting}[language=TypeScript]
+    removeAllListeners(event?: 'up' | 'down'): this;
+    protocol?: 'udp'|'tcp';
+\end{lstlisting}
+The carlo
+package\footnote{\url{https://www.npmjs.com/package/@types/carlo}}
+uses it to distinguish different 
+types of events.
+\begin{lstlisting}[language=TypeScript]
+export type AppEvent = 'exit' | 'window';
+\end{lstlisting}
+% coinbase ...
 

### 2020-09-10 programming-2021/main.tex

@@
 information. To evaluate the tool, 244 declaration files were
-generated directly from an \NPM{} module and 82 of these were
+generated directly from an \NPM{} module and 70 of these were
 compared with the corresponding declaration file provided on
-DefinitelyTyped. We obtained positive results for 52 files, which
+DefinitelyTyped. We obtained positive results for all of these files, which
 exhibited no differences at all or differences that can be resolved by
@@
 Figure \ref{fig:experiments-overall-funnel} shows that a declaration file was generated
-for 244 modules out of 6029 modules and we identified 60 modules that have only the
-features implemented by \texttt{dts-generate}. We obtained positive results for 52 of
+for 244 modules out of 6029 modules and we identified 70 modules that have only the
+features implemented by \texttt{dts-generate}. We obtained positive results for all of
 them. Section~\ref{sec:experiments-evaluation} provides a detailed explanation of the overall
@@
 \label{sec:experiments-evaluation}
-We analyzed 60 of the 244 generated files. The remaining 184 files contained at least one
+We analyzed 70 of the 244 generated files. The remaining 174 files contained at least one
 of the unimplemented TypeScript features so that they could not be considered. These
@@
 \secref{sec:dts-compare}. Modules containing only \texttt{solvable} differences were
-considered as positive results. We identified 52 such modules. We manually completed the
+considered as positive results. All of the analyzed files contained only \texttt{solvable} differences. We manually completed the
 examples for 10 of those modules. Applying \texttt{dts-generate} with the completed
@@
 \end{figure}
-
-We discovered only 8 modules with \texttt{unsolvable} differences. Most of them were
-caused by the same error: 7 modules incorrectly declared an interface for the
-\texttt{string} type. This interface contains standard properties and methods of the
-JavaScript \texttt{String} object, such as \texttt{length} or \texttt{indexOf()}.
 

### 2020-09-24 programming-2021/main.tex
Remove changethis blocks

@@
   \item \url{https://github.com/proglang/run-time-information-gathering/tree/v1.1}
-  \item \url{https://github.com/proglang/ts-declaration-file-generator/tree/v1.1}
-  \item \url{https://github.com/proglang/ts-declaration-file-generator-service/tree/v1.1}
+  \item \url{https://github.com/proglang/ts-declaration-file-generator/tree/v1.2}
+  \item \url{https://github.com/proglang/ts-declaration-file-generator-service/tree/v1.2}
   \item \url{https://github.com/proglang/dts-generate-method/tree/v1.1}
-  \item \url{https://github.com/proglang/dts-generate-results/tree/v1.1}
+  \item \url{https://github.com/proglang/dts-generate-results/tree/v1.2}
 \end{itemize}
@@
 
-\begin{changethis}
-  Question: check how many files on DT are just generated by \texttt{dts-gen} (e.g., the TypeScript tool)
-\end{changethis}
-
 \begin{figure}[tp]
@@
     \lstinline{export = XXX}.
-    \begin{changethis}
-      Q: how severe is that? That's not an issue, right?
-    \end{changethis}
   \item \texttt{FunctionMissingDifference}: A function declaration is present in the expected

### 2020-09-26 programming-2021/main.tex
final changes pre submission

@@
 
-\keywords{JavaScript, TypeScript, Dynamic Analysis, Declaration
-  Files}%TODO mandatory; please add comma-separated list of 1--5 keywords
+\keywords{JavaScript, TypeScript, Dynamic Analysis, Declaration Files}%TODO mandatory; please add comma-separated list of 1--5 keywords
 
@@
 After generating a declaration file for an \NPM{} package, we need to
-evaluate its quality. To this end, we created \texttt{dts-compare}, a tool to compute the
+evaluate its quality.
+As we extracted code examples written by developers, the quality of the generated
+declaration file is heavily dependent on these examples. For
+example, if a property of an interface is not accessed because
+a code example does not explore a particular execution path, then that
+property will not appear in the generated declaration file.
+
+As a measure of the quality of a generated file we use its distance to
+the declaration file provided on DefinitelyTyped for the same module. 
+To this end, we created \texttt{dts-compare}, a tool to compute the
 differences between two TypeScript declaration files.
@@
 of the generated files is comparable with the accuracy of the files on DefinitelyTyped,
-then \texttt{dts-generate} is a viable alternative to manually created definition files.
+then \texttt{dts-generate} is a viable alternative to hand-crafted definition files.
 
@@
 \label{sec:dts-parse}
-We need a means of comparing two declaration files. Of
-course, we are not interested in a textual comparison, but in a
-comparison of the structures described by the files.
-
-The first step is to parse the declaration files, which can be
-achieved by using the TypeScript Compiler API, a library which traverses the
+
+We are not interested in a textual comparison of declaration files, but in a
+comparison of the structures described by the files. Hence,
+the first step is to parse the declaration files using the TypeScript
+Compiler API to build and traverse the 
 Abstract Syntax Tree of a TypeScript program
 \cite{typescript-compiler-api}. 
-The step also performs a sanity check of the generated declaration
-files as it rejects files with syntactic or semantics errors.
-The output of \texttt{dts-parse} is a structure where declared
+This step also performs a sanity check of the generated declaration
+files as it rejects files with syntactic or semantic errors.
+The output of \texttt{dts-parse} is a structure similar to a symbol table where declared
 {interfaces}, {functions}, {classes}, and
@@
 JSON file containing the result of the comparison as shown in
-\coderef{code:dts-compare-example}. Following the standard naming convention of unit
+\coderef{code:dts-compare-example}. Following the naming convention of unit
 testing frameworks, the input files are marked as \lstinline{expected} or
@@
 \end{lstlisting}
+% $
 
@@
 of an abstract entity \texttt{Difference}. 
-A difference is \texttt{solvable} if it can be resolved by adding more code
-examples. For example, invoking a function with different parameters so that a missing
-interface property gets accessed. Otherwise it is  \texttt{unsolvable}.
-
-The concrete evaluated differences are as follows:
+We classify a difference as \texttt{solvable} if it can be resolved by adding more code
+examples. For example, we might add a function invocation with different parameters so that a missing
+interface property gets accessed. Otherwise it is classified as  \texttt{unsolvable}.
+
+We consider the following differences:
 \begin{itemize}
@@
 
-\begin{changethis}
-
-To measure the impact of incomplete code examples on the
-quality of generated declaration files we categorized each difference as follows:
-
-  We extracted code examples written by developers to execute the code
-  and gather run-time information. The quality of the generated
-  declaration file is heavily dependent on the code examples. For
-  example, if a property of an interface does not get accessed due to
-  a code example not exploring a particular execution path, that
-  property will not exist in the generated declaration file.
-\end{changethis}
-
 \section{Results}

### 2021-04-27 mplr-2021/sample-sigconf.tex
from sample

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf]{acmart}
\settopmatter{printfolios=true}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{10.1145/1122445.1122456}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
  June 03--05, 2018, Woodstock, NY}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{The Name of the Title is Hope}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Ben Trovato}
\authornote{Both authors contributed equally to this research.}
\email{trovato@corporation.com}
\orcid{1234-5678-9012}
\author{G.K.M. Tobin}
\authornotemark[1]
\email{webmaster@marysville-ohio.com}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \streetaddress{P.O. Box 1212}
  \city{Dublin}
  \state{Ohio}
  \country{USA}
  \postcode{43017-6221}
}

\author{Lars Th{\o}rv{\"a}ld}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
  \city{Hekla}
  \country{Iceland}}
\email{larst@affiliation.org}

\author{Valerie B\'eranger}
\affiliation{%
  \institution{Inria Paris-Rocquencourt}
  \city{Rocquencourt}
  \country{France}
}

\author{Aparna Patel}
\affiliation{%
 \institution{Rajiv Gandhi University}
 \streetaddress{Rono-Hills}
 \city{Doimukh}
 \state{Arunachal Pradesh}
 \country{India}}

\author{Huifen Chan}
\affiliation{%
  \institution{Tsinghua University}
  \streetaddress{30 Shuangqing Rd}
  \city{Haidian Qu}
  \state{Beijing Shi}
  \country{China}}

\author{Charles Palmer}
\affiliation{%
  \institution{Palmer Research Laboratories}
  \streetaddress{8600 Datapoint Drive}
  \city{San Antonio}
  \state{Texas}
  \country{USA}
  \postcode{78229}}
\email{cpalmer@prl.com}

\author{John Smith}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
  \city{Hekla}
  \country{Iceland}}
\email{jsmith@affiliation.org}

\author{Julius P. Kumquat}
\affiliation{%
  \institution{The Kumquat Consortium}
  \city{New York}
  \country{USA}}
\email{jpkumquat@consortium.net}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  A clear and well-documented \LaTeX\ document is presented as an
  article formatted for publication by ACM in a conference proceedings
  or journal publication. Based on the ``acmart'' document class, this
  article presents and explains many of the common variations, as well
  as many of the formatting elements an author may use in the
  preparation of the documentation of their work.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{datasets, neural networks, gaze detection, text tagging}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  GRAPHICS REMOVED % \includegraphics[width=\textwidth]{sampleteaser}
  \caption{Seattle Mariners at Spring Training, 2010.}
  \Description{Enjoying the baseball game from the third-base
  seats. Ichiro Suzuki preparing to bat.}
  \label{fig:teaser}
\end{teaserfigure}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
ACM's consolidated article template, introduced in 2017, provides a
consistent \LaTeX\ style for use across ACM publications, and
incorporates accessibility and metadata-extraction functionality
necessary for future Digital Library endeavors. Numerous ACM and
SIG-specific \LaTeX\ templates have been examined, and their unique
features incorporated into this single new template.

If you are new to publishing with ACM, this document is a valuable
guide to the process of preparing your work for publication. If you
have published with ACM before, this document provides insight and
instruction into more recent changes to the article template.

The ``\verb|acmart|'' document class can be used to prepare articles
for any ACM publication --- conference or journal, and for any stage
of publication, from review to final ``camera-ready'' copy, to the
author's own version, with {\itshape very} few changes to the source.

\section{Template Overview}
As noted in the introduction, the ``\verb|acmart|'' document class can
be used to prepare many different kinds of documentation --- a
double-blind initial submission of a full-length technical paper, a
two-page SIGGRAPH Emerging Technologies abstract, a ``camera-ready''
journal article, a SIGCHI Extended Abstract, and more --- all by
selecting the appropriate {\itshape template style} and {\itshape
  template parameters}.

This document will explain the major features of the document
class. For further information, the {\itshape \LaTeX\ User's Guide} is
available from
\url{https://www.acm.org/publications/proceedings-template}.

\subsection{Template Styles}

The primary parameter given to the ``\verb|acmart|'' document class is
the {\itshape template style} which corresponds to the kind of publication
or SIG publishing the work. This parameter is enclosed in square
brackets and is a part of the {\verb|documentclass|} command:
\begin{verbatim}
  \documentclass[STYLE]{acmart}
\end{verbatim}

Journals use one of three template styles. All but three ACM journals
use the {\verb|acmsmall|} template style:
\begin{itemize}
\item {\verb|acmsmall|}: The default journal template style.
\item {\verb|acmlarge|}: Used by JOCCH and TAP.
\item {\verb|acmtog|}: Used by TOG.
\end{itemize}

The majority of conference proceedings documentation will use the {\verb|acmconf|} template style.
\begin{itemize}
\item {\verb|acmconf|}: The default proceedings template style.
\item{\verb|sigchi|}: Used for SIGCHI conference articles.
\item{\verb|sigchi-a|}: Used for SIGCHI ``Extended Abstract'' articles.
\item{\verb|sigplan|}: Used for SIGPLAN conference articles.
\end{itemize}

\subsection{Template Parameters}

In addition to specifying the {\itshape template style} to be used in
formatting your work, there are a number of {\itshape template parameters}
which modify some part of the applied template style. A complete list
of these parameters can be found in the {\itshape \LaTeX\ User's Guide.}

Frequently-used parameters, or combinations of parameters, include:
\begin{itemize}
\item {\verb|anonymous,review|}: Suitable for a ``double-blind''
  conference submission. Anonymizes the work and includes line
  numbers. Use with the \verb|\acmSubmissionID| command to print the
  submission's unique ID on each page of the work.
\item{\verb|authorversion|}: Produces a version of the work suitable
  for posting by the author.
\item{\verb|screen|}: Produces colored hyperlinks.
\end{itemize}

This document uses the following string as the first command in the
source file:
\begin{verbatim}
\documentclass[sigconf]{acmart}
\end{verbatim}

\section{Modifications}

Modifying the template --- including but not limited to: adjusting
margins, typeface sizes, line spacing, paragraph and list definitions,
and the use of the \verb|\vspace| command to manually adjust the
vertical spacing between elements of your work --- is not allowed.

{\bfseries Your document will be returned to you for revision if
  modifications are discovered.}

\section{Typefaces}

The ``\verb|acmart|'' document class requires the use of the
``Libertine'' typeface family. Your \TeX\ installation should include
this set of packages. Please do not substitute other typefaces. The
``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
as they will override the built-in typeface families.

\section{Title Information}

The title of your work should use capital letters appropriately -
\url{https://capitalizemytitle.com/} has useful rules for
capitalization. Use the {\verb|title|} command to define the title of
your work. If your work has a subtitle, define it with the
{\verb|subtitle|} command.  Do not insert line breaks in your title.

If your title is lengthy, you must define a short version to be used
in the page headers, to prevent overlapping text. The \verb|title|
command has a ``short title'' parameter:
\begin{verbatim}
  \title[short title]{full title}
\end{verbatim}

\section{Authors and Affiliations}

Each author must be defined separately for accurate metadata
identification. Multiple authors may share one affiliation. Authors'
names should not be abbreviated; use full first names wherever
possible. Include authors' e-mail addresses whenever possible.

Grouping authors' names or e-mail addresses, or providing an ``e-mail
alias,'' as shown below, is not acceptable:
\begin{verbatim}
  \author{Brooke Aster, David Mehldau}
  \email{dave,judy,steve@university.edu}
  \email{firstname.lastname@phillips.org}
\end{verbatim}

The \verb|authornote| and \verb|authornotemark| commands allow a note
to apply to multiple authors --- for example, if the first two authors
of an article contributed equally to the work.

If your author list is lengthy, you must define a shortened version of
the list of authors to be used in the page headers, to prevent
overlapping text. The following command should be placed just after
the last \verb|\author{}| definition:
\begin{verbatim}
  \renewcommand{\shortauthors}{McCartney, et al.}
\end{verbatim}
Omitting this command will force the use of a concatenated list of all
of the authors' names, which may result in overlapping text in the
page headers.

The article template's documentation, available at
\url{https://www.acm.org/publications/proceedings-template}, has a
complete explanation of these commands and tips for their effective
use.

Note that authors' addresses are mandatory for journal articles.

\section{Rights Information}

Authors of any work published by ACM will need to complete a rights
form. Depending on the kind of work, and the rights management choice
made by the author, this may be copyright transfer, permission,
license, or an OA (open access) agreement.

Regardless of the rights management choice, the author will receive a
copy of the completed rights form once it has been submitted. This
form contains \LaTeX\ commands that must be copied into the source
document. When the document source is compiled, these commands and
their parameters add formatted text to several areas of the final
document:
\begin{itemize}
\item the ``ACM Reference Format'' text on the first page.
\item the ``rights management'' text on the first page.
\item the conference information in the page header(s).
\end{itemize}

Rights information is unique to the work; if you are preparing several
works for an event, make sure to use the correct set of commands with
each of the works.

The ACM Reference Format text is required for all articles over one
page in length, and is optional for one-page articles (abstracts).

\section{CCS Concepts and User-Defined Keywords}

Two elements of the ``acmart'' document class provide powerful
taxonomic tools for you to help readers find your work in an online
search.

The ACM Computing Classification System ---
\url{https://www.acm.org/publications/class-2012} --- is a set of
classifiers and concepts that describe the computing
discipline. Authors can select entries from this classification
system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
commands to be included in the \LaTeX\ source.

User-defined keywords are a comma-separated list of words and phrases
of the authors' choosing, providing a more flexible way of describing
the research being presented.

CCS concepts and user-defined keywords are required for for all
articles over two pages in length, and are optional for one- and
two-page articles (or abstracts).

\section{Sectioning Commands}

Your work should use standard \LaTeX\ sectioning commands:
\verb|section|, \verb|subsection|, \verb|subsubsection|, and
\verb|paragraph|. They should be numbered; do not remove the numbering
from the commands.

Simulating a sectioning command by setting the first word or words of
a paragraph in boldface or italicized text is {\bfseries not allowed.}

\section{Tables}

The ``\verb|acmart|'' document class includes the ``\verb|booktabs|''
package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
high-quality tables.

Table captions are placed {\itshape above} the table.

Because tables cannot be split across pages, the best placement for
them is typically the top of the page nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and the
table caption.  The contents of the table itself must go in the
\textbf{tabular} environment, to be aligned properly in rows and
columns, with the desired horizontal and vertical rules.  Again,
detailed instructions on \textbf{tabular} material are found in the
\textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table~\ref{tab:freq} is included in the input file; compare the
placement of the table here with the table in the printed output of
this document.

\begin{table}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \textbf{table*} to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more
desirable. Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table*}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table*}

Always use midrule to separate table header rows from data rows, and
use it only for this purpose. This enables assistive technologies to
recognise table headers and support their users in navigating tables
more easily.

\section{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of the three are
discussed in the next sections.

\subsection{Inline (In-text) Equations}
A formula that appears in the running text is called an inline or
in-text formula.  It is produced by the \textbf{math} environment,
which can be invoked with the usual
\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
and structures, from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
examples of in-text equations in context. Notice how this equation:
\begin{math}
  \lim_{n\rightarrow \infty}x=0
\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsection{Display Equations}
A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \textbf{equation}
environment. An unnumbered display equation is produced by the
\textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols and
structures available in \LaTeX\@; this section will just give a couple
of examples of display equations in context.  First, consider the
equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty}x=0
\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  \sum_{i=0}^{\infty} x + 1
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\section{Figures}

The ``\verb|figure|'' environment should be used for figures. One or
more images can be placed within a figure. If your figure contains
third-party material, you must clearly identify it as such, as shown
in the example below.
\begin{figure}[h]
  \centering
  GRAPHICS REMOVED %\includegraphics[width=\linewidth]{sample-franklin}
  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
    Ewing, Inc. [Public domain], via Wikimedia
    Commons. (\url{https://goo.gl/VLCRBB}).}
  \Description{A woman and a girl in white dresses sit in an open car.}
\end{figure}

Your figures should contain a caption which describes the figure to
the reader.

Figure captions are placed {\itshape below} the figure.

Every figure should also have a figure description unless it is purely
decorative. These descriptions convey what’s in the image to someone
who cannot see it. They are also used by search engine crawlers for
indexing images, and when images cannot be loaded.

A figure description must be unformatted plain text less than 2000
characters long (including spaces).  {\bfseries Figure descriptions
  should not repeat the figure caption – their purpose is to capture
  important information that is not already provided in the caption or
  the main text of the paper.} For figures that convey important and
complex new information, a short text description may not be
adequate. More complex alternative descriptions can be placed in an
appendix and referenced in a short figure description. For example,
provide a data table capturing the information in a bar chart, or a
structured list representing a graph.  For additional information
regarding how best to write figure descriptions and why doing this is
so important, please see
\url{https://www.acm.org/publications/taps/describing-figures/}.

\subsection{The ``Teaser Figure''}

A ``teaser figure'' is an image, or set of images in one figure, that
are placed after all author and affiliation information, and before
the body of the article, spanning the page. If you wish to have such a
figure in your article, place the command immediately before the
\verb|\maketitle| command:
\begin{verbatim}
  \begin{teaserfigure}
    \includegraphics[width=\textwidth]{sampleteaser}
    \caption{figure caption}
    \Description{figure description}
  \end{teaserfigure}
\end{verbatim}

\section{Citations and Bibliographies}

The use of \BibTeX\ for the preparation and formatting of one's
references is strongly recommended. Authors' names should be complete
--- use full first names (``Donald E. Knuth'') not initials
(``D. E. Knuth'') --- and the salient identifying features of a
reference should be included: title, year, volume, number, pages,
article DOI, etc.

The bibliography is included in your source document with these two
commands, placed just before the \verb|\end{document}| command:
\begin{verbatim}
  \bibliographystyle{ACM-Reference-Format}
  \bibliography{bibfile}
\end{verbatim}
where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
suffix, of the \BibTeX\ file.

Citations and references are numbered by default. A small number of
ACM publications have citations and references formatted in the
``author year'' style; for these exceptions, please include this
command in the {\bfseries preamble} (before the command
``\verb|\begin{document}|'') of your \LaTeX\ source:
\begin{verbatim}
  \citestyle{acmauthoryear}
\end{verbatim}

  Some examples.  A paginated journal article \cite{Abril07}, an
  enumerated journal article \cite{Cohen07}, a reference to an entire
  issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
  monograph/whole book in a series (see 2a in spec. document)
  \cite{Harel79}, a divisible-book such as an anthology or compilation
  \cite{Editor00} followed by the same example, however we only output
  the series if the volume number is given \cite{Editor00a} (so
  Editor00a's series should NOT be present since it has no vol. no.),
  a chapter in a divisible book \cite{Spector90}, a chapter in a
  divisible book in a series \cite{Douglass98}, a multi-volume work as
  book \cite{Knuth97}, a couple of articles in a proceedings (of a
  conference, symposium, workshop for example) (paginated proceedings
  article) \cite{Andler79, Hagerup1993}, a proceedings article with
  all possible elements \cite{Smith10}, an example of an enumerated
  proceedings article \cite{VanGundy07}, an informally published work
  \cite{Harel78}, a couple of preprints \cite{Bornmann2019,
    AnzarootPBM14}, a doctoral dissertation \cite{Clarkson85}, a
  master's thesis: \cite{anisi03}, an online document / world wide web
  resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game
  (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05}
  and (Case 3) a patent \cite{JoeScientist001}, work accepted for
  publication \cite{rous08}, 'YYYYb'-test for prolific author
  \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
  contain 'duplicate' DOI and URLs (some SIAM articles)
  \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
  multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
  couple of citations with DOIs:
  \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
  citations: \cite{TUGInstmem, Thornburg01, CTANacmart}. Artifacts:
  \cite{R} and \cite{UMassCitations}.

\section{Acknowledgments}

Identification of funding sources and other support, and thanks to
individuals and groups that assisted in the research and the
preparation of the work should be included in an acknowledgment
section, which is placed just before the reference section in your
document.

This section has a special environment:
\begin{verbatim}
  \begin{acks}
  ...
  \end{acks}
\end{verbatim}
so that the information contained therein can be more easily collected
during the article metadata extraction phase, and to ensure
consistency in the spelling of the section heading.

Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.

\section{Appendices}

If your work needs an appendix, add it before the
``\verb|\end{document}|'' command at the conclusion of your source
document.

Start the appendix with the ``\verb|appendix|'' command:
\begin{verbatim}
  \appendix
\end{verbatim}
and note that in the appendix, sections are lettered, not
numbered. This document has two appendices, demonstrating the section
and subsection identification method.

\section{SIGCHI Extended Abstracts}

The ``\verb|sigchi-a|'' template style (available only in \LaTeX\ and
not in Word) produces a landscape-orientation formatted article, with
a wide left margin. Three environments are available for use with the
``\verb|sigchi-a|'' template style, and produce formatted output in
the margin:
\begin{itemize}
\item {\verb|sidebar|}:  Place formatted text in the margin.
\item {\verb|marginfigure|}: Place a figure in the margin.
\item {\verb|margintable|}: Place a table in the margin.
\end{itemize}

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
To Robert, for the bagels and explaining CMYK and color spaces.
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\section{Research Methods}

\subsection{Part One}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
lacinia dolor. Integer ultricies commodo sem nec semper.

\subsection{Part Two}

Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
eros. Vivamus non purus placerat, scelerisque diam eu, cursus
ante. Etiam aliquam tortor auctor efficitur mattis.

\section{Online Resources}

Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
enim maximus. Vestibulum gravida massa ut felis suscipit
congue. Quisque mattis elit a risus ultrices commodo venenatis eget
dui. Etiam sagittis eleifend elementum.

Nam interdum magna at lectus dignissim, ac dignissim lorem
rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
massa et mattis lacinia.

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.

### 2021-05-08 mplr-2021/main.tex
Migrate progamming2021 to mplr-2021

--- 2020-09-26 programming-2021/main.tex
@@
-% -*- coding: utf-8; -*-
-\documentclass[english,cleveref,autoref,submission]{programming}
-% Programming:
-% The main part of the paper should not exceed 22 pages (in the
-% provided style of the journal), but there is no limit for
-% bibliography and appendices. The page limit for the main part of the
-% paper is in place in order to keep the paper on focus and to avoid
-% overloading the reviewers. Authors are encouraged to move important
-% details to appendices, which may be consulted by the reviewers. In
-% some cases, if authors feel that the main part requires
-% substantially more pages, they should explain the reasons why in the
-% additional comments field of the submission form; examples of these
-% cases may include papers with substantial source code listings, and
-% essays. Papers whose length is incommensurate with their
-% contribution will be rejected.
-
-\usepackage[backend=biber]{biblatex}
-\addbibresource{main.bib}
+%%
+%% This is file `sample-sigconf.tex',
+%% generated with the docstrip utility.
+%%
+%% The original source files were:
+%%
+%% samples.dtx  (with options: `sigconf')
+%% 
+%% IMPORTANT NOTICE:
+%% 
+%% For the copyright see the source file.
+%% 
+%% Any modified versions of this file must be renamed
+%% with new filenames distinct from sample-sigconf.tex.
+%% 
+%% For distribution of the original source see the terms
+%% for copying and modification in the file samples.dtx.
+%% 
+%% This generated file may be distributed as long as the
+%% original source files, as listed above, are part of the
+%% same distribution. (The sources need not necessarily be
+%% in the same archive or directory.)
+%%
+%% The first command in your LaTeX source must be the \documentclass command.
+
+\input{macros}
+\newcommand{\secref}[1]{Section~\ref{#1} - \nameref{#1}}
+\newcommand{\figref}[1]{Figure~\ref{#1}}
+\newcommand{\coderef}[1]{Listing~\ref{#1}}
+
+\documentclass[sigconf]{acmart}
+\usepackage{listings}
 
@@
 
-\newenvironment{changethis}{%
-  \begin{tcolorbox}[breakable,notitle,boxrule=0pt,colback=blue!20,colframe=blue!20]}{%
-  \end{tcolorbox}}
-
-\input{macros}
+\usepackage{subcaption}
+
+\usepackage[htt]{hyphenat}
 
 \definecolor{backgroundgray}{rgb}{.94,.94,.94}
-\definecolor{lightgray}{rgb}{.35,.35,.5}
-\definecolor{darkgray}{rgb}{.2, .2, .2}
-\definecolor{purple}{rgb}{0.65, 0.12, 0.82}
-
-\newcommand{\secref}[1]{Section~\ref{#1} - \nameref{#1}}
-\newcommand{\figref}[1]{Figure~\ref{#1}}
-\newcommand{\coderef}[1]{Listing~\ref{#1}}
 
@@
 
-%%\bibliographystyle{plainurl}% the mandatory bibstyle
-
+\settopmatter{printfolios=true}
+%%
+%% \BibTeX command to typeset BibTeX logo in the docs
+\AtBeginDocument{%
+  \providecommand\BibTeX{{%
+    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
+
+%% Rights management information.  This information is sent to you
+%% when you complete the rights form.  These commands have SAMPLE
+%% values in them; it is your responsibility as an author to replace
+%% the commands and values with those provided to you when you
+%% complete the rights form.
+\setcopyright{acmcopyright}
+\copyrightyear{2018}
+\acmYear{2018}
+\acmDOI{10.1145/1122445.1122456}
+
+%% These commands are for a PROCEEDINGS abstract or paper.
+\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
+  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
+\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
+  June 03--05, 2018, Woodstock, NY}
+\acmPrice{15.00}
+\acmISBN{978-1-4503-XXXX-X/18/06}
+
+%%
+%% Submission ID.
+%% Use this when submitting an article to a sponsored event. You'll
+%% receive a unique submission ID from the organizers
+%% of the event, and this ID should be used as the parameter to this command.
+%%\acmSubmissionID{123-A56-BU3}
+
+%%
+%% The majority of ACM publications use numbered citations and
+%% references.  The command \citestyle{authoryear} switches to the
+%% "author year" style.
+%%
+%% If you are preparing content for an event
+%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
+%% citations and references.
+%% Uncommenting
+%% the next command will enable that style.
+%%\citestyle{acmauthoryear}
+
+%%
+%% end of the preamble, start of the body of the document source.
+\begin{document}
+
+%%
+%% The "title" command has an optional parameter,
+%% allowing the author to define a "short title" to be used in page headers.
 \title{Generation of TypeScript Declaration Files from JavaScript Code}
-\subtitle{Catching the Developer's Intent from Examples}
-% \titlerunning{Dummy short title}%optional, please use if title is longer than one line
+
+%%
+%% The "author" command and its associated commands are used to define
+%% the authors and their affiliations.
+%% Of note is the shared affiliation of the first two authors, and the
+%% "authornote" and "authornotemark" commands
+%% used to denote shared contribution to the research.
 
 \author{Fernando Cristiani}
-\authorinfo{\email{fernando.cristiani@hs-karlsruhe.de}}
-\affiliation{Hochschule Karlsruhe, Germany}
+\affiliation{%
+  \institution{Hochschule Karlsruhe}
+  \city{Karlsruhe}
+  \country{Germany}}
+\email{fernando.cristiani@hs-karlsruhe.de}
 
 \author{Peter Thiemann}
-\authorinfo{\email{thiemann@informatik.uni-freiburg.de}}
-\affiliation{Albert-Ludwigs-Universität Freiburg, Germany}
-
-\keywords{JavaScript, TypeScript, Dynamic Analysis, Declaration Files}%TODO mandatory; please add comma-separated list of 1--5 keywords
-
-%%%%%%%%%%%%%%%%%%
-%% These data MUST be filled for your submission. (see 5.3)
-\paperdetails{
-  %% perspective options are: art, sciencetheoretical, scienceempirical, engineering.
-  %% Choose exactly the one that best describes this work. (see 2.1)
-  perspective=science-empirical,
-  %% State one or more areas, separated by a comma. (see 2.2)
-  %% Please see list of areas in http://programming-journal.org/cfp/
-  %% The list is open-ended, so use other areas if yours is/are not listed.
-  area={General-purpose programming, Web programming},
-  %% You may choose the license for your paper (see 3.)
-  %% License options include: cc-by (default), cc-by-nc
-  % license=cc-by,
-}
-%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%
-%% These data are provided by the editors. May be left out on submission.
-%\paperdetails{
-%  submitted=2016-08-10,
-%  published=2016-10-11,
-%  year=2016,
-%  volume=1,
-%  issue=1,
-%  articlenumber=1,
-%}
-%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Please go to https://dl.acm.org/ccs/ccs.cfm and generate your Classification
-% System [view CCS TeX Code] stanz and copy _all of it_ to this place.
-%% From HERE
-%%% !!! REVISE
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10002944.10011122.10003459</concept_id>
-<concept_desc>General and reference~Computing standards, RFCs and guidelines</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-<concept>
-<concept_id>10010405.10010476.10010477</concept_id>
-<concept_desc>Applied computing~Publishing</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[300]{Software and its engineering~Software notations and tools}
-\ccsdesc[100]{Software and its engineering~General programming languages}
-
-% To HERE
-%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-\maketitle
-
-%TODO mandatory: add short abstract of the document
+\affiliation{%
+  \institution{Albert-Ludwigs-Universit{\"a}t Freiburg}
+  \city{Freiburg}
+  \country{Germany}
+}
+\email{thiemann@informatik.uni-freiburg.de}
+
+%%
+%% By default, the full list of authors will be used in the page
+%% headers. Often, this list is too long, and will overlap
+%% other information printed in the page headers. This command allows
+%% the author to define a more concise list
+%% of authors' names for this purpose.
+% \renewcommand{\shortauthors}{Trovato and Tobin, et al.}
+
+%%
+%% The abstract is a short summary of the work to be presented in the
+%% article.
 \begin{abstract}
@@
 \end{abstract}
+
+%%
+%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
+%% Please copy and paste the code instead of the example below.
+%%
+\begin{CCSXML}
+<ccs2012>
+ <concept>
+  <concept_id>10010520.10010553.10010562</concept_id>
+  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+  <concept_significance>500</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010575.10010755</concept_id>
+  <concept_desc>Computer systems organization~Redundancy</concept_desc>
+  <concept_significance>300</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010553.10010554</concept_id>
+  <concept_desc>Computer systems organization~Robotics</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10003033.10003083.10003095</concept_id>
+  <concept_desc>Networks~Network reliability</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Computer systems organization~Embedded systems}
+\ccsdesc[300]{Computer systems organization~Redundancy}
+\ccsdesc{Computer systems organization~Robotics}
+\ccsdesc[100]{Networks~Network reliability}
+
+%%
+%% Keywords. The author(s) should pick words that accurately describe
+%% the work being presented. Separate the keywords with commas.
+\keywords{JavaScript, TypeScript, Dynamic Analysis, Declaration Files}
+
+%% A "teaser" image appears between the author and affiliation
+%% information and the body of the document, and typically spans the
+%% page.
+% \begin{teaserfigure}
+%   GRAPHICS REMOVED % \includegraphics[width=\textwidth]{sampleteaser}
+%   \caption{Seattle Mariners at Spring Training, 2010.}
+%   \Description{Enjoying the baseball game from the third-base
+%   seats. Ichiro Suzuki preparing to bat.}
+%   \label{fig:teaser}
+% \end{teaserfigure}
+
+%%
+%% This command processes the author and affiliation and title
+%% information and builds the first part of the formatted document.
+\maketitle
 
@@
 \end{lstlisting}
-% "01" < "00100"; // false 
-% pjt: why unexpected?
 
@@
 of the following three examples\footnote{\url{https://www.npmjs.com/package/abs}}.
+
 \begin{lstlisting}[language=JavaScript,numbers=none]
@@
 DefinitelyTyped\footnote{\url{https://github.com/DefinitelyTyped/DefinitelyTyped/blob/master/types/abs/index.d.ts}}:
+
 \begin{lstlisting}[language=TypeScript,numbers=none]
@@
 \end{lstlisting}
+
 % \begin{description}
@@
   \begin{center}
-    \includegraphics[width=0.8\linewidth]{motivating-example-glob-to-regexp-vscode.png}
+    \includegraphics[width=\linewidth]{motivating-example-glob-to-regexp-vscode.png}
   \end{center}
 
-  \caption{\textbf{Declaration file for \texttt{glob-to-regexp}
-      generated with \texttt{dts-generate}} - The interface is
-    detected correctly. Optional parameters are detected. The
-    declaration file is usable in Visual Studio
-    Code\footnote{https://code.visualstudio.com}.} 
+  \caption{Declaration file for \texttt{glob-to-regexp} 
+    generated with \texttt{dts-generate} - The interface is
+    detected correctly. Optional parameters are detected. The declaration 
+    file is usable in Visual Studio Code\protect\footnote{\url{https://code.visualstudio.com}}} 
   \label{fig:motivating-example-glob-to-regexp-vscode}
@@
   \includegraphics[width=1\linewidth]{dts-generate-block-diagram.pdf}
-    \caption[dts-generate - Architecture overview]{\textbf{dts-generate - Architecture
-        overview} - The initial code base is retrieved from the \NPM{} package's
+  \caption[dts-generate - Architecture overview]{dts-generate - Architecture
+      overview - The initial code base is retrieved from the \NPM{} package's
     repository.
@@
 \begin{table}[tp]
+  \caption{Distribution of TypeScript features in declaration files uploaded to DefinitelyTyped (DT).
+The 2nd column shows how many files in DT make use of a specific TypeScript feature.
+The 3rd column shows the corresponding percentage for a total of 6648 declaration files in DT.
+Numbers were obtained from commit \texttt{da93d7b13094bacb170ead3f4a289f3b8687e4e5} (April 4, 2020).}
   \begin{center}
@@
   \end{center}
-  \caption{Distribution of TypeScript features in declaration files uploaded to DefinitelyTyped (DT).
-The 2nd column shows how many files in DT make use of a specific TypeScript feature.
-The 3rd column shows the corresponding percentage for a total of 6648 declaration files in DT.
-Numbers were obtained from commit \texttt{da93d7b13094bacb170ead3f4a289f3b8687e4e5} (April 4, 2020).}
   \label{tab:dts-parse-stats}
@@
 
-% \paragraph*{Interfaces}
 Finally, interfaces are created by exploring \texttt{getField} and \texttt{methodCall} interactions from the run-time information. We
@@
   \begin{centering}
-        {\includegraphics[width=1\textwidth]{evaluation-diagram.pdf}}
+      {\includegraphics[width=1\linewidth]{evaluation-diagram.pdf}}
       \caption[Evaluation against DefinitelyTyped Repository]{\textbf{Evaluation of
@@
 
-\definecolor{myellow}{RGB}{228,212,0}
-\definecolor{mgreen}{RGB}{5,104,57}
-
-\newcommand\funnel[3]{%
-\pgfmathsetmacro\mwid{(0.3+\val*0.06)}
-\pgfmathsetmacro\mradius{(\val*0.01 + 1)}
-\pgfmathsetmacro\mheight{(\val*0.003 + 0.4)}
-\pgfmathsetmacro\marc{\mwid-.4}
-    \begin{scope}[%
-        shift={(0,#1)}, 
-        line width=.05pt, 
-        %x=5mm, 
-        scale=0.7,
-        yshift=\xi*0.05
-        ]
-    \draw[black,bottom color=#2, top color=#2] (-\mwid,0) -- (-\mwid+.4,-\mheight) arc (190:350:\marc cm and \mradius mm) -- (\mwid,0);
-    \draw[black,fill=#3] (0,0) ellipse (\mwid cm and \mradius mm);
-    \path (-\mwid,0) -- (-\mwid+.4,-\mheight) coordinate[midway] (a\xi);
-    \end{scope}
-}
-
 \begin{figure}[tp]
-	\hspace*{-0.16\textwidth}
   \centering
-	\begin{tikzpicture}
-		\foreach \val
-				[%
-				count=\xi starting from 0, 
-				evaluate=\xi as \shadecolor using int(25*\xi),
-				evaluate=\xi as \coord using int(\xi-12)
-				]
-			in {
-				4.05,
-				7.33,
-				18.02,
-				37.49,
-				71.19,
-				82.50,
-				100.00
-			}{
-				\funnel{\coord}{mgreen!\shadecolor !myellow}{mgreen!\shadecolor !myellow}
-			}
-
-		\node[left=0.02\textwidth of a0] {Generated Declaration Files};
-		\node[right=0.04\textwidth of a0] {\textbf{244}};
-
-		\node[left=0.02\textwidth of a1] {Run-time Information};
-		\node[right=0.06\textwidth of a1] {\textbf{436}};
-
-		\node[left=0.02\textwidth of a2] {Working Examples};
-		\node[right=0.17\textwidth of a2] {\textbf{946}};
-
-		\node[left=0.02\textwidth of a3] {Code Examples};
-		\node[right=0.25\textwidth of a3] {\textbf{2260}};
-
-		\node[left=0.02\textwidth of a4] {README file};
-		\node[right=0.45\textwidth of a4] {\textbf{4292}};
-
-		\node[left=0.02\textwidth of a5] {Github Repository};
-		\node[right=0.52\textwidth of a5] {\textbf{4974}};
-
-		\node[align=right,left=0.02\textwidth of a6, text width=0.25\textwidth] {Definitely Typed Modules};
-		\node[right=0.62\textwidth of a6] {\textbf{6029}};
-
-	\end{tikzpicture}
-	\caption[Number of analyzed modules for each stage of the experiment]{\textbf{Number of analyzed modules for each stage of the experiment} - A TypeScript Declaration File was generated for only 244 modules, out of 6029 modules in the DefinitelyTyped Repository.
-It was possible to gather valid run-time information for only 25\% of the modules for which a Code Example was extracted.}
+  \includegraphics[width=1\linewidth]{funnel.png}
+  \caption[Number of analyzed modules for each stage of the experiment]{
+    \textbf{Number of analyzed modules for each stage of the experiment} - 
+  A TypeScript Declaration File was generated for only 244 modules, out of 6029 modules in the DefinitelyTyped Repository. 
+  It was possible to gather valid run-time information for only 25\% of the modules for which a Code Example was extracted.
+  } 
   \label{fig:experiments-overall-funnel}
@@
 for generating a declaration file.
-
-\subsection{Declaration Files Generation}
-\label{sec:experiments-declaration-files-generation}
-
-%%% !!! TBC !!!
 
@@
 
-\printbibliography
-%%\bibliography{main}
+%%
+%% The next two lines define the bibliography style to be used, and
+%% the bibliography file.
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{main}
 
 \end{document}
+\endinput
+%%
+%% End of file `sample-sigconf.tex'.

### 2021-05-10 mplr-2021/main.tex
Add macro \CountModulesOnlySolvableDifferences{}

@@
   information. To evaluate the tool, 244 declaration files were
-  generated directly from an \NPM{} module and 70 of these were
+  generated directly from an \NPM{} module and \CountModulesOnlySolvableDifferences{} of these were
   compared with the corresponding declaration file provided on
@@
 \begin{itemize}
-  \item \url{https://github.com/proglang/run-time-information-gathering/tree/v1.1}
-  \item \url{https://github.com/proglang/ts-declaration-file-generator/tree/v1.2}
+  \item \url{https://github.com/proglang/run-time-information-gathering/tree/v1.1.0}
+  \item \url{https://github.com/proglang/ts-declaration-file-generator/tree/v1.5.0}
   \item \url{https://github.com/proglang/ts-declaration-file-generator-service/tree/v1.2}
-  \item \url{https://github.com/proglang/dts-generate-method/tree/v1.1}
-  \item \url{https://github.com/proglang/dts-generate-results/tree/v1.2}
+  \item \url{https://github.com/proglang/dts-generate-method/tree/v1.3.1}
+  \item \url{https://github.com/proglang/dts-generate-results/tree/v1.4.0}
 \end{itemize}
@@
 Figure \ref{fig:experiments-overall-funnel} shows that a declaration file was generated
-for 244 modules out of 6029 modules and we identified 70 modules that have only the
+for 244 modules out of 6029 modules and we identified \CountModulesOnlySolvableDifferences{} modules that have only the
 features implemented by \texttt{dts-generate}. We obtained positive results for all of
@@
 \label{sec:experiments-evaluation}
-We analyzed 70 of the 244 generated files. The remaining 174 files contained at least one
+We analyzed \CountModulesOnlySolvableDifferences{} of the 244 generated files. The remaining 174 files contained at least one
 of the unimplemented TypeScript features so that they could not be considered. These

### 2021-05-10 mplr-2021/main.tex
Add macros for result numbers

@@
   run-time information, and generates a declaration file based on this
-  information. To evaluate the tool, 244 declaration files were
+  information. To evaluate the tool, \CountModulesGeneratedDeclarationFile{} declaration files were
   generated directly from an \NPM{} module and \CountModulesOnlySolvableDifferences{} of these were
@@
 
-% Finally, we generated declaration files for 244 JavaScript libraries
+% Finally, we generated declaration files for \CountModulesGeneratedDeclarationFile{} JavaScript libraries
 % and evaluated them against the DefinitelyTyped repository.
@@
 \item An evaluation of our framework (Sections~\ref{sec:dts-generate-evaluation}
-  and~\ref{sec:results}). We examined all 6029 entries in 
-  the DefinitelyTyped repository and found 244 sufficiently
+  and~\ref{sec:results}). We examined all \CountTotalModulesDefinitelyTyped{} entries in 
+  the DefinitelyTyped repository and found \CountModulesGeneratedDeclarationFile{} sufficiently
   well documented \NPM{} packages, on which we ran \texttt{dts-generate}
@@
 Figure \ref{fig:experiments-overall-funnel} shows that a declaration file was generated
-for 244 modules out of 6029 modules and we identified \CountModulesOnlySolvableDifferences{} modules that have only the
+for \CountModulesGeneratedDeclarationFile{} modules out of \CountTotalModulesDefinitelyTyped{} modules and we identified \CountModulesOnlySolvableDifferences{} modules that have only the
 features implemented by \texttt{dts-generate}. We obtained positive results for all of
@@
     \textbf{Number of analyzed modules for each stage of the experiment} - 
-  A TypeScript Declaration File was generated for only 244 modules, out of 6029 modules in the DefinitelyTyped Repository. 
+  A TypeScript Declaration File was generated for only \CountModulesGeneratedDeclarationFile{} modules, out of \CountTotalModulesDefinitelyTyped{} modules in the DefinitelyTyped Repository. 
   It was possible to gather valid run-time information for only 25\% of the modules for which a Code Example was extracted.
@@
 shown in \figref{fig:experiments-overall-funnel}, it was only possible
-to obtain working code examples for 2260 packages. The
+to obtain working code examples for \CountModulesWithCodeExamples{} packages. The
 process of getting a valid code example for a module is divided in four
@@
 \paragraph*{Repository URL}
-The URL of the source repository could be retrieved for only 4974
+The URL of the source repository could be retrieved for only \CountModulesWithRepositoryUrl{}
 packages. More than 1000 packages on \NPM{} do not have a repository
@@
 \paragraph*{README Files}
-700 packages do not have a README file in their repository, although
+\CountModulesWithoutReadmeFile{} packages do not have a README file in their repository, although
 the implementation checks for several naming conventions like
@@
 using the \texttt{javascript} or \texttt{js} tags. However, as we are
-still left with code examples for 2260 modules, we did not further
+still left with code examples for \CountModulesWithCodeExamples{} modules, we did not further
 look into code extraction as this number was considered sufficient for
@@
 \paragraph*{Execution of Code Examples}
-We executed the remaining 2260 extracted code examples by installing the
+We executed the remaining \CountModulesWithCodeExamples{} extracted code examples by installing the
 required packages and running the code as a \NodeJS{}
-application. Unfortunately, the code examples only worked for 946
-modules. 1314 modules did not run correctly and had to be
+application. Unfortunately, the code examples only worked for \CountModulesWorkingCodeExamples{}
+modules. \CountModulesNotWorkingCodeExamples{} modules did not run correctly and had to be
 discarded. Some failing samples were analyzed and there were mainly
@@
 \paragraph*{Run-time Information}
-Run-time information was extracted for only 436 out of 946 modules with working code
+Run-time information was extracted for only \CountModulesRunTimeInfoExtracted{} out of \CountModulesWorkingCodeExamples{} modules with working code
 examples. As explained, to extract the run-time information, the behavior of the code
@@
 JavaScript features that are not supported by Jalangi. As a result, run-time information
-could not be extracted for 510 modules. An instrumentation without user-defined analysis
+could not be extracted for \CountModulesRunTimeInfoCouldNotBeExtracted{} modules. An instrumentation without user-defined analysis
 modules was not applied, so it was not possible to determine which modules were failing
@@
 \paragraph*{Generated Declaration Files}
-A declaration file was generated for 244 out of 436 modules. Despite the correct execution
-of the instrumented code examples, 192 modules did not yield suitable run-time information
+A declaration file was generated for \CountModulesGeneratedDeclarationFile{} out of \CountModulesRunTimeInfoExtracted{} modules. Despite the correct execution
+of the instrumented code examples, \CountModulesDeclarationFileCouldNotBeGenerated{} modules did not yield suitable run-time information
 for generating a declaration file. The extracted code examples for these modules did not
@@
 
-This section exhibits some examples of the 244 generated
+\subsection{Declaration Files Generation}
+\label{sec:experiments-declaration-files-generation}
+
+This section exhibits some examples of the \CountModulesGeneratedDeclarationFile{} generated
 declaration files. It shows some results for each of the implemented
@@
 \label{sec:experiments-evaluation}
-We analyzed \CountModulesOnlySolvableDifferences{} of the 244 generated files. The remaining 174 files contained at least one
+We analyzed \CountModulesOnlySolvableDifferences{} of the \CountModulesGeneratedDeclarationFile{} generated files. The remaining 174 files contained at least one
 of the unimplemented TypeScript features so that they could not be considered. These
@@
 We built a mechanism to automatically create declaration files for potentially every
-module uploaded to DefinitelyTyped. We managed to generate declaration files for 244
-modules. We compared the results against the corresponding files uploaded to
+module uploaded to DefinitelyTyped. We managed to generate declaration files for
+\CountModulesGeneratedDeclarationFile{} modules. We compared the results against the corresponding 
+files uploaded to
 DefinitelyTyped by creating \texttt{dts-parse}, a TypeScript declaration files parser and

### 2021-05-11 mplr-2021/main.tex
first-round

@@
 %% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2018}
-\acmYear{2018}
-\acmDOI{10.1145/1122445.1122456}
+
+% \setcopyright{acmcopyright}
+% \copyrightyear{2018}
+% \acmYear{2018}
+% \acmDOI{10.1145/1122445.1122456}
 
 %% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
-  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
-\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-  June 03--05, 2018, Woodstock, NY}
-\acmPrice{15.00}
-\acmISBN{978-1-4503-XXXX-X/18/06}
+% \acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
+%   Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
+% \acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
+%   June 03--05, 2018, Woodstock, NY}
+% \acmPrice{15.00}
+% \acmISBN{978-1-4503-XXXX-X/18/06}
 
@@
   compared with the corresponding declaration file provided on
-  DefinitelyTyped. We obtained positive results for all of these files, which
+  DefinitelyTyped.  All these files either
   exhibited no differences at all or differences that can be resolved by
-  modifying the developer-provided examples.
+  extending the developer-provided examples.
 \end{abstract}
@@
 %%
-\begin{CCSXML}
-<ccs2012>
- <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-  <concept_significance>500</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
+% \begin{CCSXML}
+% <ccs2012>
+%  <concept>
+%   <concept_id>10010520.10010553.10010562</concept_id>
+%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+%   <concept_significance>500</concept_significance>
+%  </concept>
+%  <concept>
+%   <concept_id>10010520.10010575.10010755</concept_id>
+%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
+%   <concept_significance>300</concept_significance>
+%  </concept>
+%  <concept>
+%   <concept_id>10010520.10010553.10010554</concept_id>
+%   <concept_desc>Computer systems organization~Robotics</concept_desc>
+%   <concept_significance>100</concept_significance>
+%  </concept>
+%  <concept>
+%   <concept_id>10003033.10003083.10003095</concept_id>
+%   <concept_desc>Networks~Network reliability</concept_desc>
+%   <concept_significance>100</concept_significance>
+%  </concept>
+% </ccs2012>
+% \end{CCSXML}
+
+% \ccsdesc[500]{Computer systems organization~Embedded systems}
+% \ccsdesc[300]{Computer systems organization~Redundancy}
+% \ccsdesc{Computer systems organization~Robotics}
+% \ccsdesc[100]{Networks~Network reliability}
 
@@
 checking of types, these  discrepancies can lead to unexpected 
-behavior and crashes. Such an experience can lead to developer
-frustration, longish debugging sessions, and decreasing confidence in
-the tool chain.
+behavior and crashes. This kind of experience can lead to longish debugging sessions,  developer
+frustration, and decreasing confidence in the tool chain.
 
@@
 We aim to improve on this situation by providing a tool that generates
-sound TypeScript declaration files automatically. To do so, we rely on the examples provided
+TypeScript declaration files automatically. To do so, we rely on the examples provided
 by the developer as part of the documentation of a library. If these
 examples are sufficiently rich, then our toolchain can generate
-high-quality declaration files for the library. This approach takes
+high-quality declaration files for the library. This way we take
 advantage of 
@@
 \texttt{abs} that
-``computes the absolute path of an input''. Its documentation consists
-of the following three examples\footnote{\url{https://www.npmjs.com/package/abs}}.
+``computes the absolute path of an input''. Its documentation contains
+the following three
+examples\footnote{\url{https://www.npmjs.com/package/abs}}.
 
@@
 \end{lstlisting}
-From the examples, our tool generates a TypeScript declaration file,
-which is equivalent to the declaration file provided in
-DefinitelyTyped\footnote{\url{https://github.com/DefinitelyTyped/DefinitelyTyped/blob/master/types/abs/index.d.ts}}:
-
+From the examples, our tool generates the following TypeScript declaration file.
 \begin{lstlisting}[language=TypeScript,numbers=none]
@@
 \end{lstlisting}
+It used to be equivalent to the declaration file provided in the
+DefinitelyTyped
+repository\footnote{\url{https://github.com/DefinitelyTyped/DefinitelyTyped/blob/master/types/abs/index.d.ts}},
+but the maintainer indicates (since May 10, 2021) that the parameter is optional:
+\begin{lstlisting}[language=TypeScript,numbers=none]
+declare function abs(input?: string): string;
+export = abs;
+\end{lstlisting}
+This case is quite frequent. The information in our declaration file is
+correct, but incomplete, because the developers did not document the
+fact that the input is optional. If the developer had provided the
+call  \lstinline/abs()/ as an additional 
+example, then our tool would have generated exactly
+the declaration on DefinitelyTyped! We call this case a solvable
+difference between our tool's output and the DefinitelyTyped
+declaration file.
 
@@
 \item
-  Instead we extract example code from the programmer's library
+  We extract example code from the programmer's library
   documentation and rely on dynamic analysis to extract typed usage
@@
 
-The implementation and the results are available under the following repositories:
+The implementation and the results are available in the following repositories:
 
@@
 \item Design and implementation of the tool \texttt{dts-generate}, a command line
-  application that generates a valid TypeScript declaration file for a
-  specific \NPM{} package using run-time information (Section~\ref{sec:typescr-decl-file}).
-
-\item A comparator for TypeScript declaration files (Section~\ref{sec:dts-compare}). This tool is
-  necessary for evaluating our framework and also useful to detect
-  incompatibilities when evolving JavaScript modules.
+  application that generates a valid TypeScript declaration file for
+  an \NPM{} package using run-time information
+  (Section~\ref{sec:typescr-decl-file}).
+
+\item A comparator for TypeScript declaration files
+  (Section~\ref{sec:dts-compare}) for evaluating our framework and to
+  detect   incompatibilities when evolving JavaScript modules.
 \item An evaluation of our framework (Sections~\ref{sec:dts-generate-evaluation}
@@
 \label{sec:motivating-example}
-The \NPM{} package \texttt{glob-to-regexp} is a simple JavaScript library, which turns a
+The \NPM{} package \texttt{glob-to-regexp} provides functionality to turn a
 glob expression for matching filenames in the shell into a regular
 expression\footnote{\url{https://www.npmjs.com/package/glob-to-regexp}}. It 
-has about 9,900,000 weekly downloads and 188 \NPM{} packages depend on it. If
+has about 9.9M weekly downloads and 188 \NPM{} packages depend on it. If
 a developer creates or extends TypeScript code that depends on the
 \texttt{glob-to-regexp} library, the TypeScript
-compiler and IDE requires a declaration file for that library to
+compiler and IDE require a declaration file for that library to
 perform static checking and code completion, respectively. With
 \texttt{dts-generate} we automatically generate a TypeScript
-declaration file for \texttt{glob-to-regexp}. The tool downloads the \NPM{} 
+declaration file for \texttt{glob-to-regexp}. Our tool downloads the \NPM{} 
 package, runs the examples extracted from its documentation, gathers
@@
 
-declare function GlobToRegexp(glob: string, opts?: GlobToRegexp.I__opts): RegExp;
+declare function GlobToRegexp(
+    glob: string, 
+    opts?: GlobToRegexp.I__opts): RegExp;
 declare namespace GlobToRegexp {
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
 \centering
@@
 \label{fig:typescript-templates-by-example}
-\end{figure}
+\end{figure*}
 
@@
 employ differing (or no) testing tools, or do not have tests
-at all. Moreover, developers try to cover edge cases or to trigger errors by using illegal
+at all. Moreover, developers try to cover corner cases or to trigger errors by using illegal
 input values in their tests. Such tests are not really useful for describing a library
@@
 appropriate and pragmatic way of extracting the developer's
-intention and provides a useful initial code base with meaningful
-examples, thus avoiding possible cold start problems.
+intention and provides an initial code base with meaningful
+examples.
+%? , thus avoiding possible cold start problems.
 
@@
   \item Parameter \lstinline{a} of function \lstinline{f} was used as
-    operand for operator \lstinline{==}. 
+    operand for \lstinline{==}. 
 \end{itemize}
@@
 The dynamic analysis framework Jalangi is used for gathering this kind of
-information. The configurable analysis modules enable
+information. Jalangi's configurable analysis modules enable
 programming custom callbacks that can get triggered with virtually any
@@
 
-We wrap each function's argument in a wrapper object, which enables to store meta-information and greatly simplifies the mapping of an observation with the corresponding \texttt{ArgumentContainer} or \texttt{interaction}.
+We wrap each function's argument in a wrapper object, which  stores meta-information and greatly simplifies the mapping of an observation with the corresponding \texttt{ArgumentContainer} or \texttt{interaction}.
 We use the original values for critical operators such as \texttt{===} or \texttt{typeof}, for which wrapper objects would modify the behavior of the code.
@@
 
-\begin{figure}[tp]
+\begin{figure*}[tp]
   \centering
@@
   \label{fig:experiments-results-module-is-uuid}
-\end{figure}
-
-We implemented only basic TypeScript features and we focused our effort in building an
-end-to-end solution that generates valid and useful declaration files. Only the basic
+\end{figure*}
+
+We implemented the most common TypeScript features first and we focused our effort in building an
+end-to-end solution that generates useful declaration files. The
 types \texttt{string}, \texttt{number}, \texttt{boolean}, unions
-thereof, and optional elements were considered
+thereof, arrays, function callbacks, and optional elements were considered
 both for function parameters and interface properties. We also
@@
 Common types such as
-\texttt{any}, arrays, function callbacks, tuples, intersections (that
+\texttt{any}, tuples, intersections (that
 is, types formed using the intersection operator \lstinline[language=TypeScript]/&/) and features such as
-generics, index signatures were not implemented.
+generics and index signatures were not implemented.
 
@@
 evaluate its quality.
-As we extracted code examples written by developers, the quality of the generated
-declaration file is heavily dependent on these examples. For
-example, if a property of an interface is not accessed because
-a code example does not explore a particular execution path, then that
-property will not appear in the generated declaration file.
+Clearly, the quality of the generated
+declaration file is heavily dependent on the quality of the examples
+provided by the developers. For
+example, if there is no code example exploring an execution path that would
+access a particular property, then that
+property will not appear in the gathered run-time data and
+consequently will not appear in the generated declaration file.
 
 As a measure of the quality of a generated file we use its distance to
-the declaration file provided on DefinitelyTyped for the same module. 
+the declaration file provided for the same module on
+DefinitelyTyped. In other words, we use DefinitelyTyped as our ground truth.
 To this end, we created \texttt{dts-compare}, a tool to compute the
@@
 \figref{fig:evaluation-diagram}. While this approach does not provide an absolute measure
-of quality, it gives us at least an indication of the pragmatics of \texttt{dts-generate}:
-The files on DefinitelyTyped are perceived to be useful by the community. If the accuracy
+of quality, it gives us at least an indication of the usefulness of \texttt{dts-generate}:
+The files on DefinitelyTyped are perceived to be valuable to the community. If the accuracy
 of the generated files is comparable with the accuracy of the files on DefinitelyTyped,
@@
 We are not interested in a textual comparison of declaration files, but in a
-comparison of the structures described by the files. Hence,
-the first step is to parse the declaration files using the TypeScript
+comparison of the structures described by the
+files. \texttt{dts-parse} receives the file as an input
+parameter and returns an AST serialized in JSON.
+
+The first step is to parse the declaration files using the TypeScript
 Compiler API to build and traverse the 
@@
 {namespaces} are stored separately. Function arguments are
-correctly described, identifying complex types like union types or
+described, identifying complex types like union types or
 callbacks. Optional parameters are also identified. For
 {classes}, a distinction between constructors and methods
-is made. Declaration files are tagged according to their features,
-which is useful for identifying and filtering out declaration files
+is made. Declaration files are tagged according to their features to identify and filter out declaration files
 that contain unimplemented features.
-
-\texttt{dts-parse} is itself written in TypeScript. It receives the file as input
-parameter and returns an AST serialized in JSON.
 
@@
 \label{sec:results}
-We analyzed the obtained results focusing on the following aspects:
+We analyzed the generated files focusing on the following aspects:
 \begin{itemize}
   \item The quality of the inferred types, interfaces, and module structure using run-time information.
-  \item The benefits of using code examples provided by developers as a first approximation to execute the libraries and avoid a cold start problem.
+  \item The benefits of using code examples provided by developers as
+    a first approximation to execute the libraries. %? and avoid a cold start problem.
   \item The usability of the generated declaration file and whether \texttt{dts-generate} can be used in a proper development environment.
@@
 Figure \ref{fig:experiments-overall-funnel} shows that a declaration file was generated
-for \CountModulesGeneratedDeclarationFile{} modules out of \CountTotalModulesDefinitelyTyped{} modules and we identified \CountModulesOnlySolvableDifferences{} modules that have only the
+for \CountModulesGeneratedDeclarationFile{} modules out of
+\CountTotalModulesDefinitelyTyped{} modules on DefinitelyTyped and we identified \CountModulesOnlySolvableDifferences{} modules that have only the
 features implemented by \texttt{dts-generate}. We obtained positive results for all of
@@
   A TypeScript Declaration File was generated for only \CountModulesGeneratedDeclarationFile{} modules, out of \CountTotalModulesDefinitelyTyped{} modules in the DefinitelyTyped Repository. 
-  It was possible to gather valid run-time information for only 25\% of the modules for which a Code Example was extracted.
+  % It was only possible to gather valid run-time information for about 25\% of the modules for which a Code Example was extracted.
   } 
@@
 Retrieving the code examples for the JavaScript libraries proved to be
-a pragmatic way of driving the type gathering at run time. However, as
-shown in \figref{fig:experiments-overall-funnel}, it was only possible
-to obtain working code examples for \CountModulesWithCodeExamples{} packages. The
-process of getting a valid code example for a module is divided in four
+a pragmatic way of driving the type gathering at run time.
+\figref{fig:experiments-overall-funnel} gives an overview of the
+process to extract code examples. To understand why it was only possible
+to obtain working code examples for \CountModulesWithCodeExamples{}
+packages, we explain the steps in the process and analyze the losses
+in each step.
+
+The process of getting a valid code example for a module is divided in four
 stages: 
@@
 explained because some developers do not wrap their code in a block
-using the \texttt{javascript} or \texttt{js} tags. However, as we are
-still left with code examples for \CountModulesWithCodeExamples{} modules, we did not further
-look into code extraction as this number was considered sufficient for
+using the \texttt{javascript} or \texttt{js} tags. As we are
+still left with code examples for \CountModulesWithCodeExamples{} modules, we did not look
+further into code extraction as this number was considered sufficient for
 evaluating the generation of declaration files.
@@
 required packages and running the code as a \NodeJS{}
-application. Unfortunately, the code examples only worked for \CountModulesWorkingCodeExamples{}
-modules. \CountModulesNotWorkingCodeExamples{} modules did not run correctly and had to be
-discarded. Some failing samples were analyzed and there were mainly
+application. Unfortunately,
+\CountModulesNotWorkingCodeExamples{} modules did not run correctly and had to be
+discarded.
+The code examples worked for the remaining \CountModulesWorkingCodeExamples{}
+modules.
+Some failing samples were analyzed and there were mainly
 two reasons for the failure: 
@@
 examples. As explained, to extract the run-time information, the behavior of the code
-under analysis was explicitly modified by wrapping the arguments. Furthermore, Jalangi’s
+under analysis was explicitly modified by wrapping the arguments. Furthermore, Jalangi's
 instrumentation itself caused some executions to fail, because the modules contained
@@
 modules was not applied, so it was not possible to determine which modules were failing
-only because of Jalangi’s own limitations.
+only because of Jalangi's own limitations.
 
@@
 for generating a declaration file. The extracted code examples for these modules did not
-execute the library. Hence, the collected run-time information was not useful
+execute the library. Hence, the collected run-time information was not sufficient
 for generating a declaration file.
@@
 The different names in the export assignment by modules \texttt{dirname-regex} and
-\texttt{escape-html} do not matter because the module can receive an arbitrary 
-name when it is imported: in
+\texttt{escape-html} do not matter because the module can be named arbitrarily when it is imported: in
 \lstinline[language=TypeScript]/import Abs = require('abs');/ the
@@
 
-\begin{figure}[tp]
+\begin{figure*}[tp]
     \centering
@@
     \label{fig:experiments-results-module-function}
-\end{figure}
-
-\begin{figure}[tp]
+\end{figure*}
+
+\begin{figure*}[tp]
   \centering
@@
   \label{fig:experiments-results-module-is-uuid}
-\end{figure}
+\end{figure*}
 
@@
   \label{fig:experiments-results-module-glob-base}
-\end{figure}
+\end{figure*}
 
@@
 We analyzed \CountModulesOnlySolvableDifferences{} of the \CountModulesGeneratedDeclarationFile{} generated files. The remaining 174 files contained at least one
-of the unimplemented TypeScript features so that they could not be considered. These
-features were identified by \texttt{dts-parse} as tags.
+of the unimplemented TypeScript features so that they could not be
+considered.
+% ? These features were identified by \texttt{dts-parse} as tags.
 
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
   \centering
@@
   \label{fig:experiments-results-manually-completed-examples}
-\end{figure}
+\end{figure*}
 

### 2021-05-13 mplr-2021/main.tex
Add examples for interactions

@@
 \texttt{interactions}. They are used for translating, modifying, and
-aggregating Jalangi's raw event information to get an application-specific data representation.
-Each function invocation is stored as a \texttt{FunctionContainer} which contains an \texttt{ArgumentContainer} for each argument.
-Each \texttt{ArgumentContainer} contains the name and index of the argument and a collection of \texttt{interactions}.
-The interaction \texttt{getField} gets triggered whenever a property of an object gets accessed and it tracks the name of the accessed field.
-The invocation of an object's method is tracked with the \texttt{methodCall} interaction.
-The \texttt{usedAsArgument} interaction gets triggered when an argument is used in the invocation of a function.
-The \texttt{followingInteractions} property is used for inferring nested interfaces.
-It is an array that recursively records \texttt{interactions} on the return value of \texttt{getField} or \texttt{methodCall}.
-
-We wrap each function's argument in a wrapper object, which  stores meta-information and greatly simplifies the mapping of an observation with the corresponding \texttt{ArgumentContainer} or \texttt{interaction}.
-We use the original values for critical operators such as \texttt{===} or \texttt{typeof}, for which wrapper objects would modify the behavior of the code.
-
-The property \texttt{requiredModule} stores the name of the module that declared the invoked function.
-If a function is explicitly exported by the module, the property \texttt{isExported} will be set to \texttt{true} when it is invoked.
-If a function of an exported object is invoked, \texttt{isExported} will be \texttt{false} and \texttt{requiredModule} will contain the name of the required module.
+aggregating Jalangi's raw event information to get an application-specific data representation. Each function invocation is stored as
+a \texttt{FunctionContainer} which contains an \texttt{ArgumentContainer} for each argument. Each \texttt{ArgumentContainer} contains 
+the name and index of the argument and a collection of \texttt{interactions}. As shown in \figref{fig:example-get-field-interaction}, the interaction \texttt{getField} gets triggered whenever 
+a property of an object gets accessed and it tracks the name of the accessed field and the type of the returned value. The invocation of an object's method is tracked with 
+the \texttt{methodCall} interaction. The \texttt{followingInteractions} property is used for inferring nested interfaces.
+It is an array that recursively records \texttt{interactions} on the return value of \texttt{getField} or 
+\texttt{methodCall}. \figref{fig:example-method-call-interaction} provides an example for both the \texttt{methodCall} interaction
+and the \texttt{followingInteractions} property. Finally, the \texttt{usedAsArgument} interaction
+gets triggered when an argument is used in the invocation of a function.
+
+\begin{figure*}[t]
+  \centering
+  \begin{subfigure}[t]{0.48\linewidth}
+    \begin{lstlisting}[language=JavaScript,numbers=none]
+function foo(hello) {
+  if (hello.world) {
+    // ...
+  }
+}
+
+foo({ world: "..." });
+    \end{lstlisting}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+    \begin{lstlisting}[language=JavaScript,numbers=none,escapechar=!]
+{
+  functionId_1: {
+    //...
+    functionName: 'foo',
+    args: {
+      0: {
+        //...
+        argumentName: "hello",
+        interactions: [
+          // ...
+          !\textbf{\{}!
+            !\textbf{code: "getField",}!
+            !\textbf{field: "world",}!
+            !\textbf{followingInteractions: [],}!
+            !\textbf{returnTypeOf: "string",}!
+          !\textbf{\},}!
+        ],
+      },
+    },
+  },
+}
+    \end{lstlisting}
+  \end{subfigure}
+  \caption{
+      Example code that for \texttt{getField} interaction
+    }
+  \label{fig:example-get-field-interaction}
+\end{figure*}
+
+\begin{figure*}[t]
+  \begin{subfigure}[t]{0.48\linewidth}
+      \begin{lstlisting}[language=JavaScript,numbers=none]
+function foo(bar) {
+  if (bar.hello().world) {
+    // ...
+  }
+}
+
+foo({
+  hello: function () {
+    return {
+      world: "xxx",
+    };
+  },
+});    
+      \end{lstlisting}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.48\linewidth}
+      \begin{lstlisting}[language=JavaScript,numbers=none,escapechar=!]
+{
+  functionId_1: {
+    //...
+    functionName: "foo",
+    args: {
+      0: {
+        // ...
+        argumentName: "bar",
+        interactions: [
+          // ...
+          !\textbf{\{}!
+            !\textbf{code: "methodCall",}!
+            !\textbf{methodName: "hello",}!
+            !\textbf{functionId: "functionId\_2",}!
+            !\textbf{followingInteractions: [}!
+              !\textbf{\{}!
+                !\textbf{code: "getField",}!
+                !\textbf{field: "world",}!
+                !\textbf{followingInteractions: [],}!
+                !\textbf{returnTypeOf: "string",}!
+              !\textbf{\},}!
+            !\textbf{],}!
+          !\textbf{\},}!
+        ],
+      },
+    },
+  },
+}        
+      \end{lstlisting}
+    \end{subfigure}
+  \caption{Example code for \texttt{}{methodCall} and \texttt{followingInteractions}}
+  \label{fig:example-method-call-interaction}
+  \end{figure*}
+
+We wrap each function's argument in a wrapper object, which  stores meta-information and greatly simplifies the mapping of an observation 
+with the corresponding \texttt{ArgumentContainer} or \texttt{interaction}. We use the original values for critical operators such as 
+\texttt{===} or \texttt{typeof}, for which wrapper objects would modify the behavior of the code.
+
+The property \texttt{requiredModule} stores the name of the module that declared the invoked function. If a function is explicitly 
+exported by the module, the property \texttt{isExported} will be set to \texttt{true} when it is invoked. If a function of an 
+exported object is invoked, \texttt{isExported} will be \texttt{false} and \texttt{requiredModule} will contain the name of 
+the required module.
 

### 2021-05-17 mplr-2021/main.tex
Comments May 18

@@
 
+% I think the reader gets the impression from the text that obtaining the code examples
+% is a "main feature" of our tool, where in fact it is only a way of easily obtaining
+% runtime information to execute our tool. We could've followed the approach of obtaining
+% the runtime information from the tests.
+% This can be useful also to divert the attention from how many repositories had README files, etc.
+% After using the code examples, we also found out that they provided a useful approximation
+% not only on the input types but also on how to use the library (i.e. we don't analyze the structure of the library).
+% This is a useful insight which can be used later to infer the correct types.
+%
+% Should we change this to 'we rely on runtime information'. 'We used the examples
+% provided by the developer as part of the documentation of a library as a first
+% pragmatic way of obtaining runtime information'.
 We aim to improve on this situation by providing a tool that generates
@@
 generating useful declaration files from code
-examples. \texttt{dts-generate} comes with a framework that 
+examples.
+% Should we change it to 'explore the possibilities for
+% generating useful declaration files from runtime information'.
+
+\texttt{dts-generate} comes with a framework that 
 supports the generation of declaration files for an existing
@@
 
-Like \texttt{dts-gen}, a TypeScript package that generates a template for a declaration
-file, we determine the kind of declaration file by examining the usage
+Unlike \texttt{dts-gen}, a TypeScript package that generates a template for a declaration
+file which only analyzes the structure of the module, we determine the kind of declaration file by examining the usage
 of the module. If the
@@
 
+% I think we should rephrase this, specially now that we know
+% it might be 'doable'.
+
 We considered option~\ref{item:2} under the assumption that most
@@
 
-In the end, option~\ref{item:3} was the most viable option even though there
-is no standard for documentation, either. However, almost all repositories
+% We can maybe change the focus to 'Option 3' was easier to implement.
+% This can be another opportunity to mention that  
+In the end, option~\ref{item:3} was the most viable option to easily gather runtime information
+even though there is no standard for documentation, either. However, almost all repositories
 contain README files where the library authors briefly describe in prose what
@@
 
+\begin{figure*}[t]
+  \begin{subfigure}[t]{0.48\linewidth}
+    \begin{lstlisting}[language=JavaScript,numbers=none]
+const foo = require('foo');
+
+foo.doSomething();
+    \end{lstlisting}
+  \caption{Example code for \texttt{module} template}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+    \begin{lstlisting}[language=JavaScript,numbers=none,escapechar=!]
+{
+  functionId_1: {
+    functionName: 'doSomething',
+    //...
+    !\textbf{isExported: false,}!
+    !\textbf{requiredModule: 'foo'}!
+  }
+}    
+    \end{lstlisting}
+    \caption{Runtime information fragment for \texttt{module} template}
+  \end{subfigure}
+
+  \begin{subfigure}[t]{0.48\linewidth}
+    \begin{lstlisting}[language=JavaScript,numbers=none]
+const foo = require('foo');
+
+foo();
+    \end{lstlisting}
+    \caption{Example code for \texttt{module-function} template}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+    \begin{lstlisting}[language=JavaScript,numbers=none,escapechar=!]
+{
+  functionId_1: {
+    functionName: 'foo',
+    //...
+    !\textbf{isExported: true,}!
+    !\textbf{requiredModule: 'foo'}!
+  }
+}    
+    \end{lstlisting}
+    \caption{Runtime information for \texttt{module-function} template}
+  \end{subfigure}
+  \caption{Example code for properties \texttt{isExported} and \texttt{requiredModule} - Property \texttt{isExported} is
+  \texttt{false} for \texttt{module} and \texttt{true} for \texttt{module-function}. Additionally, \texttt{requiredModule}
+  is \texttt{foo} for both cases.
+  }
+  \label{fig:example-is-exported-required-module}
+\end{figure*}
+
 We wrap each function's argument in a wrapper object, which  stores meta-information and greatly simplifies the mapping of an observation 
@@
 exported object is invoked, \texttt{isExported} will be \texttt{false} and \texttt{requiredModule} will contain the name of 
-the required module.
+the required module, as shown in \figref{fig:example-is-exported-required-module}.
 
@@
 
+% This is no longer true. I'll add one result to the report.
 We were unable to  generate a declaration file based on the

### 2021-05-18 mplr-2021/programming21-reviews-answered.txt
keeping track of the revisions

> ----------------------- REVIEW 1 ---------------------
> SUBMISSION: 28
> TITLE: Generation of TypeScript Declaration Files from JavaScript Code
> AUTHORS: Fernando Cristiani and Peter Thiemann
> 
> ----------- Overall evaluation -----------
> SCORE: -1 (major revision)
> ----- TEXT:
> Automatically generating TypeScript type signatures is a worthwhile
> goal; a mechanical process for generating such declarations would be
> less tedious, less error-prone, and more scalable. Pursuing a run-time
> trace-based technique is also compelling, given the significant
> difficulties that dynamically typed languages pose for static
> analysis. The initial results of the tool are promising: 244
> TypeScript declarations were automatically generated from existing
> module definitions, and 70 of these could be compared and validated
> against manually written declarations.
> 
> However, I believe the paper falls short both in terms of technical
> formulation as well as evidence of practical utility.
> 
> With respect to technical approach, the paper says (p.3/4) that the
> use of examples and dynamic analysis, rather than static analysis, is
> a novelty of the proposed tool. But a similar approach has been
> pursued previously:
> 
> * Trace Typing: An Approach for Evaluating Retrofitted Type Systems (Andreasen et al. ECOOP 2016)
> 
> * Just-in-Time Static Type Checking for Dynamic Languages (Ren et al. PLDI 2016)
> 
> * Dynamic Inference of Static Types for Ruby (An et al. POPL 2011)
> 
> * JSTrace: Run-time Type-Discovery for JavaScript (Saftoiu et al. 2010, > http://cs.brown.edu/research/plt/dl/flowtyping/v1/saftoiu-typediscovery-thesis.pdf)
> 
> How do the proposed techniques compare? Among these, trace typing is
> particularly relevant: like the current paper, it uses Jalangi for
> instrumentation.

Added discussion to the related work section.

> Besides not discussing this related work, the descriptions in Sections
> 4.2 and the end of 4.3 are rather informal: What exactly is the subset
> of JavaScript/TypeScript that is supported? What is the instrumented
> evaluator computing? How are types --- which include unions and
> intersections (at least for function types) --- being computed from
> the run-time information? The lack of precise details makes it
> difficult to understand what the tool currently does, how plausible it
> might be to extend to support more of JavaScript/TypeScript in the
> future, and how it relates the related work described above.
> 
> With respect to the implemented tool, it is not clear how close it is
> to being useful in practice. Many TypeScript features, as described in
> Table 1 and in the Evaluation, are not yet supported. For the 244
> modules for which declarations could be generated (and for the 70
> which could be compared to existing manual declarations): How many
> lines of code are in the modules? How many definitions are exported?
> How many examples were provided? How many examples were added to the
> 10 modules that were marked solvable? And how do these characteristics
> compare to the remaining Definitely Typed modules not currently
> supported? Especially because of the questions about technical
> formulation and novelty, it seems important for the system to come
> with a more comprehensive practical evaluation.
> 
> Overall, there is a lot of potential for this work --- especially on
> the practical side, for building a useful system --- but I think the
> results are preliminary to publish in their current form.
> 
> 
> *** Additional Comments and Questions ***
> 
> Fig 1: Does the tool generate declarations like this one? (If so, where
> does the name l__opts come from?)

Added explanation and a little cleanup.

> p.6: I think the reliance on Jalangi, as a black box, should be
> described earlier, along with related work.

Done.

> Fig 2: I spent some time wondering how the module names in (b) and (f)
> were chosen (the former appears to be camel case version of the module
> name, and the latter follows the name in the example program). Some of
> this explanation comes later, but perhaps it's worth adding more
> explanation here as this seems to be the primary place in the paper
> where the different kind of templates are explained.

Said that for Globtoregexp.I__opts already!

> Fig 3: Why not depict the example extraction step --- which sometimes
> includes manual addition of examples --- here? And consider
> representing Jalangi's role in this architecture.

Fernando?

> Sec 4.2/4.3: As mentioned, I think a lot more detail is warranted
> about the instrumentation and the process for computing types from
> traces.
> 
> "Execution of Code Examples": Interesting to see how many examples
> were perhaps out-of-date or were not proper examples. Were there any
> interesting patterns worth discussing?
> 
> p.17: It's encouraging to see how the proposed tool can fix some
> errors found in manual type declarations!
> 
> "Literal types vs string": If considering asking developers to mark
> strings, instead why not ask them to directly use, e.g. the "Enums" or
> "Discriminated Unions" (https://www.typescriptlang.org/docs/handbook)?
> 
> "Any": Ditto above; why not ask developers to choose the any type?
> 
> "Further types": Inferring indexed types, dependent types, etc. from
> run-time traces would seem to require quite significant technical
> extensions to the approach described here.
> 
> TSInfer: Is it fundamental that TSInfer apparently generates
> signatures for helpers? Could there not be a simple syntactic analysis
> that identifies top-level definitions, to be include in a declaration
> file?
> 
> 
> *** Minor Comments and Typos ***
> 
> Fig 1: Maybe I'm missing it, but I don't see Footnote 6 anywhere.j
> 
> p.5: "globToRegexp" ==> "GlobToRegexp" twice
> 
> p.8: I'm not sure why "syntax highlighting is done correctly" is
> noteworthy here.
> 
> Fig 4: Is this referenced in the text?
> 
> p.15: "loose" ==> "lose"
> 
> 
> 
> ----------------------- REVIEW 2 ---------------------
> SUBMISSION: 28
> TITLE: Generation of TypeScript Declaration Files from JavaScript Code
> AUTHORS: Fernando Cristiani and Peter Thiemann
> 
> ----------- Overall evaluation -----------
> SCORE: -1 (major revision)
> ----- TEXT:
> # Summary
> 
> The paper describes the tool dts-generate and its necessary companions to generate TypeScript declaration files for > JavaScript packages.
> These JavaScript packages need to provide a repository with a dedicated README that includes documentation efforts in the > form of exemplary usages.
> Based on the run-time information gathered by running these example, `dts-generate` produces the final TypeScript > declaration file after an additional analysis step that determines the scope of the file at hand (module, module-class, > module-function).
> The tool was evaluated based on all NPM packages that fulfilled the documentation requirement sufficiently and > community-driven declaration files; these declaration files are compared against the generated ones using an additional > tooling machinery developed by the authors. 
> 
> # My two cents
> 
> The general idea of the paper as well as the benefit of this idea (as well as the engineering effort) is worthwhile.
> The writing, on the other hand, does not feel mature enough for an acceptance.
> Unfortunately, I have problems myself to pin the exact reasons for this feeling.
> Let me try.
> 
> For an "empirical perspective" the evaluation (or here: result) section does not has the necessary depth to convince the > reader with, nor does it contain empirical methods (that I'm also not an expert for) that would strengthen the choice of this > perspective.
> Instead the results include further explanations of the engineering efforts necessary for the evaluation (how code examples > were extracted), single example extracts from the generated and original declaration files.
> The section concludes with a rather small paragraph (6.3 Evaluation) that only states that 70 files could be tested and all > qualified for a positive result.
> One deficit of this shallow evaluation origins from the small number of final declaration files that qualified for the evaluation.
> See some of the major and minor comments for more direct questions/remarks relating to specific sentences tackling this > deficiency.
> 
> The preceding part of the paper follows this rather shallow evaluation.
> The description of the engineering effort is given in detail, but not in depth.
> I guess these shallow descriptions are sufficient for a paper with an empirical emphasis, but since the focus of the current > version of paper is not the empirical part, I was wondering if an engineering perspective would be more suitable.
> In its current form, the description of the engineering effort needs additional work to make the shift to such an perspective > (what's the output the run-time information provided by Jalangi that dts-generate makes use of?).
> 
> All in all, I think that the paper needs a major revision (and shift of focus) in order to classify as empirical perspective or > needs to change the perspective to an engineering paper with additional efforts needed to fulfil the reader's wish to get more > in-depth insight to the technical part of this work.
> 
> ## Major comments
> 
> p. 1: I have the feeling that the abstract was not prepared in accordance to the specifications of the journal including the > keywords context, inquiry, approach, knowledge, grounding and importance.
> It's not that I do not get an impression of the work presented by the paper, but it just feels like an "ordinary abstract" and > not one that explicitly tackles the key questions behing the keywords above.
> 
> p. 4: "We examined all 6029 entries in the DefinitelyTyped repository and found 244 sufficiently well documented NPM > packages," - it's a shame that the final portion of repositories that can be tested is less than 5% of the starting value.
> 
> p. 15: "In this step, we loose another 50% of modules! This loss is mainly explained because some developers do not wrap > their code in a block using the javascript or js tags. However, as we are still left with code examples for 2260 modules, we did > not further look into code extraction as this number was considered sufficient for evaluating the generation of declaration > files." - Given the low value for the final number of repositories, you should definitely reconsider adapting your extraction > mechanism to be more robust!
> 
> p. 17: "We were unable to generate a declaration file based on the module-class template. All modules corresponding to this > template exhibited at least one of the unimplemented TypeScript features." - I was wondering if could make sense to adapt > the tooling machinery to leave out only the definitions that consist of unimplemented features, instead of leaving out a > whole module if one of its definition.
> 
> p. 18: "The remaining 174 files contained at least one of the unimplemented TypeScript features so that they could not be > considered." - This sentence sounds like you indeed generate as much as possible and filter out only layer --- how does this > statement relate my previous concern?
> 
> ## Minor comments
> 
> p. 5: "The choice of the template depends on the structure of the underlying JavaScript library:..." - here the following > listing/itemize environment seems to be missing, at least, I think that a itemize would be the better layout to use here.
> 
> p. 14: "dts-compare does not contemplate differences in function return types, because interface inference of function > return types is not covered in the current version of dts-generate." - Why is it that the inference for return types is that > different from parameters? You infer the parameters through example code, by running example code, you get concrete return > values that you can analyse in the same fashion. Furthermore, Figure 4 (for example) shows a generated file with functions > with return types. What am I missing here?
> ## Typos
> 
> all over the paper: runtime vs run time vs run-time!
> 
> p. 1: "..., which is error prone and time consuming." - error-prone
> p. 2: "..., which is time consuming and error prone." - error-prone
> p. 8: "A TypeScript Declaration File" - declaration file (why title-case?)
> p. 8: "... so that test cases were difficult to reap from the NPM packages:..." - read
> p. 9: "The Runtime Information block in Figure 3" - run-time information (why title-case?)
> p. 9: "For example, Function f ... Property foo ... Parameter a ... - function ... property ... parameter (it seems weird to start > with upper-case letters for the itemization)
> p. 11: " and traverse the Abstract Syntax Tree of a TypeScript program" - abstract syntax tree
> p. 12: "DefinitelyTyped Repository" - repository (why title-case?)
> p. 12: "uses the TypeScript Compiler API" - TypeScript compiler API (why title-case?)
> p. 18: "Excerpt of JS implementation that throws an Error when in- put parameter is not a string" - error
> p. 19: "Properties of interface Result were ignored for readability" - result
> p. 22: "future incorporation of a Symbolic Execution Engine that refines the" - symbolic execution engine (why title-case?)
> 
> ## Others
> p. 1: "from the TypeScript IDE and thus to incorrect uses of" - and, thus,  to
> p. 2: "a scripting language and thus lacks features for maintaining" - and, thus, lacks
> p. 4: "The implementation and the results are available under the following repositories:" -
> ...following repositories.
> p. 6: "It infers the overall structure of the JavaScript library,..." - non-referential _it_, why not write "the block infers..."
> p. 9: "...meaningful examples, thus avoiding possible cold start problems." - examples, thus, avoiding
> 
> 
> 
> ----------------------- REVIEW 3 ---------------------
> SUBMISSION: 28
> TITLE: Generation of TypeScript Declaration Files from JavaScript Code
> AUTHORS: Fernando Cristiani and Peter Thiemann
> 
> ----------- Overall evaluation -----------
> SCORE: -1 (major revision)
> ----- TEXT:
> The paper presents a tool that automatically generates TypeScript definition files for JavaScript projects. The tool uses > dynamic analysis. It runs the code samples found in the README file in the project repository, records information about > method invocations and argument values and uses those to generate type definitions. The paper evaluates the tool by > comparing generated definitions with hand-written definitions from the DefinitelyTyped repository - this is done for libraries > which have code samples in README, i.e. about ~270 out of ~6000.
> 
> The paper presents a tool that is useful, although it has a fairly limited scope. The approach presented by the authors - to > run a code sample using a library and use runtime monitoring to deduce the structure of the library - is definitely a promising > one. The authors make a good first step towards building a practical tool. However, the current implementation of the tool is > quite basic in a number of ways. First, it relies on code samples in README, which is not exactly common (it will work only > for quite basic libraries). Second, it ignores many interesting types that a library could use - most prominently arrays, but > those are just one example.
> 
> I believe the work presented in the paper is still interesting and useful, but it would need to be much more complete to > justify writing an extensive report about it. Although the journal does not have a specific category of "Tool" papers, I imagine > that if the authors presented the work in a shorter paper and clearly acknowledged the limitations and contributions, then it > would be interesting enough for publication. (Following the idea that the length of the paper should be commensurable with > the contribution.) The current version spends a quite a lot of time discussing the high-level structure of the tool and some > practical details which I did not find particularly important.
> 
> DETAILED COMMENTS
> 
> * In the introduction, you say that manually writing definitions is error-prone, but your paper then uses manually written > definitions from DefinitelyTyped as ground truth. I find this a bit paradoxical - maybe you could instead say that this is a lot > of work and there are many libraries that are not covered by DefinitelyTyped?
> 
> * I have to admit that I was a bit confused by the idea of "templates" that is used in the paper. Do the templates determine > just how you structure the definition, or are they somehow limiting what kind of constructs can be used in a given file? I think > this would benefit from some clarification in the paper.
> 
> * There is very little detail in the paper about how you infer types from the sample values that are observed during the > runtime execution. I guess the mechanism here is quite basic, but it is still worth discussing. (You might be interested in the > related paper [1] which does this for more complex types, but also covers things like records with optional fields.)
> 
> * I think your evaluation is somewhat limited by the fact that the libraries that have a complete documentation in the > README are the ones with very simple API. It would be interesting to evaluate how your approach works on larger libraries > (even if that meant copying some examples into a readme manually).
> 
> * I think the phrasing that "more advanced types are overblown" is a bit unfortunate, especially if the tool does not support > some quite basic types like arrays.
> 
> * The paper should give a much more extensive overview of related work - for example, it should discuss what has been > done using static analyses, whether similar tools exist for other languages and there might also be other interesting related > approaches (perhaps some based on machine learning methods). The references in the submission are fairly minimal.
> 
> [1] T. Petricek, G. Guera, D. Syme. Types from data: making structured data first-class citizens in F#, PLDI 2016

### 2021-05-18 mplr-2021/main.tex
reviewer #1 - related work

--- 2021-05-13 mplr-2021/main.tex
@@
+%% MPLR Format: up to 12 pages, excluding bibliography and appendix)
 %%
@@
 }
+%"
 
@@
 existing JavaScript libraries can be used in a TypeScript project by
-adding a declaration file that contains a description of the library's
+adding a declaration file that describes the library's
 API in terms of types. 
@@
 
-The novelty of our tool is twofold:
+The main contribution of our tool is twofold:
 \begin{enumerate}
@@
 
+Our implementation of code instrumentation  to gather data flow
+information and type information at run time is based on
+Jalangi \cite{DBLP:conf/sigsoft/SenKBG13}, a configurable
+framework for dynamic analysis for JavaScript.
 The implementation and the results are available in the following repositories:
 
-\begin{itemize}
+\begin{itemize}\footnotesize
   \item \url{https://github.com/proglang/run-time-information-gathering/tree/v1.1.0}
@@
 shown in \figref{fig:motivating-example-glob-to-regexp-vscode} and it is
-ready for use in a TypeScript project. For example,  Visual Studio's
-code completion runs properly with the generated file
-(\figref{fig:motivating-example-glob-to-regexp-vscode}). If the
+ready for use in a TypeScript project:  Visual Studio's
+code completion runs properly with the generated file. If the
 \texttt{glob-to-regexp} package gets modified in the future, a new declaration
@@
 
-The declaration file shown in \figref{fig:motivating-example-glob-to-regexp-vscode}
-describes a package with a single exported function. The first parameter is of type
-\texttt{string} and the second one is an optional object described by an interface. TypeScript namespaces
-organize types declared in a declaration file and avoid name clashes with other types
-\cite{typescript-namespaces}. For example, the interface
+The generated declaration file shown in \figref{fig:motivating-example-glob-to-regexp-vscode}
+describes a package with a single exported function. The module name
+is derived from the NPM module name by camlization\footnote{i.e.,
+  transforming to camel case.}. The first parameter is of type
+\texttt{string} and the second one is an optional object described by
+an interface. To avoid name clashes, \texttt{dts-generate} creates a
+namespace that corresponds to the module name. This pattern is a best
+practice to  organize types declared in a declaration file
+\cite{typescript-namespaces}. In the example, the interface
 \lstinline[language=TypeScript]{I__opts} is declared in namespace
-\lstinline[language=TypeScript]{globToRegexp} in
-\figref{fig:motivating-example-glob-to-regexp-vscode}. Hence its uses
+\lstinline[language=TypeScript]{globToRegexp}. Hence, its uses
 must be qualified by the namespace as in
-\lstinline[language=TypeScript]{globToRegexp.I__opts}.
+\lstinline[language=TypeScript]{globToRegexp.I__opts}. The name of the
+interface is derived from the name of the formal parameter.
 
@@
 Jalangi \cite{DBLP:conf/sigsoft/SenKBG13} to gather data flow
-information and type information at run time. Jalangi is a configurable
-framework for dynamic analysis of JavaScript. It provides several
+information and type information at run time. Jalangi provides several
 analysis modules that we extended as needed to retrieve the required
@@
 \label{sec:related-work}
+\paragraph*{Dynamic analysis techniques} Trace Typing
+\cite{DBLP:conf/ecoop/AndreasenGCSTS16} is a framework for evaluating
+retrofitted type systems. It involves gathering traces of JavaScript program
+executions, just like in our system. However, the goal of the type
+systems that they study is quite different from ours. They are
+interested in type information to improve run-time performance and
+traditional compilation, giving information about object layout and
+tag tests. Contrary to their approach, we do not just observe the
+types of arguments and return values. In addition, we also trace
+property accesses which translate to requirements on objects. Also, we
+have no generalization step to infer polymorphic types.
+
+Hummingbird \cite{DBLP:conf/pldi/RenF16} is a system for type checking
+Ruby programs at run time in the presence of
+metaprogramming. Metaprograms are extended to additionally generate
+type annotations for the generated code. Once a generated method is
+called, the method's body is statically checked against the actual
+argument types as well as the generated annotations. This system does
+not involve run-time tracing.
+
+Rubydust \cite{DBLP:conf/popl/AnCFH11} implements dynamic
+constraint-based type inference for Ruby. Rubydust runs an
+instrumented Ruby program that gathers subtyping constraints for
+parameters and return types of methods as well as for fields, which is
+quite different from the interaction that we collect. After
+completion, the solution of these constraints is proposed as a typing
+for the program. These types are sound if the runs cover sufficiently
+many paths in the program. Rubydust is reasonably effective because
+Ruby's type system is nominal. For JavaScript, this approach is much
+less effective as (run-time) typing is structural
+\cite{DBLP:conf/sfp/NausT16}.
+
+JSTrace \cite{saftoiu21:_jstrac} also performs dynamic type inference
+for JavaScript by gathering information at run time. It inspects the
+run-time types of arguments and return values of suitably wrapped
+functions. While the resulting information is treated similarly as in
+our work, we also gather interactions that result from property
+accesses and method calls and derive further typing constraints from
+those. Moreover, JSTrace performs an (expensive) deep inspection of
+the types of objects and arrays whereas our inspection is shallow and
+thus more efficient.
+
+None of the works tackles the question how to obtain a sufficient
+sample of run-time information. One part of our contribution is to
+systematically uncover the README files as a driver for running the
+JavaScript code under inspection.
+
 \paragraph*{dts-gen}

### 2021-05-18 mplr-2021/main.tex
reviewer#1

@@
 \lstinline[language=TypeScript]{I__opts} is declared in namespace
-\lstinline[language=TypeScript]{globToRegexp}. Hence, its uses
+\lstinline[language=TypeScript]{GlobToRegexp}. Hence, its uses
 must be qualified by the namespace as in
-\lstinline[language=TypeScript]{globToRegexp.I__opts}. The name of the
+\lstinline[language=TypeScript]{GlobToRegexp.I__opts}. The name of the
 interface is derived from the name of the formal parameter.
@@
   code examples are presented in code blocks labeled with the programming
-  language, so that syntax highlighting is done correctly. Hence, we
+  language. Hence, we
   retrieve the code examples from code blocks labeled \texttt{js} or
@@
 \paragraph*{Extraction of Code Examples}
-In this step, we loose another 50\% of modules! This loss is mainly
+In this step, we lose another 50\% of modules! This loss is mainly
 explained because some developers do not wrap their code in a block

### 2021-05-18 mplr-2021/programming21-reviews-answered.txt
reviewer#1

@@
 >
-> 
+
+DONE.
+
 > 

### 2021-05-20 mplr-2021/main.tex
snapshot

--- 2021-05-18 mplr-2021/main.tex
@@
 
-As \texttt{dts-generate} is based on run-time information, 
-exemplary code fragments that execute the JavaScript library are
-needed to obtain 
-significant run-time information from running the instrumented library
+As \texttt{dts-generate} is based on run-time information, code
+fragments are needed that import and use the JavaScript library to
+obtain run-time information from running the instrumented library
 code.

### 2021-05-20 mplr-2021/main.tex
streamlining section 4

@@
 
-As \texttt{dts-generate} is based on run-time information, code
-fragments are needed that import and use the JavaScript library to
-obtain run-time information from running the instrumented library
-code.
-
-The examples and the code base of the library are instrumented with
-Jalangi \cite{DBLP:conf/sigsoft/SenKBG13} to gather data flow
-information and type information at run time. Jalangi provides several
-analysis modules that we extended as needed to retrieve the required
-run-time information.
+As \texttt{dts-generate} is based on run-time information, we need code
+fragments that import and use the JavaScript library and we need to be
+able to obtain this run-time information while running the code.
+We explain in Section~\ref{sec:initial-code-base} how we obtain code
+fragments. 
+To obtain run-time information, we instrument
+the code fragments and the entire code base of the library with
+Jalangi \cite{DBLP:conf/sigsoft/SenKBG13}. We extend Jalangi's
+analysis modules to gather data flow and type information as required
+for our purposes and explained in Section~\ref{sec:run-time-information}.
+% Jalangi provides several
+% analysis modules that we extended as needed to retrieve the required
+% run-time information.
 %, which is then saved in a JSON file.
 
-A second independent block uses the run-time information to generate a
-TypeScript declaration file. It infers the overall structure of the JavaScript
-library, its interfaces, and the types from the run-time information. 
-The resulting declaration file is ready for use in the development
-process. Its contents mimic the usage of the library in the example
-code fragments and match the structure of the
-JavaScript library under analysis, so that the JavaScript code
+A second independent block explained in
+Section~\ref{sec:typescr-decl-file} uses the run-time information to
+generate a TypeScript declaration file. It infers the overall
+structure of the JavaScript library, its interfaces, and the types
+from the run-time information.  The resulting declaration file is
+ready for use in the development process. Its contents mimic the usage
+of the library in the example code fragments and match the structure
+of the JavaScript library under analysis, so that the JavaScript code
 generated after compiling the TypeScript code runs without
@@
 
-To extract run-time information from a JavaScript library, it is
-necessary, by definition, to execute the code, because the
-analysis modules provided by Jalangi can only gather information if the instrumented code gets executed.
-
-There are several options to obtain code fragments that drive the
-library code:
+% To extract run-time information from a JavaScript library, it is
+% necessary, by definition, to execute the code, because the
+% analysis modules provided by Jalangi can only gather information if the instrumented code gets executed.
+
+There are several options to obtain code fragments that exercise the
+library code in a meaningful way. In order to exercise library $A$,
+one might
 \begin{enumerate}
-\item\label{item:1} execute code that imports the library;
-\item\label{item:2} execute the test cases that come with the library;
-\item\label{item:3} execute code fragments extracted from the library documentation.
+\item\label{item:1} execute code that imports library $A$;
+\item\label{item:2} execute the test cases that come with library $A$;
+\item\label{item:3} execute code fragments extracted from library $A$'s documentation.
 \end{enumerate}
 
-Option~\ref{item:1} does not solve our problem, it just delegates it
-to the importing library, which also needs code to drive it. Moreover,
-we need to identify a dependent package, which is costly to download and instrument.
+Option~\ref{item:1} requires finding a library $B$ that depends on
+$A$. While this information is straightforward to obtain from \NPM, it
+is potentially costly to download and instrument $B$ for use with
+Jalangi. Moreover, this option delegates our problem
+to library $B$, which also needs code to drive it.
 
@@
 % This can be another opportunity to mention that  
-In the end, option~\ref{item:3} was the most viable option to easily gather runtime information
+In the end, option~\ref{item:3} was the most viable option to gather run-time information
 even though there is no standard for documentation, either. However, almost all repositories
@@
 library works and how to use it. These code fragments showcase common use cases rather
-than stress-testing the implementation (the problem of
+than stress-testing the implementation (a problem of
 option~\ref{item:2}) because they are meant to be instructive to users of the package.
@@
   written using Markdown\footnote{\url{https://www.markdownguide.org}}, a
-  popular markup language, where
+  popular markup language, where most
   code examples are presented in code blocks labeled with the programming
@@
   retrieve the code examples from code blocks labeled \texttt{js} or
-  \texttt{javascript}, which both stand for JavaScript in
-  Markdown.
+  \texttt{javascript}, which both stand for JavaScript.
 \end{itemize}
@@
 \begin{lstlisting}[
-  caption={Extracted example code for the \texttt{glob-to-regexp} package},
+  caption={Code extracted from the \texttt{glob-to-regexp} package},
   language=JavaScript,
@@
 appropriate and pragmatic way of extracting the developer's
-intention and provides an initial code base with meaningful
+intention. It provides an initial code base with meaningful
 examples.
@@
     \lstinline{b} a value of type \lstinline{number}. 
-  \item Property \lstinline{foo} of parameter \lstinline{a} of
-    function \lstinline{f} was accessed within the function. 
+  \item Function \lstinline{foo} was invoked. Its parameter
+    \lstinline{hello} was accessed as an object within the
+    function. Its property \lstinline{world} was found to have a
+    string value (see Figure~\ref{fig:example-get-field-interaction}).
+  \item Function \lstinline{foo} was invoked. Its parameter
+    \lstinline{bar} was accessed as an object. Its method
+    \lstinline{hello} was invoked. The return value of this call was
+    accessed as an object with property \lstinline{world}, which
+    returned a string (see Figure~\ref{fig:example-method-call-interaction}).
   \item Parameter \lstinline{a} of function \lstinline{f} was used as
@@
 
-The dynamic analysis framework Jalangi is used for gathering this kind of
-information. Jalangi's configurable analysis modules enable
+Jalangi's configurable analysis modules enable
 programming custom callbacks that can get triggered with virtually any
@@
 The implementation stores these observations as entities called
-\texttt{interactions}. They are used for translating, modifying, and
+\texttt{interactions}, which document each operation applied. They are used for translating, modifying, and
 aggregating Jalangi's raw event information to get an application-specific data representation. Each function invocation is stored as
 a \texttt{FunctionContainer} which contains an \texttt{ArgumentContainer} for each argument. Each \texttt{ArgumentContainer} contains 
-the name and index of the argument and a collection of \texttt{interactions}. As shown in \figref{fig:example-get-field-interaction}, the interaction \texttt{getField} gets triggered whenever 
-a property of an object gets accessed and it tracks the name of the accessed field and the type of the returned value. The invocation of an object's method is tracked with 
-the \texttt{methodCall} interaction. The \texttt{followingInteractions} property is used for inferring nested interfaces.
-It is an array that recursively records \texttt{interactions} on the return value of \texttt{getField} or 
-\texttt{methodCall}. \figref{fig:example-method-call-interaction} provides an example for both the \texttt{methodCall} interaction
-and the \texttt{followingInteractions} property. Finally, the \texttt{usedAsArgument} interaction
-gets triggered when an argument is used in the invocation of a function.
-
-\begin{figure*}[t]
+the name and index of the argument and a collection of
+\texttt{interactions} on the argument. As shown in \figref{fig:example-get-field-interaction}, the interaction \texttt{getField} gets triggered whenever 
+a property of an object gets accessed. It tracks the name of the
+accessed field and the type of the returned value. The invocation of a
+method is tracked with the \texttt{methodCall} interaction. An
+interaction can have a \texttt{followingInteractions} property if
+there are further operations on the return value of the interaction. It is used for inferring nested interfaces.
+That is, the \texttt{followingInteractions} recursively record \texttt{interactions} on the return value of \texttt{getField} or 
+\texttt{methodCall}, taking care of object identities to avoid
+looping. \figref{fig:example-method-call-interaction} provides an
+example for both the \texttt{methodCall} interaction 
+and the \texttt{followingInteractions} property. There is a
+\texttt{usedAsArgument} interaction (not shown) that
+gets triggered when an argument is used in the invocation of another function.
+
+\begin{figure}[t]
   \centering
-  \begin{subfigure}[t]{0.48\linewidth}
     \begin{lstlisting}[language=JavaScript,numbers=none]
@@
     // ...
-  }
-}
+} }
 
@@
     \end{lstlisting}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.48\linewidth}
     \begin{lstlisting}[language=JavaScript,numbers=none,escapechar=!]
-{
-  functionId_1: {
+{ functionId_1: {
     //...
@@
             !\textbf{returnTypeOf: "string",}!
-          !\textbf{\},}!
-        ],
-      },
-    },
-  },
-}
-    \end{lstlisting}
-  \end{subfigure}
+          !\textbf{\},}!],},},},}
+    \end{lstlisting}
   \caption{
-      Example code that for \texttt{getField} interaction
+      Example code for \texttt{getField} interaction
     }
   \label{fig:example-get-field-interaction}
-\end{figure*}
-
-\begin{figure*}[t]
-  \begin{subfigure}[t]{0.48\linewidth}
+\end{figure}
+
+\begin{figure}[t]
       \begin{lstlisting}[language=JavaScript,numbers=none]
@@
       world: "xxx",
-    };
-  },
-});    
-      \end{lstlisting}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.48\linewidth}
+};},});    
+      \end{lstlisting}
+
       \begin{lstlisting}[language=JavaScript,numbers=none,escapechar=!]
-{
-  functionId_1: {
+{ functionId_1: {
     //...
@@
                 !\textbf{returnTypeOf: "string",}!
-              !\textbf{\},}!
-            !\textbf{],}!
-          !\textbf{\},}!
-        ],
-      },
-    },
-  },
-}        
-      \end{lstlisting}
-    \end{subfigure}
-  \caption{Example code for \texttt{}{methodCall} and \texttt{followingInteractions}}
+              !\textbf{\},}! !\textbf{],}! !\textbf{\},}!],},},},}
+      \end{lstlisting}
+  \caption{\texttt{}{methodCall} and \texttt{followingInteractions}}
   \label{fig:example-method-call-interaction}
-\end{figure*}
+\end{figure}
 
@@
 
-We wrap each function's argument in a wrapper object, which  stores meta-information and greatly simplifies the mapping of an observation 
-with the corresponding \texttt{ArgumentContainer} or \texttt{interaction}. We use the original values for critical operators such as 
-\texttt{===} or \texttt{typeof}, for which wrapper objects would modify the behavior of the code.
-
-The property \texttt{requiredModule} stores the name of the module that declared the invoked function. If a function is explicitly 
-exported by the module, the property \texttt{isExported} will be set to \texttt{true} when it is invoked. If a function of an 
+To obtain this information we wrap each function's argument in a wrapper object, which stores meta-information and provides the mapping of an observation 
+to the corresponding \texttt{ArgumentContainer} or
+\texttt{interaction}. For operators that are aware of object identity, such as 
+\texttt{===} or \texttt{typeof}, we use the original values because
+they could return different results for wrapper objects.
+
+We also gather information to determine which  module template is most
+appropriate. To this end
+the property \texttt{requiredModule} stores the name of the module that declared the invoked function. If a function is explicitly 
+exported by the module, the property \texttt{isExported} is set to \texttt{true} when it is invoked. If a function of an 
 exported object is invoked, \texttt{isExported} will be \texttt{false} and \texttt{requiredModule} will contain the name of 
-the required module, as shown in \figref{fig:example-is-exported-required-module}.
-
-The run-time information is saved as a JSON file that can be used for later
-processing. The instrumentation tool is written in JavaScript and runs in \NodeJS{} in a
-Docker container.
+the required module, as shown in
+\figref{fig:example-is-exported-required-module}. If a function is
+used as a constructor, we set its \texttt{isConstructor} flag.
+
+% The run-time information is saved as a JSON file that can be used for later
+% processing. The instrumentation tool is written in JavaScript and runs in \NodeJS{} in a
+% Docker container.
 
@@
 
-The next step in the pipeline after gathering the run-time information is
+The next step after gathering the run-time information is
 the generation of the declaration file (cf.\
-\figref{fig:tsd_generation_method_block_diagram}). It is a lightweight,
-simple, and fast application, which solely relies on the run-time
-information gathered in the previous step.
+\figref{fig:tsd_generation_method_block_diagram}).
+% It is a lightweight,
+% simple, and fast application, which solely relies on the run-time
+% information gathered in the previous step.
 % The tool itself is written in TypeScript and runs in a
 % Docker container in \NodeJS.  
-
-Instead of analyzing the shape of the exported module, we choose the
-template based on how the module is used. We analyze the properties \texttt{isExported},
+%
+Contrary to \texttt{dts-gen} which analyzes the shape of the exported module, we choose the
+template based on the way the module is used. We analyze the properties \texttt{isExported},
 \texttt{requiredModule}, and \texttt{isConstructor} from the
-information gathered at run time to distinguish
+run-time information to distinguish
 between \textbf{module-class} and \textbf{module-function}. If a function is invoked and
@@
 
-We implemented the most common TypeScript features first and we focused our effort in building an
-end-to-end solution that generates useful declaration files. The
-types \texttt{string}, \texttt{number}, \texttt{boolean}, unions
-thereof, arrays, function callbacks, and optional elements were considered
-both for function parameters and interface properties. We also
-considered overloaded functions.
-Common types such as
-\texttt{any}, tuples, intersections (that
-is, types formed using the intersection operator \lstinline[language=TypeScript]/&/) and features such as
-generics and index signatures were not implemented.
-
-The selection of typing features covered in the tool is driven by
-examination of the files in
-the DefinitelyTyped repository. We parsed each declaration file with
+Our goal is to support the  most commonly used TypeScript
+features first in our implementation. To this end,
+the selection of typing features covered in the tool is driven by
+an analysis of the 6648 declaration files in
+the DefinitelyTyped repository. \footnote{As of commit
+  \texttt{da93d7b13094bacb170ead3f4a289f3b8687e4e5} (April 4, 2020).}
+We parsed each declaration file with 
 \texttt{dts-parse} and tagged it according to the features used in the file.
 Table~\ref{tab:dts-parse-stats} summarizes the results.
-We decided to prioritize the features with the highest number of
-occurrences and support them first in our framework.
+The \textbf{Count} column indicates the number of declaration files that use a
+specific feature. The 3rd column shows the corresponding
+percentage with respect to all files in the repository.
+
+We start with a limited scope on the features with the highest number of
+occurrences. Thus, we focus our effort on building an
+end-to-end solution that generates useful declaration files. We choose
+the types \texttt{string}, \texttt{number}, \texttt{boolean}, unions
+thereof, arrays, function callbacks, and optional elements are considered
+both for function parameters and interface properties. We also
+treat aliases and overloaded functions.
+Some common types such as \texttt{any}, literals, generics, and index
+signatures are not yet implemented. 
+Rare types such as tuples and intersections (that
+is, types formed using the intersection operator
+\lstinline[language=TypeScript]/&/) were not considered.
 
 \begin{table}[tp]
-  \caption{Distribution of TypeScript features in declaration files uploaded to DefinitelyTyped (DT).
-The 2nd column shows how many files in DT make use of a specific TypeScript feature.
-The 3rd column shows the corresponding percentage for a total of 6648 declaration files in DT.
-Numbers were obtained from commit \texttt{da93d7b13094bacb170ead3f4a289f3b8687e4e5} (April 4, 2020).}
   \begin{center}
@@
       \hline
-      \textbf{TypeScript feature} & \textbf{Count} & \multicolumn{1}{|c|}{\textbf{\%}} \\ 
+      \textbf{Feature} & \textbf{Count} & \multicolumn{1}{|c|}{\textbf{\%}} \\ 
       \hline
@@
   \end{center}
+  \caption{Usage of TypeScript features on DefinitelyTyped}
   \label{tab:dts-parse-stats}

### 2021-05-20 mplr-2021/main.tex
Add example for module-class

--- 2021-05-20 mplr-2021/main.tex
@@
 
-% I think the reader gets the impression from the text that obtaining the code examples
-% is a "main feature" of our tool, where in fact it is only a way of easily obtaining
-% runtime information to execute our tool. We could've followed the approach of obtaining
-% the runtime information from the tests.
-% This can be useful also to divert the attention from how many repositories had README files, etc.
-% After using the code examples, we also found out that they provided a useful approximation
-% not only on the input types but also on how to use the library (i.e. we don't analyze the structure of the library).
-% This is a useful insight which can be used later to infer the correct types.
-%
-% Should we change this to 'we rely on runtime information'. 'We used the examples
-% provided by the developer as part of the documentation of a library as a first
-% pragmatic way of obtaining runtime information'.
 We aim to improve on this situation by providing a tool that generates
-TypeScript declaration files automatically. To do so, we rely on the examples provided
+TypeScript declaration files automatically. To do so, we rely on we rely on runtime information
+gathered from the examples provided
 by the developer as part of the documentation of a library. If these
@@
 Our tool \texttt{dts-generate} is a first step to explore the possibilities for
-generating useful declaration files from code
-examples.
-% Should we change it to 'explore the possibilities for
-% generating useful declaration files from runtime information'.
+generating useful declaration files from runtime information.
 
@@
 
-As \texttt{dts-generate} is based on run-time information, code
-fragments are needed that import and use the JavaScript library to
-obtain run-time information from running the instrumented library
+As \texttt{dts-generate} is based on run-time information, 
+exemplary code fragments that execute the JavaScript library are
+needed to obtain 
+significant run-time information from running the instrumented library
 code.
@@
 we need to identify a dependent package, which is costly to download and instrument.
-
-% I think we should rephrase this, specially now that we know
-% it might be 'doable'.
 
@@
 
+\begin{figure*}[tp]
+  \centering
+  \begin{subfigure}[t]{0.48\linewidth}
+    \begin{lstlisting}[language=TypeScript,numbers=none]
+export = SteamID;
+declare class SteamID {
+    // ...
+    isValid(): boolean;
+    isGroupChat(): boolean;
+    isLobby(): boolean;
+    getSteam2RenderedID(newerFormat?: boolean): string;
+    // ...
+}
+declare namespace SteamID {
+    export function fromIndividualAccountID(accountid: string | number): SteamID;
+}  
+    \end{lstlisting}
+    \caption{steamid/index.d.ts - Generated (simplified)}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+    \begin{lstlisting}[language=TypeScript,numbers=none]
+// ...
+type getSteam2RenderedID = (newerFormat?: boolean) => string;
+
+declare class SteamID {
+    // ...
+    isValid(): boolean;
+    // ...
+    isGroupChat(): boolean;
+    // ...
+    isLobby(): boolean;
+    getSteam2RenderedID: getSteam2RenderedID;
+    // ...
+}
+
+declare namespace SteamID {
+    // ...
+    function fromIndividualAccountID(accountid: number | string): SteamID;
+}
+
+export = SteamID;
+    \end{lstlisting}
+    \caption{steamid/index.d.ts - DT (simplified)}
+  \end{subfigure}
+
+  \caption{Results for \textbf{module-class} template for module \texttt{steamid}}
+  \label{fig:experiments-results-module-steamid}
+\end{figure*}
+
 \figref{fig:experiments-results-module-is-uuid} contains an example
@@
 
-% This is no longer true. I'll add one result to the report.
-We were unable to  generate a declaration file based on the
-\textbf{module-class} template. All modules corresponding to this template exhibited at least one of
-the unimplemented TypeScript features.
+\figref{fig:experiments-results-module-steamid} exposes a simplified example for the
+\textbf{module-class} template where we highlight the most important features. Class methods are
+correctly generated. For the \texttt{getSteam2RenderedID} method we correctly generate a signature
+that matches the type alias chosen by the author of the library. Finally, the method \texttt{fromIndividualAccountID}
+is correctly located in the namespace section.
 

### 2021-05-20 mplr-2021/main.tex

--- 2021-05-20 mplr-2021/main.tex
@@
 to library $B$, which also needs code to drive it.
-
-% I think we should rephrase this, specially now that we know
-% it might be 'doable'.
 

### 2021-05-21 mplr-2021/programming21-reviews-answered.txt
Add comments on review
computing types from run-time information

@@
 > 4.2 and the end of 4.3 are rather informal: What exactly is the subset
-> of JavaScript/TypeScript that is supported? What is the instrumented
+> of JavaScript/TypeScript that is supported?
+
+ECMAScript 5.1
+
+> What is the instrumented
 > evaluator computing? How are types --- which include unions and
@@
 > future, and how it relates the related work described above.
-> 
+
+Added some details about type computation at end of section 4.
+
 > With respect to the implemented tool, it is not clear how close it is
@@
 Fernando?
+F: Added new images:
+  - mplr-2021/graphics/dts-generate-block-diagram-with-code-examples-block.pdf
+  - mplr-2021/graphics/dts-generate-block-diagram-with-jalangi.pdf
 

### 2021-05-21 mplr-2021/main.tex
computing types from run-time information

@@
 the code fragments and the entire code base of the library with
-Jalangi \cite{DBLP:conf/sigsoft/SenKBG13}. We extend Jalangi's
+Jalangi2 \cite{DBLP:conf/sigsoft/SenKBG13}. We extend Jalangi's
 analysis modules to gather data flow and type information as required
-for our purposes and explained in Section~\ref{sec:run-time-information}.
+for our purposes and explained in
+Section~\ref{sec:run-time-information}.
+\texttt{dts-generate} supports full ECMAScript~5.1, which is
+the latest JavaScript version supported by Jalangi2. 
 % Jalangi provides several
@@
 
-Finally, interfaces are created by exploring \texttt{getField} and \texttt{methodCall} interactions from the run-time information. We
+Interfaces are created by exploring \texttt{getField} and \texttt{methodCall} interactions from the run-time information. We
 gather the interactions for a specific argument and build the
@@
 to be equivalent are merged.
+
+To construct the type signature for a function, we inspect the
+primitive types of each argument as well as the interface created from
+interactions on the argument in the function body. We do the same for
+the return type of the function. This procedure yields a candidate
+type for each function call observed during the execution.
+For instance, for the \lstinline{globToRegExp} examples shown in
+Listing~\ref{code:code-example-extracted-motivating-example} we obtain
+the following observed candidate types
+\begin{verbatim}
+globToRegExp : (glob: string, opts: undefined) -> RegExp
+globToRegExp : (glob: string, 
+                opts: { extended : bool,
+                        globstar?: any,
+                        flags?: any }) -> RegExp
+\end{verbatim}
+As the type for \lstinline{glob} and the return type
+\lstinline{RegExp} are the same, we collapse the type for
+\lstinline{opts} into a union type. The union with the
+\lstinline{undefined} type is special and converted into an optional
+argument so we arrive at a single collapsed candidate type
+\begin{verbatim}
+globToRegExp : (glob: string, 
+                opts?: { extended : bool,
+                         globstar?: any,
+                         flags?: any }) -> RegExp
+\end{verbatim}
+We proceed in this way with merging further candidate types until we
+reach a set of non-mergable types for the function. If this set
+contains more than one candidate type, then we generate an overloaded
+definition. For example, the candidate types
+\begin{verbatim}
+f(string): string
+f(number): number
+f(boolean): boolean
+\end{verbatim}
+cannot be merged as the return types are different. They give rise to
+an overloaded definition with three alternatives. However, given the candidate types
+\begin{verbatim}
+g(string): string
+g(number): string
+g(boolean): boolean
+\end{verbatim}
+we can merge the first two types on the first argument (because all
+remaining arguments and the return types are equal) by collecting \lstinline{string} and
+\lstinline{number} into a union type. The resulting set of non-mergable types
+is thus
+\begin{verbatim}
+g(string|number): string
+g(boolean): boolean
+\end{verbatim}
 

### 2021-05-21 mplr-2021/main.tex
Add manually added code example

@@
 \end{verbatim}
+After expanding manually the code base with one particular case, we obtain the following type
+\begin{verbatim}
+re = globToRegExp("*/www/{*.js,*.html}", {
+  flags: "i",
+  globstar: true,
+});
+
+globToRegExp : (glob: string, 
+                opts: { extended? : any,
+                        globstar: bool,
+                        flags: string }) -> RegExp
+\end{verbatim}
+
 As the type for \lstinline{glob} and the return type
@@
 globToRegExp : (glob: string, 
-                opts?: { extended : bool,
-                         globstar?: any,
-                         flags?: any }) -> RegExp
+                opts?: { extended? : bool,
+                         globstar?: bool,
+                         flags?: string }) -> RegExp
 \end{verbatim}

### 2021-05-26 mplr-2021/main.tex
discussion of dts-gen

@@
 
-Unlike \texttt{dts-gen}, a TypeScript package that generates a template for a declaration
-file which only analyzes the structure of the module, we determine the kind of declaration file by examining the usage
-of the module. If the
-imported entity is used solely as a function as in Figure~\ref{fig:example-module-function}, the
-generated declaration file follows the \textbf{module-function} template as shown in Figure~\ref{fig:template-module-function}.
+The Typescript project provides a package \texttt{dts-gen}, which
+generates a template for a declaration file by analyzing the structure
+of the module. The resulting file is intended as a starting point for further
+manual development of the declaration file. 
+Unlike \texttt{dts-gen} we determine the kind of declaration file by
+examining the \emph{usage} of the module. If the imported entity is
+used solely as a function as in
+Figure~\ref{fig:example-module-function}, the 
+generated declaration file follows the \textbf{module-function}
+template as shown in Figure~\ref{fig:template-module-function}. 
 If the example only accesses properties of the imported entity as in
-Figure~\ref{fig:example-module}, we generate according to the \textbf{module} template (Figure~\ref{fig:template-module}).
+Figure~\ref{fig:example-module}, we generate the declarations
+according to the \textbf{module} template
+(Figure~\ref{fig:template-module}). 
 If the imported entity is used with \lstinline/new/ to create new instances as in
@@
 by the examples in the documentation and 
-hence reflects the developers intent.
+hence reflects the developer's intent.
 % Run-time analysis proves useful in this case,

### 2021-05-26 mplr-2021/main.tex
explaining globtoregexp

@@
 run-time information, and generates a TypeScript declaration
-file. The result is 
-shown in \figref{fig:motivating-example-glob-to-regexp-vscode} and it is
+file. As the package is insufficiently documented, we provide one
+additional example to generate the result 
+shown in \figref{fig:motivating-example-glob-to-regexp-vscode} (see
+discussion in Section~\ref{sec:typescr-decl-file}). This declaration is
 ready for use in a TypeScript project:  Visual Studio's
@@
 globToRegExp : (glob: string, 
-                opts: { extended : bool,
-                        globstar?: any,
-                        flags?: any }) -> RegExp
+                opts: { extended: bool,
+                        globstar: undefined,
+                        flags: undefined }) -> RegExp
 \end{verbatim}
-After expanding manually the code base with one particular case, we obtain the following type
+These types are correct, but underapproximate the best possible type
+for the package. Our comparison tool (see Section~\ref{sec:dts-compare}) classifies the differences as
+solvable, hence we provide one more code example:
 \begin{verbatim}
@@
 });
-
+\end{verbatim}
+From this additional example, we obtain the following type:
+\begin{verbatim}
 globToRegExp : (glob: string, 
-                opts: { extended? : any,
+                opts: { extended: undefined,
                         globstar: bool,
@@
 
-As the type for \lstinline{glob} and the return type
+As the type for the argument \lstinline{glob} and the return type
 \lstinline{RegExp} are the same, we collapse the type for
@@
 \lstinline{undefined} type is special and converted into an optional
-argument so we arrive at a single collapsed candidate type
+argument. Moreover, the union of object types proceeds
+componentwise. Altogether, we arrive at a single collapsed candidate
+type.
 \begin{verbatim}
 globToRegExp : (glob: string, 
-                opts?: { extended? : bool,
+                opts?: { extended?: bool,
                          globstar?: bool,
@@
 \lstinline{number} into a union type. The resulting set of non-mergable types
-is thus
+has two elements.
 \begin{verbatim}

### 2021-05-26 mplr-2021/main.tex
details according to reviewer #2

@@
   The repository is populated and maintained manually by volunteers, which
-  is error prone and time consuming. Discrepancies between a
+  is error-prone and time consuming. Discrepancies between a
   declaration file and the JavaScript implementation lead to
-  incorrect feedback from the TypeScript IDE and thus to incorrect uses
+  incorrect feedback from the TypeScript IDE and, thus, to incorrect uses
   of the underlying JavaScript library.
@@
 JavaScript was never intended to be more than a
-scripting language and thus lacks features for maintaining and evolving large
+scripting language and, thus, lacks features for maintaining and evolving large
 codebases. However, nowadays developers create large and complex
@@
 files in DefinitelyTyped is conducted manually,
-which is time consuming and  error prone. Errors are aggravated because TypeScript takes a
+which is time consuming and  error-prone. Errors are aggravated because TypeScript takes a
 declaration file at face value. Discrepancies between declaration and
@@
 We aim to improve on this situation by providing a tool that generates
-TypeScript declaration files automatically. To do so, we rely on we rely on runtime information
+TypeScript declaration files automatically. To do so, we rely on we rely on run-time information
 gathered from the examples provided
@@
 Our tool \texttt{dts-generate} is a first step to explore the possibilities for
-generating useful declaration files from runtime information.
+generating useful declaration files from run-time information.
 
@@
 framework for dynamic analysis for JavaScript.
-The implementation and the results are available in the following repositories:
+The implementation and the results are available in the following repositories.
 
@@
 shown in \figref{fig:motivating-example-glob-to-regexp-vscode} (see
-discussion in Section~\ref{sec:typescr-decl-file}). This declaration is
-ready for use in a TypeScript project:  Visual Studio's
-code completion runs properly with the generated file. If the
+discussion in Section~\ref{sec:typescr-decl-file}).  The interface is
+detected correctly. Optional parameters are detected. This declaration is
+ready for use in a TypeScript project:  the generated file works
+properly with Visual Studio Code\footnote{\url{https://code.visualstudio.com}}. If the
 \texttt{glob-to-regexp} package gets modified in the future, a new declaration
@@
   \caption{Declaration file for \texttt{glob-to-regexp} 
-    generated with \texttt{dts-generate} - The interface is
-    detected correctly. Optional parameters are detected. The declaration 
-    file is usable in Visual Studio Code\protect\footnote{\url{https://code.visualstudio.com}}} 
+    generated with \texttt{dts-generate}} 
   \label{fig:motivating-example-glob-to-regexp-vscode}
@@
     Run-time information is gathered driven by example code.
-    A TypeScript Declaration File is generated using run-time
+    A TypeScript declaration file is generated using run-time
     information.
@@
 Section~\ref{sec:typescr-decl-file} uses the run-time information to
-generate a TypeScript declaration file. It infers the overall
+generate a TypeScript declaration file. This part infers the overall
 structure of the JavaScript library, its interfaces, and the types
@@
 \label{sec:run-time-information}
-The Runtime Information block in
+The run-time information block in
 \figref{fig:tsd_generation_method_block_diagram} gathers
 type-related information as well as usage-related information for each function parameter.
-For example,
+Here are some examples.
 
@@
     \end{lstlisting}
-    \caption{Runtime information fragment for \texttt{module} template}
+    \caption{Run-time information fragment for \texttt{module} template}
   \end{subfigure}
@@
     \end{lstlisting}
-    \caption{Runtime information for \texttt{module-function} template}
+    \caption{Run-time information for \texttt{module-function} template}
   \end{subfigure}
@@
 
-  \caption{Results for \textbf{module}  module \texttt{is-uuid}}
+  \caption{Results for \textbf{module} \texttt{is-uuid}}
   \label{fig:experiments-results-module-is-uuid}
@@
       {\includegraphics[width=1\linewidth]{evaluation-diagram.pdf}}
-      \caption[Evaluation against DefinitelyTyped Repository]{\textbf{Evaluation of
-          generated declaration files against DefinitelyTyped Repository} - The parser
+      \caption[Evaluation against DefinitelyTyped repository]{\textbf{Evaluation of
+          generated declaration files against DefinitelyTyped repository} - The parser
         uses the TypeScript Compiler           API \cite{typescript-compiler-api} to
@@
 The first step is to parse the declaration files using the TypeScript
-Compiler API to build and traverse the 
-Abstract Syntax Tree of a TypeScript program
+compiler API to build and traverse the 
+abstract syntax tree of a TypeScript program
 \cite{typescript-compiler-api}. 
@@
 not contemplate differences in function return types, because interface inference of
-function return types is not covered in the current version of \texttt{dts-generate}.
+function return types is limited in the current version of
+\texttt{dts-generate}. While the return values are inspected, a proper
+interface inference requires further operations on the return value
+(cf.\ the explanation of interactions in \secref{sec:run-time-information}).
 
@@
     \textbf{Number of analyzed modules for each stage of the experiment} - 
-  A TypeScript Declaration File was generated for only \CountModulesGeneratedDeclarationFile{} modules, out of \CountTotalModulesDefinitelyTyped{} modules in the DefinitelyTyped Repository. 
+  A TypeScript declaration file was generated for only \CountModulesGeneratedDeclarationFile{} modules, out of \CountTotalModulesDefinitelyTyped{} modules in the DefinitelyTyped repository. 
   % It was only possible to gather valid run-time information for about 25\% of the modules for which a Code Example was extracted.
@@
     \begin{lstlisting}[language=TypeScript,numbers=none]
-export function v1(str: string): boolean;
-export function v2(str: string): boolean;
-export function v3(str: string): boolean;
-export function v4(str: string): boolean;
-export function v5(str: string): boolean;
-    \end{lstlisting}
-    \caption{is-uuid/index.d.ts - Generated}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.48\linewidth}
-    \begin{lstlisting}[language=TypeScript,numbers=none]
-export function v1(value: string): boolean;
-export function v2(value: string): boolean;
-export function v3(value: string): boolean;
-export function v4(value: string): boolean;
-export function v5(value: string): boolean;
-export function nil(value: string): boolean;
-export function anyNonNil(value: string): boolean;
-    \end{lstlisting}
-    \caption{is-uuid/index.d.ts - DT}
-  \end{subfigure}
-
-  \caption{Results for \textbf{module} \texttt{is-uuid}}
-  \label{fig:experiments-results-module-is-uuid}
-\end{figure*}
-
-\begin{figure*}[tp]
-  \centering
-  \begin{subfigure}[t]{0.48\linewidth}
-    \begin{lstlisting}[language=TypeScript,numbers=none]
 export = SteamID;
@@
     \caption{glob-base/index.js - Excerpt of JS implementation
-      that throws an Error when input parameter is not a string} 
+      that throws an error when input parameter is not a string} 
     \label{fig:subfloat-globbase-js-implementation}
@@
 generalization is desired.
+
+We believe that  JavaScript developers can be taught to write their
+code examples in such a style that high-quality TypeScript signatures
+can be extracted from them without human intervention. 
+In the subsequent paragraphs we survey some approaches that rely on
+developers using special strings and other constants in their code examples.
 
@@
 package\footnote{\url{https://www.npmjs.com/package/@types/carlo}}
-uses it to distinguish different 
+uses literals to distinguish different 
 types of events.
@@
 For our framework, it is hard to distinguish a string that is meant literally from a
-string that just serves as an example. For an automatic solution, one might collect
+string that just serves as an example. An automatic solution might collect
 typical values of literal strings and generate the corresponding literal types (or
 unions thereof) when the arguments used in the example are all typical. Alternatively, one
-might always generate literal types unless the number of different examples provided is
-greater than some threshold, in which case the generator generalizes to type
+might always generate literal types until the number of different
+examples passes some threshold, in which case the generator
+generalizes to type 
 \lstinline/string/. Both alternatives come with the problem that they
-create a local over-approximation, so that it would become harder to
+create a local over-approximation, which makes it harder to
 qualify the output of the generator. A further alternative that does
@@
 Suppose the programmer supplies examples that use \lstinline/number/, \lstinline/bool/, and
-\lstinline/string/ at the same argument position. The obvious approximation to the type of
-this argument is the union type \lstinline/number|bool|string/. But what if the programmer
+\lstinline/string/ at the same argument position. In this case, our
+tool calculates the union type \lstinline/number|bool|string/. But what if the programmer
 wants to advertise the type \lstinline/any/ for an argument? This intent is hard to
@@
 
-The tool analyzes the shape of the objects at runtime after initialization without
+The tool analyzes the shape of the objects at run time after initialization without
 executing the library. This results in most parameters and results being inferred as
@@
 for specific JavaScript libraries without additional manual tasks. The architecture
-contemplates future incorporation of a Symbolic Execution Engine that refines the
+contemplates future incorporation of a symbolic execution engine that refines the
 initial code base enabling the exploration of new execution paths. Although it is not

### 2021-05-26 mplr-2021/main.tex
related work

@@
 the declaration file provided for the same module on
-DefinitelyTyped. In other words, we use DefinitelyTyped as our ground truth.
+DefinitelyTyped. In other words, we use DefinitelyTyped as our ground
+truth, because it is the only available reference against we can
+compare our results.
 To this end, we created \texttt{dts-compare}, a tool to compute the
@@
 Similar approaches could be conceived for indexed types, polymorphic types, dependent
-types, etc. However, at present we believe incorporating these features in a tool to be
-overblown because they are not widely used in DefinitelyTyped (see
+types, etc. However, at present we believe incorporating these
+features in a tool to be of limited value because they are not widely used in DefinitelyTyped (see
 Table~\ref{tab:dts-parse-stats}).
@@
 
+\paragraph{Data Type Inference}
+Petricek and coworkers describe a mechanism to infer a datatype from
+an example input in JSON, XML, or CSV format
+\cite{DBLP:conf/pldi/PetricekGS16}. Their implementation relies on
+type providers, a metaprogramming facility of F\#. The type provider reads the
+example input at compile time and translates it into a static
+type. The approach is purely data-centric---the shape of the sample
+data determines what is legal in the program. Our approach
+analyzes the program's behavior on sample data and generalizes these
+observations to a type.
+
 \paragraph*{dts-gen}

### 2021-05-26 mplr-2021/main.tex
cosmetics

@@
 Similar approaches could be conceived for indexed types, polymorphic types, dependent
-types, etc. However, at present we believe incorporating these
+types, etc. At present we believe incorporating these
 features in a tool to be of limited value because they are not widely used in DefinitelyTyped (see

### 2021-05-26 mplr-2021/programming21-reviews-answered.txt
logged

@@
 > "Discriminated Unions" (https://www.typescriptlang.org/docs/handbook)?
-> 
+
+Misconception! They write JavaScript, not TypeScript! Added a
+paragraph of clarification.
+
 > "Any": Ditto above; why not ask developers to choose the any type?
@@
 > Fig 1: Maybe I'm missing it, but I don't see Footnote 6 anywhere.j
-> 
+
+FIXED
+
 > p.5: "globToRegexp" ==> "GlobToRegexp" twice
-> 
+
+FIXED
+
 > p.8: I'm not sure why "syntax highlighting is done correctly" is
 > noteworthy here.
-> 
+
+REMOVED
+
 > Fig 4: Is this referenced in the text?
-> 
+
+FIXED
+
 > p.15: "loose" ==> "lose"
@@
 
-DONE.
+FIXED
 
@@
 > 
-> p.
-1: I have the feeling that the abstract was not prepared in accordance to the specifications of the journal including the > keywords context, inquiry, approach, knowledge, grounding and importance.
-> It's not that I do not get an impression of the work presented by the paper, but it just feels like an "ordinary abstract" and > not one that explicitly tackles the key questions behing the keywords above.
+> p. 1: I have the feeling that the abstract was not prepared in
+> accordance to the specifications of the journal including the
+> keywords context, inquiry, approach, knowledge, grounding and importance.
+> It's not that I do not get an impression of the work presented by
+> the paper, but it just feels like an "ordinary abstract" and
+> not one that explicitly tackles the key questions behing the keywords above.
 > 
-> p.
-4: "We examined all 6029 entries in the DefinitelyTyped repository and found 244 sufficiently well documented NPM > packages," - it's a shame that the final portion of repositories that can be tested is less than 5% of the starting value.
+> p. 4: "We examined all 6029 entries in the DefinitelyTyped
+> repository and found 244 sufficiently well documented NPM
+> packages," - it's a shame that the final portion of repositories that can be tested is less than 5% of the starting value.
 > 
-> p.
-15: "In this step, we loose another 50% of modules!
-This loss is mainly explained because some developers do not wrap > their code in a block using the javascript or js tags.
-However, as we are still left with code examples for 2260 modules, we did > not further look into code extraction as this number was considered sufficient for evaluating the generation of declaration > files." - Given the low value for the final number of repositories, you should definitely reconsider adapting your extraction > mechanism to be more robust!
+> p. 15: "In this step, we loose another 50% of modules! This loss is
+> mainly explained because some developers do not wrap
+> their code in a block using the javascript or js tags. However, as
+> we are still left with code examples for 2260 modules, we did
+> not further look into code extraction as this number was considered
+> sufficient for evaluating the generation of declaration
+> files." - Given the low value for the final number of repositories,
+> you should definitely reconsider adapting your extraction
+> mechanism to be more robust!
 > 
-> p.
-17: "We were unable to generate a declaration file based on the module-class template.
-All modules corresponding to this > template exhibited at least one of the unimplemented TypeScript features." - I was wondering if could make sense to adapt > the tooling machinery to leave out only the definitions that consist of unimplemented features, instead of leaving out a > whole module if one of its definition.
+> p. 17: "We were unable to generate a declaration file based on the
+> module-class template. All modules corresponding to this template
+> exhibited at least one of the unimplemented TypeScript features." -
+> I was wondering if could make sense to adapt
+> the tooling machinery to leave out only the definitions that consist
+> of unimplemented features, instead of leaving out a
+> whole module if one of its definition.
 > 
-> p.
-18: "The remaining 174 files contained at least one of the unimplemented TypeScript features so that they could not be > considered." - This sentence sounds like you indeed generate as much as possible and filter out only layer --- how does this > statement relate my previous concern?
+> p. 18: "The remaining 174 files contained at least one of the
+> unimplemented TypeScript features so that they could not be
+> considered." - This sentence sounds like you indeed generate as much
+> as possible and filter out only layer --- how does this
+> statement relate my previous concern?
 > 
@@
 > 
-> p.
-5: "The choice of the template depends on the structure of the underlying JavaScript library:..." - here the following > listing/itemize environment seems to be missing, at least, I think that a itemize would be the better layout to use here.
-> 
-> p.
-14: "dts-compare does not contemplate differences in function return types, because interface inference of function > return types is not covered in the current version of dts-generate." - Why is it that the inference for return types is that > different from parameters?
-You infer the parameters through example code, by running example code, you get concrete return > values that you can analyse in the same fashion.
-Furthermore, Figure 4 (for example) shows a generated file with functions > with return types.
-What am I missing here?
+> p. 5: "The choice of the template depends on the structure of the
+> underlying JavaScript library:..." - here the following
+> listing/itemize environment seems to be missing, at least, I think that a itemize would be the better layout to use here.
+
+DONE.
+
+> p. 14: "dts-compare does not contemplate differences in function
+> return types, because interface inference of function
+> return types is not covered in the current version of dts-generate."
+> - Why is it that the inference for return types is that
+> different from parameters? You infer the parameters through example
+> code, by running example code, you get concrete return
+> values that you can analyse in the same fashion. Furthermore, Figure
+> 4 (for example) shows a generated file with functions
+> with return types. What am I missing here?
+
+Some explanation provided.
+
 > ## Typos
@@
 > all over the paper: runtime vs run time vs run-time!
+OK...
 > 
@@
 > p. 2: "..., which is time consuming and error prone." - error-prone
-> p. 8: "A TypeScript Declaration File" - declaration file (why title-case?)
+FIXED^2
+> p. 8: "A TypeScript Declaration File" - declaration file (why
+> title-case?)
+FIXED
 > p. 8: "... so that test cases were difficult to reap from the NPM packages:..." - read
+NOT FIXED
 > p. 9: "The Runtime Information block in Figure 3" - run-time information (why title-case?)
-> p. 9: "For example, Function f ... Property foo ... Parameter a ... - function ... property ... parameter (it seems weird to start > with upper-case letters for the itemization)
+FIXED
+> p. 9: "For example, Function f ... Property foo ... Parameter a
+> ... - function ... property ... parameter (it seems weird to start
+> with upper-case letters for the itemization)
 > p. 11: " and traverse the Abstract Syntax Tree of a TypeScript program" - abstract syntax tree
+OK
 > p. 12: "DefinitelyTyped Repository" - repository (why title-case?)
-> p. 12: "uses the TypeScript Compiler API" - TypeScript compiler API (why title-case?)
+OK
+> p. 12: "uses the TypeScript Compiler API" - TypeScript compiler API
+> (why title-case?)
+OK
 > p. 18: "Excerpt of JS implementation that throws an Error when in- put parameter is not a string" - error
-> p. 19: "Properties of interface Result were ignored for readability" - result
+OK
+> p. 19: "Properties of interface Result were ignored for readability"
+> - result
+NOT FIXED
 > p. 22: "future incorporation of a Symbolic Execution Engine that refines the" - symbolic execution engine (why title-case?)
+OK
 > 
@@
 > p. 9: "...meaningful examples, thus avoiding possible cold start problems." - examples, thus, avoiding
-> 
+ALL FIXED
 > 
@@
 > ----- TEXT:
-> The paper presents a tool that automatically generates TypeScript definition files for JavaScript projects.
-The tool uses > dynamic analysis.
-It runs the code samples found in the README file in the project repository, records information about > method invocations and argument values and uses those to generate type definitions.
-The paper evaluates the tool by > comparing generated definitions with hand-written definitions from the DefinitelyTyped repository - this is done for libraries > which have code samples in README, i.e.
-about ~270 out of ~6000.
+> The paper presents a tool that automatically generates TypeScript
+> definition files for JavaScript projects. The tool uses > dynamic
+> analysis. It runs the code samples found in the README file in the
+> project repository, records information about > method invocations
+> and argument values and uses those to generate type definitions. The
+> paper evaluates the tool by > comparing generated definitions with
+> hand-written definitions from the DefinitelyTyped repository - this
+> is done for libraries > which have code samples in README,
+> i.e. about ~270 out of ~6000. 
 > 
-> The paper presents a tool that is useful, although it has a fairly limited scope.
-The approach presented by the authors - to > run a code sample using a library and use runtime monitoring to deduce the structure of the library - is definitely a promising > one.
-The authors make a good first step towards building a practical tool.
-However, the current implementation of the tool is > quite basic in a number of ways.
-First, it relies on code samples in README, which is not exactly common (it will work only > for quite basic libraries).
-Second, it ignores many interesting types that a library could use - most prominently arrays, but > those are just one example.
+> The paper presents a tool that is useful, although it has a fairly
+> limited scope. The approach presented by the authors - to > run a
+> code sample using a library and use runtime monitoring to deduce the
+> structure of the library - is definitely a promising > one. The
+> authors make a good first step towards building a practical
+> tool. However, the current implementation of the tool is > quite
+> basic in a number of ways. First, it relies on code samples in
+> README, which is not exactly common (it will work only > for quite
+> basic libraries). Second, it ignores many interesting types that a
+> library could use - most prominently arrays, but > those are just
+> one example. 
 > 
@@
 I find this a bit paradoxical - maybe you could instead say that this is a lot > of work and there are many libraries that are not covered by DefinitelyTyped?
+
+ADDED a SENTENCE
+
 > 
@@
 I think > this would benefit from some clarification in the paper.
-> 
+
+EXPLAINED in section 3
+
 > * There is very little detail in the paper about how you infer types from the sample values that are observed during the > runtime execution.
@@
 (You might be interested in the > related paper [1] which does this for more complex types, but also covers things like records with optional fields.)
-> 
+
+DONE in section 4
+
 > * I think your evaluation is somewhat limited by the fact that the libraries that have a complete documentation in the > README are the ones with very simple API.
@@
 > * I think the phrasing that "more advanced types are overblown" is a bit unfortunate, especially if the tool does not support > some quite basic types like arrays.
-> 
+
+TONED DOWN. We have arrays now.
+
 > * The paper should give a much more extensive overview of related work - for example, it should discuss what has been > done using static analyses, whether similar tools exist for other languages and there might also be other interesting related > approaches (perhaps some based on machine learning methods).

### 2021-05-26 mplr-2021/main.tex
tightening up

@@
 \input{macros}
-\newcommand{\secref}[1]{Section~\ref{#1} - \nameref{#1}}
+\newcommand{\secref}[1]{Section~\ref{#1}}
 \newcommand{\figref}[1]{Figure~\ref{#1}}
@@
 
-  \caption{Declaration file for \texttt{glob-to-regexp} 
-    generated with \texttt{dts-generate}} 
+  \caption{Generated declaration file for \texttt{glob-to-regexp}} 
   \label{fig:motivating-example-glob-to-regexp-vscode}
@@
   \caption[dts-generate - Architecture overview]{dts-generate - Architecture
-      overview - The initial code base is retrieved from the \NPM{} package's
-    repository.
-    Run-time information is gathered driven by example code.
-    A TypeScript declaration file is generated using run-time
-    information.
+    overview
+    % - The initial code base is retrieved from the \NPM{} package's
+    % repository.
+    % Run-time information is gathered driven by example code.
+    % A TypeScript declaration file is generated using run-time
+    % information.
     % A Symbolic Execution Engine creates test cases based on the generated
@@
       {\includegraphics[width=1\linewidth]{evaluation-diagram.pdf}}
-      \caption[Evaluation against DefinitelyTyped repository]{\textbf{Evaluation of
-          generated declaration files against DefinitelyTyped repository} - The parser
-        uses the TypeScript Compiler           API \cite{typescript-compiler-api} to
-        transform declaration files into an abstract syntax which is serialized in
-        JSON. The comparison is performed on the JSON format.} 
+      \caption[Evaluation against DefinitelyTyped repository]{{Evaluation of
+          generated declaration files
+          % against DefinitelyTyped          repository
+        }
+        % - The parser
+        % uses the TypeScript Compiler           API \cite{typescript-compiler-api} to
+        % transform declaration files into an abstract syntax which is serialized in
+        % JSON. The comparison is performed on the JSON format.
+} 
       \label{fig:evaluation-diagram}
@@
 \begin{lstlisting}[
-  caption={Example for \texttt{dts-compare} executed on declaration files without differences},
+  caption={Comparing declaration files without differences},
   language=bash,
@@
   \caption[Number of analyzed modules for each stage of the experiment]{
-    \textbf{Number of analyzed modules for each stage of the experiment} - 
-  A TypeScript declaration file was generated for only \CountModulesGeneratedDeclarationFile{} modules, out of \CountTotalModulesDefinitelyTyped{} modules in the DefinitelyTyped repository. 
+    {Number of analyzed modules at each stage}
   % It was only possible to gather valid run-time information for about 25\% of the modules for which a Code Example was extracted.
@@
 
-The results obtained for each step are described in the
-following sections.
+Next, we describe the results obtained for each step.
 
@@
 A declaration file was generated for \CountModulesGeneratedDeclarationFile{} out of \CountModulesRunTimeInfoExtracted{} modules. Despite the correct execution
-of the instrumented code examples, \CountModulesDeclarationFileCouldNotBeGenerated{} modules did not yield suitable run-time information
-for generating a declaration file. The extracted code examples for these modules did not
-execute the library. Hence, the collected run-time information was not sufficient
+of the instrumented code examples, the extracted code for
+\CountModulesDeclarationFileCouldNotBeGenerated{} modules did 
+not exercise the library sufficiently. Hence, the collected run-time information was not enough
 for generating a declaration file.
@@
 package\footnote{\url{https://www.npmjs.com/package/@types/carlo}}
-uses literals to distinguish different 
-types of events.
+uses literals to distinguish different events.
 \begin{lstlisting}[language=TypeScript]
@@
 types of arguments and return values. In addition, we also trace
-property accesses which translate to requirements on objects. Also, we
+property accesses which translate to requirements on objects. We
 have no generalization step to infer polymorphic types.
@@
 Ruby programs at run time in the presence of
-metaprogramming. Metaprograms are extended to additionally generate
+metaprogramming. Metaprograms additionally generate
 type annotations for the generated code. Once a generated method is
@@
 
-JSTrace \cite{saftoiu21:_jstrac} also performs dynamic type inference
-for JavaScript by gathering information at run time. It inspects the
+JSTrace \cite{saftoiu21:_jstrac} performs dynamic type inference
+for JavaScript by gathering information at run time. It inspects
 run-time types of arguments and return values of suitably wrapped
 functions. While the resulting information is treated similarly as in
-our work, we also gather interactions that result from property
+our work, we gather interactions from property
 accesses and method calls and derive further typing constraints from
@@
 
-As TSInfer analyzes the objects and functions stored in the snapshot, it faces the problem
-of including internal methods and private properties in the declaration file. Run-time
-information would have informed that the developer has no intention of exposing these
+TSInfer faces the problem
+of including internal methods and private properties from the snapshot in the declaration file. Run-time
+information would have shown that the developer has no intention of exposing these
 methods.
@@
 We have presented \texttt{dts-generate}, a tool for generating a TypeScript declaration
-file for a specific JavaScript library. The tool downloads code examples written by the
-developers from the library's repository. It uses these examples to execute the library and
-gather data flow and type information. The tool finally generates a TypeScript declaration
+file for a specific JavaScript library. It downloads the developer's
+code examples from the library's repository. It uses these examples to
+execute the library and gather data flow and type information. The tool generates a TypeScript declaration
 file based on the information gathered at run time.
 
-We developed an architecture that supports the automatic generation of declaration files
-for specific JavaScript libraries without additional manual tasks. The architecture
-contemplates future incorporation of a symbolic execution engine that refines the
-initial code base enabling the exploration of new execution paths. Although it is not
-implemented 
-in this work, its incorporation would result in small incremental modifications to the
-presented architecture as it is considered to only expand the existing code base.
+% We developed an architecture that supports the automatic generation of declaration files
+% for specific JavaScript libraries without additional manual tasks. The architecture
+% contemplates future incorporation of a symbolic execution engine that refines the
+% initial code base enabling the exploration of new execution paths. Although it is not
+% implemented 
+% in this work, its incorporation would result in small incremental modifications to the
+% presented architecture as it is considered to only expand the existing code base.
 
 Building an end-to-end solution for the generation of TypeScript declaration files was
-prioritized over type inference accuracy. Consequently, types were taken over from the
+prioritized over type inference accuracy. Hence, types were taken over from the
 values at run time. Developers express their intent how a library should be used with
-example code in the documentation. Hence,
-obtaining the types from the code examples extracted from the repositories proved to be a
+example code in the documentation. 
+Obtaining the types from the code examples extracted from the repositories proved to be a
 pragmatic and effective approximation, enabling to work on specific aspects regarding the

### 2021-06-01 mplr-2021/main.tex
Move sections to appendix

@@
 % \end{figure}
-
-\begin{figure*}[tp]
-  \centering
-  \begin{subfigure}{0.48\linewidth}
-    \begin{lstlisting}[language=TypeScript,numbers=none]
-export function v1(str: string): boolean;
-export function v2(str: string): boolean;
-export function v3(str: string): boolean;
-export function v4(str: string): boolean;
-export function v5(str: string): boolean;
-    \end{lstlisting}
-    \caption{is-uuid/index.d.ts - Generated}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{0.48\linewidth}
-    \begin{lstlisting}[language=TypeScript,numbers=none]
-export function v1(value: string): boolean;
-export function v2(value: string): boolean;
-export function v3(value: string): boolean;
-export function v4(value: string): boolean;
-export function v5(value: string): boolean;
-export function nil(value: string): boolean;
-export function anyNonNil(value: string): boolean;
-    \end{lstlisting}
-    \caption{is-uuid/index.d.ts - DefinitelyTyped}
-  \end{subfigure}
-
-  \caption{Results for \textbf{module} \texttt{is-uuid}}
-  \label{fig:experiments-results-module-is-uuid}
-\end{figure*}
 
@@
 
-This section exhibits some examples of the \CountModulesGeneratedDeclarationFile{} generated
-declaration files. It shows some results for each of the implemented
-templates: \textbf{module}, \textbf{module-function}, and
-\textbf{module-class}.
-
-\figref{fig:experiments-results-module-function} shows the generated
-declaration files for simple modules like \texttt{abs},
-\texttt{dirname-regex}, and \texttt{escape-html}. All of them were
-generated using the \textbf{module-function} template. The left side of the figure shows
-the generated declaration file with 
+This section exhibits an example of the \CountModulesGeneratedDeclarationFile{} generated
+declaration files. \figref{fig:experiments-results-module-steamid} exposes a simplified example for the
+\textbf{module-class} template where we highlight the most important features.
+We expose the results for the remaining templates \textbf{module} and \textbf{module-function} in
+\secref{sec:appendix-results-templates}.
+The left side of the figure shows
+the generated declaration file with x
 \lstinline{dts-generate}, the right side shows the corresponding file
 in the DefinitelyTyped repository.
-There are no relevant differences between the files. 
-The different names in the export assignment by modules \texttt{dirname-regex} and
-\texttt{escape-html} do not matter because the module can be named arbitrarily when it is imported: in
-\lstinline[language=TypeScript]/import Abs = require('abs');/ the
-identifier \lstinline[language=TypeScript]/Abs/ can be chosen by the
-programmer. We generate the name in the export 
-assignment by transforming the module name into camel case form, following TypeScript
-guidelines.
-
-Functions are correctly detected; input and output types are
-accurately inferred. 
-We extract the parameter names from the variable names of the JavaScript
-code.
-A difference in the names of the function parameters does not affect the correctness of
-the declaration file.
-Furthermore, neither the order of declarations nor unused namespaces
-(as in module \texttt{escape-html}) matter.
-
-\begin{figure*}[tp]
-    \centering
-    \begin{subfigure}[t]{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript,numbers=none]
-export = Abs;
-
-declare function Abs(input: string): string;
-      \end{lstlisting}
-      \caption{abs/index.d.ts - Generated}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript,numbers=none]
-declare function Abs(input: string): string;
-export = Abs;
-      \end{lstlisting}
-      \caption{abs/index.d.ts - DT}
-    \end{subfigure}
-
-    \begin{subfigure}[t]{0.48\linewidth}
-        \begin{lstlisting}[language=TypeScript,numbers=none]
-export = DirnameRegex;
-
-declare function DirnameRegex(): RegExp;
-        \end{lstlisting}
-        \caption{dirname-regex/index.d.ts - Generated}
-      \end{subfigure}
-      \hfill
-      \begin{subfigure}[t]{0.48\linewidth}
-        \begin{lstlisting}[language=TypeScript,numbers=none]
-export = dirnameRegex;
-
-declare function dirnameRegex(): RegExp;
-        \end{lstlisting}
-        \caption{dirname-regex/index.d.ts - DT}
-      \end{subfigure}
-
-      \begin{subfigure}[t]{0.48\linewidth}
-        \begin{lstlisting}[language=TypeScript,numbers=none]
-export = EscapeHtml;
-
-declare function EscapeHtml(string: string): string;
-
-        \end{lstlisting}
-        \caption{escape-html/index.d.ts - Generated}
-      \end{subfigure}
-      \hfill
-      \begin{subfigure}[t]{0.48\linewidth}
-        \begin{lstlisting}[language=TypeScript,numbers=none]
-declare function escapeHTML(text: string): string;
-declare namespace escapeHTML { }
-
-export = escapeHTML;
-        \end{lstlisting}
-        \caption{escape-html/index.d.ts - DT}
-      \end{subfigure}
-
-    \caption{Results for \textbf{module-function} template}
-    \label{fig:experiments-results-module-function}
-\end{figure*}
 
@@
 
-\figref{fig:experiments-results-module-is-uuid} contains an example
-for the \textbf{module} template, the declaration for the \texttt{is-uuid} module.
-The generated file only contain methods that were executed by the extracted
-examples. All invoked functions are correctly detected; the additional
-functions declared on DefinitelyTyped are not used in the example
-code.
-It would be easy to generate a perfectly matching declaration file by
-adding two examples using functions \texttt{nil} and
-\texttt{anyNonNil}.
-
-\figref{fig:experiments-results-module-steamid} exposes a simplified example for the
-\textbf{module-class} template where we highlight the most important features. Class methods are
-correctly generated. For the \texttt{getSteam2RenderedID} method we correctly generate a signature
-that matches the type alias chosen by the author of the library. Finally, the method \texttt{fromIndividualAccountID}
-is correctly located in the namespace section.
+There are no relevant differences between the files. 
+Class methods and functions are correctly detected; input and output types are
+accurately inferred. We extract the parameter names from the variable names of the JavaScript
+code. For the \texttt{getSteam2RenderedID} method we correctly generate a signature
+that matches the type alias chosen by the author of the library. Finally, the method \\
+\texttt{fromIndividualAccountID} is correctly located in the namespace section.
 
 It is worth mentioning that for some libraries the declaration file in DefinitelyTyped was
-not correct. For example, for the module \texttt{glob-base}, the parameter
-\texttt{basePath} is declared as optional in DefinitelyTyped. However, when invoking the
-function without the \texttt{basePath} parameter, an error is thrown at run time. The file generated by
-\texttt{dts-generate} does not mark this parameter as optional, as shown in
-\figref{fig:experiments-results-module-glob-base}. Function return type interfaces are
-not inferred by \texttt{dts-generate}.
-% The parameter name is \texttt{pattern}
-% instead of \texttt{basePath}, because the JavaScript implementation declares the parameter
-% as \texttt{pattern}, as shown in \figref{fig:subfloat-globbase-js-implementation}. 
-We also
-discovered errors in the selected template in DefinitelyTyped. The module
-\texttt{smart-truncate} in DefinitelyTyped uses the \textbf{module} template, but
-\texttt{dts-generate} generates a file using the \textbf{module-function}
-template. \coderef{code:experiments-results-module-smart-truncate} shows that the module is indeed
-exported as a function, as inferred by \texttt{dts-generate}.
-
-\begin{figure*}[tp]
-  \centering
-  \begin{subfigure}[t]{0.48\linewidth}
-    \begin{lstlisting}[language=TypeScript,numbers=none]
-'use strict';
-
-// ...
-
-module.exports = function globBase(pattern) {
-if (typeof pattern !== 'string') {
-  throw new TypeError('glob-base expects a string.');
-}
-
-// ...
-} 
-    \end{lstlisting}
-    \caption{glob-base/index.js - Excerpt of JS implementation
-      that throws an error when input parameter is not a string} 
-    \label{fig:subfloat-globbase-js-implementation}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.48\linewidth}
-    \begin{lstlisting}[language=TypeScript,numbers=none]
-declare function globbase(basePath?: string): globbase.GlobBaseResult;
-
-declare namespace globbase {
-  interface GlobBaseResult {
-      base: string;
-      isGlob: boolean;
-      glob: string;
-  }
-}
-
-export = globbase;
-    \end{lstlisting}
-    \caption{glob-base/index.d.ts - DefinitelyTyped}
-  \end{subfigure}
-
-\begin{subfigure}[t]{0.48\linewidth}
-    \begin{lstlisting}[language=JavaScript,numbers=none]
-var globBase = require('glob-base');
-
-globBase();
-
-// Throws: TypeError: glob-base expects a string.
-    \end{lstlisting}
-    \caption{Code example invoking the exported function without the first parameter, as specified in DefinitelyTyped}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.48\linewidth}
-    \begin{lstlisting}[language=TypeScript,numbers=none]
-export = GlobBase;
-
-declare function GlobBase(pattern: string): object;
-    \end{lstlisting}
-    \caption{glob-base/index.d.ts - Generated}
-  \end{subfigure}
-
-  \caption{Incorrect optional parameter for module \texttt{glob-base}}
-  \label{fig:experiments-results-module-glob-base}
-\end{figure*}
-
-\begin{lstlisting}[
-caption={Code example for \texttt{smart-truncate} module exporting a function},
-label=code:experiments-results-module-smart-truncate,
-language=JavaScript,captionpos=b,float=tp,numbers=none
-]
-var smartTruncate = require('smart-truncate');
-
-var string = 'To iterate is human, to recurse divine.';
-
-// Append an ellipsis at the end of the truncated string.
-var truncated = smartTruncate(string, 15);
-\end{lstlisting}
+not correct. For module \texttt{glob-base} we detected that a parameter was
+incorrectly marked as optional. For module \texttt{smart-truncate} we discovered that the author
+incorrectly used the \textbf{module} template.
+We provide a more detailed explanation in \secref{sec:errors-definitely-typed}.
 
@@
 
-As shown in \figref{fig:experiments-results-module-is-uuid}, the code examples have a
-direct influence on the quality of the generated declaration file. We introduced the
+The code examples have a direct influence on the quality of the generated declaration file. We introduced the
 concept of \texttt{solvable} and \texttt{unsolvable} differences in
 \secref{sec:dts-compare}. Modules containing only \texttt{solvable} differences were
-considered as positive results. All of the analyzed files contained only \texttt{solvable} differences. We manually completed the
+considered as positive results. All of the analyzed files contained only \texttt{solvable} differences.
+We manually completed the
 examples for 10 of those modules. Applying \texttt{dts-generate} with the completed
-examples generates declaration files, which are equivalent to the files from DefinitelyTyped, as
-shown in \figref{fig:experiments-results-manually-completed-examples}.
-
-\begin{figure*}[t]
-  \centering
-  \begin{subfigure}[t]{0.45\linewidth}
-    \begin{lstlisting}[language=TypeScript]
-declare namespace gh {
-  interface Options {
-      enterprise?: boolean;
-  }
-  interface Result {
-      ...
-  }
-}
-
-declare function gh(url: string | {url: string}, options?: gh.Options): gh.Result | null;
-
-export = gh;    
-    \end{lstlisting}
-    \caption{github-url-to-object/index.d.ts - DefinitelyTyped. Properties of interface \texttt{Result} were ignored for readability, since return types were not considered.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.45\linewidth}
-    \begin{lstlisting}[language=TypeScript]
-export = GithubUrlToObject;
-
-declare function GithubUrlToObject(repoUrl: string | GithubUrlToObject.I__repoUrl, opts?: GithubUrlToObject.I__opts): object | null;
-declare namespace GithubUrlToObject {
-  export interface I__repoUrl {
-    'url'?: string;
-  }
-
-  export interface I__opts {
-    'enterprise'?: boolean;
-  }
-
-}
-    \end{lstlisting}
-    \caption{github-url-to-object/index.d.ts - Generated with manually expanded code example}
-  \end{subfigure}
-
-  \begin{subfigure}{0.80\linewidth}
-    \begin{lstlisting}[language=JavaScript]
-var gh = require('github-url-to-object')
-
-gh('github:monkey/business');
-gh('https://github.com/monkey/business');
-gh('https://github.com/monkey/business/tree/master');
-gh('https://github.com/monkey/business/tree/master/nested/file.js');
-gh('https://github.com/monkey/business.git');
-gh('http://github.com/monkey/business');
-gh('git://github.com/monkey/business.git');
-
-// Manually added:
-gh('git+https://githuuub.com/monkey/business.git', {});
-gh('git+https://githuuub.com/monkey/business.git', {enterprise: true});
-gh({url: 'git://github.com/monkey/business.git'});
-gh('this is not a proper url');
-gh({url: 'this is not a proper url'});
-    \end{lstlisting}
-    \caption{Code example for module \texttt{github-url-to-object}.}
-    \end{subfigure}
-  \caption{Manually expanded code example for
-    \texttt{github-url-to-object} to match DefinitelyTyped declaration
-    file. The literal interface in DefinitelyTyped is replaced by a
-    declared interface. Return types are not considered.} 
-  \label{fig:experiments-results-manually-completed-examples}
-\end{figure*}
+examples generates declaration files which are equivalent to the files from DefinitelyTyped.
+\secref{sec:appendix-manually-completed-examples} exhibits the manually added code examples
+for module \texttt{github-url-to-object}.
 
@@
 
+\appendix
+\section{Results}
+\subsection{Templates}
+\label{sec:appendix-results-templates}
+\subsubsection{module-function}
+\figref{fig:experiments-results-module-function} shows the generated
+declaration files for simple modules like \texttt{abs},
+\texttt{dirname-regex}, and \texttt{escape-html}. All of them were
+generated using the \textbf{module-function} template.
+
+The different names in the export assignment by modules \\
+\texttt{dirname-regex} and
+\texttt{escape-html} do not matter because the module can be named arbitrarily when it is imported: in
+\lstinline[language=TypeScript]/import Abs = require('abs');/ the
+identifier \lstinline[language=TypeScript]/Abs/ can be chosen by the
+programmer. We generate the name in the export 
+assignment by transforming the module name into camel case form, following TypeScript
+guidelines.
+
+A difference in the names of the function parameters does not affect the correctness of
+the declaration file.
+Furthermore, neither the order of declarations nor unused namespaces
+(as in module \texttt{escape-html}) matter.
+
+\begin{figure*}[tp]
+  \centering
+  \begin{subfigure}[t]{0.48\linewidth}
+    \begin{lstlisting}[language=TypeScript,numbers=none]
+export = Abs;
+
+declare function Abs(input: string): string;
+    \end{lstlisting}
+    \caption{abs/index.d.ts - Generated}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+    \begin{lstlisting}[language=TypeScript,numbers=none]
+declare function Abs(input: string): string;
+export = Abs;
+    \end{lstlisting}
+    \caption{abs/index.d.ts - DT}
+  \end{subfigure}
+
+  \begin{subfigure}[t]{0.48\linewidth}
+      \begin{lstlisting}[language=TypeScript,numbers=none]
+export = DirnameRegex;
+
+declare function DirnameRegex(): RegExp;
+      \end{lstlisting}
+      \caption{dirname-regex/index.d.ts - Generated}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.48\linewidth}
+      \begin{lstlisting}[language=TypeScript,numbers=none]
+export = dirnameRegex;
+
+declare function dirnameRegex(): RegExp;
+      \end{lstlisting}
+      \caption{dirname-regex/index.d.ts - DT}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{0.48\linewidth}
+      \begin{lstlisting}[language=TypeScript,numbers=none]
+export = EscapeHtml;
+
+declare function EscapeHtml(string: string): string;
+
+      \end{lstlisting}
+      \caption{escape-html/index.d.ts - Generated}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.48\linewidth}
+      \begin{lstlisting}[language=TypeScript,numbers=none]
+declare function escapeHTML(text: string): string;
+declare namespace escapeHTML { }
+
+export = escapeHTML;
+      \end{lstlisting}
+      \caption{escape-html/index.d.ts - DT}
+    \end{subfigure}
+
+  \caption{Results for \textbf{module-function} template}
+  \label{fig:experiments-results-module-function}
+\end{figure*}
+
+\subsubsection{module}
+\figref{fig:experiments-results-module-is-uuid} contains an example
+for the \textbf{module} template, the declaration for the \texttt{is-uuid} module.
+The generated file only contain methods that were executed by the extracted
+examples. All invoked functions are correctly detected; the additional
+functions declared on DefinitelyTyped are not used in the example
+code.
+It would be easy to generate a perfectly matching declaration file by
+adding two examples using functions \texttt{nil} and
+\texttt{anyNonNil}.
+
+\begin{figure*}[tp]
+  \centering
+  \begin{subfigure}{0.48\linewidth}
+    \begin{lstlisting}[language=TypeScript,numbers=none]
+export function v1(str: string): boolean;
+export function v2(str: string): boolean;
+export function v3(str: string): boolean;
+export function v4(str: string): boolean;
+export function v5(str: string): boolean;
+    \end{lstlisting}
+    \caption{is-uuid/index.d.ts - Generated}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{0.48\linewidth}
+    \begin{lstlisting}[language=TypeScript,numbers=none]
+export function v1(value: string): boolean;
+export function v2(value: string): boolean;
+export function v3(value: string): boolean;
+export function v4(value: string): boolean;
+export function v5(value: string): boolean;
+export function nil(value: string): boolean;
+export function anyNonNil(value: string): boolean;
+    \end{lstlisting}
+    \caption{is-uuid/index.d.ts - DefinitelyTyped}
+  \end{subfigure}
+
+  \caption{Results for \textbf{module} \texttt{is-uuid}}
+  \label{fig:experiments-results-module-is-uuid}
+\end{figure*}
+
+\subsection{Manually completed examples}
+\label{sec:appendix-manually-completed-examples}
+\figref{fig:experiments-results-manually-completed-examples} exhibits the
+ manually expanded code examples for module \texttt{github-url-to-object}.
+
+\begin{figure*}[t]
+  \centering
+  \begin{subfigure}[t]{0.45\linewidth}
+    \begin{lstlisting}[language=TypeScript]
+declare namespace gh {
+  interface Options {
+      enterprise?: boolean;
+  }
+  interface Result {
+      ...
+  }
+}
+
+declare function gh(url: string | {url: string}, options?: gh.Options): gh.Result | null;
+
+export = gh;    
+    \end{lstlisting}
+    \caption{github-url-to-object/index.d.ts - DefinitelyTyped. Properties of interface \texttt{Result} were ignored for readability, since return types were not considered.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.45\linewidth}
+    \begin{lstlisting}[language=TypeScript]
+export = GithubUrlToObject;
+
+declare function GithubUrlToObject(repoUrl: string | GithubUrlToObject.I__repoUrl, opts?: GithubUrlToObject.I__opts): object | null;
+declare namespace GithubUrlToObject {
+  export interface I__repoUrl {
+    'url'?: string;
+  }
+
+  export interface I__opts {
+    'enterprise'?: boolean;
+  }
+
+}
+    \end{lstlisting}
+    \caption{github-url-to-object/index.d.ts - Generated with manually expanded code example}
+  \end{subfigure}
+
+  \begin{subfigure}{0.80\linewidth}
+    \begin{lstlisting}[language=JavaScript]
+var gh = require('github-url-to-object')
+
+gh('github:monkey/business');
+gh('https://github.com/monkey/business');
+gh('https://github.com/monkey/business/tree/master');
+gh('https://github.com/monkey/business/tree/master/nested/file.js');
+gh('https://github.com/monkey/business.git');
+gh('http://github.com/monkey/business');
+gh('git://github.com/monkey/business.git');
+
+// Manually added:
+gh('git+https://githuuub.com/monkey/business.git', {});
+gh('git+https://githuuub.com/monkey/business.git', {enterprise: true});
+gh({url: 'git://github.com/monkey/business.git'});
+gh('this is not a proper url');
+gh({url: 'this is not a proper url'});
+    \end{lstlisting}
+    \caption{Code example for module \texttt{github-url-to-object}.}
+    \end{subfigure}
+  \caption{Manually expanded code example for
+    \texttt{github-url-to-object} to match DefinitelyTyped declaration
+    file. The literal interface in DefinitelyTyped is replaced by a
+    declared interface. Return types are not considered.} 
+  \label{fig:experiments-results-manually-completed-examples}
+\end{figure*}
+
+\subsection{Errors in DefinitelyTyped}
+\label{sec:errors-definitely-typed}
+For module \texttt{glob-base} the parameter
+\texttt{basePath} is declared as optional in DefinitelyTyped. However, when invoking the
+function without the \texttt{basePath} parameter, an error is thrown at run time. The file generated by
+\texttt{dts-generate} does not mark this parameter as optional, as shown in
+\figref{fig:experiments-results-module-glob-base}. Function return type interfaces are
+not inferred by \texttt{dts-generate}.
+
+% The parameter name is \texttt{pattern}
+% instead of \texttt{basePath}, because the JavaScript implementation declares the parameter
+% as \texttt{pattern}, as shown in \figref{fig:subfloat-globbase-js-implementation}. 
+We also
+discovered errors in the selected template in DefinitelyTyped. The module
+\texttt{smart-truncate} in DefinitelyTyped uses the \textbf{module} template, but
+\texttt{dts-generate} generates a file using the \textbf{module-function}
+template. \coderef{code:experiments-results-module-smart-truncate} shows that the module is indeed
+exported as a function, as inferred by \texttt{dts-generate}.
+
+\begin{lstlisting}[
+  caption={Code example for \texttt{smart-truncate} module exporting a function},
+  label=code:experiments-results-module-smart-truncate,
+  language=JavaScript,captionpos=b,numbers=none
+  ]
+  var smartTruncate = require('smart-truncate');
+
+  var string = 'To iterate is human, to recurse divine.';
+
+  // Append an ellipsis at the end of the truncated string.
+  var truncated = smartTruncate(string, 15);
+\end{lstlisting}
+
+\begin{figure*}[tp]
+  \centering
+  \begin{subfigure}[t]{0.48\linewidth}
+    \begin{lstlisting}[language=TypeScript,numbers=none]
+'use strict';
+
+// ...
+
+module.exports = function globBase(pattern) {
+if (typeof pattern !== 'string') {
+  throw new TypeError('glob-base expects a string.');
+}
+
+// ...
+} 
+    \end{lstlisting}
+    \caption{glob-base/index.js - Excerpt of JS implementation
+      that throws an error when input parameter is not a string} 
+    \label{fig:subfloat-globbase-js-implementation}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+    \begin{lstlisting}[language=TypeScript,numbers=none]
+declare function globbase(basePath?: string): globbase.GlobBaseResult;
+
+declare namespace globbase {
+  interface GlobBaseResult {
+      base: string;
+      isGlob: boolean;
+      glob: string;
+  }
+}
+
+export = globbase;
+    \end{lstlisting}
+    \caption{glob-base/index.d.ts - DefinitelyTyped}
+  \end{subfigure}
+
+\begin{subfigure}[t]{0.48\linewidth}
+    \begin{lstlisting}[language=JavaScript,numbers=none]
+var globBase = require('glob-base');
+
+globBase();
+
+// Throws: TypeError: glob-base expects a string.
+    \end{lstlisting}
+    \caption{Code example invoking the exported function without the first parameter, as specified in DefinitelyTyped}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\linewidth}
+    \begin{lstlisting}[language=TypeScript,numbers=none]
+export = GlobBase;
+
+declare function GlobBase(pattern: string): object;
+    \end{lstlisting}
+    \caption{glob-base/index.d.ts - Generated}
+  \end{subfigure}
+
+  \caption{Incorrect optional parameter for module \texttt{glob-base}}
+  \label{fig:experiments-results-module-glob-base}
+\end{figure*}
+
 \end{document}

### 2021-06-02 mplr-2021/main.tex
Add new page for appendix

@@
 
+\clearpage
 \appendix
@@
 \label{sec:errors-definitely-typed}
+
+\begin{lstlisting}[
+  caption={Code example for \texttt{smart-truncate} module exporting a function},
+  label=code:experiments-results-module-smart-truncate,
+  language=JavaScript,captionpos=b,float,numbers=none
+  ]
+  var smartTruncate = require('smart-truncate');
+
+  var string = 'To iterate is human, to recurse divine.';
+
+  // Append an ellipsis at the end of the truncated string.
+  var truncated = smartTruncate(string, 15);
+\end{lstlisting}
+
 For module \texttt{glob-base} the parameter
@@
 
-\begin{lstlisting}[
-  caption={Code example for \texttt{smart-truncate} module exporting a function},
-  label=code:experiments-results-module-smart-truncate,
-  language=JavaScript,captionpos=b,numbers=none
-  ]
-  var smartTruncate = require('smart-truncate');
-
-  var string = 'To iterate is human, to recurse divine.';
-
-  // Append an ellipsis at the end of the truncated string.
-  var truncated = smartTruncate(string, 15);
-\end{lstlisting}
-
 \begin{figure*}[tp]

### 2021-08-15 mplr-2021/main.tex
Apply changes for camera-ready

@@
 
+\documentclass[sigplan,screen]{acmart}
+
+%%
+%% \BibTeX command to typeset BibTeX logo in the docs
+\AtBeginDocument{%
+  \providecommand\BibTeX{{%
+    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
+
 \input{macros}
@@
 
-\documentclass[sigconf]{acmart}
 \usepackage{listings}
@@
 \usepackage[htt]{hyphenat}
+
+\usepackage{microtype}
 
@@
 \settopmatter{printfolios=true}
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
 
@@
 %%
-% \begin{CCSXML}
-% <ccs2012>
-%  <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%   <concept_significance>500</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%   <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-% \ccsdesc[500]{Computer systems organization~Embedded systems}
-% \ccsdesc[300]{Computer systems organization~Redundancy}
-% \ccsdesc{Computer systems organization~Robotics}
-% \ccsdesc[100]{Networks~Network reliability}
+\begin{CCSXML}
+<ccs2012>
+  <concept>
+    <concept_id>10011007.10011006.10011008</concept_id>
+    <concept_desc>Software and its engineering~General programming languages</concept_desc>
+    <concept_significance>300</concept_significance>
+    </concept>
+  <concept>
+    <concept_id>10011007.10011006.10011066</concept_id>
+    <concept_desc>Software and its engineering~Development frameworks and environments</concept_desc>
+    <concept_significance>100</concept_significance>
+  </concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[300]{Software and its engineering~General programming languages}
+\ccsdesc[100]{Software and its engineering~Development frameworks and environments}
 
@@
   We do not rely on static analysis, which is hard to implement
-  soundly and precisely and which is prone to maintenance problems
+  soundly and precisely. It is also prone to maintenance problems
   when keeping up with JavaScript's frequent language updates.
@@
 \begin{table}[tp]
+  \caption{Usage of TypeScript features on DefinitelyTyped}
   \begin{center}
@@
   \end{center}
-  \caption{Usage of TypeScript features on DefinitelyTyped}
   \label{tab:dts-parse-stats}
@@
 \begin{verbatim}
-globToRegExp : (glob: string, opts: undefined) -> RegExp
+globToRegExp : (glob: string,
+                opts: undefined) -> RegExp
+
 globToRegExp : (glob: string, 
@@
 JSTrace \cite{saftoiu21:_jstrac} performs dynamic type inference
-for JavaScript by gathering information at run time. It inspects
+for \\ JavaScript by gathering information at run time. It inspects
 run-time types of arguments and return values of suitably wrapped
@@
 
-%%
-%% The next two lines define the bibliography style to be used, and
-%% the bibliography file.
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{main}
-
 \clearpage
@@
 
+\clearpage
+
+%%
+%% The next two lines define the bibliography style to be used, and
+%% the bibliography file.
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{main}
+
 \end{document}

### 2021-08-16 mplr-2021/main.tex
Remove typo pointed out in the feedback
Add copyright and future work

@@
 
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
 \usepackage{microtype}
@@
 
+\copyrightyear{2021}
+\acmYear{2021}
+\setcopyright{acmlicensed}\acmConference[MPLR '21]{Proceedings of the 18th ACM SIGPLAN International Conference on Managed Programming Languages and Runtimes}{September 29--30, 2021}{Münster, Germany}
+\acmBooktitle{Proceedings of the 18th ACM SIGPLAN International Conference on Managed Programming Languages and Runtimes (MPLR '21), September 29--30, 2021, Münster, Germany}
+\acmPrice{15.00}
+\acmDOI{10.1145/3475738.3480941}
+\acmISBN{978-1-4503-8675-3/21/09}
+
 %% Rights management information.  This information is sent to you
@@
 \affiliation{%
-  \institution{Albert-Ludwigs-Universit{\"a}t Freiburg}
+  \institution{Albert-Ludwigs-Universität Freiburg}
   \city{Freiburg}
@@
 We aim to improve on this situation by providing a tool that generates
-TypeScript declaration files automatically. To do so, we rely on we rely on run-time information
+TypeScript declaration files automatically. To do so, we rely on run-time information
 gathered from the examples provided
@@
 The TypeScript project provides a guide on how to write high-quality declaration
-files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/by-example.html}. The guide
+files\footnote{\url{https://www.typescriptlang.org/docs/handbook/declaration-files/by-example.html}}. The guide
 explains the main concepts through examples. We selected one example for each template
@@
 
+Return types could be defined more accurately by analyzing the interactions tracked on
+them. We consider that enhancing the run-time information by executing tests already
+present in the codebase is an approach worth following as future work.
+
 % \begin{changethis}

### 2021-08-17 mplr-2021/main.tex
Remove appendices

@@
 \usepackage{microtype}
+
+\usepackage{balance}
 
@@
 incorrectly used the \textbf{module} template.
-We provide a more detailed explanation in \secref{sec:errors-definitely-typed}.
 
@@
 examples generates declaration files which are equivalent to the files from DefinitelyTyped.
-\secref{sec:appendix-manually-completed-examples} exhibits the manually added code examples
-for module \texttt{github-url-to-object}.
 
@@
 
-\clearpage
 \appendix
-\section{Results}
-\subsection{Templates}
-\label{sec:appendix-results-templates}
-\subsubsection{module-function}
-\figref{fig:experiments-results-module-function} shows the generated
-declaration files for simple modules like \texttt{abs},
-\texttt{dirname-regex}, and \texttt{escape-html}. All of them were
-generated using the \textbf{module-function} template.
-
-The different names in the export assignment by modules \\
-\texttt{dirname-regex} and
-\texttt{escape-html} do not matter because the module can be named arbitrarily when it is imported: in
-\lstinline[language=TypeScript]/import Abs = require('abs');/ the
-identifier \lstinline[language=TypeScript]/Abs/ can be chosen by the
-programmer. We generate the name in the export 
-assignment by transforming the module name into camel case form, following TypeScript
-guidelines.
-
-A difference in the names of the function parameters does not affect the correctness of
-the declaration file.
-Furthermore, neither the order of declarations nor unused namespaces
-(as in module \texttt{escape-html}) matter.
-
-\begin{figure*}[tp]
-  \centering
-  \begin{subfigure}[t]{0.48\linewidth}
-    \begin{lstlisting}[language=TypeScript,numbers=none]
-export = Abs;
-
-declare function Abs(input: string): string;
-    \end{lstlisting}
-    \caption{abs/index.d.ts - Generated}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.48\linewidth}
-    \begin{lstlisting}[language=TypeScript,numbers=none]
-declare function Abs(input: string): string;
-export = Abs;
-    \end{lstlisting}
-    \caption{abs/index.d.ts - DT}
-  \end{subfigure}
-
-  \begin{subfigure}[t]{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript,numbers=none]
-export = DirnameRegex;
-
-declare function DirnameRegex(): RegExp;
-      \end{lstlisting}
-      \caption{dirname-regex/index.d.ts - Generated}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript,numbers=none]
-export = dirnameRegex;
-
-declare function dirnameRegex(): RegExp;
-      \end{lstlisting}
-      \caption{dirname-regex/index.d.ts - DT}
-    \end{subfigure}
-
-    \begin{subfigure}[t]{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript,numbers=none]
-export = EscapeHtml;
-
-declare function EscapeHtml(string: string): string;
-
-      \end{lstlisting}
-      \caption{escape-html/index.d.ts - Generated}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.48\linewidth}
-      \begin{lstlisting}[language=TypeScript,numbers=none]
-declare function escapeHTML(text: string): string;
-declare namespace escapeHTML { }
-
-export = escapeHTML;
-      \end{lstlisting}
-      \caption{escape-html/index.d.ts - DT}
-    \end{subfigure}
-
-  \caption{Results for \textbf{module-function} template}
-  \label{fig:experiments-results-module-function}
-\end{figure*}
-
-\subsubsection{module}
-\figref{fig:experiments-results-module-is-uuid} contains an example
-for the \textbf{module} template, the declaration for the \texttt{is-uuid} module.
-The generated file only contain methods that were executed by the extracted
-examples. All invoked functions are correctly detected; the additional
-functions declared on DefinitelyTyped are not used in the example
-code.
-It would be easy to generate a perfectly matching declaration file by
-adding two examples using functions \texttt{nil} and
-\texttt{anyNonNil}.
 
@@
 
-\subsection{Manually completed examples}
-\label{sec:appendix-manually-completed-examples}
-\figref{fig:experiments-results-manually-completed-examples} exhibits the
- manually expanded code examples for module \texttt{github-url-to-object}.
-
-\begin{figure*}[t]
-  \centering
-  \begin{subfigure}[t]{0.45\linewidth}
-    \begin{lstlisting}[language=TypeScript]
-declare namespace gh {
-  interface Options {
-      enterprise?: boolean;
-  }
-  interface Result {
-      ...
-  }
-}
-
-declare function gh(url: string | {url: string}, options?: gh.Options): gh.Result | null;
-
-export = gh;    
-    \end{lstlisting}
-    \caption{github-url-to-object/index.d.ts - DefinitelyTyped. Properties of interface \texttt{Result} were ignored for readability, since return types were not considered.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.45\linewidth}
-    \begin{lstlisting}[language=TypeScript]
-export = GithubUrlToObject;
-
-declare function GithubUrlToObject(repoUrl: string | GithubUrlToObject.I__repoUrl, opts?: GithubUrlToObject.I__opts): object | null;
-declare namespace GithubUrlToObject {
-  export interface I__repoUrl {
-    'url'?: string;
-  }
-
-  export interface I__opts {
-    'enterprise'?: boolean;
-  }
-
-}
-    \end{lstlisting}
-    \caption{github-url-to-object/index.d.ts - Generated with manually expanded code example}
-  \end{subfigure}
-
-  \begin{subfigure}{0.80\linewidth}
-    \begin{lstlisting}[language=JavaScript]
-var gh = require('github-url-to-object')
-
-gh('github:monkey/business');
-gh('https://github.com/monkey/business');
-gh('https://github.com/monkey/business/tree/master');
-gh('https://github.com/monkey/business/tree/master/nested/file.js');
-gh('https://github.com/monkey/business.git');
-gh('http://github.com/monkey/business');
-gh('git://github.com/monkey/business.git');
-
-// Manually added:
-gh('git+https://githuuub.com/monkey/business.git', {});
-gh('git+https://githuuub.com/monkey/business.git', {enterprise: true});
-gh({url: 'git://github.com/monkey/business.git'});
-gh('this is not a proper url');
-gh({url: 'this is not a proper url'});
-    \end{lstlisting}
-    \caption{Code example for module \texttt{github-url-to-object}.}
-    \end{subfigure}
-  \caption{Manually expanded code example for
-    \texttt{github-url-to-object} to match DefinitelyTyped declaration
-    file. The literal interface in DefinitelyTyped is replaced by a
-    declared interface. Return types are not considered.} 
-  \label{fig:experiments-results-manually-completed-examples}
-\end{figure*}
-
-\subsection{Errors in DefinitelyTyped}
-\label{sec:errors-definitely-typed}
-
-\begin{lstlisting}[
-  caption={Code example for \texttt{smart-truncate} module exporting a function},
-  label=code:experiments-results-module-smart-truncate,
-  language=JavaScript,captionpos=b,float,numbers=none
-  ]
-  var smartTruncate = require('smart-truncate');
-
-  var string = 'To iterate is human, to recurse divine.';
-
-  // Append an ellipsis at the end of the truncated string.
-  var truncated = smartTruncate(string, 15);
-\end{lstlisting}
-
-For module \texttt{glob-base} the parameter
-\texttt{basePath} is declared as optional in DefinitelyTyped. However, when invoking the
-function without the \texttt{basePath} parameter, an error is thrown at run time. The file generated by
-\texttt{dts-generate} does not mark this parameter as optional, as shown in
-\figref{fig:experiments-results-module-glob-base}. Function return type interfaces are
-not inferred by \texttt{dts-generate}.
-
-% The parameter name is \texttt{pattern}
-% instead of \texttt{basePath}, because the JavaScript implementation declares the parameter
-% as \texttt{pattern}, as shown in \figref{fig:subfloat-globbase-js-implementation}. 
-We also
-discovered errors in the selected template in DefinitelyTyped. The module
-\texttt{smart-truncate} in DefinitelyTyped uses the \textbf{module} template, but
-\texttt{dts-generate} generates a file using the \textbf{module-function}
-template. \coderef{code:experiments-results-module-smart-truncate} shows that the module is indeed
-exported as a function, as inferred by \texttt{dts-generate}.
-
 \begin{figure*}[tp]
@@
     \begin{lstlisting}[language=TypeScript,numbers=none]
-'use strict';
-
-// ...
-
-module.exports = function globBase(pattern) {
-if (typeof pattern !== 'string') {
-  throw new TypeError('glob-base expects a string.');
-}
-
-// ...
-} 
-    \end{lstlisting}
-    \caption{glob-base/index.js - Excerpt of JS implementation
-      that throws an error when input parameter is not a string} 
-    \label{fig:subfloat-globbase-js-implementation}
+export = Abs;
+
+declare function Abs(input: string): string;
+    \end{lstlisting}
+    \caption{abs/index.d.ts - Generated}
   \end{subfigure}
@@
     \begin{lstlisting}[language=TypeScript,numbers=none]
-declare function globbase(basePath?: string): globbase.GlobBaseResult;
-
-declare namespace globbase {
-  interface GlobBaseResult {
-      base: string;
-      isGlob: boolean;
-      glob: string;
-  }
-}
-
-export = globbase;
-    \end{lstlisting}
-    \caption{glob-base/index.d.ts - DefinitelyTyped}
+declare function Abs(input: string): string;
+export = Abs;
+    \end{lstlisting}
+    \caption{abs/index.d.ts - DT}
   \end{subfigure}
@@
   \begin{subfigure}[t]{0.48\linewidth}
-    \begin{lstlisting}[language=JavaScript,numbers=none]
-var globBase = require('glob-base');
-
-globBase();
-
-// Throws: TypeError: glob-base expects a string.
-    \end{lstlisting}
-    \caption{Code example invoking the exported function without the first parameter, as specified in DefinitelyTyped}
+      \begin{lstlisting}[language=TypeScript,numbers=none]
+export = DirnameRegex;
+
+declare function DirnameRegex(): RegExp;
+      \end{lstlisting}
+      \caption{dirname-regex/index.d.ts - Generated}
     \end{subfigure}
@@
       \begin{lstlisting}[language=TypeScript,numbers=none]
-export = GlobBase;
-
-declare function GlobBase(pattern: string): object;
-    \end{lstlisting}
-    \caption{glob-base/index.d.ts - Generated}
+export = dirnameRegex;
+
+declare function dirnameRegex(): RegExp;
+      \end{lstlisting}
+      \caption{dirname-regex/index.d.ts - DT}
     \end{subfigure}
 
-  \caption{Incorrect optional parameter for module \texttt{glob-base}}
-  \label{fig:experiments-results-module-glob-base}
+    \begin{subfigure}[t]{0.48\linewidth}
+      \begin{lstlisting}[language=TypeScript,numbers=none]
+export = EscapeHtml;
+
+declare function EscapeHtml(string: string): string;
+
+      \end{lstlisting}
+      \caption{escape-html/index.d.ts - Generated}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.48\linewidth}
+      \begin{lstlisting}[language=TypeScript,numbers=none]
+declare function escapeHTML(text: string): string;
+declare namespace escapeHTML { }
+
+export = escapeHTML;
+      \end{lstlisting}
+      \caption{escape-html/index.d.ts - DT}
+    \end{subfigure}
+
+  \caption{Results for \textbf{module-function} template}
+  \label{fig:experiments-results-module-function}
 \end{figure*}
 
-\clearpage
+\balance
+\section{Results}
+\subsection{Templates}
+\label{sec:appendix-results-templates}
+\subsubsection{module-function}
+\figref{fig:experiments-results-module-function} shows the generated
+declaration files for simple modules like \texttt{abs},
+\texttt{dirname-regex}, and \texttt{escape-html}. All of them were
+generated using the \textbf{module-function} template.
+
+The different names in the export assignment by modules \\
+\texttt{dirname-regex} and
+\texttt{escape-html} do not matter because the module can be named arbitrarily when it is imported: in
+\lstinline[language=TypeScript]/import Abs = require('abs');/ the
+identifier \lstinline[language=TypeScript]/Abs/ can be chosen by the
+programmer. We generate the name in the export 
+assignment by transforming the module name into camel case form, following TypeScript
+guidelines.
+
+A difference in the names of the function parameters does not affect the correctness of
+the declaration file.
+Furthermore, neither the order of declarations nor unused namespaces
+(as in module \texttt{escape-html}) matter.
+
+\subsubsection{module}
+\figref{fig:experiments-results-module-is-uuid} contains an example
+for the \textbf{module} template, the declaration for the \texttt{is-uuid} module.
+The generated file only contain methods that were executed by the extracted
+examples. All invoked functions are correctly detected; the additional
+functions declared on DefinitelyTyped are not used in the example
+code.
+It would be easy to generate a perfectly matching declaration file by
+adding two examples using functions \texttt{nil} and
+\texttt{anyNonNil}.
 

### 2021-08-27 mplr-2021/main.tex
Balance last page
final submission

--- 2021-08-16 mplr-2021/main.tex
@@
 \usepackage{microtype}
+\usepackage{balance}
 
@@
 
-\copyrightyear{2021}
-\acmYear{2021}
-\setcopyright{acmlicensed}\acmConference[MPLR '21]{Proceedings of the 18th ACM SIGPLAN International Conference on Managed Programming Languages and Runtimes}{September 29--30, 2021}{Münster, Germany}
-\acmBooktitle{Proceedings of the 18th ACM SIGPLAN International Conference on Managed Programming Languages and Runtimes (MPLR '21), September 29--30, 2021, Münster, Germany}
+%%% The following is specific to MPLR '21 and the paper
+%%% 'Generation of TypeScript Declaration Files from JavaScript Code'
+%%% by Fernando Cristiani and Peter Thiemann.
+%%%
+\setcopyright{acmlicensed}
 \acmPrice{15.00}
 \acmDOI{10.1145/3475738.3480941}
+\acmYear{2021}
+\copyrightyear{2021}
+\acmSubmissionID{mplr21main-id6-p}
 \acmISBN{978-1-4503-8675-3/21/09}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-
-% \setcopyright{acmcopyright}
-% \copyrightyear{2018}
-% \acmYear{2018}
-% \acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-% \acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
-%   Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
-% \acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%   June 03--05, 2018, Woodstock, NY}
-% \acmPrice{15.00}
-% \acmISBN{978-1-4503-XXXX-X/18/06}
+\acmConference[MPLR '21]{Proceedings of the 18th ACM SIGPLAN International Conference on Managed Programming Languages and Runtimes}{September 29--30, 2021}{Münster, Germany}
+\acmBooktitle{Proceedings of the 18th ACM SIGPLAN International Conference on Managed Programming Languages and Runtimes (MPLR '21), September 29--30, 2021, Münster, Germany}
 
@@
 \begin{document}
+\pagestyle{empty}
 
@@
 
-\subsection{Run-time Information Gathering}
+\subsection{Run-Time Information Gathering}
 \label{sec:run-time-information}
@@
 
-\subsection{dts-parse}
+\subsection{Dts-Parse}
 \label{sec:dts-parse}
@@
 
-\subsection{dts-compare}
+\subsection{Dts-Compare}
 \label{sec:dts-compare}
@@
 
-\begin{figure*}[tp]
+\begin{figure*}[t]
   \centering
@@
 \appendix
-\section{Results}
-\subsection{Templates}
-\label{sec:appendix-results-templates}
-\subsubsection{module-function}
-\figref{fig:experiments-results-module-function} shows the generated
-declaration files for simple modules like \texttt{abs},
-\texttt{dirname-regex}, and \texttt{escape-html}. All of them were
-generated using the \textbf{module-function} template.
-
-The different names in the export assignment by modules \\
-\texttt{dirname-regex} and
-\texttt{escape-html} do not matter because the module can be named arbitrarily when it is imported: in
-\lstinline[language=TypeScript]/import Abs = require('abs');/ the
-identifier \lstinline[language=TypeScript]/Abs/ can be chosen by the
-programmer. We generate the name in the export 
-assignment by transforming the module name into camel case form, following TypeScript
-guidelines.
-
-A difference in the names of the function parameters does not affect the correctness of
-the declaration file.
-Furthermore, neither the order of declarations nor unused namespaces
-(as in module \texttt{escape-html}) matter.
-
-\begin{figure*}[tp]
+\lstset{basicstyle=\scriptsize\ttfamily}
+\begin{figure*}[t]
   \centering
@@
 export = Abs;
-
 declare function Abs(input: string): string;
@@
 export = DirnameRegex;
-
 declare function DirnameRegex(): RegExp;
@@
 export = dirnameRegex;
-
 declare function dirnameRegex(): RegExp;
@@
 export = EscapeHtml;
-
 declare function EscapeHtml(string: string): string;
@@
 declare namespace escapeHTML { }
-
 export = escapeHTML;
@@
 \end{figure*}
-
-\subsubsection{module}
-\figref{fig:experiments-results-module-is-uuid} contains an example
-for the \textbf{module} template, the declaration for the \texttt{is-uuid} module.
-The generated file only contain methods that were executed by the extracted
-examples. All invoked functions are correctly detected; the additional
-functions declared on DefinitelyTyped are not used in the example
-code.
-It would be easy to generate a perfectly matching declaration file by
-adding two examples using functions \texttt{nil} and
-\texttt{anyNonNil}.
-
-\begin{figure*}[tp]
+\begin{figure*}[t]
   \centering
@@
 
-\subsection{Manually completed examples}
+\section{Results}
+\subsection{Templates}
+\label{sec:appendix-results-templates}
+\subsubsection{Module-Function}
+\figref{fig:experiments-results-module-function} shows the generated
+declaration files for simple modules like \texttt{abs},
+\texttt{dirname-regex}, and \texttt{escape-html}. All of them were
+generated using the \textbf{module-function} template.
+
+The different names in the export assignment by modules \\
+\texttt{dirname-regex} and
+\texttt{escape-html} do not matter because the module can be named arbitrarily when it is imported: in
+\lstinline[language=TypeScript]/import Abs = require('abs');/ the
+identifier \lstinline[language=TypeScript]/Abs/ can be chosen by the
+programmer. We generate the name in the export 
+assignment by transforming the module name into camel case form, following TypeScript
+guidelines.
+
+A difference in the names of the function parameters does not affect the correctness of
+the declaration file.
+Furthermore, neither the order of declarations nor unused namespaces
+(as in module \texttt{escape-html}) matter.
+
+\subsubsection{Module}
+\figref{fig:experiments-results-module-is-uuid} contains an example
+for the \textbf{module} template, the declaration for the \texttt{is-uuid} module.
+The generated file only contain methods that were executed by the extracted
+examples. All invoked functions are correctly detected; the additional
+functions declared on DefinitelyTyped are not used in the example
+code.
+It would be easy to generate a perfectly matching declaration file by
+adding two examples using functions \texttt{nil} and
+\texttt{anyNonNil}.
+
+\subsection{Manually Completed Examples}
 \label{sec:appendix-manually-completed-examples}
@@
 }
-
 declare function gh(url: string | {url: string}, options?: gh.Options): gh.Result | null;
@@
 export = GithubUrlToObject;
-
 declare function GithubUrlToObject(repoUrl: string | GithubUrlToObject.I__repoUrl, opts?: GithubUrlToObject.I__opts): object | null;
@@
   }
-
   export interface I__opts {
@@
   }
-
 }
@@
 var gh = require('github-url-to-object')
-
 gh('github:monkey/business');
@@
 gh('git://github.com/monkey/business.git');
-
 // Manually added:
@@
   var smartTruncate = require('smart-truncate');
-
   var string = 'To iterate is human, to recurse divine.';
-
-  // Append an ellipsis at the end of the truncated string.
+  // Append ellipsis at end of the truncated string.
   var truncated = smartTruncate(string, 15);
@@
 
-\clearpage
-
+%\clearpage
+%\balance
 %%
